### 2018-07-26 digitisation.tex
first commit

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[apj]{emulateapj}
%\documentclass[preprint2]{aastex61}
%\documentclass[12pt,preprint]{aastex}
\graphicspath{{figures/}}
\DeclareGraphicsExtensions{.jpg,.pdf,.png,.eps,.ps}

\usepackage[table,usenames,dvipsnames]{xcolor}
%\usepackage{amsmath}
%\usepackage{subfigure}
\usepackage[backref,breaklinks,colorlinks,citecolor=blue]{hyperref}
\usepackage{natbib}
%\usepackage{natbib}
\bibliographystyle{fapj}
%\usepackage{graphicx}
%\usepackage{multirow}
\usepackage{soul}

%\newcommand{\jcap}{JCAP}

\newcommand{\sqdeg}{deg$^2$ }
\newcommand{\omb}{\ensuremath{\Omega_b h^2}}
\newcommand{\omc}{\ensuremath{\Omega_c h^2}}
\newcommand{\clpp}{\ensuremath{C_{L}^{\phi\phi}}}
\newcommand{\cpmf}{\ensuremath{C_{\ell}^{\rm PMF}}}

\newcommand{\cpmftens}{\ensuremath{C_{\ell}^{\rm PMF,\,tens}}}
\newcommand{\cpmfvec}{\ensuremath{C_{\ell}^{\rm PMF,\,vec}}}
\newcommand{\apmf}{\ensuremath{A_{\rm PMF}}}
\newcommand{\bpmf}{\ensuremath{B_{\rm 1\,Mpc}}}
\newcommand{\alens}{\ensuremath{A_{\rm lens}}}
\newcommand{\lcdm}{\ensuremath{\Lambda}CDM}
\newcommand{\nrun}{\ensuremath{n_{\rm run}}}
\newcommand{\neff}{\ensuremath{N_{\rm eff}}}
\newcommand{\ho}{H\ensuremath{_0}}
\newcommand{\mnu}{\ensuremath{\sum m_\nu}}
\newcommand{\ukarcmin}{\ensuremath{\mu}{\rm K-arcmin}}
\newcommand{\lknee}{\ensuremath{\ell_{\rm knee}}}
\newcommand{\fermilat}{\textit{Fermi}-LAT}

\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\planck}{{\sl Planck}}
\newcommand{\wmap}{{\sl WMAP}}
\newcommand{\bicepkeck}{BICEP2/Keck Array}
\newcommand{\sptnew}{SPT-3G}
\newcommand{\pb}{\textsc{Polarbear}}
\newcommand{\simons}{Simons Array}
\newcommand{\sptpol}{SPTpol}
\newcommand{\advactpol}{Adv.~ACTpol}

\newcommand{\tbd}[1]{\textcolor{Red}{{\bf TBD}: #1}}
\newcommand{\gab}[1]{\textcolor{Orchid}{[{\bf GS}: #1]}}
\newcommand{\changed}[1]{\textcolor{Red}{#1}}
\newcommand{\removed}[1]{\textcolor{Red}{}}
\include{number_list}

%

% ref to section \S\ref{sec:label}

%\submitjournal{ApJ}
\def\Melbourne{1}
\def\uci{2}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\title{Digitisation}
\author{L.~Balkenhol\altaffilmark{\Melbourne} and C.~L.~Reichardt\altaffilmark{\Melbourne}}
\altaffiltext{\Melbourne}{School of Physics, University of Melbourne, Parkville, VIC 3010, Australia}
\email{christian.reichardt@unimelb.edu.au}

\begin{abstract}

It's awesome!
\end{abstract}

\keywords{ cosmic background radiation --- polarization }
\section{Introduction}
\label{sec:intro}

CMB is great; One reason is that there is a history of compression and computational techniques that reduce the load of large datasets. ie maps; bandpowers; pseudo-cls.

satellites have also reduced bits on TOD; ground based haven't had to yet

however as we discuss building ever larger arrays at remote sites, we are starting to be limited: spt example.

in this work we present digitisation for ground-based cmb polarisation measurements.
teaser results

The outline of this paper is as follows. 
We present the digisation schemes in \S\ref{sec:dig}, and their performance in \S\ref{sec:results}
We summarize our findings in \S\ref{sec:conclusions}.

\section{Digitisation}
\label{sec:dig}

overview

\subsection{Problem}
\label{subsec:problem}

why is it worth considering

\subsection{Methods}
\label{subsec:method}

\begin{figure*}[htb]\centering
\includegraphics[width=0.9\textwidth,clip,trim={1.5cm 12.5cm 5cm 3.8cm}]{pretty.pdf}
  \caption[Current ]{
  Current 
           \label{fig:ig}
  }
\end{figure*}

\section{Results}
\label{sec:results}

\begin{table}[tbh]
\begin{center}
\caption{\label{tab:noise} Noise levels}
\small
\begin{tabular}{l | c c c }
Model   & XX&XX&$r$\\
\hline

\end{tabular}
\tablecomments{ 
go go
} \normalsize
\end{center}
\end{table}

\begin{table*}[tbh]
\begin{center}
\caption{\label{tab:experiments} Assumed survey parameters}
\small
\begin{tabular}{l || c c c c c }
Experiment & Sky coverage & Polarized Noise level  & 1/$f$ knee & Beam FWHM \\
& &($\mu$K-arcmin)&&(arcmin.)\\
\hline
\tiny \\ \small
CMB Stage III & & & & \\
~~~~~SPT-3G & 6\% & 3.0 & 200 & 1.2 \\
~~~~~Simons Array & 36\% & 9.5 & 200 & 3.5 \\ 
\tiny \\ \small
%\hline
CMB Stage IV & 55\% & 1.3 & 100 & 4.0 \\
\end{tabular}
\tablecomments{ 
Key numbers about the planned stage III and IV experiments. 
The sky coverage percentages are after galactic cuts. 
Unless otherwise noted,  the Fisher matrix forecasts in this work use these numbers. 
All forecasts also allow for beam and calibration uncertainties as noted in the text. 
} \normalsize
\end{center}
\end{table*}

\section{Conclusions}
\label{sec:conclusions}
\acknowledgments

We thank the \changed{referee as well as} Srinivasan Raghunathan and Federico Bianchini for valuable feedback on the manuscript. 
We acknowledge support from an Australian Research Council Future Fellowship (FT150100074), and also from the University of Melbourne. 
This research used resources of the National Energy Research Scientific Computing Center, which is supported by the Office of Science of the U.S. Department of Energy under Contract No. DE-AC02-05CH11231. 
We acknowledge the use of the Legacy Archive for Microwave Background Data Analysis (LAMBDA). Support for LAMBDA is provided by the NASA Office of Space Science.

\bibliography{digitisation}

\end{document}

### 2018-08-02 digitisation.tex
Structure and Fillers
Wrote more detailed structure and filled in key ideas to be communicated in each paragraph.

@@
 
-\begin{abstract}
-
-It's awesome!
+\begin{abstract} % copied off ASA for now
+Observations of the Cosmic Microwave Background (CMB) are of immense value to modern cosmology.
+However, future CMB experiments must confront challenges in mission planning, hardware and analysis that arise from the sheer size of the time-ordered-data being recorded.
+These challenges are particularly significant for Antarctic and satellite experiments which depend on satellite links to transmit the data.
+We investigate using extreme digitisation to address these challenges.
+Unlike lossless compression, extreme digitisation introduces additional noise into the data.
+We present an optimal 1, 2 and 3 bit digitisation schemes and lay out how the added noise varies between the schemes and as a function of hits per pixel in the map for temperature and polarisation observations.
+We show that the noise penalty of 3 bit digitisation is at the percent-level for CMB power spectrum measurements.
+This is impressive considering that it would reduce the data volume by an order of magnitude.
+We argue that extreme digitisation is a promising strategy for upcoming experiments.
 \end{abstract}
@@
 
-CMB is great; One reason is that there is a history of compression and computational techniques that reduce the load of large datasets. ie maps; bandpowers; pseudo-cls.
+Observations of the Cosmic Microwave Background (CMB) have played a key role since 1964 (discovery paper).
+Current and future CMB experiments will continue to deliver new insights by studying the temperature and polarisation information contained in the CMB.
+These will constrain help putting tighter constraints on cosmological models.
+The most prominent science goal is the discovery of the imprint left by inflationary gravitational waves.
+Additionally studies of CMB lensing and the SZ effects will deliver new insight.
+CMB experiments however also provide a valuable counterpart to ground-based particle physics experiments, as they probe the relativisitc number of species, the helium fraction and the neutrino mass sum.
 
-satellites have also reduced bits on TOD; ground based haven't had to yet
+The outstanding contribution of CMB science to modern physics has demanded a high standard of data analysis.
+The CMB community has developed a variety of compression and computational techniques to manage the increasing influx of data, while maximising the science output.
+These include the compression of time-ordered data into maps, bandpower estimation and the pseudo Cl method.
 
-however as we discuss building ever larger arrays at remote sites, we are starting to be limited: spt example.
+A growing hurdle for experiments at remote locations are the transmission limitations of satellite links.
+Space-based experiments have employed a combination of lossless and lossy compression techniques, including reduced bits in the time-ordered-data (TOD).
+Antarctica based experiments that transmit a portion of their data via a satellite link have downsampled their TOD in the past to meet their telemetry requirements.
+They have yet to exploit few bit digitisation of the TOD.
 
-in this work we present digitisation for ground-based cmb polarisation measurements.
-teaser results
+As we approach the next generation of ground-based experiments, Stage-4, and the launch of a new generation of space-based missions (liteBIRD, PIXIE, COrE+), we must treat the transmission bottleneck carefully.
+Without a review of the current compression techniques employed we are sure to lose information.
 
-The outline of this paper is as follows. 
-We present the digisation schemes in \S\ref{sec:dig}, and their performance in \S\ref{sec:results}
+In this work we present the method of extreme, i.e.
+few-bit digitisation, and detail its effect when applied to the TOD.
+To highlight the consequences we investigate temperature and polarisation powerspectra.
+We find that an optimal 3-bit digitisatio scheme adds as little as $\sim 2\%$ to the map noise level.
+While the digtisation schemes described here are primarily laid out for ground-based Stage-4 experiments, future space-based missions must employ a similar digitisation as a step of their compression algorithm.
+
+This work is structured as follows.
+We present the digitisation schemes in \S\ref{sec:dig}, by formulating them mathematically, motivating their use and describing the framework employed to test their performance.
+In \S\ref{sec:results} we present the results obtained.
 We summarize our findings in \S\ref{sec:conclusions}.
+
+%CMB is great; One reason is that there is a history of compression and computational techniques that reduce the load of large datasets. ie maps; bandpowers; pseudo-cls.
+
+%satellites have also reduced bits on TOD; ground based haven't had to yet
+
+%however as we discuss building ever larger arrays at remote sites, we are starting to be limited: spt example.
+
+%in this work we present digitisation for ground-based cmb polarisation measurements.
+%teaser results
+
+%The outline of this paper is as follows. 
+%We present the digisation schemes in \S\ref{sec:dig}, and their performance in \S\ref{sec:results}
+%We summarize our findings in \S\ref{sec:conclusions}.
 
@@
 
-overview
-
 \subsection{Problem}
@@
 
-why is it worth considering
+%Data influx + Transmission
+
+The science goals of upcoming CMB experiments naturally lead to a large influx of data.
+To achieve them longer observations with more detectors are needed.
+However, the best observations sites for CMB observations are at remote locations: in space and in the Antarctica.
+Experimetns at these locations depend on satellite transmission The next generation of ground-based experiments, Stage-4, will aim to collect 2 million detector years worth of data.
+A Stage-4 style experiment will face a data influx of $\sim O(10)Tb/d$.
+However, the current transmission allocation for the South Pole Telescope (SPT) is at $150Gb/d$.
+While this will likely see an increase going into the enxt stage of experimetns it will not be able to over the entire observation.
+In fact we expect that compression rates for transmission must increase by an order of magnitude.
+Similarly space-based missions aim to increase their detector count by at least an order of magnitude compared to PLANCK.
+It is questionable whether their telemetry specifications will allow for transmission of the data with PLANCK-style compression.
+
+%Hardware (other than maybe storage cost for satellites?)
+
+Telemetry drawbacks can be sought to be combated by increasing storage capacity. However, storage space becomes a considerable price point when considering the amount of data to be recorded.
+
+%Planning
+
+Beyond these transmission hurdles detailed mission planning is becoming exceedingly expensive.
+As noted by (S4 science book) a full simulation of TOD over the entire parameter space of detection scenarios for numerous set-ups is the desired way to plan Stage-4 configuration.
+Similarly space-based missions must carry out a similar analysis to optimise their science out-put.
+However, given the sheer size of TOD this is not possible.
+We must turn to different planning strategies or aim to reduce the size of the TOD in order to maximise the productivity of planning and development stages.
+
+Operations on the TOD, such as noise-removal or map-making are a vital part of CMB data analysis. Through the expontentially growing size of TOD CMB data analysis is becoming increasingly expensive.
+
+%why is it worth considering
+
+Extreme Digitisation would tackle the challenges mentioned above by recuding the size of the TOD by an order of magnitude.
+Together with already exploited lossless compression techniques (such as FLAC) this will directly tackle transmission hurdles.
+While it needs to be investigated to what extent existing algorithms can be carried over, extreme digitisation has the possiblity of solving planning and analysis problems.
+
+\subsection{Extreme Digitisation}
+\label{subsec:extremedigitisation}
+
+Digitisation is a lossy compression technique.
+However the induced noise depends on the number of bits used as well as the the digitisation thresholds and output levels chosen.
+It is possible to calculate the an optimal set of digitisation thresholds and output levels for any number of bits  given the nature of the input signal.
+This was laid out by Max in 1978.
+
+For 1 bit digitisation we apply the sign function to the TOD
+
+\[ \hat{x}(t) = \left\{ \begin{array}{lr}
+1, & \text{for } x(t) > 0\\
+-1, & \text{for } x(t) \leq 0
+\end{array} \right. \]
+
+For 2 bit digitisation we apply the optimal digitisation scheme suggested by Max as
+
+\[ \hat{x}(t) = \left\{ \begin{array}{lr}
+1.51 \sigma, & \text{for } x(t) \geq 0.9816 \sigma\\
+0.4528 \sigma, & \text{for } 0 \leq x(t) < 0.9816 \sigma\\
+-0.4528 \sigma, & \text{for } 0.9816 \sigma \leq x(t) < 0\\
+-1.51 \sigma, & \text{for } 0.9816 \sigma < x(t)\\
+\end{array} \right. \]
+
+Finally the optimal 3 bit digitisation scheme is
+
+\[ \hat{x}(t) = \left\{ \begin{array}{lr}
+2.152 \sigma, & \text{for } x(t) \geq 1.748 \sigma\\
+1.344 \sigma, & \text{for } 1.05 \sigma \leq x(t) < 1.748 \sigma\\
+0.756 \sigma, & \text{for } 0.501 \sigma \leq x(t) < 1.05 \sigma\\
+0.245 \sigma, & \text{for } 0 \leq x(t) < 0.501 \sigma\\
+-0.245 \sigma, & \text{for } 0.501 \sigma \leq x(t) < 0\\
+-0.756 \sigma, & \text{for } 1.05 \sigma \leq x(t) < 0.501 \sigma\\
+-1.344 \sigma, & \text{for } 1.748 \sigma \leq x(t) < 1.05 \sigma\\
+-2.152 \sigma, & \text{for } 1.748 \sigma < x(t)\\
+\end{array} \right. \]
+
+% Theoretical lvls for gaussian white noise
 
@@
 
-\begin{figure*}[htb]\centering
-\includegraphics[width=0.9\textwidth,clip,trim={1.5cm 12.5cm 5cm 3.8cm}]{pretty.pdf}
-  \caption[Current ]{
-  Current 
-           \label{fig:ig}
-  }
-\end{figure*}
+Template maps:
+
+To investigate the performance of the aforementioned digitisation schemes we simulate many scans over CMB template maps at the timstream level.
+We begin by creating said template maps by creating a realisation of one $I, Q, U$ map, for which we use the power spectrum provieded by the Planck collaboration XXX.
+The underlying cosmological parameters are summarised in table \ref{tab:inputcosparams}.
+
+Scan strategy:
+
+These template maps are formulated in the Healpix framework.
+We then proceed to simulate constant elevation scans (CES) over a square patch of sky using a single detector.
+The scan speed is tweaked to match a desired number of hits-per-pixel.
+We allow ourselves to repeat the entire scan 100 times slightly offset in RA and DEC each time, so that we may hit all pixels roughly uniformly.
+We assume that the noise is purely gaussian white noise.
+The simulation parameters are summarised in table \ref{tab:modelparams}.
+
+Noise addition and digitisation:
+
+While performing each CES the appropriate pixels that are being targeted are determined.
+The corresponding values from the template maps are then accessed and added to realisations of the detector noise of appropriate length.
+The constructed timestream for each channels is then compressed in 4 different ways: averaging of the timestream, respecting each datapoint as a 64bit floating point number and applying the introduced digitisation schemes before averaging the timestream, again withing a 64bit floating point framework.
+Through this procedure we slowly build up 4 different maps: one map that corresponds to a normal scan over the CMB and 3 maps that reconstruct the observed path from extremely reduced time-ordered data.
+
+Compression:
+
+To obtain maps from the digitised time-stream we simply use our knowledge of the scan strategy to recognise what pixel each recorded point hits.
+All hits of a pixel are then simply averaged to obtain the pixel value.
+
+\begin{table}[tbh]
+\begin{center}
+\caption{\label{tab:inputcosparams} Input Cosmological Parameters}
+\small
+\begin{tabular}{l | c c c }
+$\Omega$ & XX&XX&$r$\\
+\hline
+
+\end{tabular}
+\tablecomments{ 
+comments
+} \normalsize
+\end{center}
+\end{table}
+
+\begin{table}[tbh]
+\begin{center}
+\caption{\label{tab:modelparams} Model Parameters}
+\small
+\begin{tabular}{c c c c c}
+$\mathrm{NSIDE}$ & $f_{readout}$ & $f_{sky}$ & $\sigma^{\mathrm{T}}_{detector}$ & $\sigma^{\mathrm{Pol}}_{detector}$\\
+\hline
+$4096$ & $200\mathrm{Hz}$ & $\sim 0.014$ & $500\mathrm{K}\sqrt{\mathrm{s}}$ & $\sqrt{2}\times500\mathrm{K}\sqrt{\mathrm{s}}$\\
+
+\end{tabular}
+\tablecomments{ 
+comments
+} \normalsize
+\end{center}
+\end{table}
+
+%\begin{figure*}[htb]\centering
+%\includegraphics[width=0.9\textwidth,clip,trim={1.5cm 12.5cm 5cm 3.8cm}]{pretty.pdf}
+%  \caption[Current ]{
+%  Current 
+%           \label{fig:ig}
+%  }
+%\end{figure*}
 
@@
 \label{sec:results}
+
+Get power spectra
+
+Analysis procedure
+
+Analysis results: Nbits to map noise
+
+Analysis results: Nhits per pixel to map noise
+
+Analysis result: Noise disappears in cross-spectra -> actually cannot check this in current framework.
+
+Any other results that are outstanding
+
+? Analysis result: To uncertainty in cosmological parameters?
+
+This is why it works well: low S/N many hits per pixel, jumps in CMB noise level are order of magnitude.
 
@@
 \label{sec:conclusions}
+
+In this work we have demonstrated the power of digitising CMB TOD from 64-bit down to few-bits.
+The additional noise induced through this step was at the percent-level for chosen optimal digitisation schemes for temperature and polarisation observations.
+
+The motivation behind employing extreme digitisation is the reduction of TOD by an order of magnitude.
+This primarily addresses challenges in data transmission that will face upcoming CMB experiments at remote locations.
+Benefits in planning and analysis seem feasible.
+
+Future work on investigating this compression technique must aim to understand the nature of the induced noise better. For this the performance of cluster-finding algorithms is interesting.
+
+While the X added noise is already impressive it should be laid out how this changes when moving to a more realistic noise profile.
+
 \acknowledgments

### 2018-08-03 digitisation.tex
Added theoretical digitisation levels, editing
Added the theoretical digitisation threshold/levels calculation, mostly as presented by Max/Patrick. Edited bits here and there.

@@
 
-\title{Digitisation}
+\title{Constructing Cosmic Microwave Background Power Spectra from Few Bit Timestreams}
 \author{L.~Balkenhol\altaffilmark{\Melbourne} and C.~L.~Reichardt\altaffilmark{\Melbourne}}
@@
 The most prominent science goal is the discovery of the imprint left by inflationary gravitational waves.
-Additionally studies of CMB lensing and the SZ effects will deliver new insight.
+Additionally studies of CMB lensing and the Sunyaev-Zeldovich (SZ) effects will deliver new insight.
 CMB experiments however also provide a valuable counterpart to ground-based particle physics experiments, as they probe the relativisitc number of species, the helium fraction and the neutrino mass sum.
@@
 
-In this work we present the method of extreme, i.e.
-few-bit digitisation, and detail its effect when applied to the TOD.
-To highlight the consequences we investigate temperature and polarisation powerspectra.
-We find that an optimal 3-bit digitisatio scheme adds as little as $\sim 2\%$ to the map noise level.
-While the digtisation schemes described here are primarily laid out for ground-based Stage-4 experiments, future space-based missions must employ a similar digitisation as a step of their compression algorithm.
+In this work we present the method of extreme digitisation, which compresses a rich digital input signal and compresses it into a few bits.
+We apply extreme digitisation to the TOD and detail its effect on temperature and polarisation observations.
+We find that an optimal 3-bit digitisation scheme adds as little as $\sim 2\%$ to the map noise level.
+While the digtisation schemes described here are primarily laid out for ground-based Stage-4 experiments, future space-based missions must inevitably study lossy compression techniques.
 
 This work is structured as follows.
-We present the digitisation schemes in \S\ref{sec:dig}, by formulating them mathematically, motivating their use and describing the framework employed to test their performance.
+We detail the arising challenges in handling large TOD in \S\ref{subsec:problem}.
+We subsequently formulate extreme digitisation formerly in \S\ref{subsec:extremedigitisation} and lay out the framework used to test the performance of this compression technique in \S\ref{subsec:method}.
 In \S\ref{sec:results} we present the results obtained.
@@
 A Stage-4 style experiment will face a data influx of $\sim O(10)Tb/d$.
-However, the current transmission allocation for the South Pole Telescope (SPT) is at $150Gb/d$.
-While this will likely see an increase going into the enxt stage of experimetns it will not be able to over the entire observation.
-In fact we expect that compression rates for transmission must increase by an order of magnitude.
-Similarly space-based missions aim to increase their detector count by at least an order of magnitude compared to PLANCK.
+However, the current transmission allocation for the SPT3G is at $150Gb/d$.
+SPT3G currently deals with its transmission bottleneck by downsampling the TOD, i.e.
+discarding high frequency information.
+While this will likely see an increase going into the next stage of experimetns it will not be able to over the entire observation.
+Going into Stage-4 we anticipate that compression rates for transmission must increase by an order of magnitude.
+Continued use of heavy downsampling will narrow the informatino window decisively - prohibiting high l science to be carried out on the transmitted dataset.
+This also means that any potential faults or errors in the experiment that only become visible in high frequencies will go unnoticed for longer.
+
+Similarly, future space-based missions (PIXIE, COrE+, liteBIRD) aim to increase their detector count by at least an order of magnitude compared to PLANCK.
 It is questionable whether their telemetry specifications will allow for transmission of the data with PLANCK-style compression.
-
-%Hardware (other than maybe storage cost for satellites?)
-
-Telemetry drawbacks can be sought to be combated by increasing storage capacity. However, storage space becomes a considerable price point when considering the amount of data to be recorded.
+Methods of storing large amounts of data on upcoming satellites will likely be prohibited by financial decisions: the amount of storage space required becomes financially relevant at the scales targeted.
 
@@
 
+% rephrase these, basically off Max1978
+
 Digitisation is a lossy compression technique.
-However the induced noise depends on the number of bits used as well as the the digitisation thresholds and output levels chosen.
-It is possible to calculate the an optimal set of digitisation thresholds and output levels for any number of bits  given the nature of the input signal.
-This was laid out by Max in 1978.
+However the induced noise depends on the number of bits used, the digitisation thresholds and the output levels chosen.
+To minimise the noise induced through this process one must know the nature of input signal.
+A theoretical framework to obtain these levels was laid out by Max in 1978.
+We review the key aspects of his work relevant to our findings below.
 
-For 1 bit digitisation we apply the sign function to the TOD
+We divide the space $-\infty \to \infty$ into $N$ ranges.
+One then has endpoints $x_k$ and output levels $y_k$, such that an input between $x_i$ and $x_{i+1}$ produces an output at $y_i$.
+Conventionally one chooses $x_{1} = -\infty$ and $x_{N+1} = \infty$.
+In order to determine an optimal set of $x_k$ and $y_k$ we would like to minimise the distortion
+
+\[ D = \left\langle  \left( s_{in} - s_{out} \right)^2 \right\rangle \]
+
+For a given input amplitude probability density $p(x)$ we may rewrite the above as
+
+\[ D = \sum_{i = 1}^N \int_{x_i}^{x_{i+1}} \left(x-y_i\right)^2 p(x) dx \]
+
+Seeing as we wish to minimise the distortion we differentiate the above with respect to $x_i$ and $y_i$ and set the derivatives to zero.
+
+\[ \frac{\partial D}{\partial x_i} = \left(x_i-y_{i-1}\right)^2 p(x_i) - \left(x_i - y_i\right)^2 p(x_i) = 0 \]
+
+\[ \frac{\partial D}{\partial x_j} = -2 \int_{x_i}^{x_{i+1}} \left( x-y_i \right) p(x) dx = 0 \]
+
+From the first equation above we deduce that
+
+\begin{equation} \label{eq:digitequalspacecondition}
+x_i = \frac{y_i+y_{i+1}}{2}
+\end{equation}
+
+The second one
+
+\begin{equation} \label{eq:digitareacondition}
+\int_{x_i}^{x_{i+1}} \left( x-y_i \right) p(x) dx = 0
+\end{equation}
+
+implies that we should choose $y_i$, such that it halves the area underneath $p(x)$ in the interval $x_i \to x_{i+1}$.
+
+We are interested in the case where our input signal can fairly be approximated by gaussian white noise, i.e.
+$p(x) = 1/sqrt{2\pi} e^{-x^2/2}$.
+Even with these assumptions we cannot solve the resulting equations analytically.
+Instead one can begin by picking $y_1$ and calculating the remaining $x_i$ and $y_i$ using equation \ref{eq:digitequalspacecondition}.
+One then observes weather this choice of values satisfy the conditions given by equation \ref{eq:digitareacondition}.
+If that is the case the $x_k$ and $y_k$ were chosen appropriately.
+This is unlikely to succeed on the first try.
+We rather solve the above system iteratively.
+
+This was carried out by Max. We incorporate his results by formulating the multi-level functions we use for our 1, 2 and 3 bit digitisation process. For 1 bit digitisation we apply the sign function
 
@@
 
-For 2 bit digitisation we apply the optimal digitisation scheme suggested by Max as
+to the TOD. For 2 bit digitisation we apply the four-level function
 
@@
 
-Finally the optimal 3 bit digitisation scheme is
+Finally the optimal 3 bit digitisation is described by the eight-level function
 
@@
 
+Summary:
+
+To investigate the performance of the aforementioned digitisation schemes we simulate many scans over CMB template maps at the timestream level, where we add detector noise, for which we use gaussian white noise .
+We obtain maps that use 64bit TOD and maps that have undergone 1, 2 and 3 bit digitisatino at the timestream level.
+We obtain the $TT, EE, BB, TE$ powerspectra of each map.
+Through comparison we then determine the additional noise induced through the extreme digitisation process.
+
 Template maps:
 
-To investigate the performance of the aforementioned digitisation schemes we simulate many scans over CMB template maps at the timstream level.
-We begin by creating said template maps by creating a realisation of one $I, Q, U$ map, for which we use the power spectrum provieded by the Planck collaboration XXX.
+The $I, Q, U$ template maps used were created through healpy's synfast.
+We rely on the power spectrum provieded by the Planck collaboration XXX.
 The underlying cosmological parameters are summarised in table \ref{tab:inputcosparams}.
@@
 
-These template maps are formulated in the Healpix framework.
-We then proceed to simulate constant elevation scans (CES) over a square patch of sky using a single detector.
-The scan speed is tweaked to match a desired number of hits-per-pixel.
-We allow ourselves to repeat the entire scan 100 times slightly offset in RA and DEC each time, so that we may hit all pixels roughly uniformly.
-We assume that the noise is purely gaussian white noise.
+We simulate observing a $\sim 600$ \sqdeg path of the sky.
+To do so we perform constant elevation scans over the observation region.
+We repeat the observation strategy $100$ times with a slight offset in RA and DEC each time, such that all pixels within the path are hit roughly uniformly.
 The simulation parameters are summarised in table \ref{tab:modelparams}.
@@
 The corresponding values from the template maps are then accessed and added to realisations of the detector noise of appropriate length.
-The constructed timestream for each channels is then compressed in 4 different ways: averaging of the timestream, respecting each datapoint as a 64bit floating point number and applying the introduced digitisation schemes before averaging the timestream, again withing a 64bit floating point framework.
-Through this procedure we slowly build up 4 different maps: one map that corresponds to a normal scan over the CMB and 3 maps that reconstruct the observed path from extremely reduced time-ordered data.
+The constructed timestreams for $I$, $Q$ and $U$ are then compressed in 4 different ways: averaging of the timestream, representing each datapoint as a 64bit floating point number and applying the three introduced digitisation schemes before averaging the timestream.
 
@@
 
-To obtain maps from the digitised time-stream we simply use our knowledge of the scan strategy to recognise what pixel each recorded point hits.
-All hits of a pixel are then simply averaged to obtain the pixel value.
+We average all datapoints that lie within one pixel for each observation scheme and each channel.
+Through this procedure we slowly build up 12 different maps: three maps that corresponds to a normal scan over the polarised CMB and 9 maps that reconstruct the observed patch from an extremely compressed time-ordered data.
+Additionally we keep hold of the hitmap.
 
@@
 \tablecomments{ 
-comments
+Cosmological parameters used to create the template maps.
 } \normalsize
@@
 \begin{center}
-\caption{\label{tab:modelparams} Model Parameters}
+\caption{\label{tab:modelparams} Assumed Survey Parameters}
 \small
@@
 \tablecomments{ 
-comments
+Parameters used in the simulated sky strategy. The RA speed is tweaked to match the desired hits per pixel.
 } \normalsize
@@
 
-\acknowledgments
+\acknowledgments % thank patrick for discussion/starting point? Andrew?
 

### 2018-08-07 digitisation.tex
Editing.
Editing.

@@
 
-\title{Constructing Cosmic Microwave Background Power Spectra from Few Bit Timestreams}
+\title{Cosmic Microwave Background Power Spectra from Few Bit Timestreams}
 \author{L.~Balkenhol\altaffilmark{\Melbourne} and C.~L.~Reichardt\altaffilmark{\Melbourne}}
@@
 
-Summary:
-
-To investigate the performance of the aforementioned digitisation schemes we simulate many scans over CMB template maps at the timestream level, where we add detector noise, for which we use gaussian white noise .
-We obtain maps that use 64bit TOD and maps that have undergone 1, 2 and 3 bit digitisatino at the timestream level.
-We obtain the $TT, EE, BB, TE$ powerspectra of each map.
-Through comparison we then determine the additional noise induced through the extreme digitisation process.
-
-Template maps:
+To investigate the performance of the aforementioned digitisation schemes we simulate many scans over CMB template maps at the timestream level.
+Each scan is performed by a single detector.
+We obtain maps that use 64bit TOD and maps that have undergone 1, 2 and 3 bit digitisation at the timestream level.
+We calculate the temperature and polarisation power spectra of each map and determine the additional noise induced through the extreme digitisation process.
 
 The $I, Q, U$ template maps used were created through healpy's synfast.
-We rely on the power spectrum provieded by the Planck collaboration XXX.
+As the input power spectra we use the results by Placnk 2015.
 The underlying cosmological parameters are summarised in table \ref{tab:inputcosparams}.
 
-Scan strategy:
-
-We simulate observing a $\sim 600$ \sqdeg path of the sky.
-To do so we perform constant elevation scans over the observation region.
-We repeat the observation strategy $100$ times with a slight offset in RA and DEC each time, such that all pixels within the path are hit roughly uniformly.
-The simulation parameters are summarised in table \ref{tab:modelparams}.
-
-Noise addition and digitisation:
+We simulate observing a $\sim 600$ \sqdeg patch of the sky.
+To do so we perform constant elevation scans (CES) over the observation region.
+We repeat the observation strategy $100$ times with a slight offset in RA and DEC each time, such that all pixels within the path are hit approximately uniformly.
 
@@
 The corresponding values from the template maps are then accessed and added to realisations of the detector noise of appropriate length.
+We assume the detector noise to be gaussian white noise.
+
 The constructed timestreams for $I$, $Q$ and $U$ are then compressed in 4 different ways: averaging of the timestream, representing each datapoint as a 64bit floating point number and applying the three introduced digitisation schemes before averaging the timestream.
+The simulation parameters are summarised in table \ref{tab:modelparams}.
+Through this procedure we slowly build up 12 different maps: three maps that corresponds to a normal scan over the polarised CMB and 9 maps that reconstruct the observed patch from an extremely compressed time-ordered data.
+Additionally we retain the hitmap.
 
-Compression:
-
-We average all datapoints that lie within one pixel for each observation scheme and each channel.
-Through this procedure we slowly build up 12 different maps: three maps that corresponds to a normal scan over the polarised CMB and 9 maps that reconstruct the observed patch from an extremely compressed time-ordered data.
-Additionally we keep hold of the hitmap.
+The simulation described above was implemented in Python 2.7.
+It has considerable computational requirements if we want to reach up to $\sim 10^8$ hits per pixel.
+To carry out this simulation we make use of parallelisation and the computing power provided by NERSC.
 
@@
 
-%\begin{figure*}[htb]\centering
-%\includegraphics[width=0.9\textwidth,clip,trim={1.5cm 12.5cm 5cm 3.8cm}]{pretty.pdf}
-%  \caption[Current ]{
-%  Current 
-%           \label{fig:ig}
-%  }
-%\end{figure*}
-
 \section{Results}
@@
 
-Get power spectra
+\subsection{Analysis}
+\label{subsec:analysis}
 
-Analysis procedure
+We use PolSpice to compute the $TT, EE, BB$ power spectra of the obtained maps.
+When doing so we apodise the observed skypatch using a cosine mask to minimise cut sky effects on the spectra.
+Please see \ref{tab:polspiceparams} for an overview of the parameters used in this process.
 
-Analysis results: Nbits to map noise
+\begin{table}[tbh]
+\begin{center}
+\caption{\label{tab:polspiceparams} PolSpice Parameters}
+\small
+\begin{tabular}{c c c c}
+weightfile & apodizesigma & apodizetype & polarization \\
+\hline
+cosine mask & $\sqrt{600}/2$ & 1 & YES \\
 
-Analysis results: Nhits per pixel to map noise
+\end{tabular}
+\tablecomments{ 
+Parameters used when calling PolSpice to calculate the power spectra. Remaining parameters have been left at their default value.
+} \normalsize
+\end{center}
+\end{table}
 
-Analysis result: Noise disappears in cross-spectra -> actually cannot check this in current framework.
+The obtained power spectra originating from digitised time streams are normalised against the power spectra that use the full time stream data.
+We use results that originate from $\sim 10^8hpp$ maps.
+We separately normalise the TT, EE and BB spectra.
+The normalisation constant obtained from this calculation are then applied to all lower hit per pixel runs.
 
-Any other results that are outstanding
+We restrict ourselves to windows in multipole space where we trust that the spectra are within appropriate multipole moment range given the patch size and are signal dominated.
+We therefore place a lower limit on the normalisation window for each channel from
 
-? Analysis result: To uncertainty in cosmological parameters?
+\[ l = \frac{2}{\pi} \frac{(180^\circ )^2}{600 sqdeg} \approx 35 \]
 
-This is why it works well: low S/N many hits per pixel, jumps in CMB noise level are order of magnitude.
+where we have rounded up to the next highest integer. We find the upper bound on the normalisation window by demanding that
+
+\[ \frac{C_l^{S}}{C_l^{N}} \geq 10 \]
+
+\subsection{Additional Noise}
+\label{subsec:additionalnoise}
+
+To determine how much the quality of the power spectra suffers due to the digitisation we would like to compare the map noise levels.
+We would like to put this value into context with the map noise level induced through the detector noise.
+We therefore formulate
+
+\[ \frac{\Delta \sigma}{\sigma} = \frac{\sigma_{MAP}^{D}-\sigma_{MAP}^{SN}}{\sigma_{MAP}^{SN}} \]
+
+where $\sigma_{MAP}^{SN}$ is the map noise level of the full power spectra (signal plus noise) and $\sigma_{MAP}^{D}$ is the map noise level obtained from the power spectra originating from extremely digitised TOD.
+We assume that the digitisation process results in adding some constant  to the powerspectrum, i.e.
+
+\[ C_l^D = C_l^N + C_l^X \]
+
+in the limit where we are dominated by noise.
+Here $C_l^D$ is the power spectrum originating from a digitised timestream, $C_l^N$ is the detector noise level and $C_l^X$ the additional noise induced.
+We may now progress with
+
+\[\frac{\Delta \sigma}{\sigma} = \sqrt{\frac{C_l^D}{C_l^{N}}} - 1  = \sqrt{\frac{C_l^N + C_l^X}{C_l^{N}}} - 1 = \sqrt{1 + \frac{C_l^X}{C_l^{N}}} - 1  \]
+
+Before analysing $C_l^X/C_l^N$ we rebin the power spectra to $\Delta l = 123$.
+This guarantees that the points in the noise tail are independent of one another, allowing us to extract an uncertainty for the above quantity.
+
+The results for the 1, 2 and 3 bit digitisation schmes for the TT, EE and BB power spectra are summarised in \ref{fig:extranoise}.
+We would like to points out three key results.
+Firstly, 3 Bit digitisation performs the best, followed by 2 Bit and finally 1 Bit digitisation.
+This is what we expect, given that with each additional bit we may retain more information.
+Secondly, for a fixed detector noise level, the additional percentage to the map noise level due to digitisation is independent of the number of hits per pixel.
+This is a valuable result, as it means that there is not a certain point at which surveys start becoming viable for this compression techniques, but they rather perform equally well throughout the range of hits per pixel tested.
+Lastly, the added noise levels are astonishingly low.
+
+%Any other results that are outstanding
+
+%? Analysis result: To uncertainty in cosmological parameters?
+
+We present the average additional percentage to the map noise level through digitisation in \ref{tab:noise}.
+These results are obtained by calculating the weighted average across all hits per pixel for a given digitisation scheme and channel.
+We note that an optimal 3 bit digitisation scheme adds as little as $\sim2\%$ to the map noise level.
+
+Given that advances in CMB observations happen in order magnitude jumps between generations of experiments this is a negligible noise addition, considering that it reduces the TOD by an order of magnitude.
+
+Extreme digitisation is a viable lossy compression technique when dealing with large low signal to noise, slowly varying datasets.
+Within this limit digitisation relies on the noise to push the small signal towards the correct side of the locally stationary signal.
+A low signal to noise ratio prevents us from a signal saturating the digitisation levels.
+For large datasets are needed to provide many samplings of each locally stationary signal.
+An increase in bit number does not necessarily improve the recovery of the mean, but rather advances narrowing down on its uncertainty, by improving the nulling out of the noise.
+Improving the number of samples in a dataset improves the finding of the mean, as it allows for more correctly pushed digitised points to be considered.
+
+\begin{figure*}[htb]\centering
+\includegraphics[width=0.9\textwidth,clip]{Plots/addednoise.png}
+  \caption[Current ]{
+  Addition to the map noise level due to digitisation. 
+           \label{fig:extranoise}
+  }
+\end{figure*}
 
@@
 \begin{tabular}{l | c c c }
-Model   & XX&XX&$r$\\
+ Channel & 1 Bit & 2 Bit & 3 Bit\\
  \hline
+ TT & & & \\
+ EE & & & \\
+ BB & & & \\
 
@@
 \tablecomments{ 
-go go
+Added percentage to the map noise level due to digitisation.
 } \normalsize
@@
 
-While the X added noise is already impressive it should be laid out how this changes when moving to a more realistic noise profile.
+It should be laid out how this changes when moving to a more realistic noise profile.
+We do not expect this to alter the practicality of our results - even if a moderate change in the noise profile doubles the additional percentage to the map noise level extreme digitisation is still practical.
+Ideas on how to deal with 1/f noise, e.g.
+chunking of the data before applying extreme digitisation have already been investigated by Planck.
+These thoughts should be considered when designing the compression schemes of future space-based CMB missions, which will be unable to recover their full data with latency, but rather fully rely on the transmitted data.
 

### 2018-08-08 digitisation.tex
Editing.
Done a lot of editing.

@@
 
-\begin{abstract} % copied off ASA for now
+\begin{abstract}
 Observations of the Cosmic Microwave Background (CMB) are of immense value to modern cosmology.
@@
 Unlike lossless compression, extreme digitisation introduces additional noise into the data.
-We present an optimal 1, 2 and 3 bit digitisation schemes and lay out how the added noise varies between the schemes and as a function of hits per pixel in the map for temperature and polarisation observations.
-We show that the noise penalty of 3 bit digitisation is at the percent-level for CMB power spectrum measurements.
+We present optimal 1, 2 and 3 bit digitisation schemes and determine the degradation in temperature and polarisation power spectra caused by this process.
+In particular we show that 3 bit digitisation has a percent-level contribution to the map noise level.
 This is impressive considering that it would reduce the data volume by an order of magnitude.
@@
 
-\keywords{ cosmic background radiation --- polarization }
+\keywords{ cosmic background radiation --- polarization --- data compression}
 \section{Introduction}
@@
 
-Observations of the Cosmic Microwave Background (CMB) have played a key role since 1964 (discovery paper).
+Observations of the Cosmic Microwave Background (CMB) have played a key role in physics since 1964 (discovery paper).
 Current and future CMB experiments will continue to deliver new insights by studying the temperature and polarisation information contained in the CMB.
-These will constrain help putting tighter constraints on cosmological models.
-The most prominent science goal is the discovery of the imprint left by inflationary gravitational waves.
-Additionally studies of CMB lensing and the Sunyaev-Zeldovich (SZ) effects will deliver new insight.
-CMB experiments however also provide a valuable counterpart to ground-based particle physics experiments, as they probe the relativisitc number of species, the helium fraction and the neutrino mass sum.
+These will put tight constraints on cosmological models.
+The most prominent science goal is the discovery of the imprint of inflationary gravitational waves.
+Additionally studies of CMB lensing and the Sunyaev-Zeldovich (SZ) effects will give us access to previously unseen information.
+Through probing the relativistic number of species, the helium fraction and the neutrino mass sum, CMB experiments are a valuable counter-part to ground-based particle physics experiments.
 
@@
 The CMB community has developed a variety of compression and computational techniques to manage the increasing influx of data, while maximising the science output.
-These include the compression of time-ordered data into maps, bandpower estimation and the pseudo Cl method.
+These include the compression of time-ordered data into maps, bandpower estimation and the pseudo $C_l$ method.
 
@@
 We find that an optimal 3-bit digitisation scheme adds as little as $\sim 2\%$ to the map noise level.
-While the digtisation schemes described here are primarily laid out for ground-based Stage-4 experiments, future space-based missions must inevitably study lossy compression techniques.
+While the digtisation schemes described here are primarily laid out for ground-based experiments, future space-based missions must inevitably study lossy compression techniques.
 
+% update this paragraph towards the end
 This work is structured as follows.
@@
 The science goals of upcoming CMB experiments naturally lead to a large influx of data.
-To achieve them longer observations with more detectors are needed.
-However, the best observations sites for CMB observations are at remote locations: in space and in the Antarctica.
-Experimetns at these locations depend on satellite transmission The next generation of ground-based experiments, Stage-4, will aim to collect 2 million detector years worth of data.
-A Stage-4 style experiment will face a data influx of $\sim O(10)Tb/d$.
-However, the current transmission allocation for the SPT3G is at $150Gb/d$.
-SPT3G currently deals with its transmission bottleneck by downsampling the TOD, i.e.
-discarding high frequency information.
-While this will likely see an increase going into the next stage of experimetns it will not be able to over the entire observation.
+To achieve the targeted sensitivity longer observations with more detectors are needed.
+In fact the number of detectors of ground-based experiments has been following a Moore's law like trend, doubling approximately every 2 years.
+This directly translates into an exponential growth in data volume.
+
+The best observations sites for CMB measurements are at remote locations: in space and in the Antarctica.
+Experiments at these locations depend on satellite transmission.
+The next generation of ground-based experiments will aim to collect 2 million detector years worth of data.
+An experiment contributing to Stage-4 at the South Pole will face a data influx of $\sim O(10)Tb/d$.
+However, the current transmission allocation for the SPT3G is at $150Gb/d$, which will likely only see a moderate increase in the coming years.
+The transmission bottleneck is currently overcome by recovering the full data with some latency and by transmitting a downsampled version of the data.
+The downsampling process loses high frequency information.
 Going into Stage-4 we anticipate that compression rates for transmission must increase by an order of magnitude.
-Continued use of heavy downsampling will narrow the informatino window decisively - prohibiting high l science to be carried out on the transmitted dataset.
+Continued use of downsampling will narrow the information window decisively - prohibiting high multipole moment science to be carried out on the transmitted dataset.
 This also means that any potential faults or errors in the experiment that only become visible in high frequencies will go unnoticed for longer.
 
-Similarly, future space-based missions (PIXIE, COrE+, liteBIRD) aim to increase their detector count by at least an order of magnitude compared to PLANCK.
-It is questionable whether their telemetry specifications will allow for transmission of the data with PLANCK-style compression.
+Future space-based missions aim to exceed the detector count of Planck by at least an order of magnitude.
+It is questionable whether their telemetry specifications will allow for transmission of the data with Planck-style compression.
 Methods of storing large amounts of data on upcoming satellites will likely be prohibited by financial decisions: the amount of storage space required becomes financially relevant at the scales targeted.
+Missions will likely be left to design their own compression algorithms which will incorporate a combination of lossless and lossy compression techniques.
 
@@
 
-Beyond these transmission hurdles detailed mission planning is becoming exceedingly expensive.
-As noted by (S4 science book) a full simulation of TOD over the entire parameter space of detection scenarios for numerous set-ups is the desired way to plan Stage-4 configuration.
-Similarly space-based missions must carry out a similar analysis to optimise their science out-put.
-However, given the sheer size of TOD this is not possible.
-We must turn to different planning strategies or aim to reduce the size of the TOD in order to maximise the productivity of planning and development stages.
+Beyond transmission challenges, mission planning is becoming exceedingly expensive.
+As noted by (S4 science book) a full simulation of TOD over the entire parameter space of detection scenarios for numerous set-ups is the desired way to decide on Stage-4 configurations.
+Space-based missions must aim to carry out a similar analysis to optimise their science output.
+We must rely on different planning strategies or aim to reduce the size of the TOD in order to maximise the productivity of planning and development stages.
 
-Operations on the TOD, such as noise-removal or map-making are a vital part of CMB data analysis. Through the expontentially growing size of TOD CMB data analysis is becoming increasingly expensive.
+% Analysis
+
+Operations on the TOD, such as noise-removal or map-making are a vital part of CMB data analysis. Through the exponentially growing size of TOD CMB data analysis is becoming increasingly expensive.
 
@@
 
-Extreme Digitisation would tackle the challenges mentioned above by recuding the size of the TOD by an order of magnitude.
+Extreme Digitisation would tackle the challenges mentioned above by reducing the size of the TOD by an order of magnitude.
 Together with already exploited lossless compression techniques (such as FLAC) this will directly tackle transmission hurdles.
-While it needs to be investigated to what extent existing algorithms can be carried over, extreme digitisation has the possiblity of solving planning and analysis problems.
+While it needs to be investigated to what extent existing algorithms can be carried over, extreme digitisation has the possibility of solving planning and analysis problems.
+Other areas of science have already demonstrated that extreme digitisation can be a viable compression technique.
+For example Jenet and Anderson have shown the advantages of this compression technique to pulsar timing measurements.
+The computationally challenging searches for continuous gravitational wave searcher may also profit from few-bit digitisation as laid out by Clearwater et al.
+(in prep.).
 
@@
 A theoretical framework to obtain these levels was laid out by Max in 1978.
-We review the key aspects of his work relevant to our findings below.
+We review the key aspects of his work relevant for us below.
 
-We divide the space $-\infty \to \infty$ into $N$ ranges.
-One then has endpoints $x_k$ and output levels $y_k$, such that an input between $x_i$ and $x_{i+1}$ produces an output at $y_i$.
+Digitisation discretises an input signal by sorting it into $N$ appropriate ranges, such that an input between $x_i$ and $x_{i+1}$ produces an output at $y_i$.
+A digitisation scheme is described the the number of ranges, $N$, the endpoints of these ranges $x_k$ and the output levels $y_k$.
 Conventionally one chooses $x_{1} = -\infty$ and $x_{N+1} = \infty$.
-In order to determine an optimal set of $x_k$ and $y_k$ we would like to minimise the distortion
+In order to quantify the performance of a given digitisation scheme that produces a signal out $s_{out}$ form an input signal $s_{in}$ we define the distortion as
 
@@
 
-Seeing as we wish to minimise the distortion we differentiate the above with respect to $x_i$ and $y_i$ and set the derivatives to zero.
+Seeing as we wish to minimise the distortion we differentiate the above with respect to $x_i$ and $y_i$ and set the derivatives to zero. We obtain the two equations
 
-\[ \frac{\partial D}{\partial x_i} = \left(x_i-y_{i-1}\right)^2 p(x_i) - \left(x_i - y_i\right)^2 p(x_i) = 0 \]
+\begin{equation} \label{eq:distderiv1}
+\frac{\partial D}{\partial x_i} = \left(x_i-y_{i-1}\right)^2 p(x_i) - \left(x_i - y_i\right)^2 p(x_i) = 0
+\end{equation}
 
-\[ \frac{\partial D}{\partial x_j} = -2 \int_{x_i}^{x_{i+1}} \left( x-y_i \right) p(x) dx = 0 \]
+\begin{equation} \label{eq:distderiv2}
+\frac{\partial D}{\partial x_j} = -2 \int_{x_i}^{x_{i+1}} \left( x-y_i \right) p(x) dx = 0
+\end{equation}
 
-From the first equation above we deduce that
+Rearranging equation \ref{eq:distderiv1} we deduce
 
@@
 
-The second one
+which informs us that an output level $y_i$ must lie halfway between is delimiting thresholds $x_i$ and $x_{i+1}$. We gain an additional condition from equation \ref{eq:distderiv2}
 
@@
 
-implies that we should choose $y_i$, such that it halves the area underneath $p(x)$ in the interval $x_i \to x_{i+1}$.
+This implies that we should choose $y_i$, such that it halves the area underneath $p(x)$ in the interval from $x_i$ to $x_{i+1}$.
 
-We are interested in the case where our input signal can fairly be approximated by gaussian white noise, i.e.
-$p(x) = 1/sqrt{2\pi} e^{-x^2/2}$.
-Even with these assumptions we cannot solve the resulting equations analytically.
-Instead one can begin by picking $y_1$ and calculating the remaining $x_i$ and $y_i$ using equation \ref{eq:digitequalspacecondition}.
-One then observes weather this choice of values satisfy the conditions given by equation \ref{eq:digitareacondition}.
-If that is the case the $x_k$ and $y_k$ were chosen appropriately.
-This is unlikely to succeed on the first try.
-We rather solve the above system iteratively.
+To progress further we have to make an assumption about the distribution of input signals, $p(x)$.
+For our purposes we may safely assume that ground-based CMB observations operate at low signal to noise.
+Furthermore we assume that the noise profile is Gaussian white noise\footnote{Please see the conclusion for a discussion of the effect of more realistic noise profile}, i.e.
+$p(x) = 1/\sqrt{2\pi} e^{-x^2/2}$.
+Given this assumptions we cannot solve proceed analytically, but can the problem using a numerical iterative procedure.
+One begins by picking $y_1$ and calculating the remaining $x_i$ and $y_i$ using equation \ref{eq:digitequalspacecondition}.
+Afterwards one observes weather this choice of values satisfy the conditions given by equation \ref{eq:digitareacondition}.
+If that is the case, the $x_k$ and $y_k$ were chosen appropriately.
 
-This was carried out by Max. We incorporate his results by formulating the multi-level functions we use for our 1, 2 and 3 bit digitisation process. For 1 bit digitisation we apply the sign function
+This was carried out by Max.
+We incorporate his results by formulating the multi-level functions we use for our 1, 2 and 3 bit digitisation process.
+Given an input signal $x(t)$ a digitisation scheme using $N$ bits returns $\hat{x}_N(t)$.
+For 1 bit digitisation we apply
 
-\[ \hat{x}(t) = \left\{ \begin{array}{lr}
+\[ \hat{x}_1(t) = \left\{ \begin{array}{lr}
 1, & \text{for } x(t) > 0\\
@@
 
-\[ \hat{x}(t) = \left\{ \begin{array}{lr}
+\[ \hat{x}_2(t) = \left\{ \begin{array}{rl}
 1.51 \sigma, & \text{for } x(t) \geq 0.9816 \sigma\\
@@
 
-Finally the optimal 3 bit digitisation is described by the eight-level function
+to the input signal. Finally the optimal 3 bit digitisation is described by the eight-level function
 
-\[ \hat{x}(t) = \left\{ \begin{array}{lr}
+\[ \hat{x}_3(t) = \left\{ \begin{array}{rl}
 2.152 \sigma, & \text{for } x(t) \geq 1.748 \sigma\\
@@
 
-% Theoretical lvls for gaussian white noise
+Other digitisation schemes can be thought of that place the digitisation thresholds and output levels in a different way.
+However, these are not of interest for now, given the ease at which the devised optimal schemes can be implemented and the fact that they will lead to worse performance.
 
@@
 
-To investigate the performance of the aforementioned digitisation schemes we simulate many scans over CMB template maps at the timestream level.
+To investigate the performance of the derived digitisation schemes we simulate many scans over CMB template maps at the timestream level.
 Each scan is performed by a single detector.
@@
 
-The $I, Q, U$ template maps used were created through healpy's synfast.
-As the input power spectra we use the results by Placnk 2015.
-The underlying cosmological parameters are summarised in table \ref{tab:inputcosparams}.
+To create the template maps we use the healpix framework and the wealth of support available for it.
+We generate a realisation of I, Q and U maps based on the results of Planck 2015.
+The key cosmological parameters are summarised in \ref{tab:inputcosparams}.
 
 We simulate observing a $\sim 600$ \sqdeg patch of the sky.
-To do so we perform constant elevation scans (CES) over the observation region.
+To do so we perform a number of equally spaced constant elevation scans (CES) over the observation region.
 We repeat the observation strategy $100$ times with a slight offset in RA and DEC each time, such that all pixels within the path are hit approximately uniformly.
+The speed at which we sweep across the survey field is set to a constant to produce a desired number of hits per pixel in the output maps.
 
@@
 The corresponding values from the template maps are then accessed and added to realisations of the detector noise of appropriate length.
-We assume the detector noise to be gaussian white noise.
+We assume the detector noise to be Gaussian white noise.
 
-The constructed timestreams for $I$, $Q$ and $U$ are then compressed in 4 different ways: averaging of the timestream, representing each datapoint as a 64bit floating point number and applying the three introduced digitisation schemes before averaging the timestream.
+At this point we apply the digitisation schemes to the TOD.
+We compress the timestream into maps by averaging all hits falling into the same pixel.
+We produce 13 maps in total: three maps that have not undergone few-bit digitisation at the timestream level for comparison purposes, nine maps with three each corresponding to each digitisation scheme and the hitmap.
+
 The simulation parameters are summarised in table \ref{tab:modelparams}.
-Through this procedure we slowly build up 12 different maps: three maps that corresponds to a normal scan over the polarised CMB and 9 maps that reconstruct the observed patch from an extremely compressed time-ordered data.
-Additionally we retain the hitmap.
-
-The simulation described above was implemented in Python 2.7.
 It has considerable computational requirements if we want to reach up to $\sim 10^8$ hits per pixel.
@@
 \small
-\begin{tabular}{l | c c c }
-$\Omega$ & XX&XX&$r$\\
+\begin{tabular}{l | c}
+Parameter & Planck 2015\\
 \hline
-
+$100\theta_{MC}$ & $ 1.04086 \pm 0.00048$\\
+$\Omega_b h^2$ & $  0.02222 \pm 0.00023$\\
+$\Omega_c h^2$ & $ 0.1199 \pm 0.0022$\\
+$H_0$ & $67.26 \pm 0.98$\\
+$n_s$ & $0.9652 \pm 0.0062$\\
+$\Omega_m$ & $0.316 \pm 0.014$\\
+$\sigma_8$ & $0.830 \pm 0.015 $\\
+$\tau$ & $0.078 \pm 0.019$\\
+$10^9 A_s e^{-2\tau}$ & $1.881 \pm 0.014 $\\
 \end{tabular}
 \tablecomments{ 
-Cosmological parameters used to create the template maps.
+Cosmological parameters used to create the template maps. Taken from Planck 2015: Cosmological Parameters.
 } \normalsize
@@
 \small
-\begin{tabular}{c c c c c}
-$\mathrm{NSIDE}$ & $f_{readout}$ & $f_{sky}$ & $\sigma^{\mathrm{T}}_{detector}$ & $\sigma^{\mathrm{Pol}}_{detector}$\\
+\begin{tabular}{l c c c c}
+$\mathrm{NSIDE}$ & $f_{\mathrm{readout}} [\mathrm{Hz}]$ & $f_{{\mathrm{sky}}}$ & $\sigma^{\mathrm{T}}_{\mathrm{det}} [\mathrm{\mu K \sqrt{s}}]$ & $\sigma^{\mathrm{Pol}}_{\mathrm{det}} [\mathrm{\mu K \sqrt{s}}]$\\
 \hline
-$4096$ & $200\mathrm{Hz}$ & $\sim 0.014$ & $500\mathrm{K}\sqrt{\mathrm{s}}$ & $\sqrt{2}\times500\mathrm{K}\sqrt{\mathrm{s}}$\\
+$4096$ & $200$ & $\sim 0.014$ & $500$ & $\sqrt{2}\times500$\\
 
@@
 
-\subsection{Analysis}
-\label{subsec:analysis}
+\subsection{Power Spectrum Estimation}
+\label{subsec:psestimation}
 
-We use PolSpice to compute the $TT, EE, BB$ power spectra of the obtained maps.
+We use PolSpice to compute the $TT, EE, BB$ power spectra of the reconstructed maps.
 When doing so we apodise the observed skypatch using a cosine mask to minimise cut sky effects on the spectra.
@@
 
-The obtained power spectra originating from digitised time streams are normalised against the power spectra that use the full time stream data.
-We use results that originate from $\sim 10^8hpp$ maps.
-We separately normalise the TT, EE and BB spectra.
-The normalisation constant obtained from this calculation are then applied to all lower hit per pixel runs.
+To normalise the the obtained power spectra we focus on the $\sim 10^8 \mathrm{hpp}$ run.
+We normalise each power spectrum originating from few bit TOD against its corresponding 64bit TOD counterpart.
+We place a lower limit on the normalisation window in multipole space by considering the size of the observed patch via
 
-We restrict ourselves to windows in multipole space where we trust that the spectra are within appropriate multipole moment range given the patch size and are signal dominated.
-We therefore place a lower limit on the normalisation window for each channel from
-
-\[ l = \frac{2}{\pi} \frac{(180^\circ )^2}{600 sqdeg} \approx 35 \]
+\[ l = \frac{2}{\pi} \frac{32400}{600} \approx 35 \]
 
@@
 
+we close the normalisation window the first instance the above condition is met.
+The normalisation constants for each digitisation scheme and channel obtained through this way are then applied to all lower hit per pixel simulations.
+
 \subsection{Additional Noise}
@@
 
-To determine how much the quality of the power spectra suffers due to the digitisation we would like to compare the map noise levels.
-We would like to put this value into context with the map noise level induced through the detector noise.
-We therefore formulate
+To quantify how much the quality of the power spectra suffers due to the digitisation compare the map noise levels inferred from the power spectra.
+These are put into the context of the map noise levels deduced from the 64bit TOD power spectra.
+We formulate
 
-\[ \frac{\Delta \sigma}{\sigma} = \frac{\sigma_{MAP}^{D}-\sigma_{MAP}^{SN}}{\sigma_{MAP}^{SN}} \]
+\begin{equation} \label{eq:extramapnoise}
+\frac{\Delta \sigma}{\sigma} = \frac{\sigma_{\mathrm{map}}^{\mathrm{D}}-\sigma_{\mathrm{map}}^{\mathrm{F}}}{\sigma_{\mathrm{map}}^{\mathrm{F}}}
+\end{equation}
 
-where $\sigma_{MAP}^{SN}$ is the map noise level of the full power spectra (signal plus noise) and $\sigma_{MAP}^{D}$ is the map noise level obtained from the power spectra originating from extremely digitised TOD.
-We assume that the digitisation process results in adding some constant  to the powerspectrum, i.e.
+where $\sigma_{\mathrm{map}}^{\mathrm{F}}$ is the map noise level of the control power spectra and $\sigma_{\mathrm{map}}^{\mathrm{D}}$ is the map noise level obtained from the power spectra originating from extremely digitised TOD.
+We assume that the digitisation process results in adding some constant independent of multipole moment to the power spectrum.
+If we are dominated by noise we can write
 
@@
 
-in the limit where we are dominated by noise.
-Here $C_l^D$ is the power spectrum originating from a digitised timestream, $C_l^N$ is the detector noise level and $C_l^X$ the additional noise induced.
-We may now progress with
+Here $C_l^D$ is the power spectrum originating from a digitised timestream, $C_l^N$ is the detector noise level and $C_l^X$ the additional noise induced through digitisation.
+We now progress equation \ref{eq:extramapnoise} to
 
 \[\frac{\Delta \sigma}{\sigma} = \sqrt{\frac{C_l^D}{C_l^{N}}} - 1  = \sqrt{\frac{C_l^N + C_l^X}{C_l^{N}}} - 1 = \sqrt{1 + \frac{C_l^X}{C_l^{N}}} - 1  \]
+
+The value of $l$ at which we can safely assume to be noise dominated and apply the above framework varies between simulated observations of different hits per pixel and for different channels.
+The range considered involves any datapoints at multipole moments larger than the last point at which the detector noise is at least than 10 times the input power spectrum.
 
@@
 
-The results for the 1, 2 and 3 bit digitisation schmes for the TT, EE and BB power spectra are summarised in \ref{fig:extranoise}.
+%plot these X/N ?
+
+The results for the 1, 2 and 3 bit digitisation schemes are summarised in figure \ref{fig:extranoise}.
 We would like to points out three key results.
@@
 Secondly, for a fixed detector noise level, the additional percentage to the map noise level due to digitisation is independent of the number of hits per pixel.
-This is a valuable result, as it means that there is not a certain point at which surveys start becoming viable for this compression techniques, but they rather perform equally well throughout the range of hits per pixel tested.
+The added noise therefore scales in the same fashion as the map noise level with the number of hits per pixel, given a fixed detector noise level.
 Lastly, the added noise levels are astonishingly low.
+Keeping in mind that CMB detector sensitivity improves in steps of order of magnitude every few years adding an extra percent level noise term does not deteriorate the results appreciably.
+This is impressive, given that the use of an optimal 3 bit digitisation scheme will save approximately an order of magnitude in TOD volume.
 
-%Any other results that are outstanding
-
-%? Analysis result: To uncertainty in cosmological parameters?
-
-We present the average additional percentage to the map noise level through digitisation in \ref{tab:noise}.
-These results are obtained by calculating the weighted average across all hits per pixel for a given digitisation scheme and channel.
-We note that an optimal 3 bit digitisation scheme adds as little as $\sim2\%$ to the map noise level.
-
-Given that advances in CMB observations happen in order magnitude jumps between generations of experiments this is a negligible noise addition, considering that it reduces the TOD by an order of magnitude.
-
-Extreme digitisation is a viable lossy compression technique when dealing with large low signal to noise, slowly varying datasets.
-Within this limit digitisation relies on the noise to push the small signal towards the correct side of the locally stationary signal.
-A low signal to noise ratio prevents us from a signal saturating the digitisation levels.
-For large datasets are needed to provide many samplings of each locally stationary signal.
-An increase in bit number does not necessarily improve the recovery of the mean, but rather advances narrowing down on its uncertainty, by improving the nulling out of the noise.
-Improving the number of samples in a dataset improves the finding of the mean, as it allows for more correctly pushed digitised points to be considered.
+Extreme digitisation is a viable lossy compression technique when dealing with large volume, low signal to noise, slowly varying datasets.
+Making these assumptions the signal small and constant for a considerable number of datapoints.
+For such an interval the many samplings of the noise result in producing appropriate numbers in each output bin.
+The addition of more bits helps by classifying the size of the noise more and more appropriately and hence approaching a good estimate for the value of the signal faster.
+It is necessary to remain in the low signal to noise regime, as otherwise a dominant signal would saturate the digitisation scheme and for a given scheme only values in the range $\pm |N_{\mathrm{hits}}y_N|$ can be represented, given that we have $N_{\mathrm{hits}}$ datapoints.
+We must demand the data to be slowly varying such that many datapoints can be used to reconstruct a single significant step of the input signal.
+Finally a large data volume is clearly favourable for the reasons illustrated above: it allows for repeated sampling and therefore more accurate reconstruction of the weak signal.
 
@@
 \end{figure*}
-
-\begin{table}[tbh]
-\begin{center}
-\caption{\label{tab:noise} Noise levels}
-\small
-\begin{tabular}{l | c c c }
- Channel & 1 Bit & 2 Bit & 3 Bit\\
- \hline
- TT & & & \\
- EE & & & \\
- BB & & & \\
-
-\end{tabular}
-\tablecomments{ 
-Added percentage to the map noise level due to digitisation.
-} \normalsize
-\end{center}
-\end{table}
 

### 2018-08-09 digitisation.tex
Editing
Completed another editing pass.

@@
 However, future CMB experiments must confront challenges in mission planning, hardware and analysis that arise from the sheer size of the time-ordered-data being recorded.
-These challenges are particularly significant for Antarctic and satellite experiments which depend on satellite links to transmit the data.
-We investigate using extreme digitisation to address these challenges.
+These challenges are particularly significant for Antarctic and satellite experiments which depend on satellite links to transmit their data.
+We investigate using extreme digitisation to address these issues.
 Unlike lossless compression, extreme digitisation introduces additional noise into the data.
-We present optimal 1, 2 and 3 bit digitisation schemes and determine the degradation in temperature and polarisation power spectra caused by this process.
-In particular we show that 3 bit digitisation has a percent-level contribution to the map noise level.
+We present optimal 1, 2 and 3 bit digitisation schemes and determine the degradation in temperature and polarisation power spectra caused by applying this process to the time-ordered-data (TOD).
+We find that 3 bit digitisation has a percent-level contribution to the map noise level.
 This is impressive considering that it would reduce the data volume by an order of magnitude.
@@
 
-Observations of the Cosmic Microwave Background (CMB) have played a key role in physics since 1964 (discovery paper).
+Observations of the Cosmic Microwave Background (CMB) have played a key role in physics since 1964.
 Current and future CMB experiments will continue to deliver new insights by studying the temperature and polarisation information contained in the CMB.
@@
 The most prominent science goal is the discovery of the imprint of inflationary gravitational waves.
-Additionally studies of CMB lensing and the Sunyaev-Zeldovich (SZ) effects will give us access to previously unseen information.
-Through probing the relativistic number of species, the helium fraction and the neutrino mass sum, CMB experiments are a valuable counter-part to ground-based particle physics experiments.
+Additionally studies of CMB lensing and the Sunyaev-Zeldovich (SZ) effects will open up access to unprecedented insight.
+Moreover, through probing the relativistic number of species, the helium fraction and the neutrino mass sum, CMB experiments are a valuable counter-part to ground-based particle physics experiments.
 
@@
 The CMB community has developed a variety of compression and computational techniques to manage the increasing influx of data, while maximising the science output.
-These include the compression of time-ordered data into maps, bandpower estimation and the pseudo $C_l$ method.
+These include the compression of time-ordered data (TOD) into maps, bandpower estimation and the pseudo $C_l$ method.
 
 A growing hurdle for experiments at remote locations are the transmission limitations of satellite links.
-Space-based experiments have employed a combination of lossless and lossy compression techniques, including reduced bits in the time-ordered-data (TOD).
-Antarctica based experiments that transmit a portion of their data via a satellite link have downsampled their TOD in the past to meet their telemetry requirements.
+Space-based experiments have employed a combination of lossless and lossy compression techniques, including reduced bits in the time-ordered-data.
+Antarctica-based experiments that transmit a portion of their data via a satellite link have downsampled their TOD in the past to meet their telemetry requirements.
 They have yet to exploit few bit digitisation of the TOD.
-
 As we approach the next generation of ground-based experiments, Stage-4, and the launch of a new generation of space-based missions (liteBIRD, PIXIE, COrE+), we must treat the transmission bottleneck carefully.
@@
 We find that an optimal 3-bit digitisation scheme adds as little as $\sim 2\%$ to the map noise level.
-While the digtisation schemes described here are primarily laid out for ground-based experiments, future space-based missions must inevitably study lossy compression techniques.
+While the digtisation schemes described here are primarily laid out for ground-based experiments, the results we present here should also be considered by future space-based missions, which inevitably must incorporate lossy compression.
 
-% update this paragraph towards the end
 This work is structured as follows.
 We detail the arising challenges in handling large TOD in \S\ref{subsec:problem}.
-We subsequently formulate extreme digitisation formerly in \S\ref{subsec:extremedigitisation} and lay out the framework used to test the performance of this compression technique in \S\ref{subsec:method}.
-In \S\ref{sec:results} we present the results obtained.
+We subsequently formulate extreme digitisation formerly in \S\ref{subsec:extremedigitisation} and lay out the framework used to test its performance in \S\ref{subsec:method}.
+We detail how we estimate the power spectra from simulated maps in \S\ref{subsec:psestimation}.
+We continue by presenting the noise induced through the dicitisation process in \S\ref{subsec:additionalnoise}.
 We summarize our findings in \S\ref{sec:conclusions}.
@@
 The next generation of ground-based experiments will aim to collect 2 million detector years worth of data.
-An experiment contributing to Stage-4 at the South Pole will face a data influx of $\sim O(10)Tb/d$.
-However, the current transmission allocation for the SPT3G is at $150Gb/d$, which will likely only see a moderate increase in the coming years.
-The transmission bottleneck is currently overcome by recovering the full data with some latency and by transmitting a downsampled version of the data.
+An experiment contributing to Stage-4 at the South Pole will face a data influx of $\sim O(10)\mathrm{Tb/d}$.
+However, the current transmission allocation for SPT3G is at $150\mathrm{Gb/d}$, which will likely only see a moderate increase in the coming years.
+The transmission bottleneck is currently overcome by recovering the full data on hard drives with some latency and by transmitting a downsampled version of the data.
 The downsampling process loses high frequency information.
@@
 
-Beyond transmission challenges, mission planning is becoming exceedingly expensive.
+Beyond transmission challenges, mission planning is becoming exceedingly difficult.
 As noted by (S4 science book) a full simulation of TOD over the entire parameter space of detection scenarios for numerous set-ups is the desired way to decide on Stage-4 configurations.
 Space-based missions must aim to carry out a similar analysis to optimise their science output.
-We must rely on different planning strategies or aim to reduce the size of the TOD in order to maximise the productivity of planning and development stages.
+Given the shear size of TOD expected, this is not possible.
+We must rely on different planning strategies or aim to reduce the size of the TOD in order to maximise the productivity of planning and development stages and guarantee maximal science output.
 
@@
 
-Operations on the TOD, such as noise-removal or map-making are a vital part of CMB data analysis. Through the exponentially growing size of TOD CMB data analysis is becoming increasingly expensive.
+Operations on the TOD, such as noise-removal or map-making are a vital part of CMB data analysis.
+While we have not experienced the limitations of the accessible computational assets, the exponential growth of CMB data makes its analysis increasingly expensive.
 
@@
 Extreme Digitisation would tackle the challenges mentioned above by reducing the size of the TOD by an order of magnitude.
-Together with already exploited lossless compression techniques (such as FLAC) this will directly tackle transmission hurdles.
+Together with already established lossless compression techniques (such as FLAC, run-length coding, Huffman coding, etc.) this will directly tackle transmission hurdles.
 While it needs to be investigated to what extent existing algorithms can be carried over, extreme digitisation has the possibility of solving planning and analysis problems.
-Other areas of science have already demonstrated that extreme digitisation can be a viable compression technique.
-For example Jenet and Anderson have shown the advantages of this compression technique to pulsar timing measurements.
-The computationally challenging searches for continuous gravitational wave searcher may also profit from few-bit digitisation as laid out by Clearwater et al.
-(in prep.).
+
+Other science areas have demonstrated that extreme digitisation is a valuable compression technique.
+Jenet and Anderson (cite) explored the application of such compression to radio pulsar timing measurements with success.
+Recently Clearwater et al.
+(in prep.) have investigated the advantages of using 1 and 2 bit data when searching for continuous gravitational waves using the Laser Interferometer Gravitational-wave Observatory (LIGO) performance of searches for continuous gravitational wave searches using 1 and 2 bit data.
 
@@
 Digitisation is a lossy compression technique.
-However the induced noise depends on the number of bits used, the digitisation thresholds and the output levels chosen.
+However the induced noise depends on the number of bits used, the digitisation thresholds, and the output levels chosen.
 To minimise the noise induced through this process one must know the nature of input signal.
@@
 Digitisation discretises an input signal by sorting it into $N$ appropriate ranges, such that an input between $x_i$ and $x_{i+1}$ produces an output at $y_i$.
-A digitisation scheme is described the the number of ranges, $N$, the endpoints of these ranges $x_k$ and the output levels $y_k$.
+A digitisation scheme is described by the number of ranges, $N$, the endpoints of these ranges, $x_k$, and the output levels, $y_k$.
 Conventionally one chooses $x_{1} = -\infty$ and $x_{N+1} = \infty$.
-In order to quantify the performance of a given digitisation scheme that produces a signal out $s_{out}$ form an input signal $s_{in}$ we define the distortion as
+In order to quantify the performance of a given digitisation scheme we define the distortion as
 
-\[ D = \left\langle  \left( s_{in} - s_{out} \right)^2 \right\rangle \]
+\[ D = \left\langle  \left( s - \hat{s} \right)^2 \right\rangle \]
 
-For a given input amplitude probability density $p(x)$ we may rewrite the above as
+where $s$ is the input and $\hat{s}$ the output signal.
+For an input signal that has at least some stochastic element to it we introduce the input amplitude probability density $p(x)$.
+This allows us to rewrite the above as
 
@@
 For our purposes we may safely assume that ground-based CMB observations operate at low signal to noise.
-Furthermore we assume that the noise profile is Gaussian white noise\footnote{Please see the conclusion for a discussion of the effect of more realistic noise profile}, i.e.
+Furthermore we assume that the noise profile is Gaussian white noise\footnote{Please see the conclusion for a discussion of the effect of more realistic noise profiles.}, i.e.
 $p(x) = 1/\sqrt{2\pi} e^{-x^2/2}$.
-Given this assumptions we cannot solve proceed analytically, but can the problem using a numerical iterative procedure.
-One begins by picking $y_1$ and calculating the remaining $x_i$ and $y_i$ using equation \ref{eq:digitequalspacecondition}.
-Afterwards one observes weather this choice of values satisfy the conditions given by equation \ref{eq:digitareacondition}.
-If that is the case, the $x_k$ and $y_k$ were chosen appropriately.
+Given this assumption we can solve the problem using a numerical iterative procedure.
+One begins by picking $y_1$ and calculating the remaining $x_i$'s and $y_i$'s using equation \ref{eq:digitequalspacecondition}.
+Afterwards one observes whether this choice of values satisfy the conditions given by equation \ref{eq:digitareacondition}.
+If that is the case, the $x_k$'s and $y_k$'s were chosen appropriately.
 
@@
 We incorporate his results by formulating the multi-level functions we use for our 1, 2 and 3 bit digitisation process.
-Given an input signal $x(t)$ a digitisation scheme using $N$ bits returns $\hat{x}_N(t)$.
-For 1 bit digitisation we apply
+Given an input signal $s(t)$ a digitisation scheme using $N$ bits returns the output $\hat{s}_N(t)$.
+For 1 bit digitisation we apply the function
 
-\[ \hat{x}_1(t) = \left\{ \begin{array}{lr}
-1, & \text{for } x(t) > 0\\
--1, & \text{for } x(t) \leq 0
+\[ \hat{s}_1(t) = \left\{ \begin{array}{lr}
+1, & \text{for } s(t) > 0\\
+-1, & \text{for } s(t) \leq 0
 \end{array} \right. \]
@@
 
-\[ \hat{x}_2(t) = \left\{ \begin{array}{rl}
-1.51 \sigma, & \text{for } x(t) \geq 0.9816 \sigma\\
-0.4528 \sigma, & \text{for } 0 \leq x(t) < 0.9816 \sigma\\
--0.4528 \sigma, & \text{for } 0.9816 \sigma \leq x(t) < 0\\
--1.51 \sigma, & \text{for } 0.9816 \sigma < x(t)\\
+\[ \hat{s}_2(t) = \left\{ \begin{array}{rl}
+1.51 \sigma, & \text{for } s(t) \geq 0.9816 \sigma\\
+0.4528 \sigma, & \text{for } 0 \leq s(t) < 0.9816 \sigma\\
+-0.4528 \sigma, & \text{for } 0.9816 \sigma \leq s(t) < 0\\
+-1.51 \sigma, & \text{for } 0.9816 \sigma < s(t)\\
 \end{array} \right. \]
@@
 
-\[ \hat{x}_3(t) = \left\{ \begin{array}{rl}
-2.152 \sigma, & \text{for } x(t) \geq 1.748 \sigma\\
-1.344 \sigma, & \text{for } 1.05 \sigma \leq x(t) < 1.748 \sigma\\
-0.756 \sigma, & \text{for } 0.501 \sigma \leq x(t) < 1.05 \sigma\\
-0.245 \sigma, & \text{for } 0 \leq x(t) < 0.501 \sigma\\
--0.245 \sigma, & \text{for } 0.501 \sigma \leq x(t) < 0\\
--0.756 \sigma, & \text{for } 1.05 \sigma \leq x(t) < 0.501 \sigma\\
--1.344 \sigma, & \text{for } 1.748 \sigma \leq x(t) < 1.05 \sigma\\
--2.152 \sigma, & \text{for } 1.748 \sigma < x(t)\\
+\[ \hat{s}_3(t) = \left\{ \begin{array}{rl}
+2.152 \sigma, & \text{for } s(t) \geq 1.748 \sigma\\
+1.344 \sigma, & \text{for } 1.05 \sigma \leq s(t) < 1.748 \sigma\\
+0.756 \sigma, & \text{for } 0.501 \sigma \leq s(t) < 1.05 \sigma\\
+0.245 \sigma, & \text{for } 0 \leq s(t) < 0.501 \sigma\\
+-0.245 \sigma, & \text{for } 0.501 \sigma \leq s(t) < 0\\
+-0.756 \sigma, & \text{for } 1.05 \sigma \leq s(t) < 0.501 \sigma\\
+-1.344 \sigma, & \text{for } 1.748 \sigma \leq s(t) < 1.05 \sigma\\
+-2.152 \sigma, & \text{for } 1.748 \sigma < s(t)\\
 \end{array} \right. \]
 
-Other digitisation schemes can be thought of that place the digitisation thresholds and output levels in a different way.
-However, these are not of interest for now, given the ease at which the devised optimal schemes can be implemented and the fact that they will lead to worse performance.
+Other digitisation schemes can be thought of, which place the digitisation thresholds and output levels in a different way.
+However, given the assumptions made the schemes derived above are optimal.
+Additionally, they are simple enough to be easily implemented computationally.
 
@@
 
-We simulate observing a $\sim 600$ \sqdeg patch of the sky.
-To do so we perform a number of equally spaced constant elevation scans (CES) over the observation region.
-We repeat the observation strategy $100$ times with a slight offset in RA and DEC each time, such that all pixels within the path are hit approximately uniformly.
-The speed at which we sweep across the survey field is set to a constant to produce a desired number of hits per pixel in the output maps.
+We simulate observing a $\sim 600 \mathrm{deg^2}$ patch of the sky.
+To do so we perform a number of constant elevation scans (CES), equally spaced in declination (DEC).
+We repeat the observation strategy $100$ times with a slight offset in right ascension (RA) and DEC each time, such that all pixels within the patch are hit approximately uniformly.
+The speed at which we sweep across the survey area is adjusted to to produce a desired number of hits per pixel (hpp) in the output maps.
 
-While performing each CES the appropriate pixels that are being targeted are determined.
+While performing each CES the pixels being targeted are determined.
 The corresponding values from the template maps are then accessed and added to realisations of the detector noise of appropriate length.
@@
 We compress the timestream into maps by averaging all hits falling into the same pixel.
-We produce 13 maps in total: three maps that have not undergone few-bit digitisation at the timestream level for comparison purposes, nine maps with three each corresponding to each digitisation scheme and the hitmap.
+We produce 13 maps in total: three maps (I, Q, U) that have not undergone few-bit digitisation at the timestream level for comparison purposes, nine maps with three each corresponding to each digitisation scheme and the hitmap.
 
+This process is carried out 7 times, producing maps with different numbers of hpp.
 The simulation parameters are summarised in table \ref{tab:modelparams}.
-It has considerable computational requirements if we want to reach up to $\sim 10^8$ hits per pixel.
+The calculation outlined above has considerable computational requirements if we want to reach up to $\sim 10^8$ hpp.
 To carry out this simulation we make use of parallelisation and the computing power provided by NERSC.
@@
 
-\begin{table}[tbh]
+\begin{table}[tbh] % transpose and include hpp!
 \begin{center}
@@
 \tablecomments{ 
-Parameters used in the simulated sky strategy. The RA speed is tweaked to match the desired hits per pixel.
+Parameters used in the simulated observation. The RA speed is tweaked to match the desired hpp.
 } \normalsize
@@
 
-We use PolSpice to compute the $TT, EE, BB$ power spectra of the reconstructed maps.
+We use PolSpice to compute the TT, EE and BB power spectra of the reconstructed maps I, Q, U maps.
 When doing so we apodise the observed skypatch using a cosine mask to minimise cut sky effects on the spectra.
@@
 \begin{tabular}{c c c c}
-weightfile & apodizesigma & apodizetype & polarization \\
+weights & $\sigma_{\mathrm{APOD}} [\mathrm{deg^2}]$ & apodistation type & polarisation \\
 \hline
-cosine mask & $\sqrt{600}/2$ & 1 & YES \\
+cosine mask & $\sqrt{600}/2$ & cosine & Yes \\
 
@@
 
-To normalise the the obtained power spectra we focus on the $\sim 10^8 \mathrm{hpp}$ run.
+To normalise the the obtained power spectra we focus on the $\sim 10^8 \mathrm{hpp}$ maps.
 We normalise each power spectrum originating from few bit TOD against its corresponding 64bit TOD counterpart.
-We place a lower limit on the normalisation window in multipole space by considering the size of the observed patch via
+We place a lower limit on the normalisation window in multipole space by considering the size of the observed patch with
 
-\[ l = \frac{2}{\pi} \frac{32400}{600} \approx 35 \]
+\[ l = \frac{2}{\pi} \left( \frac{32400}{600} \right) \approx 35 \]
 
-where we have rounded up to the next highest integer. We find the upper bound on the normalisation window by demanding that
+where we have rounded up to the next largest integer. We find the upper bound on the normalisation window by demanding that
 
-\[ \frac{C_l^{S}}{C_l^{N}} \geq 10 \]
+\[ \frac{C_l^{\mathrm{S}}}{C_l^{\mathrm{N}}} \leq 10 \]
+
+where $C_l^{\mathrm{S}}$ is the power spectrum of the input template maps and $C_l^{\mathrm{N}}$ the detector noise level.
 
 we close the normalisation window the first instance the above condition is met.
-The normalisation constants for each digitisation scheme and channel obtained through this way are then applied to all lower hit per pixel simulations.
+The normalisation constants for each digitisation scheme and channel obtained through this way are then applied to all lower hpp simulations.
 
@@
 
-To quantify how much the quality of the power spectra suffers due to the digitisation compare the map noise levels inferred from the power spectra.
-These are put into the context of the map noise levels deduced from the 64bit TOD power spectra.
+To quantify how much the quality of the power spectra suffers from the digitisation process we compare the map noise levels inferred from the power spectra.
+These are put into the context of the map noise levels deduced from the 64bit TOD (control) power spectra.
 We formulate
@@
 \begin{equation} \label{eq:extramapnoise}
-\frac{\Delta \sigma}{\sigma} = \frac{\sigma_{\mathrm{map}}^{\mathrm{D}}-\sigma_{\mathrm{map}}^{\mathrm{F}}}{\sigma_{\mathrm{map}}^{\mathrm{F}}}
+\frac{\Delta \sigma}{\sigma} = \frac{\sigma_{\mathrm{map}}^{\mathrm{D}}-\sigma_{\mathrm{map}}^{\mathrm{C}}}{\sigma_{\mathrm{map}}^{\mathrm{C}}}
 \end{equation}
 
-where $\sigma_{\mathrm{map}}^{\mathrm{F}}$ is the map noise level of the control power spectra and $\sigma_{\mathrm{map}}^{\mathrm{D}}$ is the map noise level obtained from the power spectra originating from extremely digitised TOD.
-We assume that the digitisation process results in adding some constant independent of multipole moment to the power spectrum.
+where $\sigma_{\mathrm{map}}^{\mathrm{C}}$ is the map noise level of the control power spectra and $\sigma_{\mathrm{map}}^{\mathrm{D}}$ is the map noise level obtained from the power spectra originating from extremely digitised TOD.
+We assume that the digitisation process results in adding a constant noise term to the power spectrum.
 If we are dominated by noise we can write
 
-\[ C_l^D = C_l^N + C_l^X \]
+\[ C_l^D \approx C_l^N + C_l^X \]
 
@@
 
-The value of $l$ at which we can safely assume to be noise dominated and apply the above framework varies between simulated observations of different hits per pixel and for different channels.
-The range considered involves any datapoints at multipole moments larger than the last point at which the detector noise is at least than 10 times the input power spectrum.
+The value of $l$ at which we can safely assume to be noise dominated and apply the above framework varies between simulated observations of different hpp and TT, EE and BB power spectra.
+The range considered involves any datapoints at multipole moments larger than the last point at which the detector noise is at least an order of magnitude larger than the template power spectrum, i.e.
+
+\[ \frac{C_l^{\mathrm{N}}}{C_l^{\mathrm{S}}} \geq 10 \]
 
@@
 The results for the 1, 2 and 3 bit digitisation schemes are summarised in figure \ref{fig:extranoise}.
-We would like to points out three key results.
+We would like to point out three key results.
 Firstly, 3 Bit digitisation performs the best, followed by 2 Bit and finally 1 Bit digitisation.
-This is what we expect, given that with each additional bit we may retain more information.
-Secondly, for a fixed detector noise level, the additional percentage to the map noise level due to digitisation is independent of the number of hits per pixel.
-The added noise therefore scales in the same fashion as the map noise level with the number of hits per pixel, given a fixed detector noise level.
+This is what we expect, given that with each additional bit we retain more information.
+Secondly, for a fixed detector noise level, the additional percentage to the map noise level due to digitisation is independent of the number of hpp.
+The added noise therefore scales in the same fashion as the map noise level with the number of hpp, given a fixed detector noise level.
 Lastly, the added noise levels are astonishingly low.
-Keeping in mind that CMB detector sensitivity improves in steps of order of magnitude every few years adding an extra percent level noise term does not deteriorate the results appreciably.
+Keeping in mind that CMB detector sensitivity improves in steps of order of magnitude every few years adding an extra percent-level noise term does not deteriorate the results appreciably.
 This is impressive, given that the use of an optimal 3 bit digitisation scheme will save approximately an order of magnitude in TOD volume.
 
-Extreme digitisation is a viable lossy compression technique when dealing with large volume, low signal to noise, slowly varying datasets.
-Making these assumptions the signal small and constant for a considerable number of datapoints.
-For such an interval the many samplings of the noise result in producing appropriate numbers in each output bin.
-The addition of more bits helps by classifying the size of the noise more and more appropriately and hence approaching a good estimate for the value of the signal faster.
+Extreme digitisation is a viable lossy compression technique when dealing with low signal to noise, large datasets with a signal that is slowly varying with respect to the sampling rate.
+Under these conditions we have a locally flat, low signal, on top of which we add many noise realisations.
+In the limit of many samples these allow us to reconstruct the signal well.
+Furthermore a small signal will prevent saturation of the output, i.e.
+the inability of the digitised output to communicate any information where in the range $\infty$ to $x_N$ the input signal lies.
+
+%Extreme digitisation is a viable lossy compression technique when dealing with large volume, low signal to noise, slowly varying datasets.
+Making these assumptions the signal is small and constant for a considerable number of datapoints.
+For such an interval many samplings of the signal plus noise result in producing appropriate numbers in each output bin.
+The addition of more bits helps by classifying the size of the noise more accurately and hence approaching an improved estimate for the value of the signal faster.
 It is necessary to remain in the low signal to noise regime, as otherwise a dominant signal would saturate the digitisation scheme and for a given scheme only values in the range $\pm |N_{\mathrm{hits}}y_N|$ can be represented, given that we have $N_{\mathrm{hits}}$ datapoints.
@@
 
-In this work we have demonstrated the power of digitising CMB TOD from 64-bit down to few-bits.
-The additional noise induced through this step was at the percent-level for chosen optimal digitisation schemes for temperature and polarisation observations.
+In this work we have motivated the investigation of extreme digitisation as a technique in combating arising data challenges in CMB data analysis.
+The reduction of the TOD by an order of magnitude directly addresses the issues in data transmission faced by remote location observations.
+Benefits in mission planning and data analysis are possible.
 
-The motivation behind employing extreme digitisation is the reduction of TOD by an order of magnitude.
-This primarily addresses challenges in data transmission that will face upcoming CMB experiments at remote locations.
-Benefits in planning and analysis seem feasible.
+We have derived a set of optimal digitisation schemes and presented the level at which the induce noise into the temperate and polarisation power spectra.
+We find that an optimal 3 bit digitisation adds as little as $<2\%$ to the map noise level.
 
-Future work on investigating this compression technique must aim to understand the nature of the induced noise better. For this the performance of cluster-finding algorithms is interesting.
+Future work investigating this compression technique must aim to understand the nature of the induced noise better.
+It is of great value to find the higher statistical moments, i.e.
+skewness and kurtosis of the added noise term.
+For this an analysis of the performance of cluster-finding algorithms on the digitised datasets is useful.
 
@@
 chunking of the data before applying extreme digitisation have already been investigated by Planck.
-These thoughts should be considered when designing the compression schemes of future space-based CMB missions, which will be unable to recover their full data with latency, but rather fully rely on the transmitted data.
+These thoughts should be considered when designing the compression schemes of future space-based CMB missions, which will be unable to recover their full data with latency, but rather entirely rely on the transmitted data.
 

### 2018-08-13 digitisation.tex
Figures
Added some reliminary figures.

@@
 %\usepackage{graphicx}
-%\usepackage{multirow}
+\usepackage{multirow}
 \usepackage{soul}
@@
 \small
-\begin{tabular}{l c c c c}
-$\mathrm{NSIDE}$ & $f_{\mathrm{readout}} [\mathrm{Hz}]$ & $f_{{\mathrm{sky}}}$ & $\sigma^{\mathrm{T}}_{\mathrm{det}} [\mathrm{\mu K \sqrt{s}}]$ & $\sigma^{\mathrm{Pol}}_{\mathrm{det}} [\mathrm{\mu K \sqrt{s}}]$\\
+\begin{tabular}{l | c}
+Parameter & Value\\
 \hline
-$4096$ & $200$ & $\sim 0.014$ & $500$ & $\sqrt{2}\times500$\\
-
+$\mathrm{NSIDE}$ &$4096$\\
+$f_{\mathrm{readout}} [\mathrm{Hz}]$ &$200$\\
+$f_{{\mathrm{sky}}}$ & $\sim 0.014$\\
+$\sigma^{\mathrm{T}}_{\mathrm{det}} [\mathrm{\mu K \sqrt{s}}]$ &$500$\\
+$\sigma^{\mathrm{Pol}}_{\mathrm{det}} [\mathrm{\mu K \sqrt{s}}]$ & $\sqrt{2}\times500$\\
+\multirow{2}{*}{$\mathrm{hpp}$} & $\left(1, 10, 100 \right) \times 800$\\
+& $\left( 1, 10, 100 \right) \times 1024000$ \\
 \end{tabular}
 \tablecomments{ 
-Parameters used in the simulated observation. The RA speed is tweaked to match the desired hpp.
+Parameters used in the simulated observation. The RA speed is adjusted to match the desired hpp.
 } \normalsize
@@
 The normalisation constants for each digitisation scheme and channel obtained through this way are then applied to all lower hpp simulations.
+A sample of the obtained power spectra are shown in figures \ref{fig:ttzoom}, \ref{fig:eezoom}, and \ref{fig:bbzoom}.
+
+\begin{figure}[htb]\centering
+\includegraphics[width=0.5\textwidth,clip]{Plots/ttzoom.png}
+  \caption[Current ]{
+  $\mathrm{TT}$ power spectrum reconstructed from a $\sim 10^8\mathrm{hpp}$ map.
+           \label{fig:ttzoom}
+  }
+\end{figure}
+
+\begin{figure}[htb]\centering
+\includegraphics[width=0.5\textwidth,clip]{Plots/eezoom.png}
+  \caption[Current ]{
+  $\mathrm{EE}$ power spectrum reconstructed from a $\sim 10^8\mathrm{hpp}$ map.
+\label{fig:eezoom}
+}
+\end{figure}
+
+\begin{figure}[htb]\centering
+\includegraphics[width=0.5\textwidth,clip]{Plots/bbzoom.png}
+  \caption[Current ]{
+  $\mathrm{BB}$ power spectrum reconstructed from a $\sim 10^8\mathrm{hpp}$ map.
+\label{fig:bbzoom}
+}
+\end{figure}
 
@@
 This guarantees that the points in the noise tail are independent of one another, allowing us to extract an uncertainty for the above quantity.
+Plots for $C_l^X/C_l^N$ are shown in figures \ref{fig:clxclntt}, \ref{fig:clxclnee}, and \ref{fig:clxclnbb} for $TT$, $EE$, and $BB$ respectively.
 
-%plot these X/N ?
+\begin{figure}[htb]\centering
+\includegraphics[width=0.5\textwidth,clip]{Plots/ttratio.png}
+  \caption[Current ]{
+  Calculated ratio of $C_l^X/C_l^N$ of the rebinned $\mathrm{TT}$ power spectrum. The vertical black line indicates from which point onwards data is used in calculating the equivalent noise level.
+           \label{fig:clxclntt}
+  }
+\end{figure}
 
-The results for the 1, 2 and 3 bit digitisation schemes are summarised in figure \ref{fig:extranoise}.
+\begin{figure}[htb]\centering
+\includegraphics[width=0.5\textwidth,clip]{Plots/eeratio.png}
+  \caption[Current ]{
+  Calculated ratio of $C_l^X/C_l^N$ of the rebinned $\mathrm{EE}$ power spectrum.
+\label{fig:clxclnee}
+}
+\end{figure}
+
+\begin{figure}[htb]\centering
+\includegraphics[width=0.5\textwidth,clip]{Plots/bbratio.png}
+  \caption[Current ]{
+  Calculated ratio of $C_l^X/C_l^N$ of the rebinned $\mathrm{BB}$ power spectrum.
+\label{fig:clxclnbb}
+}
+\end{figure}
+
+The deduced additional noise for 1, 2 and 3 bit digitisation schemes are shown with respect to the hits per pixel in the maps in figure \ref{fig:extranoise}.
 We would like to point out three key results.
@@
 This is impressive, given that the use of an optimal 3 bit digitisation scheme will save approximately an order of magnitude in TOD volume.
+
+\begin{figure}[htb]\centering
+\includegraphics[width=0.5\textwidth,clip]{Plots/addednoise.png}
+  \caption[Current ]{
+  Addition to the map noise level due to digitisation. 
+           \label{fig:extranoise}
+  }
+\end{figure}
 
@@
 
-\begin{figure*}[htb]\centering
-\includegraphics[width=0.9\textwidth,clip]{Plots/addednoise.png}
-  \caption[Current ]{
-  Addition to the map noise level due to digitisation. 
-           \label{fig:extranoise}
-  }
-\end{figure*}
-
-\begin{table*}[tbh]
-\begin{center}
-\caption{\label{tab:experiments} Assumed survey parameters}
-\small
-\begin{tabular}{l || c c c c c }
-Experiment & Sky coverage & Polarized Noise level  & 1/$f$ knee & Beam FWHM \\
-& &($\mu$K-arcmin)&&(arcmin.)\\
-\hline
-\tiny \\ \small
-CMB Stage III & & & & \\
-~~~~~SPT-3G & 6\% & 3.0 & 200 & 1.2 \\
-~~~~~Simons Array & 36\% & 9.5 & 200 & 3.5 \\ 
-\tiny \\ \small
+%\begin{table*}[tbh]
+%\begin{center}
+%\caption{\label{tab:experiments} Assumed survey parameters}
+%\small
+%\begin{tabular}{l || c c c c c }
+%Experiment & Sky coverage & Polarized Noise level  & 1/$f$ knee & Beam FWHM \\
+%& &($\mu$K-arcmin)&&(arcmin.)\\
 %\hline
-CMB Stage IV & 55\% & 1.3 & 100 & 4.0 \\
-\end{tabular}
-\tablecomments{ 
-Key numbers about the planned stage III and IV experiments. 
-The sky coverage percentages are after galactic cuts. 
-Unless otherwise noted,  the Fisher matrix forecasts in this work use these numbers. 
-All forecasts also allow for beam and calibration uncertainties as noted in the text. 
-} \normalsize
-\end{center}
-\end{table*}
+%\tiny \\ \small
+%CMB Stage III & & & & \\
+%~~~~~SPT-3G & 6\% & 3.0 & 200 & 1.2 \\
+%~~~~~Simons Array & 36\% & 9.5 & 200 & 3.5 \\ 
+%\tiny \\ \small
+%\hline
+%CMB Stage IV & 55\% & 1.3 & 100 & 4.0 \\
+%\end{tabular}
+%\tablecomments{ 
+%Key numbers about the planned stage III and IV experiments. 
+%The sky coverage percentages are after galactic cuts. 
+%Unless otherwise noted,  the Fisher matrix forecasts in this work use these numbers. 
+%All forecasts also allow for beam and calibration uncertainties as noted in the text. 
+%} \normalsize
+%\end{center}
+%\end{table*}
 

### 2018-08-14 digitisation.tex
Editing
Done another editing pass and adjusted the placement of the figures.

@@
 %\usepackage{amsmath}
-%\usepackage{subfigure}
+\usepackage{subfigure}
 \usepackage[backref,breaklinks,colorlinks,citecolor=blue]{hyperref}
@@
 \bibliographystyle{fapj}
-%\usepackage{graphicx}
+\usepackage{graphicx}
 \usepackage{multirow}
@@
 Observations of the Cosmic Microwave Background (CMB) are of immense value to modern cosmology.
-However, future CMB experiments must confront challenges in mission planning, hardware and analysis that arise from the sheer size of the time-ordered-data being recorded.
+However, future CMB experiments must confront challenges in mission planning, hardware and analysis that arise from the sheer size of the time-ordered-data (TOD) being recorded.
 These challenges are particularly significant for Antarctic and satellite experiments which depend on satellite links to transmit their data.
@@
 Unlike lossless compression, extreme digitisation introduces additional noise into the data.
-We present optimal 1, 2 and 3 bit digitisation schemes and determine the degradation in temperature and polarisation power spectra caused by applying this process to the time-ordered-data (TOD).
+We present optimal 1, 2 and 3 bit digitisation schemes and determine the degradation in temperature and polarisation power spectra caused by applying this process to the TOD.
 We find that 3 bit digitisation has a percent-level contribution to the map noise level.
@@
 The most prominent science goal is the discovery of the imprint of inflationary gravitational waves.
-Additionally studies of CMB lensing and the Sunyaev-Zeldovich (SZ) effects will open up access to unprecedented insight.
+Additionally, studies of CMB lensing and the Sunyaev-Zeldovich (SZ) effects will open up access to unprecedented insight.
 Moreover, through probing the relativistic number of species, the helium fraction and the neutrino mass sum, CMB experiments are a valuable counter-part to ground-based particle physics experiments.
@@
 A growing hurdle for experiments at remote locations are the transmission limitations of satellite links.
-Space-based experiments have employed a combination of lossless and lossy compression techniques, including reduced bits in the time-ordered-data.
-Antarctica-based experiments that transmit a portion of their data via a satellite link have downsampled their TOD in the past to meet their telemetry requirements.
+Space-based experiments have employed a combination of lossless and lossy compression techniques, including reduced bits in the TOD.
+Antarctica-based experiments that transmit a portion of their data via a satellite link have downsampled their data in the past to meet their telemetry requirements.
 They have yet to exploit few bit digitisation of the TOD.
@@
 In this work we present the method of extreme digitisation, which compresses a rich digital input signal and compresses it into a few bits.
-We apply extreme digitisation to the TOD and detail its effect on temperature and polarisation observations.
-We find that an optimal 3-bit digitisation scheme adds as little as $\sim 2\%$ to the map noise level.
+We apply extreme digitisation to the TOD and detail its effect on temperature and polarisation power spectra.
+We find that an optimal 3-bit digitisation scheme adds as little as $< 2\%$ to the map noise level.
 While the digtisation schemes described here are primarily laid out for ground-based experiments, the results we present here should also be considered by future space-based missions, which inevitably must incorporate lossy compression.
@@
 We detail the arising challenges in handling large TOD in \S\ref{subsec:problem}.
-We subsequently formulate extreme digitisation formerly in \S\ref{subsec:extremedigitisation} and lay out the framework used to test its performance in \S\ref{subsec:method}.
-We detail how we estimate the power spectra from simulated maps in \S\ref{subsec:psestimation}.
-We continue by presenting the noise induced through the dicitisation process in \S\ref{subsec:additionalnoise}.
-We summarize our findings in \S\ref{sec:conclusions}.
+We subsequently formulate extreme digitisation in \S\ref{subsec:extremedigitisation} and lay out the framework used to test its performance in \S\ref{subsec:method}.
+The details of the power spectrum estimation used are laid out in \S\ref{subsec:psestimation}.
+We continue by presenting the noise induced through the dicitisation process in \S\ref{subsec:additionalnoise} and summarise our findings in \S\ref{sec:conclusions}.
 
@@
 Experiments at these locations depend on satellite transmission.
-The next generation of ground-based experiments will aim to collect 2 million detector years worth of data.
-An experiment contributing to Stage-4 at the South Pole will face a data influx of $\sim O(10)\mathrm{Tb/d}$.
-However, the current transmission allocation for SPT3G is at $150\mathrm{Gb/d}$, which will likely only see a moderate increase in the coming years.
+The next generation of ground-based experiments will aim to collect 2 million detector years of data.
+An experiment contributing to Stage-4 at the South Pole will face a data influx of $\sim \mathrm{O}(10)\mathrm{Tb/d}$.
+However, the current transmission allocation for SPT3G is at $150\mathrm{Gb/d}$, which will likely only see a moderate increase in coming years.
 The transmission bottleneck is currently overcome by recovering the full data on hard drives with some latency and by transmitting a downsampled version of the data.
@@
 Given the shear size of TOD expected, this is not possible.
-We must rely on different planning strategies or aim to reduce the size of the TOD in order to maximise the productivity of planning and development stages and guarantee maximal science output.
+We must rely on different planning strategies or aim to reduce the size of the TOD in order to maximise the productivity of planning and development stages and guarantee scientific excellence.
 
@@
 Extreme Digitisation would tackle the challenges mentioned above by reducing the size of the TOD by an order of magnitude.
-Together with already established lossless compression techniques (such as FLAC, run-length coding, Huffman coding, etc.) this will directly tackle transmission hurdles.
-While it needs to be investigated to what extent existing algorithms can be carried over, extreme digitisation has the possibility of solving planning and analysis problems.
+Together with already established lossless compression techniques (e.g.
+FLAC, run-length coding, Huffman coding, etc.) this will directly tackle transmission hurdles.
+While it needs to be investigated to what extent existing algorithms can be carried over, extreme digitisation has the potential of solving planning and analysis problems.
 
@@
 Recently Clearwater et al.
-(in prep.) have investigated the advantages of using 1 and 2 bit data when searching for continuous gravitational waves using the Laser Interferometer Gravitational-wave Observatory (LIGO) performance of searches for continuous gravitational wave searches using 1 and 2 bit data.
+(in prep.) have investigated the advantages of using 1 and 2 bit data when searching for continuous gravitational waves using the Laser Interferometer Gravitational-wave Observatory (LIGO).
 
@@
 However the induced noise depends on the number of bits used, the digitisation thresholds, and the output levels chosen.
-To minimise the noise induced through this process one must know the nature of input signal.
+To minimise the noise induced through this process one must know the nature of the input signal.
 A theoretical framework to obtain these levels was laid out by Max in 1978.
@@
 Digitisation discretises an input signal by sorting it into $N$ appropriate ranges, such that an input between $x_i$ and $x_{i+1}$ produces an output at $y_i$.
-A digitisation scheme is described by the number of ranges, $N$, the endpoints of these ranges, $x_k$, and the output levels, $y_k$.
+A digitisation scheme is described by the number of ranges, $N$, the endpoints of these ranges, $x_i$, and the output levels, $y_i$.
 Conventionally one chooses $x_{1} = -\infty$ and $x_{N+1} = \infty$.
@@
 
-which informs us that an output level $y_i$ must lie halfway between is delimiting thresholds $x_i$ and $x_{i+1}$. We gain an additional condition from equation \ref{eq:distderiv2}
+which informs us that an output level $y_i$ must lie halfway between its delimiting thresholds $x_i$ and $x_{i+1}$. We gain an additional condition from equation \ref{eq:distderiv2}
 
@@
 To progress further we have to make an assumption about the distribution of input signals, $p(x)$.
-For our purposes we may safely assume that ground-based CMB observations operate at low signal to noise.
+For our purposes we assume that CMB observations operate at low signal to noise.
 Furthermore we assume that the noise profile is Gaussian white noise\footnote{Please see the conclusion for a discussion of the effect of more realistic noise profiles.}, i.e.
@@
 Afterwards one observes whether this choice of values satisfy the conditions given by equation \ref{eq:digitareacondition}.
-If that is the case, the $x_k$'s and $y_k$'s were chosen appropriately.
+If that is the case, the $x_i$'s and $y_i$'s were chosen appropriately.
 
@@
 
-to the input signal. Finally the optimal 3 bit digitisation is described by the eight-level function
+where $\sigma$ is the standard deviation of the input signal. Finally the optimal 3 bit digitisation is described by the eight-level function
 
@@
 Each scan is performed by a single detector.
-We obtain maps that use 64bit TOD and maps that have undergone 1, 2 and 3 bit digitisation at the timestream level.
+We obtain control maps that use 64bit TOD and maps that have undergone 1, 2 and 3 bit digitisation at the timestream level.
 We calculate the temperature and polarisation power spectra of each map and determine the additional noise induced through the extreme digitisation process.
@@
 We generate a realisation of I, Q and U maps based on the results of Planck 2015.
-The key cosmological parameters are summarised in \ref{tab:inputcosparams}.
+The key cosmological parameters are summarised in table \ref{tab:inputcosparams}.
 
@@
 We compress the timestream into maps by averaging all hits falling into the same pixel.
-We produce 13 maps in total: three maps (I, Q, U) that have not undergone few-bit digitisation at the timestream level for comparison purposes, nine maps with three each corresponding to each digitisation scheme and the hitmap.
+We produce 12 CMB maps in total: three control maps (I, Q, U) and nine maps with three each obtained from each digitisation scheme.
 
-This process is carried out 7 times, producing maps with different numbers of hpp.
+This process is carried out 6 times, producing maps with different numbers of hpp.
 The simulation parameters are summarised in table \ref{tab:modelparams}.
 The calculation outlined above has considerable computational requirements if we want to reach up to $\sim 10^8$ hpp.
-To carry out this simulation we make use of parallelisation and the computing power provided by NERSC.
+To carry out this simulation we make use of parallelisation and the computing facilities provided by NERSC.
 
@@
 Parameter & Planck 2015\\
-\hline
+\hline\\[-1.5ex]
 $100\theta_{MC}$ & $ 1.04086 \pm 0.00048$\\
@@
 Parameter & Value\\
-\hline
+\hline \\[-1.5ex]
 $\mathrm{NSIDE}$ &$4096$\\
@@
 $\sigma^{\mathrm{Pol}}_{\mathrm{det}} [\mathrm{\mu K \sqrt{s}}]$ & $\sqrt{2}\times500$\\
-\multirow{2}{*}{$\mathrm{hpp}$} & $\left(1, 10, 100 \right) \times 800$\\
+\multirow{2}{*}{$\mathrm{hpp}$} & $\left(1, 10, 100 \right) \times 800,$\\
 & $\left( 1, 10, 100 \right) \times 1024000$ \\
@@
 When doing so we apodise the observed skypatch using a cosine mask to minimise cut sky effects on the spectra.
-Please see \ref{tab:polspiceparams} for an overview of the parameters used in this process.
+Please see table \ref{tab:polspiceparams} for an overview of the parameters used in this process.
 
@@
 \begin{tabular}{c c c c}
-weights & $\sigma_{\mathrm{APOD}} [\mathrm{deg^2}]$ & apodistation type & polarisation \\
+weights & $\sigma_{\mathrm{APOD}} [\mathrm{deg}]$ & apodistation type & polarisation \\
 \hline
@@
 To normalise the the obtained power spectra we focus on the $\sim 10^8 \mathrm{hpp}$ maps.
-We normalise each power spectrum originating from few bit TOD against its corresponding 64bit TOD counterpart.
+We normalise each power spectrum originating from few bit TOD against its corresponding control counterpart.
 We place a lower limit on the normalisation window in multipole space by considering the size of the observed patch with
@@
 
-\[ \frac{C_l^{\mathrm{S}}}{C_l^{\mathrm{N}}} \leq 10 \]
+\[ \frac{C_l^{\mathrm{S}}}{C_l^{\mathrm{N}}} \geq 10 \]
 
@@
 
-we close the normalisation window the first instance the above condition is met.
+we close the normalisation window the first instance the above condition is violated.
 The normalisation constants for each digitisation scheme and channel obtained through this way are then applied to all lower hpp simulations.
-A sample of the obtained power spectra are shown in figures \ref{fig:ttzoom}, \ref{fig:eezoom}, and \ref{fig:bbzoom}.
+A sample of the obtained power spectra are shown in figure \ref{fig:powspeczoom}.
 
@@
 \includegraphics[width=0.5\textwidth,clip]{Plots/ttzoom.png}
-  \caption[Current ]{
-  $\mathrm{TT}$ power spectrum reconstructed from a $\sim 10^8\mathrm{hpp}$ map.
-           \label{fig:ttzoom}
-  }
-\end{figure}
-
-\begin{figure}[htb]\centering
 \includegraphics[width=0.5\textwidth,clip]{Plots/eezoom.png}
-  \caption[Current ]{
-  $\mathrm{EE}$ power spectrum reconstructed from a $\sim 10^8\mathrm{hpp}$ map.
-\label{fig:eezoom}
-}
-\end{figure}
-
-\begin{figure}[htb]\centering
 \includegraphics[width=0.5\textwidth,clip]{Plots/bbzoom.png}
   \caption[Current ]{
-  $\mathrm{BB}$ power spectrum reconstructed from a $\sim 10^8\mathrm{hpp}$ map.
-\label{fig:bbzoom}
+   Reconstructed $\mathrm{TT}$, $\mathrm{EE}$, and $\mathrm{BB}$ power spectra.
+These originate from an observation with $102400000 \mathrm{hpp}$.
+We observe that digitisation appears to add some constant to the noise level.
+As one would expect $3 \mathrm{Bit}$ digitisation performs the best, followed by $2 \mathrm{Bit}$- and finally $1 \mathrm{Bit}$ digitisation.
+\label{fig:powspeczoom}
 }
@@
 To quantify how much the quality of the power spectra suffers from the digitisation process we compare the map noise levels inferred from the power spectra.
-These are put into the context of the map noise levels deduced from the 64bit TOD (control) power spectra.
+These are put into the context of the map noise levels deduced from the control power spectra.
 We formulate
@@
 where $\sigma_{\mathrm{map}}^{\mathrm{C}}$ is the map noise level of the control power spectra and $\sigma_{\mathrm{map}}^{\mathrm{D}}$ is the map noise level obtained from the power spectra originating from extremely digitised TOD.
-We assume that the digitisation process results in adding a constant noise term to the power spectrum.
+We assume that the digitisation process results in adding a constant noise term to each power spectrum.
 If we are dominated by noise we can write
@@
 
-The value of $l$ at which we can safely assume to be noise dominated and apply the above framework varies between simulated observations of different hpp and TT, EE and BB power spectra.
+The value of $l$ at which we can safely assume to be noise dominated and apply the above framework varies between simulated observations of different hpp and temperature and polarisation power spectra.
 The range considered involves any datapoints at multipole moments larger than the last point at which the detector noise is at least an order of magnitude larger than the template power spectrum, i.e.
@@
 This guarantees that the points in the noise tail are independent of one another, allowing us to extract an uncertainty for the above quantity.
-Plots for $C_l^X/C_l^N$ are shown in figures \ref{fig:clxclntt}, \ref{fig:clxclnee}, and \ref{fig:clxclnbb} for $TT$, $EE$, and $BB$ respectively.
+Plots for $C_l^X/C_l^N$ are shown in figure \ref{fig:clxcln}.
 
@@
 \includegraphics[width=0.5\textwidth,clip]{Plots/ttratio.png}
-  \caption[Current ]{
-  Calculated ratio of $C_l^X/C_l^N$ of the rebinned $\mathrm{TT}$ power spectrum. The vertical black line indicates from which point onwards data is used in calculating the equivalent noise level.
-           \label{fig:clxclntt}
-  }
-\end{figure}
-
-\begin{figure}[htb]\centering
 \includegraphics[width=0.5\textwidth,clip]{Plots/eeratio.png}
-  \caption[Current ]{
-  Calculated ratio of $C_l^X/C_l^N$ of the rebinned $\mathrm{EE}$ power spectrum.
-\label{fig:clxclnee}
-}
-\end{figure}
-
-\begin{figure}[htb]\centering
 \includegraphics[width=0.5\textwidth,clip]{Plots/bbratio.png}
   \caption[Current ]{
-  Calculated ratio of $C_l^X/C_l^N$ of the rebinned $\mathrm{BB}$ power spectrum.
-\label{fig:clxclnbb}
+  Calculated ratio of $C_l^X/C_l^N$ of the rebinned $\mathrm{TT}$, $\mathrm{EE}$, and $\mathrm{BB}$ power spectra.
+The vertical black line indicates from which point onwards data is used to calculate the equivalent noise level.
+The above power spectra originate from a $80000\mathrm{hpp}$ map.
+Increasing the hpp in the map shifts the plateau to the right.
+\label{fig:clxcln}
 }
@@
 We would like to point out three key results.
-Firstly, 3 Bit digitisation performs the best, followed by 2 Bit and finally 1 Bit digitisation.
+Firstly, 3 Bit digitisation performs the best, followed by 2 Bit- and finally 1 Bit digitisation.
 This is what we expect, given that with each additional bit we retain more information.
-Secondly, for a fixed detector noise level, the additional percentage to the map noise level due to digitisation is independent of the number of hpp.
-The added noise therefore scales in the same fashion as the map noise level with the number of hpp, given a fixed detector noise level.
+Secondly, for a fixed detector noise level the deterioration of the map quality scales in the same fashion as the number of hpp.
+We see that $\Delta \sigma / \sigma$ is independent of the hpp in the maps.
 Lastly, the added noise levels are astonishingly low.
-Keeping in mind that CMB detector sensitivity improves in steps of order of magnitude every few years adding an extra percent-level noise term does not deteriorate the results appreciably.
+Keeping in mind that CMB detector sensitivity improves in steps of order of magnitude every few years, adding an extra percent-level noise term does not deteriorate the results appreciably.
 This is impressive, given that the use of an optimal 3 bit digitisation scheme will save approximately an order of magnitude in TOD volume.
@@
 Furthermore a small signal will prevent saturation of the output, i.e.
-the inability of the digitised output to communicate any information where in the range $\infty$ to $x_N$ the input signal lies.
+the inability of the digitised output to communicate any information where in the range $\infty$ to $x_N$ an input signal lies.
 
@@
 
-It should be laid out how this changes when moving to a more realistic noise profile.
-We do not expect this to alter the practicality of our results - even if a moderate change in the noise profile doubles the additional percentage to the map noise level extreme digitisation is still practical.
+It should be laid out how the results changes when moving to a more realistic noise profile.
+We do not expect this to alter the practicality of our results - even if a different noise profile doubles the additional percentage to the map noise level extreme digitisation is still practical.
 Ideas on how to deal with 1/f noise, e.g.
 chunking of the data before applying extreme digitisation have already been investigated by Planck.
-These thoughts should be considered when designing the compression schemes of future space-based CMB missions, which will be unable to recover their full data with latency, but rather entirely rely on the transmitted data.
+These thoughts should be considered when designing the compression schemes of future space-based CMB missions, which will be unable to recover their full data with latency, but rely entirely on the transmitted data.
 
@@
 
-\bibliography{digitisation}
+%\bibliography{digitisation}
 

### 2018-08-30 digitisation.tex
Editing and References.
Done more editing in line with what christian suggested. Began to include some references.

@@
 \begin{abstract}
-Observations of the Cosmic Microwave Background (CMB) are of immense value to modern cosmology.
+
+Observations of the Cosmic Microwave Background (CMB) are of significant value to modern cosmology and particle physics.
+Future CMB experiments face issues in hardware design, mission planning and analysis that arise from the vast size of time-ordered-data (TOD) being recorded.
+These challenges are particularly significant for Antarctic and satellite experiments which depend on satellite links to transmit their data.
+We explore the viability of reducing the TOD to few-bit numbers to address these issues.
+Unlike lossless compression, this technique introduces additional noise into the data.
+We present 1, 2 and 3 bit digitisation schemes and determine the degradation this compression causes in temperature and polarisation power spectra.
+We find that 3 bit digitisation has a percent-level contribution to the map noise level.
+We argue that such digitisation is a promising strategy for upcoming experiments.
+
+%Observations of the Cosmic Microwave Background (CMB) are of immense value to modern cosmology.
 However, future CMB experiments must confront challenges in mission planning, hardware and analysis that arise from the sheer size of the time-ordered-data (TOD) being recorded.
@@
 
-\keywords{ cosmic background radiation --- polarization --- data compression}
+\keywords{cosmic background radiation --- polarization --- data compression}
 \section{Introduction}
@@
 
-Observations of the Cosmic Microwave Background (CMB) have played a key role in physics since 1964.
+Observations of the Cosmic Microwave Background (CMB) have played a key role in physics since its discovery \citep{penzias1965}.
 Current and future CMB experiments will continue to deliver new insights by studying the temperature and polarisation information contained in the CMB.
 These will put tight constraints on cosmological models.
-The most prominent science goal is the discovery of the imprint of inflationary gravitational waves.
+Upcoming science goals include the discovery of inflationary gravitational waves and thorough studies of CMB lensing and the Sunyaev-Zeldovich (SZ) effects.
+Moreover, CMB experiments will measure the relativistic number of species and the neutrino mass sum \citep{s4sciencebook}.
+
+%The most prominent science goal is the discovery of the imprint of inflationary gravitational waves.
 Additionally, studies of CMB lensing and the Sunyaev-Zeldovich (SZ) effects will open up access to unprecedented insight.
@@
 
-The outstanding contribution of CMB science to modern physics has demanded a high standard of data analysis.
-The CMB community has developed a variety of compression and computational techniques to manage the increasing influx of data, while maximising the science output.
-These include the compression of time-ordered data (TOD) into maps, bandpower estimation and the pseudo $C_l$ method.
+High fidelity CMB science demands an excellent standard of data analysis.
+The CMB community has developed a variety of compression and computational techniques to manage the increasing influx of data, while maximising the science output \citep{tristam2007}.
+These include the compression of time-ordered data (TOD) into maps \citep{tegmark1997}, bandpower estimation \citep{tegmark1998} and the pseudo $C_l$ method \citep{brown2005}.
 
 A growing hurdle for experiments at remote locations are the transmission limitations of satellite links.
-Space-based experiments have employed a combination of lossless and lossy compression techniques, including reduced bits in the TOD.
+Space-based experiments have employed a combination of lossless and lossy compression techniques, including reduced bits in the TOD \citep{gaztanaga1998, maris2003}.
 Antarctica-based experiments that transmit a portion of their data via a satellite link have downsampled their data in the past to meet their telemetry requirements.
@@
 
-In this work we present the method of extreme digitisation, which compresses a rich digital input signal and compresses it into a few bits.
+In this work we present the method of extreme digitisation, which compresses a rich digital input signal into a few bits.
 We apply extreme digitisation to the TOD and detail its effect on temperature and polarisation power spectra.
 We find that an optimal 3-bit digitisation scheme adds as little as $< 2\%$ to the map noise level.
-While the digtisation schemes described here are primarily laid out for ground-based experiments, the results we present here should also be considered by future space-based missions, which inevitably must incorporate lossy compression.
+The digtisation schemes described here are laid out for ground-based experiments.
+Space-based missions must study carefully the effect of digitisation in their specific compression algorithms.% of t the results we present here should also be considered by future space-based missions, which inevitably must incorporate lossy compression.
 
 This work is structured as follows.
+In \S\ref{sec:dig} we detail the challenges originating from handling large TOD, formulate the process of extreme digitisation and lay out the framework used to test its performance.
+Subsequently, we describe the power spectrum estimation employed and interpret the results obtained through this process in section \S\ref{sec:results}.
+We summarise our findings in \S\ref{sec:conclusions}.
+
+%This work is structured as follows.
 We detail the arising challenges in handling large TOD in \S\ref{subsec:problem}.
@@
 
-The science goals of upcoming CMB experiments naturally lead to a large influx of data.
-To achieve the targeted sensitivity longer observations with more detectors are needed.
+The science goals of upcoming CMB experiments naturally lead to a large influx of data To achieve the targeted sensitivity longer observations with more detectors are needed.
 In fact the number of detectors of ground-based experiments has been following a Moore's law like trend, doubling approximately every 2 years.
-This directly translates into an exponential growth in data volume.
+This directly translates into an exponential growth in data volume \citep{s4sciencebook}..
 
-The best observations sites for CMB measurements are at remote locations: in space and in the Antarctica.
-Experiments at these locations depend on satellite transmission.
-The next generation of ground-based experiments will aim to collect 2 million detector years of data.
+Excellent observation conditions for CMB observations can not only be found in space, but also at various locations on earth \citep{li2017, kovac2007} among which is the South Pole.
+Space- and Antarctica-based experiments depend on satellite transmission.
+The next generation of ground-based instruments will aim to collect 2 million detector years of data.
 An experiment contributing to Stage-4 at the South Pole will face a data influx of $\sim \mathrm{O}(10)\mathrm{Tb/d}$.
@@
 Continued use of downsampling will narrow the information window decisively - prohibiting high multipole moment science to be carried out on the transmitted dataset.
-This also means that any potential faults or errors in the experiment that only become visible in high frequencies will go unnoticed for longer.
+This also means that any potential faults or errors in the experiment that only become visible in said range will go unnoticed for longer.
 
-Future space-based missions aim to exceed the detector count of Planck by at least an order of magnitude.
-It is questionable whether their telemetry specifications will allow for transmission of the data with Planck-style compression.
+The Planck mission has demonstrated the merits of carrying out CMB observations from space \citep{planck2018}.
+Upcoming missions aim to exceed the detector count of Planck by at least an order of magnitude \citep{litebird2014, pixie2011, core2018}.
+However, strategies to meet the telemetry specifications of each mission appear to be non-settled.
+It is not clear whether the data compression knowledge developed during the Planck mission will guarantee optimal performance for future satellites.
+Each mission will have to carefully construct a compression algorithm through a combination of lossless and lossy techniques.
+
+%It is questionable whether their telemetry specifications will allow for transmission of the data with Planck-style compression.
 Methods of storing large amounts of data on upcoming satellites will likely be prohibited by financial decisions: the amount of storage space required becomes financially relevant at the scales targeted.
@@
 Beyond transmission challenges, mission planning is becoming exceedingly difficult.
-As noted by (S4 science book) a full simulation of TOD over the entire parameter space of detection scenarios for numerous set-ups is the desired way to decide on Stage-4 configurations.
+A full simulation of TOD over the entire parameter space of detection scenarios for numerous set-ups is the desired way to decide on Stage-4 configurations.
 Space-based missions must aim to carry out a similar analysis to optimise their science output.
-Given the shear size of TOD expected, this is not possible.
+Given the shear size of TOD expected, this is not possible \citep{s4sciencebook}.
 We must rely on different planning strategies or aim to reduce the size of the TOD in order to maximise the productivity of planning and development stages and guarantee scientific excellence.
@@
 While we have not experienced the limitations of the accessible computational assets, the exponential growth of CMB data makes its analysis increasingly expensive.
+The community depends on the excellent facilities provided by the National Energy Research Scientific Computing Center (NERSC).
 
@@
 FLAC, run-length coding, Huffman coding, etc.) this will directly tackle transmission hurdles.
-While it needs to be investigated to what extent existing algorithms can be carried over, extreme digitisation has the potential of solving planning and analysis problems.
+Beyond that extreme digitisation has the potential of solving planning and analysis problems.
 
-Other science areas have demonstrated that extreme digitisation is a valuable compression technique.
-Jenet and Anderson (cite) explored the application of such compression to radio pulsar timing measurements with success.
-Recently Clearwater et al.
-(in prep.) have investigated the advantages of using 1 and 2 bit data when searching for continuous gravitational waves using the Laser Interferometer Gravitational-wave Observatory (LIGO).
+The Planck mission has demonstrated the advantages of using few bit compression of the TOD \citep{maris2003}.
+Other science areas have also shown the power of extreme digtisation.
+\citep{jenet1998} explored the application of such compression to radio pulsar timing measurements with success.
+Recently \citep{clearwater2018} have investigated the advantages of using 1 and 2 bit data when searching for continuous gravitational waves using the Laser Interferometer Gravitational-Wave Observatory (LIGO).
 
@@
 
-% rephrase these, basically off Max1978
-
 Digitisation is a lossy compression technique.
-However the induced noise depends on the number of bits used, the digitisation thresholds, and the output levels chosen.
+However, the induced noise depends on the number of bits used, the digitisation thresholds, and the output levels chosen.
 To minimise the noise induced through this process one must know the nature of the input signal.
-A theoretical framework to obtain these levels was laid out by Max in 1978.
-We review the key aspects of his work relevant for us below.
+There is usually little concern around finding an optimal set of digitisation parameters for a given input signal.
+It is common to have access to large numbers of bits to store information, where changes in the digitisation scheme become insignificant to the distortion induced.
+It is in our interest however to consider extreme, i.e.
+few-bit, digitisation.
+To do so we review the key ideas laid out by \cite{max1960} below.
 
@@
 
-This was carried out by Max.
-We incorporate his results by formulating the multi-level functions we use for our 1, 2 and 3 bit digitisation process.
+We use the findings of Max to formulate the multi-level functions we use for our 1, 2 and 3 bit digitisation process.
 Given an input signal $s(t)$ a digitisation scheme using $N$ bits returns the output $\hat{s}_N(t)$.
-For 1 bit digitisation we apply the function
+For 1 bit digitisation we apply the sign function
 
@@
 
-Other digitisation schemes can be thought of, which place the digitisation thresholds and output levels in a different way.
+The fact that the single bit digitisation scheme makes no references to the standard deviation of the signal is irrelevant.
+This affects the normalisation of the signal, which we can account for later.
+What is important is the spacing between digitisation thresholds and relative distance between the output levels chosen.
+
+Completely different digitisation schemes can be thought of, which would for example reference the most drastic outlier in the dataset, or seek to place equal numbers of points into each digitisation level.
 However, given the assumptions made the schemes derived above are optimal.
@@
 To create the template maps we use the healpix framework and the wealth of support available for it.
-We generate a realisation of I, Q and U maps based on the results of Planck 2015.
-The key cosmological parameters are summarised in table \ref{tab:inputcosparams}.
+We generate a realisation of I, Q and U maps with $\mathrm{NSIDE} = 4096$ based on the results of Planck 2015.
+The key cosmological parameters assumed are $\Omega_b \approx 0.049 $, $\Omega_c \approx 0.265$, $\Omega_m \approx 0.316$ and $h \approx 0.67$.
 
@@
 We repeat the observation strategy $100$ times with a slight offset in right ascension (RA) and DEC each time, such that all pixels within the patch are hit approximately uniformly.
-The speed at which we sweep across the survey area is adjusted to to produce a desired number of hits per pixel (hpp) in the output maps.
+The speed at which we sweep across the survey area is adjusted to to produce a desired number of hits per pixel (hits per pixel) in the output maps.
 
@@
 
-This process is carried out 6 times, producing maps with different numbers of hpp.
-The simulation parameters are summarised in table \ref{tab:modelparams}.
-The calculation outlined above has considerable computational requirements if we want to reach up to $\sim 10^8$ hpp.
+This process is carried out 6 times.
+We produce maps with approximately $800$, $8,000$, $80,000$, $1,024,000$, $10,240,000$ and $102,400,000$ hits per pixel.
+We assume that the detector is read out $200$ times per second and has a noise level of $500 \mu K \sqrt{s}$ for temperature and $\sqrt{2} \times 500 \mu K \sqrt{s}$ for polarisation observations.
+The calculation outlined above has considerable computational requirements if we want to reach up to $\sim 10^8$ hits per pixel.
 To carry out this simulation we make use of parallelisation and the computing facilities provided by NERSC.
 
-\begin{table}[tbh]
-\begin{center}
-\caption{\label{tab:inputcosparams} Input Cosmological Parameters}
-\small
-\begin{tabular}{l | c}
-Parameter & Planck 2015\\
-\hline\\[-1.5ex]
-$100\theta_{MC}$ & $ 1.04086 \pm 0.00048$\\
-$\Omega_b h^2$ & $  0.02222 \pm 0.00023$\\
-$\Omega_c h^2$ & $ 0.1199 \pm 0.0022$\\
-$H_0$ & $67.26 \pm 0.98$\\
-$n_s$ & $0.9652 \pm 0.0062$\\
-$\Omega_m$ & $0.316 \pm 0.014$\\
-$\sigma_8$ & $0.830 \pm 0.015 $\\
-$\tau$ & $0.078 \pm 0.019$\\
-$10^9 A_s e^{-2\tau}$ & $1.881 \pm 0.014 $\\
-\end{tabular}
-\tablecomments{ 
-Cosmological parameters used to create the template maps. Taken from Planck 2015: Cosmological Parameters.
-} \normalsize
-\end{center}
-\end{table}
+%\begin{table}[tbh]
+%\begin{center}
+%\caption{\label{tab:inputcosparams} Input Cosmological Parameters}
+%\small
+%\begin{tabular}{l | c}
+%Parameter & Planck 2015\\
+%\hline\\[-1.5ex]
+%$100\theta_{MC}$ & $ 1.04086 \pm 0.00048$\\
+%$\Omega_b h^2$ & $  0.02222 \pm 0.00023$\\
+%$\Omega_c h^2$ & $ 0.1199 \pm 0.0022$\\
+%$H_0$ & $67.26 \pm 0.98$\\
+%$n_s$ & $0.9652 \pm 0.0062$\\
+%$\Omega_m$ & $0.316 \pm 0.014$\\
+%$\sigma_8$ & $0.830 \pm 0.015 $\\
+%$\tau$ & $0.078 \pm 0.019$\\
+%$10^9 A_s e^{-2\tau}$ & $1.881 \pm 0.014 $\\
+%\end{tabular}
+%\tablecomments{ 
+%Cosmological parameters used to create the template maps. Taken from Planck 2015: Cosmological Parameters.
+%} \normalsize
+%\end{center}
+%\end{table}
 
-\begin{table}[tbh] % transpose and include hpp!
-\begin{center}
-\caption{\label{tab:modelparams} Assumed Survey Parameters}
-\small
-\begin{tabular}{l | c}
-Parameter & Value\\
-\hline \\[-1.5ex]
-$\mathrm{NSIDE}$ &$4096$\\
-$f_{\mathrm{readout}} [\mathrm{Hz}]$ &$200$\\
-$f_{{\mathrm{sky}}}$ & $\sim 0.014$\\
-$\sigma^{\mathrm{T}}_{\mathrm{det}} [\mathrm{\mu K \sqrt{s}}]$ &$500$\\
-$\sigma^{\mathrm{Pol}}_{\mathrm{det}} [\mathrm{\mu K \sqrt{s}}]$ & $\sqrt{2}\times500$\\
-\multirow{2}{*}{$\mathrm{hpp}$} & $\left(1, 10, 100 \right) \times 800,$\\
-& $\left( 1, 10, 100 \right) \times 1024000$ \\
-\end{tabular}
-\tablecomments{ 
-Parameters used in the simulated observation. The RA speed is adjusted to match the desired hpp.
-} \normalsize
-\end{center}
-\end{table}
+%\begin{table}[tbh] % transpose and include hits per pixel!
+%\begin{center}
+%\caption{\label{tab:modelparams} Assumed Survey Parameters}
+%\small
+%\begin{tabular}{l | c}
+%Parameter & Value\\
+%\hline \\[-1.5ex]
+%$\mathrm{NSIDE}$ &$4096$\\
+%$f_{\mathrm{readout}} [\mathrm{Hz}]$ &$200$\\
+%$f_{{\mathrm{sky}}}$ & $\sim 0.014$\\
+%$\sigma^{\mathrm{T}}_{\mathrm{det}} [\mathrm{\mu K \sqrt{s}}]$ &$500$\\
+%$\sigma^{\mathrm{Pol}}_{\mathrm{det}} [\mathrm{\mu K \sqrt{s}}]$ & $\sqrt{2}\times500$\\
+%\multirow{2}{*}{$\mathrm{hits per pixel}$} & $\left(1, 10, 100 \right) \times 800,$\\
+%& $\left( 1, 10, 100 \right) \times 1024000$ \\
+%\end{tabular}
+%\tablecomments{ 
+%Parameters used in the simulated observation. The RA speed is adjusted to match the desired hits per pixel.
+%} \normalsize
+%\end{center}
+%\end{table}
 
@@
 We use PolSpice to compute the TT, EE and BB power spectra of the reconstructed maps I, Q, U maps.
-When doing so we apodise the observed skypatch using a cosine mask to minimise cut sky effects on the spectra.
-Please see table \ref{tab:polspiceparams} for an overview of the parameters used in this process.
+When doing so we apodise the observed skypatch using a cosine mask with $\sigma_{\mathrm{APOD}} = \sqrt{600}/2 \mathrm{deg}$ to minimise cut sky effects on the spectra.
 
-\begin{table}[tbh]
-\begin{center}
-\caption{\label{tab:polspiceparams} PolSpice Parameters}
-\small
-\begin{tabular}{c c c c}
-weights & $\sigma_{\mathrm{APOD}} [\mathrm{deg}]$ & apodistation type & polarisation \\
-\hline
-cosine mask & $\sqrt{600}/2$ & cosine & Yes \\
+%\begin{table}[tbh]
+%\begin{center}
+%\caption{\label{tab:polspiceparams} PolSpice Parameters}
+%\small
+%\begin{tabular}{c c c c}
+%weights & $\sigma_{\mathrm{APOD}} [\mathrm{deg}]$ & apodistation type & polarisation \\
+%\hline
+%cosine mask & $\sqrt{600}/2$ & cosine & Yes \\
+%
+%\end{tabular}
+%\tablecomments{ 
+%Parameters used when calling PolSpice to calculate the power spectra. Remaining parameters have been left at their default value.
+%} \normalsize
+%\end{center}
+%\end{table}
 
-\end{tabular}
-\tablecomments{ 
-Parameters used when calling PolSpice to calculate the power spectra. Remaining parameters have been left at their default value.
-} \normalsize
-\end{center}
-\end{table}
+Normalisation of the obtained power spectra is necessary.
+The digitisation schemes introduced in section \S\ref{sec:dig} are designed to minimise the distortion of the TOD, not to preserve power at the map level.
+Calibration to the CMB is common for many experiments (e.g.
+SPT).
 
-To normalise the the obtained power spectra we focus on the $\sim 10^8 \mathrm{hpp}$ maps.
-We normalise each power spectrum originating from few bit TOD against its corresponding control counterpart.
-We place a lower limit on the normalisation window in multipole space by considering the size of the observed patch with
+For the normalisation procedure we focus on the $\sim 10^8 \mathrm{hits per pixel}$ maps.
+We normalise the $\mathrm{TT}$, $\mathrm{EE}$, and $\mathrm{BB}$ power spectra from few bit TOD against their control counterparts.
+We consider the normalisation window cut out by the conditions
 
-\[ l = \frac{2}{\pi} \left( \frac{32400}{600} \right) \approx 35 \]
-
-where we have rounded up to the next largest integer. We find the upper bound on the normalisation window by demanding that
-
-\[ \frac{C_l^{\mathrm{S}}}{C_l^{\mathrm{N}}} \geq 10 \]
+\[ l > 35 \> ; \> \frac{C_l^{\mathrm{S}}}{C_l^{\mathrm{N}}} \geq 10 \]
 
 where $C_l^{\mathrm{S}}$ is the power spectrum of the input template maps and $C_l^{\mathrm{N}}$ the detector noise level.
+The lower limit is placed in accordance to the patch-size surveyed, the upper limit ensures appropriate signal to noise in the normalisation window.
 
-we close the normalisation window the first instance the above condition is violated.
-The normalisation constants for each digitisation scheme and channel obtained through this way are then applied to all lower hpp simulations.
+The normalisation constants for each digitisation scheme and channel are obtained through this way are then applied to all lower hits per pixel simulations.
 A sample of the obtained power spectra are shown in figure \ref{fig:powspeczoom}.
@@
    Reconstructed $\mathrm{TT}$, $\mathrm{EE}$, and $\mathrm{BB}$ power spectra.
-These originate from an observation with $102400000 \mathrm{hpp}$.
+These originate from an observation with $102400000 \mathrm{hits per pixel}$.
 We observe that digitisation appears to add some constant to the noise level.
@@
 
-The value of $l$ at which we can safely assume to be noise dominated and apply the above framework varies between simulated observations of different hpp and temperature and polarisation power spectra.
+The value of $l$ at which we can safely assume to be noise dominated and apply the above framework varies between simulated observations of different hits per pixel and temperature and polarisation power spectra.
 The range considered involves any datapoints at multipole moments larger than the last point at which the detector noise is at least an order of magnitude larger than the template power spectrum, i.e.
@@
 The vertical black line indicates from which point onwards data is used to calculate the equivalent noise level.
-The above power spectra originate from a $80000\mathrm{hpp}$ map.
-Increasing the hpp in the map shifts the plateau to the right.
+The above power spectra originate from a $80000\mathrm{hits per pixel}$ map.
+Increasing the hits per pixel in the map shifts the plateau to the right.
 \label{fig:clxcln}
@@
 This is what we expect, given that with each additional bit we retain more information.
-Secondly, for a fixed detector noise level the deterioration of the map quality scales in the same fashion as the number of hpp.
-We see that $\Delta \sigma / \sigma$ is independent of the hpp in the maps.
+Secondly, for a fixed detector noise level the deterioration of the map quality scales in the same fashion as the number of hits per pixel.
+We see that $\Delta \sigma / \sigma$ is independent of the hits per pixel in the maps.
 Lastly, the added noise levels are astonishingly low.
@@
 The reduction of the TOD by an order of magnitude directly addresses the issues in data transmission faced by remote location observations.
-Benefits in mission planning and data analysis are possible.
+Benefits in mission planning, data analysis and hardware requirements are possible.
 
 We have derived a set of optimal digitisation schemes and presented the level at which the induce noise into the temperate and polarisation power spectra.
-We find that an optimal 3 bit digitisation adds as little as $<2\%$ to the map noise level.
+We find that an optimal 3 bit digitisation adds as little as $<2\%$ to the map noise level for temperature and polarisation observations.
+As mentioned above this addition is insignificant given that the sensitivity of CMB experiments follows a Moore's law like trend.
+The benefit of this compression technique is the reduction of the TOD volume by an order of magnitude.
 
@@
 
-%\bibliography{digitisation}
+\bibliography{digitisation}
 

### 2018-09-04 digitisation.tex
Editing
Condinuted Editing.

@@
 Future CMB experiments face issues in hardware design, mission planning and analysis that arise from the vast size of time-ordered-data (TOD) being recorded.
-These challenges are particularly significant for Antarctic and satellite experiments which depend on satellite links to transmit their data.
+These challenges are particularly significant for Antarctica- and space-based experiments which depend on satellite links to transmit their data.
 We explore the viability of reducing the TOD to few-bit numbers to address these issues.
@@
 
-%Observations of the Cosmic Microwave Background (CMB) are of immense value to modern cosmology.
-However, future CMB experiments must confront challenges in mission planning, hardware and analysis that arise from the sheer size of the time-ordered-data (TOD) being recorded.
-These challenges are particularly significant for Antarctic and satellite experiments which depend on satellite links to transmit their data.
-We investigate using extreme digitisation to address these issues.
-Unlike lossless compression, extreme digitisation introduces additional noise into the data.
-We present optimal 1, 2 and 3 bit digitisation schemes and determine the degradation in temperature and polarisation power spectra caused by applying this process to the TOD.
-We find that 3 bit digitisation has a percent-level contribution to the map noise level.
-This is impressive considering that it would reduce the data volume by an order of magnitude.
-We argue that extreme digitisation is a promising strategy for upcoming experiments.
 \end{abstract}
 
-\keywords{cosmic background radiation --- polarization --- data compression}
+%C: maybe sub techniques:misc for methods: data analysis ? unfortunately no data compression tag
+\keywords{cosmic background radiation --- polarization --- techniques: miscellaneous}
 \section{Introduction}
@@
 These will put tight constraints on cosmological models.
-Upcoming science goals include the discovery of inflationary gravitational waves and thorough studies of CMB lensing and the Sunyaev-Zeldovich (SZ) effects.
+Upcoming science goals include the discovery of inflationary gravitational waves, the construction of precise lensing maps and thorough studies of the Sunyaev-Zeldovich (SZ) effects.
 Moreover, CMB experiments will measure the relativistic number of species and the neutrino mass sum \citep{s4sciencebook}.
-
-%The most prominent science goal is the discovery of the imprint of inflationary gravitational waves.
-Additionally, studies of CMB lensing and the Sunyaev-Zeldovich (SZ) effects will open up access to unprecedented insight.
-Moreover, through probing the relativistic number of species, the helium fraction and the neutrino mass sum, CMB experiments are a valuable counter-part to ground-based particle physics experiments.
+%L: reference to science goals by satellite missions as well
 
 High fidelity CMB science demands an excellent standard of data analysis.
-The CMB community has developed a variety of compression and computational techniques to manage the increasing influx of data, while maximising the science output \citep{tristam2007}.
+The CMB community has developed a variety of compression techniques and computational approaches to manage the increasing influx of data, while maximising the science output \citep{tristam2007}.
 These include the compression of time-ordered data (TOD) into maps \citep{tegmark1997}, bandpower estimation \citep{tegmark1998} and the pseudo $C_l$ method \citep{brown2005}.
@@
 As we approach the next generation of ground-based experiments, Stage-4, and the launch of a new generation of space-based missions (liteBIRD, PIXIE, COrE+), we must treat the transmission bottleneck carefully.
-Without a review of the current compression techniques employed we are sure to lose information.
+Without a thorough review of compression techniques we are sure to lose information.
+%L:ref to non-planck satellites for reduced bit TOD?
+%C: Is there a reference detailing downsampling of SPT data?
 
@@
 This work is structured as follows.
-In \S\ref{sec:dig} we detail the challenges originating from handling large TOD, formulate the process of extreme digitisation and lay out the framework used to test its performance.
-Subsequently, we describe the power spectrum estimation employed and interpret the results obtained through this process in section \S\ref{sec:results}.
+In \S\ref{sec:dig} we detail the challenges originating from handling large TOD, introduce the process of extreme digitisation and lay out the framework used to test its performance.
+Subsequently, we describe the power spectrum estimation employed and interpret the obtained results in section \S\ref{sec:results}.
 We summarise our findings in \S\ref{sec:conclusions}.
@@
 In fact the number of detectors of ground-based experiments has been following a Moore's law like trend, doubling approximately every 2 years.
-This directly translates into an exponential growth in data volume \citep{s4sciencebook}..
+This directly translates into an exponential growth in data volume \citep{s4sciencebook}.
+%L: check out the particle paper where the moores law figure comes from
 
@@
 Space- and Antarctica-based experiments depend on satellite transmission.
-The next generation of ground-based instruments will aim to collect 2 million detector years of data.
-An experiment contributing to Stage-4 at the South Pole will face a data influx of $\sim \mathrm{O}(10)\mathrm{Tb/d}$.
+An experiment at the South Pole will be part of the next generation of ground-based instruments, which aims to collect 2 million detector years in total.
+Such an experiment will face a data influx of $\sim \mathrm{O}(10)\mathrm{Tb/d}$.
 However, the current transmission allocation for SPT3G is at $150\mathrm{Gb/d}$, which will likely only see a moderate increase in coming years.
@@
 Going into Stage-4 we anticipate that compression rates for transmission must increase by an order of magnitude.
-Continued use of downsampling will narrow the information window decisively - prohibiting high multipole moment science to be carried out on the transmitted dataset.
+Continued use of downsampling will narrow the information window decisively - prohibiting high multipole moment science to be carried out on the transmitted data set.
 This also means that any potential faults or errors in the experiment that only become visible in said range will go unnoticed for longer.
+%L: reference for observing conditions in chile or a general review of sites %C: Is there a paper detailing trasmission limits for SPT and that the data is currently downsampled?
 
@@
 Upcoming missions aim to exceed the detector count of Planck by at least an order of magnitude \citep{litebird2014, pixie2011, core2018}.
-However, strategies to meet the telemetry specifications of each mission appear to be non-settled.
+However, strategies to meet the telemetry specifications of each mission appear to be in development.
 It is not clear whether the data compression knowledge developed during the Planck mission will guarantee optimal performance for future satellites.
-Each mission will have to carefully construct a compression algorithm through a combination of lossless and lossy techniques.
-
-%It is questionable whether their telemetry specifications will allow for transmission of the data with Planck-style compression.
-Methods of storing large amounts of data on upcoming satellites will likely be prohibited by financial decisions: the amount of storage space required becomes financially relevant at the scales targeted.
-Missions will likely be left to design their own compression algorithms which will incorporate a combination of lossless and lossy compression techniques.
+Most likely each mission will have to carefully construct a compression algorithm through a combination of lossless and lossy techniques, specific to their requirements.
 
@@
 While we have not experienced the limitations of the accessible computational assets, the exponential growth of CMB data makes its analysis increasingly expensive.
-The community depends on the excellent facilities provided by the National Energy Research Scientific Computing Center (NERSC).
+The community depends on the facilities provided by the National Energy Research Scientific Computing Center (NERSC).
 
@@
 FLAC, run-length coding, Huffman coding, etc.) this will directly tackle transmission hurdles.
-Beyond that extreme digitisation has the potential of solving planning and analysis problems.
+Beyond that extreme digitisation has the potential of easing issues in planning, analysis and hardware requirements.
 
@@
 It is in our interest however to consider extreme, i.e.
-few-bit, digitisation.
-To do so we review the key ideas laid out by \cite{max1960} below.
+few-bit, digitisation, where a digitisation scheme must be carefully constructed.
+We review the key ideas regarding construction of digitisation schemes as laid out by \cite{max1960} below.
 
 Digitisation discretises an input signal by sorting it into $N$ appropriate ranges, such that an input between $x_i$ and $x_{i+1}$ produces an output at $y_i$.
-A digitisation scheme is described by the number of ranges, $N$, the endpoints of these ranges, $x_i$, and the output levels, $y_i$.
+The set of parameters $N, x_i, y_i$ fully specify a digitisation scheme.
 Conventionally one chooses $x_{1} = -\infty$ and $x_{N+1} = \infty$.
@@
 
-\[ D = \left\langle  \left( s - \hat{s} \right)^2 \right\rangle \]
+\begin{equation} D = \left\langle  \left( s - \hat{s} \right)^2 \right\rangle, \end{equation}
 
@@
 
-\[ D = \sum_{i = 1}^N \int_{x_i}^{x_{i+1}} \left(x-y_i\right)^2 p(x) dx \]
+\begin{equation} D = \sum_{i = 1}^N \int_{x_i}^{x_{i+1}} \left(x-y_i\right)^2 p(x) dx. \end{equation}
 
-Seeing as we wish to minimise the distortion we differentiate the above with respect to $x_i$ and $y_i$ and set the derivatives to zero. We obtain the two equations
+Since we wish to minimise the distortion we differentiate the above with respect to $x_i$ and $y_i$ and set the derivatives to zero. We obtain the two equations
 
 \begin{equation} \label{eq:distderiv1}
-\frac{\partial D}{\partial x_i} = \left(x_i-y_{i-1}\right)^2 p(x_i) - \left(x_i - y_i\right)^2 p(x_i) = 0
+\frac{\partial D}{\partial x_i} = \left(x_i-y_{i-1}\right)^2 p(x_i) - \left(x_i - y_i\right)^2 p(x_i) = 0,
 \end{equation}
@@
 \begin{equation} \label{eq:distderiv2}
-\frac{\partial D}{\partial x_j} = -2 \int_{x_i}^{x_{i+1}} \left( x-y_i \right) p(x) dx = 0
+\frac{\partial D}{\partial x_j} = -2 \int_{x_i}^{x_{i+1}} \left( x-y_i \right) p(x) dx = 0.
 \end{equation}
@@
 \begin{equation} \label{eq:digitequalspacecondition}
-x_i = \frac{y_i+y_{i+1}}{2}
+x_i = \frac{y_i+y_{i+1}}{2},
 \end{equation}
 
-which informs us that an output level $y_i$ must lie halfway between its delimiting thresholds $x_i$ and $x_{i+1}$. We gain an additional condition from equation \ref{eq:distderiv2}
+which informs us that an output level $y_i$ must lie halfway between its delimiting thresholds $x_i$ and $x_{i+1}$. From equation \ref{eq:distderiv2} we gain the additional condition
 
 \begin{equation} \label{eq:digitareacondition}
-\int_{x_i}^{x_{i+1}} \left( x-y_i \right) p(x) dx = 0
+\int_{x_i}^{x_{i+1}} \left( x-y_i \right) p(x) dx = 0.
 \end{equation}
@@
 For our purposes we assume that CMB observations operate at low signal to noise.
-Furthermore we assume that the noise profile is Gaussian white noise\footnote{Please see the conclusion for a discussion of the effect of more realistic noise profiles.}, i.e.
-$p(x) = 1/\sqrt{2\pi} e^{-x^2/2}$.
+Furthermore we assume that the noise profile is Gaussian white noise with standard deviation $\sigma$, i.e.
+$p(x) = (1/\sqrt{2\pi\sigma^2}) e^{-x^2/2\sigma^2}$.
 Given this assumption we can solve the problem using a numerical iterative procedure.
 One begins by picking $y_1$ and calculating the remaining $x_i$'s and $y_i$'s using equation \ref{eq:digitequalspacecondition}.
-Afterwards one observes whether this choice of values satisfy the conditions given by equation \ref{eq:digitareacondition}.
+Afterwards one checks whether this choice of values satisfy the conditions given by equation \ref{eq:digitareacondition}.
 If that is the case, the $x_i$'s and $y_i$'s were chosen appropriately.
 
-We use the findings of Max to formulate the multi-level functions we use for our 1, 2 and 3 bit digitisation process.
+We use the findings of \cite{max1960} to formulate the multi-level functions we use for our 1, 2 and 3 bit digitisation process.
 Given an input signal $s(t)$ a digitisation scheme using $N$ bits returns the output $\hat{s}_N(t)$.
@@
 
-\[ \hat{s}_1(t) = \left\{ \begin{array}{lr}
+\begin{equation} \hat{s}_1(t) = \left\{ \begin{array}{lr}
 1, & \text{for } s(t) > 0\\
 -1, & \text{for } s(t) \leq 0
-\end{array} \right. \]
+\end{array} \right. \end{equation}
 
@@
 
-\[ \hat{s}_2(t) = \left\{ \begin{array}{rl}
+\begin{equation} \hat{s}_2(t) = \left\{ \begin{array}{rl}
 1.51 \sigma, & \text{for } s(t) \geq 0.9816 \sigma\\
@@
 -1.51 \sigma, & \text{for } 0.9816 \sigma < s(t)\\
-\end{array} \right. \]
+\end{array} \right. . \end{equation}
+%C: should i round to equal significant figures here? Unfortunately these are the exact values that i used...
 
-where $\sigma$ is the standard deviation of the input signal. Finally the optimal 3 bit digitisation is described by the eight-level function
+Finally the optimal 3 bit digitisation is described by the eight-level function
 
-\[ \hat{s}_3(t) = \left\{ \begin{array}{rl}
+\begin{equation} \hat{s}_3(t) = \left\{ \begin{array}{rl}
 2.152 \sigma, & \text{for } s(t) \geq 1.748 \sigma\\
@@
 -2.152 \sigma, & \text{for } 1.748 \sigma < s(t)\\
-\end{array} \right. \]
+\end{array} \right. . \end{equation}
 
@@
 
-Completely different digitisation schemes can be thought of, which would for example reference the most drastic outlier in the dataset, or seek to place equal numbers of points into each digitisation level.
+Completely different digitisation schemes can be thought of, which would for example reference the most drastic outlier in the data set, or seek to place equal numbers of points into each digitisation level.
 However, given the assumptions made the schemes derived above are optimal.
@@
 We calculate the temperature and polarisation power spectra of each map and determine the additional noise induced through the extreme digitisation process.
+%C: Do I need to mention that we assume to have an infinitely fast spinning plate? %L: Check whether Python uses 32 or 64 bit under on NERSC for floats
 
-To create the template maps we use the healpix framework and the wealth of support available for it.
-We generate a realisation of I, Q and U maps with $\mathrm{NSIDE} = 4096$ based on the results of Planck 2015.
-The key cosmological parameters assumed are $\Omega_b \approx 0.049 $, $\Omega_c \approx 0.265$, $\Omega_m \approx 0.316$ and $h \approx 0.67$.
+To create the template maps we use the Hierarchical Equal Area isoLatitude Pixelization of a sphere \citep[HEALPix;][]{healpix}\footnote{Available at \url{http://healpix.sourceforge.net/}} framework.
+We generate a realisation of I, Q and U maps with $\mathrm{NSIDE} = 4096$ based on the results of the Planck collaboration with key cosmological parameters of $\Omega_b \approx 0.049 $, $\Omega_c \approx 0.265$, $\Omega_m \approx 0.316$ and $h \approx 0.67$ \citep{planck2016}.
+
+%C: Do I need to comment on the fact that this reflects no particular frequency channel, but the results come from a combination of all observed channels?
 
@@
 To do so we perform a number of constant elevation scans (CES), equally spaced in declination (DEC).
-We repeat the observation strategy $100$ times with a slight offset in right ascension (RA) and DEC each time, such that all pixels within the patch are hit approximately uniformly.
-The speed at which we sweep across the survey area is adjusted to to produce a desired number of hits per pixel (hits per pixel) in the output maps.
+We repeat the observation strategy $100$ times with a slight offset in right ascension (RA) and DEC each time, such that each pixel is sampled approximately uniformly.
+The speed at which we sweep across the survey area is adjusted to produce a desired number of hits per pixel in the output maps.
 
 While performing each CES the pixels being targeted are determined.
-The corresponding values from the template maps are then accessed and added to realisations of the detector noise of appropriate length.
+The corresponding values from the template maps are then accessed and added to realisations of the detector noise of appropriate length to generate TOD streams.
 We assume the detector noise to be Gaussian white noise.
+For temperature observations we assume the detector noise level to be $500 \mathrm{\mu K \sqrt{s}}$ and for polarisation observations $\sqrt{2} \hspace{0.7ex} 500 \mathrm{\mu K \sqrt{s}}$, i.e.
+the corresponding photon noise limit.
 
@@
 
+% LINK TO GITHUB WITH CODE HERE
+
 This process is carried out 6 times.
-We produce maps with approximately $800$, $8,000$, $80,000$, $1,024,000$, $10,240,000$ and $102,400,000$ hits per pixel.
-We assume that the detector is read out $200$ times per second and has a noise level of $500 \mu K \sqrt{s}$ for temperature and $\sqrt{2} \times 500 \mu K \sqrt{s}$ for polarisation observations.
-The calculation outlined above has considerable computational requirements if we want to reach up to $\sim 10^8$ hits per pixel.
-To carry out this simulation we make use of parallelisation and the computing facilities provided by NERSC.
+We produce maps where pixels in the survey patch are hit approximately uniformly $800$, $8,000$, $80,000$, $1,024,000$, $10,240,000$ and $102,400,000$ times.
+We assume that the detector is read out $200$ times per second.
+%The calculation outlined above has considerable computational requirements if we want to reach up to $\sim 10^8$ hits per pixel.
+This calculation was made possible through use of the NERSC facilities.
 
@@
 %Parameter & Planck 2015\\
-%\hline\\[-1.5ex]
+%\hline\\begin{equation}-1.5ex]
 %$100\theta_{MC}$ & $ 1.04086 \pm 0.00048$\\
@@
 %Parameter & Value\\
-%\hline \\[-1.5ex]
+%\hline \\begin{equation}-1.5ex]
 %$\mathrm{NSIDE}$ &$4096$\\
@@
 
-We use PolSpice to compute the TT, EE and BB power spectra of the reconstructed maps I, Q, U maps.
-When doing so we apodise the observed skypatch using a cosine mask with $\sigma_{\mathrm{APOD}} = \sqrt{600}/2 \mathrm{deg}$ to minimise cut sky effects on the spectra.
+We use the Spatially Inhomogeneous Correlation Estimator for Temperature and Polarisation \citep[PolSpice;][]{polspice} to compute the $\mathrm{TT}, \mathrm{EE}$ and $\mathrm{BB}$ power spectra of the reconstructed maps I, Q, U maps.
+To minimise cut sky effects we apodise the observed path using a cosine mask with $\sigma_{\mathrm{APOD}} = \sqrt{600}/2 \hspace{0.7ex} \mathrm{deg}$.
+%C: Should I also mention the multiplication with a cosine mask that I supply in addition to setting this option on polspice?
+This will likely be reworked when using Healpix to get the power spectra though.
+
+%%%%%%%%%%%
+% INCLUDE SECTION OF ALM CUTS (maybe a plot of the cuts too)
+%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%
+% THIS MIGHT CHANGE WHEN WORKING OUT THEORETICAL NORMALISATION
+%%%%%%%%%%%
+
 Normalisation of the obtained power spectra is necessary.
 The digitisation schemes introduced in section \S\ref{sec:dig} are designed to minimise the distortion of the TOD, not to preserve power at the map level.
-Calibration to the CMB is common for many experiments (e.g.
+Normalising obtained spectra against previous experiments is a common calibration technique in many CMB experiments (e.g.
 SPT).
+%C: Do you know a good reference for this?
 
-For the normalisation procedure we focus on the $\sim 10^8 \mathrm{hits per pixel}$ maps.
+For the normalisation procedure we focus on the $\sim 10^8$ hits per pixel maps.
 We normalise the $\mathrm{TT}$, $\mathrm{EE}$, and $\mathrm{BB}$ power spectra from few bit TOD against their control counterparts.
@@
 
-\[ l > 35 \> ; \> \frac{C_l^{\mathrm{S}}}{C_l^{\mathrm{N}}} \geq 10 \]
+\begin{equation} l > 35 \> , \> \frac{C_l^{\mathrm{S}}}{C_l^{\mathrm{N}}} \geq 10. \end{equation}
 
 where $C_l^{\mathrm{S}}$ is the power spectrum of the input template maps and $C_l^{\mathrm{N}}$ the detector noise level.
-The lower limit is placed in accordance to the patch-size surveyed, the upper limit ensures appropriate signal to noise in the normalisation window.
+The lower limit is placed in accordance to the patch-size surveyed while the upper limit ensures appropriate signal to noise in the normalisation window.
 
-The normalisation constants for each digitisation scheme and channel are obtained through this way are then applied to all lower hits per pixel simulations.
-A sample of the obtained power spectra are shown in figure \ref{fig:powspeczoom}.
+The normalisation constants for each digitisation scheme and channel obtained through this way are then applied to all lower hits per pixel simulations.
+A sample of the obtained power spectra is shown in figure X.
 
-\begin{figure}[htb]\centering
-\includegraphics[width=0.5\textwidth,clip]{Plots/ttzoom.png}
-\includegraphics[width=0.5\textwidth,clip]{Plots/eezoom.png}
-\includegraphics[width=0.5\textwidth,clip]{Plots/bbzoom.png}
-  \caption[Current ]{
-   Reconstructed $\mathrm{TT}$, $\mathrm{EE}$, and $\mathrm{BB}$ power spectra.
+%%%%%%%%%%%
+%SAMPLE POWER SPECTRA WITH THEORY SPECTRA
+% want to show that we recover the PS, including theory PS in background
+%%%%%%%%%%%
+
+%\begin{figure}[htb]\centering
+%\includegraphics[width=0.5\textwidth,clip]{Plots/ttzoom.png}
+%\includegraphics[width=0.5\textwidth,clip]{Plots/eezoom.png}
+%\includegraphics[width=0.5\textwidth,clip]{Plots/bbzoom.png}
+%  \caption[Current ]{
+%   Reconstructed $\mathrm{TT}$, $\mathrm{EE}$, and $\mathrm{BB}$ power spectra.
 These originate from an observation with $102400000 \mathrm{hits per pixel}$.
@@
 As one would expect $3 \mathrm{Bit}$ digitisation performs the best, followed by $2 \mathrm{Bit}$- and finally $1 \mathrm{Bit}$ digitisation.
-\label{fig:powspeczoom}
-}
-\end{figure}
+%\label{fig:powspeczoom}
+%}
+%\end{figure}
 
@@
 
-To quantify how much the quality of the power spectra suffers from the digitisation process we compare the map noise levels inferred from the power spectra.
-These are put into the context of the map noise levels deduced from the control power spectra.
+%C: This is where I'm planning on
+%%%%%%%%%%%
+% INSERT A COUPLE PARAGRAPHS ABOUT THE PS OF DIFFERENCE MAPS
+% INCLUDE PLOTS OF PS OF DIFFERENCE MAPS
+% want to show that beyond some calibration error, we do not observe any l dependence in the added noise
+%%%%%%%%%%%
+
+To quantify the distortion caused by extreme digitisation we infer the map noise level from the relevant power spectra.
+These are put into the context of the map noise level deduced from the control power spectra.
 We formulate
@@
 \begin{equation} \label{eq:extramapnoise}
-\frac{\Delta \sigma}{\sigma} = \frac{\sigma_{\mathrm{map}}^{\mathrm{D}}-\sigma_{\mathrm{map}}^{\mathrm{C}}}{\sigma_{\mathrm{map}}^{\mathrm{C}}}
+\frac{\Delta \sigma}{\sigma} = \frac{\sigma_{\mathrm{map}}^{\mathrm{D}}-\sigma_{\mathrm{map}}^{\mathrm{C}}}{\sigma_{\mathrm{map}}^{\mathrm{C}}},
 \end{equation}
@@
 
-\[ C_l^D \approx C_l^N + C_l^X \]
+\begin{equation} C_l^\mathrm{D} \approx C_l^\mathrm{N} + C_l^\mathrm{X}. \end{equation}
 
-Here $C_l^D$ is the power spectrum originating from a digitised timestream, $C_l^N$ is the detector noise level and $C_l^X$ the additional noise induced through digitisation.
+Here $C_l^\mathrm{D}$ is the total power spectrum originating from a digitised timestream and $C_l^\mathrm{X}$ the additional noise induced through digitisation.
 We now progress equation \ref{eq:extramapnoise} to
 
-\[\frac{\Delta \sigma}{\sigma} = \sqrt{\frac{C_l^D}{C_l^{N}}} - 1  = \sqrt{\frac{C_l^N + C_l^X}{C_l^{N}}} - 1 = \sqrt{1 + \frac{C_l^X}{C_l^{N}}} - 1  \]
+\begin{equation}\frac{\Delta \sigma}{\sigma} = \sqrt{\frac{C_l^\mathrm{D}}{C_l^{\mathrm{N}}}} - 1 = \sqrt{1 + \frac{C_l^\mathrm{X}}{C_l^{\mathrm{N}}}} - 1 . \end{equation}
+
+% = \sqrt{\frac{C_l^N + C_l^X}{C_l^{N}}} - 1
 
 The value of $l$ at which we can safely assume to be noise dominated and apply the above framework varies between simulated observations of different hits per pixel and temperature and polarisation power spectra.
-The range considered involves any datapoints at multipole moments larger than the last point at which the detector noise is at least an order of magnitude larger than the template power spectrum, i.e.
+To define the range of noise domination we choose the condition
 
-\[ \frac{C_l^{\mathrm{N}}}{C_l^{\mathrm{S}}} \geq 10 \]
+\begin{equation} \frac{C_l^{\mathrm{N}}}{C_l^{\mathrm{S}}} \geq 10. \end{equation}
 
-Before analysing $C_l^X/C_l^N$ we rebin the power spectra to $\Delta l = 123$.
-This guarantees that the points in the noise tail are independent of one another, allowing us to extract an uncertainty for the above quantity.
-Plots for $C_l^X/C_l^N$ are shown in figure \ref{fig:clxcln}.
+Before analysing $C_l^\mathrm{X}/C_l^\mathrm{N}$ we rebin the power spectra to $\Delta l \approx 100$.
+This guarantees that the points in the noise tail are independent of one another, allowing us to extract an uncertainty for the above quantity.% Plots for $C_l^\mathrm{X}/C_l^\mathrm{N}$ are shown in figure X.% l = 123 at the moment
 
-\begin{figure}[htb]\centering
-\includegraphics[width=0.5\textwidth,clip]{Plots/ttratio.png}
-\includegraphics[width=0.5\textwidth,clip]{Plots/eeratio.png}
-\includegraphics[width=0.5\textwidth,clip]{Plots/bbratio.png}
-  \caption[Current ]{
-  Calculated ratio of $C_l^X/C_l^N$ of the rebinned $\mathrm{TT}$, $\mathrm{EE}$, and $\mathrm{BB}$ power spectra.
+%\begin{figure}[htb]\centering
+%\includegraphics[width=0.5\textwidth,clip]{Plots/ttratio.png}
+%\includegraphics[width=0.5\textwidth,clip]{Plots/eeratio.png}
+%\includegraphics[width=0.5\textwidth,clip]{Plots/bbratio.png}
+%  \caption[Current ]{
+%  Calculated ratio of $C_l^\mathrm{X}/C_l^\mathrm{N}$ of the rebinned $\mathrm{TT}$, $\mathrm{EE}$, and $\mathrm{BB}$ power spectra.
 The vertical black line indicates from which point onwards data is used to calculate the equivalent noise level.
@@
 Increasing the hits per pixel in the map shifts the plateau to the right.
-\label{fig:clxcln}
-}
-\end{figure}
+%\label{fig:clxcln}
+%}
+%\end{figure}
 
-The deduced additional noise for 1, 2 and 3 bit digitisation schemes are shown with respect to the hits per pixel in the maps in figure \ref{fig:extranoise}.
-We would like to point out three key results.
-Firstly, 3 Bit digitisation performs the best, followed by 2 Bit- and finally 1 Bit digitisation.
+The deduced additional noise for 1, 2 and 3 bit digitisation schemes are shown with respect to the hits per pixel in the maps in table \ref{tab:extranoise}.
+We would like to point out the following key results.
+Firstly, 3 Bit digitisation performs the best, followed by 2 Bit and finally 1 Bit digitisation.
 This is what we expect, given that with each additional bit we retain more information.
-Secondly, for a fixed detector noise level the deterioration of the map quality scales in the same fashion as the number of hits per pixel.
+For a fixed detector noise level the deterioration of the map quality scales in the same fashion as the number of hits per pixel.
 We see that $\Delta \sigma / \sigma$ is independent of the hits per pixel in the maps.
+Furthermore, the compression works equally well for temperature as it does for polarisation observations.
 Lastly, the added noise levels are astonishingly low.
+The optimal 3 Bit digitisaiton scheme devised adds as little as $<2\%$ to the map noise level.
 Keeping in mind that CMB detector sensitivity improves in steps of order of magnitude every few years, adding an extra percent-level noise term does not deteriorate the results appreciably.
@@
 
-\begin{figure}[htb]\centering
-\includegraphics[width=0.5\textwidth,clip]{Plots/addednoise.png}
-  \caption[Current ]{
-  Addition to the map noise level due to digitisation. 
-           \label{fig:extranoise}
-  }
-\end{figure}
+\def\arraystretch{1.2}
+\begin{table*}[tbh]
+\begin{center}
+\caption{\label{tab:extranoise} Additional Noise}
+\small
+\begin{tabular}{c c c c c}
+Hits Per Pixel & Channel & 1 Bit & 2 Bit & 3 Bit \\
+\hline
+\hline
+\multirow{3}{*}{800}  & TT  & $ 0.252 \pm 0.009 $  & $ 0.064 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
+& EE  & $ 0.252 \pm 0.012 $  & $ 0.064 \pm 0.005 $  & $ 0.017 \pm 0.004 $ \\
+& BB  & $ 0.252 \pm 0.011 $  & $ 0.066 \pm 0.005 $  & $ 0.018 \pm 0.004 $ \\
+\hline
+\multirow{3}{*}{8000}  & TT  & $ 0.252 \pm 0.007 $  & $ 0.064 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
+& EE  & $ 0.252 \pm 0.007 $  & $ 0.064 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
+& BB  & $ 0.25 \pm 0.018 $  & $ 0.064 \pm 0.009 $  & $ 0.018 \pm 0.004 $ \\
+\hline
+\multirow{3}{*}{80000}  & TT  & $ 0.249 \pm 0.011 $  & $ 0.064 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
+& EE  & $ 0.25 \pm 0.011 $  & $ 0.064 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
+& BB  & $ 0.247 \pm 0.018 $  & $ 0.064 \pm 0.009 $  & $ 0.018 \pm 0.004 $ \\
+\hline
+\multirow{3}{*}{1024000}  & TT  & $ 0.25 \pm 0.009 $  & $ 0.063 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
+& EE  & $ 0.252 \pm 0.008 $  & $ 0.065 \pm 0.003 $  & $ 0.018 \pm 0.002 $ \\
+& BB  & $ 0.251 \pm 0.008 $  & $ 0.064 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
+\hline
+\multirow{3}{*}{10240000}  & TT  & $ 0.186 \pm 0.025 $  & $ 0.047 \pm 0.007 $  & $ 0.013 \pm 0.002 $ \\
+& EE  & $ 0.252 \pm 0.009 $  & $ 0.064 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
+& BB  & $ 0.25 \pm 0.012 $  & $ 0.065 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
+\hline
+\multirow{3}{*}{102400000}  & TT  & $ 0.246 \pm 0.004 $  & $ 0.063 \pm 0.002 $  & $ 0.017 \pm 0.001 $ \\
+& EE  & $ 0.251 \pm 0.008 $  & $ 0.064 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
+& BB  & $ 0.251 \pm 0.009 $  & $ 0.065 \pm 0.003 $  & $ 0.018 \pm 0.002 $ \\
+\end{tabular}
+\tablecomments{ 
+Percent addition to the map noise level $\Delta \sigma / \sigma$ by digitisation. Notice the invariance for each digitisation scheme with the number of hits per pixel and channel.
+} \normalsize
+\end{center}
+\end{table*}
 
-Extreme digitisation is a viable lossy compression technique when dealing with low signal to noise, large datasets with a signal that is slowly varying with respect to the sampling rate.
-Under these conditions we have a locally flat, low signal, on top of which we add many noise realisations.
-In the limit of many samples these allow us to reconstruct the signal well.
+Extreme digitisation is a viable lossy compression technique when dealing with low signal to noise, large data sets with a signal that is slowly varying with respect to the sampling rate.
+Under these conditions we have a locally flat, small signal, on top of which we add many large noise realisations.
+These conditions allow us to reconstruct the underlying signal well.
 Furthermore a small signal will prevent saturation of the output, i.e.
-the inability of the digitised output to communicate any information where in the range $\infty$ to $x_N$ an input signal lies.
+the inability of the digitised output to represent numbers larger than $y_N N_{\mathrm{hit}}$ or smaller than $y_1 N_{\mathrm{hit}}$, where $N_{\mathrm{hit}}$ is the number of data points in a given interval.
+
+%communicate any information where in the first and last digitisation range ( $\left(x_N, \infty \right)$ and $\left( -\infty, x_1 \right)$ )the input signal lies.
 
@@
 
-\section{Conclusions}
+\section{Conclusion}
 \label{sec:conclusions}
 
-In this work we have motivated the investigation of extreme digitisation as a technique in combating arising data challenges in CMB data analysis.
-The reduction of the TOD by an order of magnitude directly addresses the issues in data transmission faced by remote location observations.
+In this work we have motivated the investigation of extreme digitisation as a technique in combating arising challenges in CMB data analysis.
+The reduction of TOD by an order of magnitude directly addresses issues in transmission faced by remote location observations.
 Benefits in mission planning, data analysis and hardware requirements are possible.
 
-We have derived a set of optimal digitisation schemes and presented the level at which the induce noise into the temperate and polarisation power spectra.
-We find that an optimal 3 bit digitisation adds as little as $<2\%$ to the map noise level for temperature and polarisation observations.
-As mentioned above this addition is insignificant given that the sensitivity of CMB experiments follows a Moore's law like trend.
-The benefit of this compression technique is the reduction of the TOD volume by an order of magnitude.
+We derived a set of optimal 1, 2 and 3 bit digitisation schemes assuming white detector noise.
+Through simulated observation we determined the distortion this compression technique causes in temperate and polarisation power spectra.
+We find that an optimal 3 bit digitisation adds as little as $<2\%$ to the map noise level for temperature and polarisation observations alike.
+No change in the results is observed for maps of different hits per pixel.
+The additional noise is insignificant given that the sensitivity of CMB experiments follows a Moore's law like trend, improving by orders of magnitude between generations.
+%L: Comment about invariance with multipole moment
 
@@
 It is of great value to find the higher statistical moments, i.e.
-skewness and kurtosis of the added noise term.
-For this an analysis of the performance of cluster-finding algorithms on the digitised datasets is useful.
+skewness and kurtosis, of the added noise term.
+For this an analysis of the performance of cluster-finding algorithms on the digitised TOD is considered useful.
 
-It should be laid out how the results changes when moving to a more realistic noise profile.
+%It should be laid out how the results changes when moving to a more realistic noise profile.
 We do not expect this to alter the practicality of our results - even if a different noise profile doubles the additional percentage to the map noise level extreme digitisation is still practical.
@@
 chunking of the data before applying extreme digitisation have already been investigated by Planck.
-These thoughts should be considered when designing the compression schemes of future space-based CMB missions, which will be unable to recover their full data with latency, but rely entirely on the transmitted data.
+These thoughts should be considered when designing the compression schemes of future space-based CMB missions, which will be unable to recover their full data with latency, but rely entirely on the transmitted information.
 
 \acknowledgments % thank patrick for discussion/starting point? Andrew?
+
+% Melbourne CMB group, funding, NERSC.
 
@@
 
+% all code stuff
+
+This research made use of the NumPy \citep{numpy}, SciPy \citep{scipy}, Matplotlib \citep{matplotlib}, and Astropy \citep{astropy} packages available for Python.
+
 \bibliography{digitisation}

### 2018-09-04 digitisation.tex

@@
 These will put tight constraints on cosmological models.
-Upcoming science goals include the discovery of inflationary gravitational waves, the construction of precise lensing maps and thorough studies of the Sunyaev-Zeldovich (SZ) effects.
+Upcoming science goals include the discovery of inflationary gravitational waves, construction of precise lensing maps and thorough studies of the Sunyaev-Zeldovich (SZ) effects.
 Moreover, CMB experiments will measure the relativistic number of species and the neutrino mass sum \citep{s4sciencebook}.

### 2018-09-17 digitisation.tex
Included 1/f noise, editing, updated plots

@@
 \usepackage[table,usenames,dvipsnames]{xcolor}
-%\usepackage{amsmath}
+\usepackage{amsmath}
 \usepackage{subfigure}
@@
 \newcommand{\lknee}{\ensuremath{\ell_{\rm knee}}}
+\newcommand{\lcut}{\ensuremath{\ell_{\rm cut}}}
 \newcommand{\fermilat}{\textit{Fermi}-LAT}
@@
 Observations of the Cosmic Microwave Background (CMB) are of significant value to modern cosmology and particle physics.
-Future CMB experiments face issues in hardware design, mission planning and analysis that arise from the vast size of time-ordered-data (TOD) being recorded.
-These challenges are particularly significant for Antarctica- and space-based experiments which depend on satellite links to transmit their data.
+The large size of time-ordered-data (TOD) being recorded causes new challenges in hardware design, mission planning and analysis for future CMB experiments.
+These issues are particularly significant for Antarctica- and space-based experiments which depend on satellite links to transmit their data.
 We explore the viability of reducing the TOD to few-bit numbers to address these issues.
 Unlike lossless compression, this technique introduces additional noise into the data.
-We present 1, 2 and 3 bit digitisation schemes and determine the degradation this compression causes in temperature and polarisation power spectra.
-We find that 3 bit digitisation has a percent-level contribution to the map noise level.
-We argue that such digitisation is a promising strategy for upcoming experiments.
+We present a set of 1, 2 and 3 bit digitisation schemes and determine the degradation the compression causes in temperature and polarisation power spectra.
+The added noise has no dependence on the multipole moment $\ell$ for white and $1/f$ detector noise.
+We note that 3 bit digitisation has a percent-level contribution to the map noise level in all cases considered.
+Such digitisation is a promising strategy for upcoming experiments.
 
@@
 Observations of the Cosmic Microwave Background (CMB) have played a key role in physics since its discovery \citep{penzias1965}.
-Current and future CMB experiments will continue to deliver new insights by studying the temperature and polarisation information contained in the CMB.
-These will put tight constraints on cosmological models.
+Current and future study of the information it contains will put tight constraints on cosmological models.
 Upcoming science goals include the discovery of inflationary gravitational waves, construction of precise lensing maps and thorough studies of the Sunyaev-Zeldovich (SZ) effects.
-Moreover, CMB experiments will measure the relativistic number of species and the neutrino mass sum \citep{s4sciencebook}.
-%L: reference to science goals by satellite missions as well
+Additionally, the CMB probes structure formation and reionisation.
+It is a powerful test for dark matter and dark energy models.
+Lastly, the relativistic number of species and the neutrino mass sum are measured by CMB experiments \citep{s4sciencebook, core2011, litebird2016, pixie2011}.
 
-High fidelity CMB science demands an excellent standard of data analysis.
-The CMB community has developed a variety of compression techniques and computational approaches to manage the increasing influx of data, while maximising the science output \citep{tristam2007}.
+The CMB community has developed a variety of compression techniques and computational approaches to handle the increasing volume of data the science goals demand \citep{tristam2007}.
 These include the compression of time-ordered data (TOD) into maps \citep{tegmark1997}, bandpower estimation \citep{tegmark1998} and the pseudo $C_l$ method \citep{brown2005}.
 
-A growing hurdle for experiments at remote locations are the transmission limitations of satellite links.
+A potential hurdle for experiments at remote locations are the transmission limitations of satellite links.
 Space-based experiments have employed a combination of lossless and lossy compression techniques, including reduced bits in the TOD \citep{gaztanaga1998, maris2003}.
-Antarctica-based experiments that transmit a portion of their data via a satellite link have downsampled their data in the past to meet their telemetry requirements.
-They have yet to exploit few bit digitisation of the TOD.
-As we approach the next generation of ground-based experiments, Stage-4, and the launch of a new generation of space-based missions (liteBIRD, PIXIE, COrE+), we must treat the transmission bottleneck carefully.
-Without a thorough review of compression techniques we are sure to lose information.
-%L:ref to non-planck satellites for reduced bit TOD?
+Antarctica-based experiments that transmit a portion of their data via a satellite link have downsampled their data in the past to meet their telemetry requirements and have not yet to used few bit digitisation of the TOD.
+As we approach the next generation of ground-based experiments, CMB-S4, and the launch of a new generation of space-based missions (LiteBIRD, PIXIE, COrE+), we must consider potential transmission bottlenecks carefully.
 %C: Is there a reference detailing downsampling of SPT data?
 
-In this work we present the method of extreme digitisation, which compresses a rich digital input signal into a few bits.
-We apply extreme digitisation to the TOD and detail its effect on temperature and polarisation power spectra.
-We find that an optimal 3-bit digitisation scheme adds as little as $< 2\%$ to the map noise level.
-The digtisation schemes described here are laid out for ground-based experiments.
-Space-based missions must study carefully the effect of digitisation in their specific compression algorithms.% of t the results we present here should also be considered by future space-based missions, which inevitably must incorporate lossy compression.
+In this work we present the method of extreme digitisation, which reduces a many bit (often 32-or 64-bit) signal to a few bits for ground-based experiments.
+We apply extreme digitisation to the TOD and detail the resulting effects on temperature and polarisation power spectra.
+We find that an optimal 3-bit digitisation scheme adds $< 2\%$ to the map noise level.
 
 This work is structured as follows.
-In \S\ref{sec:dig} we detail the challenges originating from handling large TOD, introduce the process of extreme digitisation and lay out the framework used to test its performance.
-Subsequently, we describe the power spectrum estimation employed and interpret the obtained results in section \S\ref{sec:results}.
+In \S\ref{sec:dig} we detail the challenges that come with handling large data volumes, introduce the process of extreme digitisation and lay out the framework used to test its performance.
+We describe the power spectrum estimation and present the results in section \S\ref{sec:results}.
 We summarise our findings in \S\ref{sec:conclusions}.
@@
 
-\subsection{Problem}
+\subsection{The Challenges Of Large Data Sets}
 \label{subsec:problem}
@@
 
-The science goals of upcoming CMB experiments naturally lead to a large influx of data To achieve the targeted sensitivity longer observations with more detectors are needed.
-In fact the number of detectors of ground-based experiments has been following a Moore's law like trend, doubling approximately every 2 years.
-This directly translates into an exponential growth in data volume \citep{s4sciencebook}.
-%L: check out the particle paper where the moores law figure comes from
+The science goals of upcoming CMB experiments depend on achieving substantially faster mapping speeds.
+This has lead to the number of detectors of ground-based experiments following a Moore's law like trend, doubling approximately every 2 years \citep{s4sciencebook, Abazajian2015}.
+The growing detector count translates into an exponential growth in data volume.
 
-Excellent observation conditions for CMB observations can not only be found in space, but also at various locations on earth \citep{li2017, kovac2007} among which is the South Pole.
+Excellent observation conditions for CMB observations can be found at various locations on earth, including the South Pole \citep{kovac2007}, and in space.
 Space- and Antarctica-based experiments depend on satellite transmission.
 An experiment at the South Pole will be part of the next generation of ground-based instruments, which aims to collect 2 million detector years in total.
-Such an experiment will face a data influx of $\sim \mathrm{O}(10)\mathrm{Tb/d}$.
-However, the current transmission allocation for SPT3G is at $150\mathrm{Gb/d}$, which will likely only see a moderate increase in coming years.
-The transmission bottleneck is currently overcome by recovering the full data on hard drives with some latency and by transmitting a downsampled version of the data.
+Such an experiment will face a data influx of $\sim \mathcal{O}(10)\mathrm{Tb/d}$.
+However, the current transmission allocation for SPT3G is at $150\mathrm{Gb/d}$, which will likely see only a moderate increase in coming years.
+The transmission bottleneck is currently overcome by recovering the full data on hard drives with some latency and transmitting a downsampled version of the data.
 The downsampling process loses high frequency information.
-Going into Stage-4 we anticipate that compression rates for transmission must increase by an order of magnitude.
 Continued use of downsampling will narrow the information window decisively - prohibiting high multipole moment science to be carried out on the transmitted data set.
-This also means that any potential faults or errors in the experiment that only become visible in said range will go unnoticed for longer.
-%L: reference for observing conditions in chile or a general review of sites %C: Is there a paper detailing trasmission limits for SPT and that the data is currently downsampled?
+This also means that any potential faults or errors in the experiment visible in this range are not visible in the transmitted data.
+%C: Is there a paper detailing trasmission limits for SPT and that the data is currently downsampled?
 
@@
 Beyond transmission challenges, mission planning is becoming exceedingly difficult.
-A full simulation of TOD over the entire parameter space of detection scenarios for numerous set-ups is the desired way to decide on Stage-4 configurations.
-Space-based missions must aim to carry out a similar analysis to optimise their science output.
+Full simulation of TOD over the entire parameter space of detection scenarios is the desired way to decide on Stage-4 configurations.
 Given the shear size of TOD expected, this is not possible \citep{s4sciencebook}.
-We must rely on different planning strategies or aim to reduce the size of the TOD in order to maximise the productivity of planning and development stages and guarantee scientific excellence.
+Space-based missions carry out a similar analysis specific to their instruments to optimise their science output.
+Ground-based observations have to rely on different planning strategies or aim to reduce the size of the TOD in order to maximise the productivity of instruments.
 
@@
 While we have not experienced the limitations of the accessible computational assets, the exponential growth of CMB data makes its analysis increasingly expensive.
-The community depends on the facilities provided by the National Energy Research Scientific Computing Center (NERSC).
 
@@
 
-Extreme Digitisation would tackle the challenges mentioned above by reducing the size of the TOD by an order of magnitude.
+Extreme Digitisation would tackle the challenges mentioned above by reducing TOD volume by an order of magnitude.
 Together with already established lossless compression techniques (e.g.
@@
 
-The Planck mission has demonstrated the advantages of using few bit compression of the TOD \citep{maris2003}.
+The Planck mission has demonstrated the advantages of using few bit compression of the CMB data \citep{maris2003}.
 Other science areas have also shown the power of extreme digtisation.
-\citep{jenet1998} explored the application of such compression to radio pulsar timing measurements with success.
-Recently \citep{clearwater2018} have investigated the advantages of using 1 and 2 bit data when searching for continuous gravitational waves using the Laser Interferometer Gravitational-Wave Observatory (LIGO).
+\cite{jenet1998} explored the application of such compression to radio pulsar timing measurements with success.
+Recently Clearwater et al.
+(private communication) have investigated the advantages of using 1 and 2 bit data when searching for continuous gravitational waves using the Laser Interferometer Gravitational-Wave Observatory (LIGO).
 
@@
 Digitisation is a lossy compression technique.
-However, the induced noise depends on the number of bits used, the digitisation thresholds, and the output levels chosen.
+However, the induced noise depends on the details of the digitisation scheme.
 To minimise the noise induced through this process one must know the nature of the input signal.
-There is usually little concern around finding an optimal set of digitisation parameters for a given input signal.
-It is common to have access to large numbers of bits to store information, where changes in the digitisation scheme become insignificant to the distortion induced.
+There is usually little concern around finding an optimal set of digitisation parameters for an input signal as it is common to have access to large numbers of bits to store information, where changes in the digitisation scheme become insignificant to the distortion induced.
 It is in our interest however to consider extreme, i.e.
-few-bit, digitisation, where a digitisation scheme must be carefully constructed.
-We review the key ideas regarding construction of digitisation schemes as laid out by \cite{max1960} below.
+few-bit, digitisation.
+In this case the digitisation scheme must be carefully constructed.
+We review the key of digitisation as edtablished by \cite{max1960} below.
 
@@
 where $s$ is the input and $\hat{s}$ the output signal.
-For an input signal that has at least some stochastic element to it we introduce the input amplitude probability density $p(x)$.
+For an input signal that has a stochastic element we introduce the input amplitude probability density $p(x)$.
 This allows us to rewrite the above as
@@
 
-To progress further we have to make an assumption about the distribution of input signals, $p(x)$.
-For our purposes we assume that CMB observations operate at low signal to noise.
-Furthermore we assume that the noise profile is Gaussian white noise with standard deviation $\sigma$, i.e.
-$p(x) = (1/\sqrt{2\pi\sigma^2}) e^{-x^2/2\sigma^2}$.
-Given this assumption we can solve the problem using a numerical iterative procedure.
+To progress further we have to make an assumption about the probability distribution of input signals.
+We suppose CMB observations are low signal to noise and detector noise is Gaussian white.
+The input is then purely stochastic and  $p(x) = (1/\sqrt{2\pi\sigma^2}) e^{-x^2/2\sigma^2}$, where $\sigma$ is the standard deviation of the detector noise.
+Under these assumptions the system of equations can be solved numerically.
 One begins by picking $y_1$ and calculating the remaining $x_i$'s and $y_i$'s using equation \ref{eq:digitequalspacecondition}.
@@
 
-\begin{equation} \hat{s}_1(t) = \left\{ \begin{array}{lr}
+\begin{equation} \label{eq:1bit}
+\hat{s}_1(t) = \left\{ \begin{array}{lr}
 1, & \text{for } s(t) > 0\\
@@
 
-\begin{equation} \hat{s}_2(t) = \left\{ \begin{array}{rl}
+\begin{equation}  \label{eq:2bit}
+\hat{s}_2(t) = \left\{ \begin{array}{rl}
 1.51 \sigma, & \text{for } s(t) \geq 0.9816 \sigma\\
 0.4528 \sigma, & \text{for } 0 \leq s(t) < 0.9816 \sigma\\
--0.4528 \sigma, & \text{for } 0.9816 \sigma \leq s(t) < 0\\
--1.51 \sigma, & \text{for } 0.9816 \sigma < s(t)\\
+-0.4528 \sigma, & \text{for } -0.9816 \sigma \leq s(t) < 0\\
+-1.51 \sigma, & \text{for } -0.9816 \sigma < s(t)\\
 \end{array} \right. . \end{equation}
@@
 
-\begin{equation} \hat{s}_3(t) = \left\{ \begin{array}{rl}
+\begin{equation}  \label{eq:3bit}
+\hat{s}_3(t) = \left\{ \begin{array}{rl}
 2.152 \sigma, & \text{for } s(t) \geq 1.748 \sigma\\
@@
 0.245 \sigma, & \text{for } 0 \leq s(t) < 0.501 \sigma\\
--0.245 \sigma, & \text{for } 0.501 \sigma \leq s(t) < 0\\
--0.756 \sigma, & \text{for } 1.05 \sigma \leq s(t) < 0.501 \sigma\\
--1.344 \sigma, & \text{for } 1.748 \sigma \leq s(t) < 1.05 \sigma\\
--2.152 \sigma, & \text{for } 1.748 \sigma < s(t)\\
+-0.245 \sigma, & \text{for } -0.501 \sigma \leq s(t) < 0\\
+-0.756 \sigma, & \text{for } -1.05 \sigma \leq s(t) < -0.501 \sigma\\
+-1.344 \sigma, & \text{for } -1.748 \sigma \leq s(t) < -1.05 \sigma\\
+-2.152 \sigma, & \text{for } -1.748 \sigma < s(t)\\
 \end{array} \right. . \end{equation}
@@
 The fact that the single bit digitisation scheme makes no references to the standard deviation of the signal is irrelevant.
-This affects the normalisation of the signal, which we can account for later.
+This affects the normalisation of the signal, which we can account for during later analysis as demonstrated in the appendix.
 What is important is the spacing between digitisation thresholds and relative distance between the output levels chosen.
 
-Completely different digitisation schemes can be thought of, which would for example reference the most drastic outlier in the data set, or seek to place equal numbers of points into each digitisation level.
-However, given the assumptions made the schemes derived above are optimal.
+%Completely different digitisation schemes can be thought of, which would for example reference the most drastic outlier in the data set, or seek to place equal numbers of points into each digitisation level.
+However, given the assumptions made the schemes derived above minimise the distortion metric defined.
 Additionally, they are simple enough to be easily implemented computationally.
+We consider how to design digitisation schemes that preserve the power of the TOD and how to adjust a given digitisation scheme to ensure power conservation in the appendix.
 
@@
 To investigate the performance of the derived digitisation schemes we simulate many scans over CMB template maps at the timestream level.
-Each scan is performed by a single detector.
 We obtain control maps that use 64bit TOD and maps that have undergone 1, 2 and 3 bit digitisation at the timestream level.
-We calculate the temperature and polarisation power spectra of each map and determine the additional noise induced through the extreme digitisation process.
-%C: Do I need to mention that we assume to have an infinitely fast spinning plate? %L: Check whether Python uses 32 or 64 bit under on NERSC for floats
+We calculate the temperature and polarisation power spectra of each map and determine the noise level.
+%C: Do I need to mention that we assume to have an infinitely fast spinning plate?
 
 To create the template maps we use the Hierarchical Equal Area isoLatitude Pixelization of a sphere \citep[HEALPix;][]{healpix}\footnote{Available at \url{http://healpix.sourceforge.net/}} framework.
-We generate a realisation of I, Q and U maps with $\mathrm{NSIDE} = 4096$ based on the results of the Planck collaboration with key cosmological parameters of $\Omega_b \approx 0.049 $, $\Omega_c \approx 0.265$, $\Omega_m \approx 0.316$ and $h \approx 0.67$ \citep{planck2016}.
+We generate a realisation of I, Q and U maps with $\mathrm{NSIDE} = 4096$ based on the 2015 results of the Planck collaboration.
+The key cosmological parameters are $\Omega_b \approx 0.049 $, $\Omega_c \approx 0.265$, $\Omega_m \approx 0.316$ and $h \approx 0.67$ \citep{planck2016}.
 
@@
 
-We simulate observing a $\sim 600 \mathrm{deg^2}$ patch of the sky.
+We simulate observing a $\sim 600 \mathrm{deg^2}$ patch of the sky with a single detector.
 To do so we perform a number of constant elevation scans (CES), equally spaced in declination (DEC).
 We repeat the observation strategy $100$ times with a slight offset in right ascension (RA) and DEC each time, such that each pixel is sampled approximately uniformly.
+The detector is read out at $200 \mathrm{Hz}$.
 The speed at which we sweep across the survey area is adjusted to produce a desired number of hits per pixel in the output maps.
@@
 While performing each CES the pixels being targeted are determined.
-The corresponding values from the template maps are then accessed and added to realisations of the detector noise of appropriate length to generate TOD streams.
-We assume the detector noise to be Gaussian white noise.
+The corresponding values from the template maps are accessed and added to detector noise realisations.
 For temperature observations we assume the detector noise level to be $500 \mathrm{\mu K \sqrt{s}}$ and for polarisation observations $\sqrt{2} \hspace{0.7ex} 500 \mathrm{\mu K \sqrt{s}}$, i.e.
@@
 
-At this point we apply the digitisation schemes to the TOD.
-We compress the timestream into maps by averaging all hits falling into the same pixel.
-We produce 12 CMB maps in total: three control maps (I, Q, U) and nine maps with three each obtained from each digitisation scheme.
+Having constructed I,Q and U TOD we apply the digitisation schemes derived above.
+We compress the timestreams into maps by averaging all hits falling into the same pixel.
+We produce 24 CMB maps in total: three control maps and nine maps with three each obtained from the different digitisation schemes.
+We simulate six observations, producing maps with $800$, $8,000$, $80,000$, $1,024,000$, $10,240,000$ and $102,400,000$ hits in each pixel.
+
+%It is often commonplace to fit a polynomial to the TOD, a series of sines and cosines, subtract local averages or do a combination of these techniques \citep[e.g.][, planck?]{chown2018} to reduce $1/f$ noise.
+We do not employ any of these filtering mechanisms.
+A thorough application of such noise removal techniques would lead to similar results as the white noise case.
 
@@
 
-This process is carried out 6 times.
-We produce maps where pixels in the survey patch are hit approximately uniformly $800$, $8,000$, $80,000$, $1,024,000$, $10,240,000$ and $102,400,000$ times.
-We assume that the detector is read out $200$ times per second.
 %The calculation outlined above has considerable computational requirements if we want to reach up to $\sim 10^8$ hits per pixel.
 This calculation was made possible through use of the NERSC facilities.
-
-%\begin{table}[tbh]
-%\begin{center}
-%\caption{\label{tab:inputcosparams} Input Cosmological Parameters}
-%\small
-%\begin{tabular}{l | c}
-%Parameter & Planck 2015\\
-%\hline\\begin{equation}-1.5ex]
-%$100\theta_{MC}$ & $ 1.04086 \pm 0.00048$\\
-%$\Omega_b h^2$ & $  0.02222 \pm 0.00023$\\
-%$\Omega_c h^2$ & $ 0.1199 \pm 0.0022$\\
-%$H_0$ & $67.26 \pm 0.98$\\
-%$n_s$ & $0.9652 \pm 0.0062$\\
-%$\Omega_m$ & $0.316 \pm 0.014$\\
-%$\sigma_8$ & $0.830 \pm 0.015 $\\
-%$\tau$ & $0.078 \pm 0.019$\\
-%$10^9 A_s e^{-2\tau}$ & $1.881 \pm 0.014 $\\
-%\end{tabular}
-%\tablecomments{ 
-%Cosmological parameters used to create the template maps. Taken from Planck 2015: Cosmological Parameters.
-%} \normalsize
-%\end{center}
-%\end{table}
-
-%\begin{table}[tbh] % transpose and include hits per pixel!
-%\begin{center}
-%\caption{\label{tab:modelparams} Assumed Survey Parameters}
-%\small
-%\begin{tabular}{l | c}
-%Parameter & Value\\
-%\hline \\begin{equation}-1.5ex]
-%$\mathrm{NSIDE}$ &$4096$\\
-%$f_{\mathrm{readout}} [\mathrm{Hz}]$ &$200$\\
-%$f_{{\mathrm{sky}}}$ & $\sim 0.014$\\
-%$\sigma^{\mathrm{T}}_{\mathrm{det}} [\mathrm{\mu K \sqrt{s}}]$ &$500$\\
-%$\sigma^{\mathrm{Pol}}_{\mathrm{det}} [\mathrm{\mu K \sqrt{s}}]$ & $\sqrt{2}\times500$\\
-%\multirow{2}{*}{$\mathrm{hits per pixel}$} & $\left(1, 10, 100 \right) \times 800,$\\
-%& $\left( 1, 10, 100 \right) \times 1024000$ \\
-%\end{tabular}
-%\tablecomments{ 
-%Parameters used in the simulated observation. The RA speed is adjusted to match the desired hits per pixel.
-%} \normalsize
-%\end{center}
-%\end{table}
 
@@
 
-We use the Spatially Inhomogeneous Correlation Estimator for Temperature and Polarisation \citep[PolSpice;][]{polspice} to compute the $\mathrm{TT}, \mathrm{EE}$ and $\mathrm{BB}$ power spectra of the reconstructed maps I, Q, U maps.
+We use the Spatially Inhomogeneous Correlation Estimator for Temperature and Polarisation \citep[PolSpice;][]{polspice} to compute $\mathrm{TT}, \mathrm{EE}$ and $\mathrm{BB}$ power spectra.
 To minimise cut sky effects we apodise the observed path using a cosine mask with $\sigma_{\mathrm{APOD}} = \sqrt{600}/2 \hspace{0.7ex} \mathrm{deg}$.
@@
 This will likely be reworked when using Healpix to get the power spectra though.
-
-%%%%%%%%%%%
-% INCLUDE SECTION OF ALM CUTS (maybe a plot of the cuts too)
-%%%%%%%%%%%
-
-%\begin{table}[tbh]
-%\begin{center}
-%\caption{\label{tab:polspiceparams} PolSpice Parameters}
-%\small
-%\begin{tabular}{c c c c}
-%weights & $\sigma_{\mathrm{APOD}} [\mathrm{deg}]$ & apodistation type & polarisation \\
-%\hline
-%cosine mask & $\sqrt{600}/2$ & cosine & Yes \\
-%
-%\end{tabular}
-%\tablecomments{ 
-%Parameters used when calling PolSpice to calculate the power spectra. Remaining parameters have been left at their default value.
-%} \normalsize
-%\end{center}
-%\end{table}
 
@@
 Normalisation of the obtained power spectra is necessary.
-The digitisation schemes introduced in section \S\ref{sec:dig} are designed to minimise the distortion of the TOD, not to preserve power at the map level.
-Normalising obtained spectra against previous experiments is a common calibration technique in many CMB experiments (e.g.
+The digitisation schemes introduced in section \S\ref{sec:dig} are designed to minimise the distortion of the TOD, not to preserve power.
+We obtain the appropriate normalisation as described in the appendix.
+
+%Normalising obtained spectra against previous experiments is a common calibration technique in many CMB experiments (e.g.
 SPT).
-%C: Do you know a good reference for this?
+We again refer to the discussion in the appendix on how to modify digitisation schemes to ensure power preservation.%C: Do you know a good reference for this?
 
-For the normalisation procedure we focus on the $\sim 10^8$ hits per pixel maps.
-We normalise the $\mathrm{TT}$, $\mathrm{EE}$, and $\mathrm{BB}$ power spectra from few bit TOD against their control counterparts.
-We consider the normalisation window cut out by the conditions
-
-\begin{equation} l > 35 \> , \> \frac{C_l^{\mathrm{S}}}{C_l^{\mathrm{N}}} \geq 10. \end{equation}
-
-where $C_l^{\mathrm{S}}$ is the power spectrum of the input template maps and $C_l^{\mathrm{N}}$ the detector noise level.
-The lower limit is placed in accordance to the patch-size surveyed while the upper limit ensures appropriate signal to noise in the normalisation window.
-
-The normalisation constants for each digitisation scheme and channel obtained through this way are then applied to all lower hits per pixel simulations.
-A sample of the obtained power spectra is shown in figure X.
+All digitisation schemes are able to recover the input power spectra well. This is shown by figure \ref{fig:psrecover}.
 
@@
 
-%\begin{figure}[htb]\centering
-%\includegraphics[width=0.5\textwidth,clip]{Plots/ttzoom.png}
-%\includegraphics[width=0.5\textwidth,clip]{Plots/eezoom.png}
-%\includegraphics[width=0.5\textwidth,clip]{Plots/bbzoom.png}
-%  \caption[Current ]{
-%   Reconstructed $\mathrm{TT}$, $\mathrm{EE}$, and $\mathrm{BB}$ power spectra.
-These originate from an observation with $102400000 \mathrm{hits per pixel}$.
-We observe that digitisation appears to add some constant to the noise level.
-As one would expect $3 \mathrm{Bit}$ digitisation performs the best, followed by $2 \mathrm{Bit}$- and finally $1 \mathrm{Bit}$ digitisation.
-%\label{fig:powspeczoom}
-%}
-%\end{figure}
+\begin{figure}[htb]\centering
+\includegraphics[width=0.5\textwidth,clip]{Plots/psrecovery.pdf}
+  \caption[Current ]{
+  Input plus detector noise power spectra (black) and power spectra using 3-bit TOD (violet) for observations with $\sim 10^6$ hits per map pixel.
+Differences occur because we observe a single CMB realisation.
+\label{fig:psrecover}
+}
+\end{figure}
 
-\subsection{Additional Noise}
-\label{subsec:additionalnoise}
+\subsection{White Noise}
+\label{subsec:whitenoise}
 
-%C: This is where I'm planning on
-%%%%%%%%%%%
-% INSERT A COUPLE PARAGRAPHS ABOUT THE PS OF DIFFERENCE MAPS
-% INCLUDE PLOTS OF PS OF DIFFERENCE MAPS
-% want to show that beyond some calibration error, we do not observe any l dependence in the added noise
-%%%%%%%%%%%
+To investigate whether the noise added through digitisation has any $\ell$ dependence we subtract the template maps from the simulated observations.
+We use theoretically calculated normalisation constants to minimise calibration error.
+The power spectra of the difference maps do not show any $\ell$ dependence in the added noise as seen in figure \ref{fig:diffpswn}.
+% This is in agreement with (planck website) but does not match the results of (planck short paper).
+This is likely due to the intricities of the planck compression scheme differing to what is considered here.
+A full comparison is difficult in the absense of a detailed description of the simulations carried out in either paper.
 
-To quantify the distortion caused by extreme digitisation we infer the map noise level from the relevant power spectra.
-These are put into the context of the map noise level deduced from the control power spectra.
-We formulate
+\begin{figure}[htb]\centering
+\includegraphics[width=0.5\textwidth,clip]{Plots/diffwn.pdf}
+  \caption[Current ]{
+  Power spectra of reconstructed minus input maps for 1 Bit (green), 2 Bit (orange), 3 Bit (violet) and the control case (black).
+We do not observe any $\ell$ dependence.
+Simulations with $\sim 10^6$ hits per map pixel are shown.
+\label{fig:diffpswn}
+}
+\end{figure}
+
+To quantify the distortion caused by extreme digitisation we infer the map noise level from the relevant power spectra and compare them to the map noise level of control observations. We formulate
 
@@
 
-where $\sigma_{\mathrm{map}}^{\mathrm{C}}$ is the map noise level of the control power spectra and $\sigma_{\mathrm{map}}^{\mathrm{D}}$ is the map noise level obtained from the power spectra originating from extremely digitised TOD.
-We assume that the digitisation process results in adding a constant noise term to each power spectrum.
-If we are dominated by noise we can write
+where $\sigma_{\mathrm{map}}^{\mathrm{C}}$ is the map noise level of control observations and $\sigma_{\mathrm{map}}^{\mathrm{D}}$ is the map noise level of observations using few bit TOD.
+Because we have seen no $\ell$ dependence in the additional noise we presume digitisation to add a constant noise term to the power spectrum.
+In the regime
 
-\begin{equation} C_l^\mathrm{D} \approx C_l^\mathrm{N} + C_l^\mathrm{X}. \end{equation}
+\begin{equation} \frac{C_\ell^{\mathrm{N}}}{C_\ell^{\mathrm{S}}} \geq 10, \end{equation}
 
-Here $C_l^\mathrm{D}$ is the total power spectrum originating from a digitised timestream and $C_l^\mathrm{X}$ the additional noise induced through digitisation.
-We now progress equation \ref{eq:extramapnoise} to
+where $C_l^\mathrm{S}$ is the input CMB signal and $C_l^\mathrm{N}$ the detector noise level, power spectra are assumed to be noise dominated. Under this condition
 
-\begin{equation}\frac{\Delta \sigma}{\sigma} = \sqrt{\frac{C_l^\mathrm{D}}{C_l^{\mathrm{N}}}} - 1 = \sqrt{1 + \frac{C_l^\mathrm{X}}{C_l^{\mathrm{N}}}} - 1 . \end{equation}
+\begin{equation} C_{\ell}^\mathrm{D} = C_\ell^\mathrm{N} + C_\ell^\mathrm{X}, \end{equation}
 
-% = \sqrt{\frac{C_l^N + C_l^X}{C_l^{N}}} - 1
+where $C_\ell^\mathrm{D}$ is the total power spectrum originating from a digitised timestream and $C_l^\mathrm{X}$ the additional noise induced through digitisation.
+We progress equation \ref{eq:extramapnoise} to
 
-The value of $l$ at which we can safely assume to be noise dominated and apply the above framework varies between simulated observations of different hits per pixel and temperature and polarisation power spectra.
-To define the range of noise domination we choose the condition
+\begin{equation}\frac{\Delta \sigma}{\sigma} = \sqrt{\frac{C_l^\mathrm{D}}{C_l^{\mathrm{N}}}} - 1 = \sqrt{1 + \frac{C_\ell^\mathrm{X}}{C_\ell^{\mathrm{N}}}} - 1. \end{equation}
 
-\begin{equation} \frac{C_l^{\mathrm{N}}}{C_l^{\mathrm{S}}} \geq 10. \end{equation}
+Before analysing $C_\ell^\mathrm{X}/C_\ell^\mathrm{N}$ we rebin the power spectra to $\Delta \ell \approx 100$.
+This guarantees that the points in the noise tail are independent of one another, allowing us to extract an uncertainty for the above quantity.
 
-Before analysing $C_l^\mathrm{X}/C_l^\mathrm{N}$ we rebin the power spectra to $\Delta l \approx 100$.
-This guarantees that the points in the noise tail are independent of one another, allowing us to extract an uncertainty for the above quantity.% Plots for $C_l^\mathrm{X}/C_l^\mathrm{N}$ are shown in figure X.% l = 123 at the moment
-
-%\begin{figure}[htb]\centering
-%\includegraphics[width=0.5\textwidth,clip]{Plots/ttratio.png}
-%\includegraphics[width=0.5\textwidth,clip]{Plots/eeratio.png}
-%\includegraphics[width=0.5\textwidth,clip]{Plots/bbratio.png}
-%  \caption[Current ]{
-%  Calculated ratio of $C_l^\mathrm{X}/C_l^\mathrm{N}$ of the rebinned $\mathrm{TT}$, $\mathrm{EE}$, and $\mathrm{BB}$ power spectra.
-The vertical black line indicates from which point onwards data is used to calculate the equivalent noise level.
-The above power spectra originate from a $80000\mathrm{hits per pixel}$ map.
-Increasing the hits per pixel in the map shifts the plateau to the right.
-%\label{fig:clxcln}
-%}
-%\end{figure}
-
-The deduced additional noise for 1, 2 and 3 bit digitisation schemes are shown with respect to the hits per pixel in the maps in table \ref{tab:extranoise}.
-We would like to point out the following key results.
-Firstly, 3 Bit digitisation performs the best, followed by 2 Bit and finally 1 Bit digitisation.
-This is what we expect, given that with each additional bit we retain more information.
-For a fixed detector noise level the deterioration of the map quality scales in the same fashion as the number of hits per pixel.
-We see that $\Delta \sigma / \sigma$ is independent of the hits per pixel in the maps.
+The deduced additional noise for 1, 2 and 3 bit digitisation schemes are shown in table \ref{tab:extranoisewhite} in the appendix.
+As expected 3 Bit digitisation performs the best, followed by 2 Bit and finally 1 Bit digitisation.
+This i because with each additional bit we retain more information.
+We observe that $\Delta \sigma / \sigma$ is independent of the hits per pixel in the maps.
+This suggests that the additional noise scales in the same way as detector noise does.
 Furthermore, the compression works equally well for temperature as it does for polarisation observations.
-Lastly, the added noise levels are astonishingly low.
-The optimal 3 Bit digitisaiton scheme devised adds as little as $<2\%$ to the map noise level.
+Lastly, the added noise levels are low.
+The optimal 3 Bit digitisaiton scheme devised adds as little as $1.8\%$ to the map noise level.
 Keeping in mind that CMB detector sensitivity improves in steps of order of magnitude every few years, adding an extra percent-level noise term does not deteriorate the results appreciably.
-This is impressive, given that the use of an optimal 3 bit digitisation scheme will save approximately an order of magnitude in TOD volume.
-
-\def\arraystretch{1.2}
-\begin{table*}[tbh]
-\begin{center}
-\caption{\label{tab:extranoise} Additional Noise}
-\small
-\begin{tabular}{c c c c c}
-Hits Per Pixel & Channel & 1 Bit & 2 Bit & 3 Bit \\
-\hline
-\hline
-\multirow{3}{*}{800}  & TT  & $ 0.252 \pm 0.009 $  & $ 0.064 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
-& EE  & $ 0.252 \pm 0.012 $  & $ 0.064 \pm 0.005 $  & $ 0.017 \pm 0.004 $ \\
-& BB  & $ 0.252 \pm 0.011 $  & $ 0.066 \pm 0.005 $  & $ 0.018 \pm 0.004 $ \\
-\hline
-\multirow{3}{*}{8000}  & TT  & $ 0.252 \pm 0.007 $  & $ 0.064 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
-& EE  & $ 0.252 \pm 0.007 $  & $ 0.064 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
-& BB  & $ 0.25 \pm 0.018 $  & $ 0.064 \pm 0.009 $  & $ 0.018 \pm 0.004 $ \\
-\hline
-\multirow{3}{*}{80000}  & TT  & $ 0.249 \pm 0.011 $  & $ 0.064 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
-& EE  & $ 0.25 \pm 0.011 $  & $ 0.064 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
-& BB  & $ 0.247 \pm 0.018 $  & $ 0.064 \pm 0.009 $  & $ 0.018 \pm 0.004 $ \\
-\hline
-\multirow{3}{*}{1024000}  & TT  & $ 0.25 \pm 0.009 $  & $ 0.063 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
-& EE  & $ 0.252 \pm 0.008 $  & $ 0.065 \pm 0.003 $  & $ 0.018 \pm 0.002 $ \\
-& BB  & $ 0.251 \pm 0.008 $  & $ 0.064 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
-\hline
-\multirow{3}{*}{10240000}  & TT  & $ 0.186 \pm 0.025 $  & $ 0.047 \pm 0.007 $  & $ 0.013 \pm 0.002 $ \\
-& EE  & $ 0.252 \pm 0.009 $  & $ 0.064 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
-& BB  & $ 0.25 \pm 0.012 $  & $ 0.065 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
-\hline
-\multirow{3}{*}{102400000}  & TT  & $ 0.246 \pm 0.004 $  & $ 0.063 \pm 0.002 $  & $ 0.017 \pm 0.001 $ \\
-& EE  & $ 0.251 \pm 0.008 $  & $ 0.064 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
-& BB  & $ 0.251 \pm 0.009 $  & $ 0.065 \pm 0.003 $  & $ 0.018 \pm 0.002 $ \\
-\end{tabular}
-\tablecomments{ 
-Percent addition to the map noise level $\Delta \sigma / \sigma$ by digitisation. Notice the invariance for each digitisation scheme with the number of hits per pixel and channel.
-} \normalsize
-\end{center}
-\end{table*}
-
-Extreme digitisation is a viable lossy compression technique when dealing with low signal to noise, large data sets with a signal that is slowly varying with respect to the sampling rate.
-Under these conditions we have a locally flat, small signal, on top of which we add many large noise realisations.
-These conditions allow us to reconstruct the underlying signal well.
-Furthermore a small signal will prevent saturation of the output, i.e.
-the inability of the digitised output to represent numbers larger than $y_N N_{\mathrm{hit}}$ or smaller than $y_1 N_{\mathrm{hit}}$, where $N_{\mathrm{hit}}$ is the number of data points in a given interval.
+This is impressive, given that the use of an optimal 3 bit digitisation scheme reduces TOD volume by an order of magnitude.
 
@@
 
-%\begin{table*}[tbh]
-%\begin{center}
-%\caption{\label{tab:experiments} Assumed survey parameters}
-%\small
-%\begin{tabular}{l || c c c c c }
-%Experiment & Sky coverage & Polarized Noise level  & 1/$f$ knee & Beam FWHM \\
-%& &($\mu$K-arcmin)&&(arcmin.)\\
-%\hline
-%\tiny \\ \small
-%CMB Stage III & & & & \\
-%~~~~~SPT-3G & 6\% & 3.0 & 200 & 1.2 \\
-%~~~~~Simons Array & 36\% & 9.5 & 200 & 3.5 \\ 
-%\tiny \\ \small
-%\hline
-%CMB Stage IV & 55\% & 1.3 & 100 & 4.0 \\
-%\end{tabular}
-%\tablecomments{ 
-%Key numbers about the planned stage III and IV experiments. 
-%The sky coverage percentages are after galactic cuts. 
-%Unless otherwise noted,  the Fisher matrix forecasts in this work use these numbers. 
-%All forecasts also allow for beam and calibration uncertainties as noted in the text. 
-%} \normalsize
-%\end{center}
-%\end{table*}
+\subsection{1/f Noise}
+\label{subsec:oofnoise}
+
+We follow up on the compelling results of the white noise case by implementing a more realistic noise profile.
+Noise profiles of ground-based CMB observations have prominent $1/f$ character due to atmospheric effects \citep{lay2000}.
+We use the Fourier space noise profile
+
+\begin{equation} \label{eq:1/fnoiseform}
+\begin{aligned}
+\left| N_{1/f}(\ell) \right|^2 &= \left| N\left(0, \sigma \right) \right|^2\\
+&\times \left[ 1 + \left( \frac{\ell}{\lknee} \right)^{-3/2} \right]^2  exp\left[ -\left( \frac{\lcut}{\ell} \right)^{12} \right],
+\end{aligned}
+\end{equation}
+
+where the first term on the second line produces $1/f$ behaviour and the second term on the same line suppressed the lowest modes.
+For temperature observations we choose $\lknee = 1000$ and $\lcut = 336$, to ensure that $|N_{1/f}(0.3\lknee)|^2 = |0.95 \sigma|^2$.
+For polarisation we set $\lknee = 50$ and $\lcut = 17$ for the same reason.
+
+We use the same digitisation schemes as for white noise, but modify the parameter $\sigma$ in equations \ref{eq:1bit}, \ref{eq:2bit} and \ref{eq:3bit} by considering Parseval's theorem on the above noise profile.
+The appropriate value is obtained by evaluating the integral
+
+\begin{equation}
+\sigma^2 = \int_0^\infty \left| N_{1/f}(\ell) \right|^2 d\ell.
+\end{equation}
+
+We carry out the same simulations as described in subsection \ref{subsec:method} using the above noise profile and digitisation scheme for maps with $8,000$, $102,400$ and $1,024,000$ hits per pixel.
+HEALPix is used to calculate the spherical harmonics transform coefficients $a\ell m$.
+We mask the lowest modes in scan direction and obtain the power spectra.
+The analysis procedure of section \ref{subsec:whitenoise} is used to characterise the additional noise.
+
+We recover similar results to the white noise case.
+As before 3 bit performs the best, followed by 2 bit and 1 bit.
+The excess noise due to digitisation does not show any $\ell$ dependence.
+The noise contribution to the map noise level from digitisation for $1/f$ noise is the same as for white noise and does not vary with the number of hits per pixel in the map.
+Temperature and polarisation observations incur the same distortion.
+Table \ref{tab:extranoiseoof} in the appendix displays the values of $\Delta \sigma / \sigma$ calculated.
+
+Extreme digitisation is a viable lossy compression technique when dealing with low signal to noise, large data sets with a signal that is slowly varying with respect to the sampling rate.
+These conditions allow us to reconstruct the underlying signal well.
+A small signal will also prevent saturation of the output, i.e.
+the inability of the digitised output to represent numbers larger than $y_N N_{\mathrm{hit}}$ or smaller than $y_1 N_{\mathrm{hit}}$, where $N_{\mathrm{hit}}$ is the number of data points in a given interval.
+These conditions are satisfied by CMB data and our findings reflect this.
 
@@
 
-In this work we have motivated the investigation of extreme digitisation as a technique in combating arising challenges in CMB data analysis.
-The reduction of TOD by an order of magnitude directly addresses issues in transmission faced by remote location observations.
+In this work we have motivated the investigation of extreme digitisation as a technique in combating potential challenges of CMB experiments.
+The reduction of TOD by an order of magnitude addresses potential issues in transmission from remote locations.
 Benefits in mission planning, data analysis and hardware requirements are possible.
 
-We derived a set of optimal 1, 2 and 3 bit digitisation schemes assuming white detector noise.
+We derived a set of 1, 2 and 3 bit digitisation schemes assuming white detector noise.
 Through simulated observation we determined the distortion this compression technique causes in temperate and polarisation power spectra.
-We find that an optimal 3 bit digitisation adds as little as $<2\%$ to the map noise level for temperature and polarisation observations alike.
-No change in the results is observed for maps of different hits per pixel.
-The additional noise is insignificant given that the sensitivity of CMB experiments follows a Moore's law like trend, improving by orders of magnitude between generations.
-%L: Comment about invariance with multipole moment
+We find that an optimal 3 bit digitisation adds as little as $1.8\%$ to the map noise level for temperature and polarisation observations alike.
+No change in the results is observed for maps of different hits per pixel and no $\ell$ dependence was observed in the added noise.
+Extending the model to more realistic $1/f$ detector noise did not change these results.
+
+The additional noise is insignificant for ground-based experiments, which make order of mangitude jumps in detector count and hence sensitivity between generations.
+Space-based experiments which have their own unique issues may consider this noise penalty problematic.
 
@@
 skewness and kurtosis, of the added noise term.
-For this an analysis of the performance of cluster-finding algorithms on the digitised TOD is considered useful.
-
-%It should be laid out how the results changes when moving to a more realistic noise profile.
-We do not expect this to alter the practicality of our results - even if a different noise profile doubles the additional percentage to the map noise level extreme digitisation is still practical.
-Ideas on how to deal with 1/f noise, e.g.
-chunking of the data before applying extreme digitisation have already been investigated by Planck.
-These thoughts should be considered when designing the compression schemes of future space-based CMB missions, which will be unable to recover their full data with latency, but rely entirely on the transmitted information.
+For this an analysis of the performance of cluster-finding algorithms on the digitised TOD is recommended.
 
@@
 
-This research made use of the NumPy \citep{numpy}, SciPy \citep{scipy}, Matplotlib \citep{matplotlib}, and Astropy \citep{astropy} packages available for Python.
+This research made use of the NumPy \citep{numpy}, SciPy \citep{scipy}, Matplotlib \citep{matplotlib}, and Astropy \citep{astropy} packages.
+
+\newpage
+
+% APPENDIX
+\appendix
+
+% result tables
+\subsection{Additional Noise due to Digitisation}
+\label{subsec:appendixnoisetables}
+
+\def\arraystretch{1.3}
+\begin{table*}[tbh]
+\begin{center}
+\caption{\label{tab:extranoisewhite} Additional Noise - White Detector Noise}
+\small
+\begin{tabular}{c c c c c}
+Hits Per Pixel & Channel & 1 Bit & 2 Bit & 3 Bit \\
+\hline
+\hline
+\multirow{3}{*}{$800$}  & TT  & $ 0.252 \pm 0.009 $  & $ 0.064 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
+& EE  & $ 0.252 \pm 0.012 $  & $ 0.064 \pm 0.005 $  & $ 0.017 \pm 0.004 $ \\
+& BB  & $ 0.252 \pm 0.011 $  & $ 0.066 \pm 0.005 $  & $ 0.018 \pm 0.004 $ \\
+\hline
+\multirow{3}{*}{$8,000$}  & TT  & $ 0.252 \pm 0.007 $  & $ 0.064 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
+& EE  & $ 0.252 \pm 0.007 $  & $ 0.064 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
+& BB  & $ 0.25 \pm 0.018 $  & $ 0.064 \pm 0.009 $  & $ 0.018 \pm 0.004 $ \\
+\hline
+\multirow{3}{*}{$80,000$}  & TT  & $ 0.249 \pm 0.011 $  & $ 0.064 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
+& EE  & $ 0.25 \pm 0.011 $  & $ 0.064 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
+& BB  & $ 0.247 \pm 0.018 $  & $ 0.064 \pm 0.009 $  & $ 0.018 \pm 0.004 $ \\
+\hline
+\multirow{3}{*}{$1,024,000$}  & TT  & $ 0.25 \pm 0.009 $  & $ 0.063 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
+& EE  & $ 0.252 \pm 0.008 $  & $ 0.065 \pm 0.003 $  & $ 0.018 \pm 0.002 $ \\
+& BB  & $ 0.251 \pm 0.008 $  & $ 0.064 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
+\hline
+\multirow{3}{*}{$10,240,000$}  & TT  & $ 0.186 \pm 0.025 $  & $ 0.047 \pm 0.007 $  & $ 0.013 \pm 0.002 $ \\
+& EE  & $ 0.252 \pm 0.009 $  & $ 0.064 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
+& BB  & $ 0.25 \pm 0.012 $  & $ 0.065 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
+\hline
+\multirow{3}{*}{$102,400,000$}  & TT  & $ 0.246 \pm 0.004 $  & $ 0.063 \pm 0.002 $  & $ 0.017 \pm 0.001 $ \\
+& EE  & $ 0.251 \pm 0.008 $  & $ 0.064 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
+& BB  & $ 0.251 \pm 0.009 $  & $ 0.065 \pm 0.003 $  & $ 0.018 \pm 0.002 $ \\
+\end{tabular}
+\tablecomments{ 
+Percent addition to the map noise level $\Delta \sigma / \sigma$ by digitisation. Notice the invariance for each digitisation scheme with the number of hits per pixel and channel.
+} \normalsize
+\end{center}
+\end{table*}
+
+\def\arraystretch{1.3}
+\begin{table*}[tbh]
+\begin{center}
+\caption{\label{tab:extranoiseoof} Additional Noise - $1/f$ Detector Noise}
+\small
+\begin{tabular}{c c c c c}
+Hits Per Pixel & Channel & 1 Bit & 2 Bit & 3 Bit \\
+\hline
+\hline
+\multirow{3}{*}{$8,000$}  & TT  & $ 0.252 \pm 0.007 $  & $ 0.064 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
+& EE  & $ 0.252 \pm 0.007 $  & $ 0.064 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
+& BB  & $ 0.25 \pm 0.018 $  & $ 0.064 \pm 0.009 $  & $ 0.018 \pm 0.004 $ \\
+\hline
+\multirow{3}{*}{$102,400$}  & TT  & $ 0.249 \pm 0.011 $  & $ 0.064 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
+& EE  & $ 0.25 \pm 0.011 $  & $ 0.064 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
+& BB  & $ 0.247 \pm 0.018 $  & $ 0.064 \pm 0.009 $  & $ 0.018 \pm 0.004 $ \\
+\hline
+\multirow{3}{*}{$1,024,000$}  & TT  & $ 0.25 \pm 0.009 $  & $ 0.063 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
+& EE  & $ 0.252 \pm 0.008 $  & $ 0.065 \pm 0.003 $  & $ 0.018 \pm 0.002 $ \\
+& BB  & $ 0.251 \pm 0.008 $  & $ 0.064 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
+\end{tabular}
+\tablecomments{ 
+Percent addition to the map noise level $\Delta \sigma / \sigma$ by digitisation. Notice the invariance for each digitisation scheme with the number of hits per pixel and channel.
+} \normalsize
+\end{center}
+\end{table*}
+
+% Power preservation theoretical calculation
+\subsection{Preserving Power in Digitised Data}
+\label{subsec:appendixpreservepower}
+
+As mentioned in subsection \ref{subsec:extremedigitisation} the designed digitisation schemes do not conserve power.
+This is because the distortion metric used does not demand this.
+We can ensure power preservation by multiplying the output by a constant $\mathcal{N}$.
+In the following we will show that this is true and derive an expression for $\mathcal{N}$.
+
+We demand equal cross power in input and output, i.e.
+
+\begin{equation} \label{eq:normcrosspower}
+\langle s_1 s_2 \rangle = \mathcal{N} \langle D(s_1) D(s_2) \rangle,
+\end{equation}
+
+where timestreams $s_{1,2}$ sample the same underlying signal, $\mu$, but add different noise realisations $\xi_{1, 2}$.
+The digitisation process is denoted by $D(\dots)$.
+We begin by looking at the crosspower of non-digitised timestreams
+
+\begin{equation}
+\langle s_1 s_2 \rangle = \langle (\mu + \xi_1) (\mu + \xi_2) \rangle =  \langle \mu^2 \rangle = \mu^2.
+\end{equation}
+
+This allows us to state
+
+\begin{equation}
+cov(s_1, s_2) = \langle s_1 s_2 \rangle - \langle s_1 \rangle \langle s_2 \rangle = \mu^2 - \mu^2 = 0.
+\end{equation}
+
+We argue that this still holds true after the digitisation process, since timestreams still have independent stochastic components.
+Digitisation discretises the probability distributions at play, but does not change the underlying dynamics.
+
+\begin{equation}
+\langle D(s_1) D(s_2) \rangle = \sum_1 \sum_2 D(\mu + \xi_1) D(\mu + \xi_1) P(D(\mu + \xi_1)) P(D(\mu + \xi_2)),
+\end{equation}
+
+where $P$ denotes the probability of a certain digitisation level occurring.
+We now assume that the underlying signal is small, i.e.
+$| \mu | \ll | \xi |$.
+This is true for ground-based CMB observations.
+In this limit $D(\mu + \xi)$ is digitised to the same value as $D(\xi)$.
+The probabilities can then be expressed as integrals over the relevant sections of correctly parameterised probability distributions.
+We proceed by assuming that noise is drawn from normal distributions with mean $\mu$ and standard deviation $\sigma$, i.e.
+$N(\mu, \sigma, \xi)$.
+
+\begin{equation}
+\langle D(s_1) D(s_2) \rangle = \sum_{i=0}^N \sum_{j=0}^N  \int_{x_i}^{x_{i+1}} \int_{x_j}^{x_{j+1}} y_i y_j N(\mu, \sigma, \xi_1) N(\mu, \sigma, \xi_2) d\xi_1 d\xi_2.
+\end{equation}
+
+We separate the above sums temporarily to detail our next steps. We split up the integrals over ranges that allow us to recognise the error function evaluated at different points.
+
+\begin{equation}
+\langle D(s_1) D(s_2) \rangle = \left[ \sum_{i=0}^N  y_i \int_{x_i}^{x_{i+1}} N(\mu, \sigma)  d\xi_1 \right] \left[  \sum_{j=0}^N y_j \int_{x_j}^{x_{j+1}} N(\mu, \sigma)d\xi_2 \right].
+\end{equation}
+
+\begin{equation}
+= \left\{ \sum_{i=0}^N  y_i \left[ \int_{-\infty}^{x_{i+1}} - \int_{-\infty}^{x_{i}} \right] N(\mu, \sigma)  d\xi_1 \right\} \left\{  \sum_{j=0}^N y_j \left[ \int_{-\infty}^{x_{j+1}} - \int_{-\infty}^{x_{j}} \right] N(\mu, \sigma)d\xi_2 \right\}.
+\end{equation}
+
+\begin{equation}
+= \left\{ \sum_{i=0}^N  \frac{y_i}{2} \left[ erf \left( \frac{x_{i+1} - \mu}{\sqrt{2}\sigma} \right) - erf \left( \frac{x_{i} - \mu}{\sqrt{2}\sigma} \right) \right] \right\} \left\{  \sum_{j=0}^N \frac{y_j}{2} \left[ erf \left( \frac{x_{j+1} - \mu}{\sqrt{2}\sigma} \right) - erf \left( \frac{x_{j} - \mu}{\sqrt{2}\sigma} \right) \right] \right\}.
+\end{equation}
+
+Using the Maclaurin series of the error function
+
+\begin{equation}
+\begin{aligned}
+\langle D(s_1) D(s_2) \rangle &= \left\{ \sum_{i=0}^N  \frac{y_i}{2} \left[ \frac{2}{\sqrt{\pi}} \sum_{n = 0}^\infty \frac{(-1)^n}{n!
+(2n+1)} \frac{1}{(\sqrt{2}\sigma)^{2n+1}} \left( (x_{i+1}-\mu)^{2n+1} - (x_{i}-\mu)^{2n+1} \right) \right] \right\} \\
+&\times \left\{ \sum_{j=0}^N  \frac{y_j}{2} \left[ \frac{2}{\sqrt{\pi}} \sum_{m = 0}^\infty \frac{(-1)^m}{m!
+(2m+1)} \frac{1}{(\sqrt{2}\sigma)^{2m+1}} \left( (x_{j+1}-\mu)^{2m+1} - (x_{j}-\mu)^{2m+1} \right) \right] \right\}
+\end{aligned}.
+\end{equation}
+
+The binomial expansion is now applied to all brackets containing $\mu$ and the sums over $n$ and $m$ are recombined
+
+\begin{equation}
+\begin{aligned}
+\langle D(s_1) D(s_2) \rangle &= \left\{ \sum_{i=0}^N  \frac{y_i}{\sqrt{pi}} \left[ \sum_{n = 0}^\infty \frac{(-1)^n}{n!
+(2n+1)} \frac{1}{(\sqrt{2}\sigma)^{2n+1}} \left( \sum_{k=0}^{2n+1} {2n+1 \choose k} \mu^k ( x_{i+1}^{2n+1-k} - x_{i}^{2n+1-k} ) \right) \right] \right\} \\
+&\times \left\{ \sum_{j=0}^N  \frac{y_j}{\sqrt{2}} \left[ \sum_{m = 0}^\infty \frac{(-1)^m}{m!
+(2m+1)} \frac{1}{(\sqrt{2}\sigma)^{2m+1}} \left( \sum_{l=0}^{2m+1} {2m+1 \choose l} \mu^l ( x_{j+1}^{2m+1-l} - x_{j}^{2m+1-l} ) \right) \right] \right\}.
+\end{aligned}
+\end{equation}
+
+\begin{equation}
+\begin{aligned}
+&= \sum_{i,j=0}^N  \frac{y_i y_j}{\pi} \left[ \sum_{n,m = 0}^\infty \frac{1}{n! (2n+1)} \frac{(-1)^{n+m}}{m! (2m+1)} \frac{1}{(\sqrt{2}\sigma)^{2n+2m+2}} \right. \\
+& \left. \times \left( \sum_{k, l = 0}^{2n+1} {2n+1 \choose k} {2m+1 \choose l} \mu^{k+l} ( x_{i+1}^{2n+1-k} - x_{i}^{2n+1-k} ) ( x_{j+1}^{2m+1-l}- x_{j}^{2m+1-l}) \right) \right].
+\end{aligned}
+\end{equation}
+
+\begin{equation} \label{eq:dcompleteklsum}
+=  \frac{1}{\pi} \sum_{n,m = 0}^\infty \frac{1}{n!
+(2n+1)} \frac{(-1)^{n+m}}{m!
+(2m+1)} \frac{1}{(\sqrt{2}\sigma)^{2n+2m+2}} \sum_{k, l = 0}^{2n+1} {2n+1 \choose k} {2m+1 \choose l} \mu^{k+l} \sum_{i,j=0}^N y_i y_j \psi^{k, n}_i \psi^{l, m}_j,
+\end{equation}
+
+where in the last line we have moved $y_i$ and $y_j$ and the sum over $i,j$ as far forward as possible. Additionally, we have introduced the new variable
+
+\begin{equation}
+\psi_i^{k,n} = x_{i+1}^{2n+1-k} - x_{i}^{2n+1-k}.
+\end{equation}
+
+We now split the innermost term such that
+
+\begin{equation}
+\sum_{i,j = 0}^N y_i y_j \psi_i^{k, n} \psi_j^{l, m} = \left( \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} + \sum_{i = N/2}^{N}\sum_{j = 0}^{N/2} + \sum_{i = 0}^{N/2}\sum_{j = N/2}^{N} + \sum_{i = N/2}^{N}\sum_{j = N/2}^{N} \right) y_i y_j \psi_i^{k, n} \psi_j^{l, m}.
+\end{equation}
+
+\begin{equation}
+ = \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} y_i y_j \psi_i^{k, n} \psi_j^{l, m} + \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} y_{N-i} y_j \psi_{N-i}^{k, n} \psi_j^{l, m} + \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} y_i y_{N-j} \psi_i^{k, n} \psi_{N-j}^{l, m} + \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} y_{N-i} y_{N-j} \psi_{N-i}^{k, n} \psi_{N-j}^{l, m},
+\end{equation}
+
+where we have shifted all indices of sums that do not start at $0$ to $N-i$ and/or $N-j$ as applicable. We recombine the sums such that
+
+\begin{equation}
+\sum_{i,j = 0}^N y_i y_j \psi_i^{k, n} \psi_j^{l, m} = \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} \left( y_i y_j \psi_i^{k, n} \psi_j^{l, m} +  y_{N-i} y_j \psi_{N-i}^{k, n} \psi_j^{l, m} + y_i y_{N-j} \psi_i^{k, n} \psi_{N-j}^{l, m} + y_{N-i} y_{N-j} \psi_{N-i}^{k, n} \psi_{N-j}^{l, m} \right)
+\end{equation}
+
+A digitisation scheme designed according to the distortion metric in section \ref{subsec:extremedigitisation} for normally distributed noise will satisfy $y_i = -y_{N-i}$ and $x_i = -x_{N+1-i}$.
+This simply states symmetry of the digitisation thresholds and levels about zero or after a mean has been subtracted.
+With this information we can condense the above to
+
+\begin{equation} \label{eq:ijsum}
+\sum_{i,j = 0}^N y_i y_j \psi_i^{k, n} \psi_j^{l, m} = \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} y_i y_j \left( \psi_i^{k, n} \psi_j^{l, m} - \psi_{N-i}^{k, n} \psi_j^{l, m} - \psi_i^{k, n} \psi_{N-j}^{l, m} + \psi_{N-i}^{k, n} \psi_{N-j}^{l, m} \right)
+\end{equation}
+
+We keep hold of this result and move onto the double sum over $k,l$ in equation \ref{eq:dcompleteklsum}.
+Since we assumed that the signal is small we would like to calculate the leading order contributions in $\mu$.
+We split up the double sum such that
+
+\begin{equation} \label{eq:ksums}
+\sum_{k, l = 0}^{2n+1} = \sum_{\mathrm{even} \> k}^{2n+1} \sum_{\mathrm{even} \> l}^{2n+1}  + \sum_{\mathrm{odd} \> k}^{2n+1} \sum_{\mathrm{odd} \> l}^{2n+1} + \sum_{\mathrm{odd} \> k}^{2n+1} \sum_{\mathrm{even} \> l}^{2n+1}  + \sum_{\mathrm{even} \> k}^{2n+1} \sum_{\mathrm{odd} \> l}^{2n+1}.
+\end{equation}
+
+The first two terms in the above will lead to even powers of $\mu$ whereas the last two will lead to odd powers of $\mu$.
+Before we can look at these terms in more detail we need to see how the index shift $i \to N-i$ affects $\psi_{i}^{k, n}$.
+
+\begin{equation}
+\begin{aligned}
+\psi_{N-i}^{k,n} &= x_{N+1-i}^{2n+1-k} - x_{N-i}^{2n+1-k} = (-x_{i})^{2n+1-k} - (-x_{i+1})^{2n+1-k}\\
+&= (-1)^{2n+1-k} (x_{i})^{2n+1-k} - x_{i+1})^{2n+1-k} ) = (-1)^{2n+2-k} (x_{i+1})^{2n+1-k} - x_{i1})^{2n+1-k} )\\
+&= (-1)^{k} \psi_{i}^{k,n}.
+\end{aligned}
+\end{equation}
+
+We are now ready to put all pieces together.
+We start by investigating even powers of $\mu$ in equation \ref{eq:dcompleteklsum} that arise from even $k$ and $l$.
+Under the assumptions made \ref{eq:ijsum} reduces to
+
+\begin{equation}
+\sum_{i,j = 0}^N y_i y_j \psi_i^{k, n} \psi_j^{l, m} = \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} y_i y_j  \left( \psi_i^{k, n} \psi_j^{l, m} - \psi_{i}^{k, n} \psi_j^{l, m} - \psi_i^{k, n} \psi_{j}^{l, m} + \psi_{i}^{k, n} \psi_{j}^{l, m} \right) = 0.
+\end{equation}
+
+The term vanishes. The other way of producing even powers of $\mu$ in equation \ref{eq:dcompleteklsum} is with odd $k$ and $l$. These assumptions simplify equation \ref{eq:ijsum} to
+
+\begin{equation}
+\sum_{i,j = 0}^N y_i y_j \psi_i^{k, n} \psi_j^{l, m} = \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} y_i y_j \left( \psi_i^{k, n} \psi_j^{l, m} + \psi_{i}^{k, n} \psi_j^{l, m} + \psi_i^{k, n} \psi_{j}^{l, m} + \psi_{i}^{k, n} \psi_{j}^{l, m} \right) = 4 \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} y_i y_j \psi_i^{k, n} \psi_j^{l, m}.
+\end{equation}
+
+This term survives and leads to even powers of $\mu$ in equation \ref{eq:dcompleteklsum}. Moving on to odd powers of $\mu$ we consider the case of odd $k$ and even $l$. We observe
+
+\begin{equation}
+\sum_{i,j = 0}^N y_i y_j \psi_i^{k, n} \psi_j^{l, m} = \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} y_i y_j \left( \psi_i^{k, n} \psi_j^{l, m} + \psi_{i}^{k, n} \psi_j^{l, m} - \psi_i^{k, n} \psi_{j}^{l, m} - \psi_{i}^{k, n} \psi_{j}^{l, m} \right) = 0
+\end{equation}
+
+and similarly
+
+\begin{equation}
+\sum_{i,j = 0}^N y_i y_j \psi_i^{k, n} \psi_j^{l, m} = \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} y_i y_j \left( \psi_i^{k, n} \psi_j^{l, m} - \psi_{i}^{k, n} \psi_j^{l, m} + \psi_i^{k, n} \psi_{j}^{l, m} - \psi_{i}^{k, n} \psi_{j}^{l, m} \right) = 0\\
+\end{equation}
+
+for odd $k$ and even $l$.
+All odd powers of $\mu$ in equation \ref{eq:dcompleteklsum} vanish.
+Moreover we can also state that there is no $\mu$ independent term, since this could only be produced by $k=l=0$, but even $k$ and $l$ terms are shown to vanish.
+Therefore
+
+\begin{equation}
+\langle D(s_1) D(s_2) \rangle = c \mu^2 + \mathcal{O}(\mu^4)
+\end{equation}
+
+where $c$ is some constants that depend on the specific digitisation scheme chosen.
+We now have prove that in the small signal limit and under the assumptions about the digitisation scheme stated normalisation can indeed be achieved by a constant.
+Returning to equation \ref{eq:normcrosspower} we see
+
+\begin{equation}
+\mathcal{N}  = \frac{\langle s_1 s_2 \rangle}{\langle D(s_1) D(s_2) \rangle} = \frac{\mu^2}{c \mu^2 + \mathcal{O}(\mu^4)} = \frac{1}{c + \mathcal{O}(\mu^2)} = \frac{1}{c},
+\end{equation}
+
+where we have ignored higher order contributions. From equation \ref{eq:dcompleteklsum} we obtain an expression for $\mathcal{N}$. We focus on $k=l=1$ terms.
+
+\begin{equation}
+\begin{aligned}
+c = \sum_{i,j=0}^N  \frac{y_i y_j}{\pi} \left[ \sum_{n,m = 0}^\infty \frac{1}{n!
+(2n+1)} \frac{(-1)^{n+m}}{m!
+(2m+1)} \frac{1}{(\sqrt{2}\sigma)^{2n+2m+2}} \left( {2n+1 \choose 1} {2m+1 \choose 1} ( x_{i+1}^{2n} - x_{i}^{2n} ) ( x_{j+1}^{2m} - x_{j}^{2m} ) \right) \right]
+\end{aligned}
+\end{equation}
+
+\begin{equation}
+c = \sum_{i,j=0}^N  \frac{y_i y_j}{\pi} \left[ \sum_{n,m = 0}^\infty \frac{1}{n!
+(2n+1)} \frac{(-1)^{n+m}}{m!
+(2m+1)} \frac{1}{(\sqrt{2}\sigma)^{2n+2m+2}} \left( \frac{(2n+1)!}{(2n)!} \frac{(2m+1)!}{(2m)!} ( x_{i+1}^{2n} - x_{i}^{2n} ) ( x_{j+1}^{2m} - x_{j}^{2m} ) \right) \right]
+\end{equation}
+
+\begin{equation}
+c = \sum_{i,j=0}^N \frac{y_i y_j}{2\pi\sigma^2} \left[ \sum_{n = 0}^\infty \frac{(-1)^n}{n!} \frac{1}{(\sqrt{2}\sigma)^{2n}}  ( x_{i+1}^{2n} - x_{i}^{2n} ) \right] \left[ \sum_{m = 0}^\infty \frac{(-1)^m}{m!} \frac{1}{(\sqrt{2}\sigma)^{2m}} ( x_{j+1}^{2m} - x_{j}^{2m} ) \right].
+\end{equation}
+
+We recognise the exponential function in the above.
+
+\begin{equation}
+c = \sum_{i,j=0}^N \frac{y_i y_j}{2\pi\sigma^2} \left\{ exp \left[ - \left(\frac{x_{i+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{i}}{\sqrt{2}\sigma} \right)^2 \right] \right\} \left\{ exp \left[ - \left(\frac{x_{j+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{j}}{\sqrt{2}\sigma} \right)^2 \right] \right\}
+\end{equation}
+
+The normalisation constant therefore is given by
+
+\begin{equation}
+\mathcal{N} = \left( \sum_{i,j=0}^N \frac{y_i y_j}{2\pi\sigma^2} \left\{ exp \left[ - \left(\frac{x_{i+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{i}}{\sqrt{2}\sigma} \right)^2 \right] \right\} \left\{ exp \left[ - \left(\frac{x_{j+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{j}}{\sqrt{2}\sigma} \right)^2 \right] \right\} \right)^{-1}
+\end{equation}
+
+This can be calculated for a given digitisation scheme $N, x_i, y_i$ and standard deviation of noise $\sigma$.
+Notice that if the digitisation scheme is chosen such that $y_i$ and $x_i$ depend linearly on $\sigma$ the normalisation constant is independent of $\sigma$.
+
+\newpage
 

### 2018-09-17 digitisation.tex
Editing

@@
 
-Excellent observation conditions for CMB observations can be found at various locations on earth, including the South Pole \citep{kovac2007}, and in space.
+Excellent observation conditions for CMB observations can be found in space and at various locations on earth, including the South Pole \citep{kovac2007}.
 Space- and Antarctica-based experiments depend on satellite transmission.
-An experiment at the South Pole will be part of the next generation of ground-based instruments, which aims to collect 2 million detector years in total.
-Such an experiment will face a data influx of $\sim \mathcal{O}(10)\mathrm{Tb/d}$.
-However, the current transmission allocation for SPT3G is at $150\mathrm{Gb/d}$, which will likely see only a moderate increase in coming years.
+The South Pole will continue its contribution to CMB science in the next generation of ground-based instruments, which aims to collect 2 million detector years in total.
+During Stage-4 operation the Sout Pole site will face a data influx of $\sim \mathcal{O}(10)\mathrm{Tb/d}$.
+However, the current transmission allocation for SPT3G is at $150\mathrm{Gb/d}$, which will likely see only a moderate increase.
 The transmission bottleneck is currently overcome by recovering the full data on hard drives with some latency and transmitting a downsampled version of the data.
-The downsampling process loses high frequency information.
-Continued use of downsampling will narrow the information window decisively - prohibiting high multipole moment science to be carried out on the transmitted data set.
-This also means that any potential faults or errors in the experiment visible in this range are not visible in the transmitted data.
+The downsampling process loses high frequency information and continued usage of it will narrow the information window decisively - prohibiting high multipole moment science to be carried out on the transmitted data set.
+This also means that any potential faults or errors in the experiment apparent in this range are not visible in the transmitted data.
 %C: Is there a paper detailing trasmission limits for SPT and that the data is currently downsampled?
@@
 Upcoming missions aim to exceed the detector count of Planck by at least an order of magnitude \citep{litebird2014, pixie2011, core2018}.
-However, strategies to meet the telemetry specifications of each mission appear to be in development.
+Strategies to meet the telemetry specifications of each mission appear to be in development.
 It is not clear whether the data compression knowledge developed during the Planck mission will guarantee optimal performance for future satellites.
@@
 Space-based missions carry out a similar analysis specific to their instruments to optimise their science output.
-Ground-based observations have to rely on different planning strategies or aim to reduce the size of the TOD in order to maximise the productivity of instruments.
+Ground-based observations have to rely on different planning strategies or aim to reduce the size of the TOD in order to maximise the productivity of the design stage.
 
@@
 
-The Planck mission has demonstrated the advantages of using few bit compression of the CMB data \citep{maris2003}.
+The Planck mission has demonstrated the advantages of using few bit CMB data \citep{maris2003}.
 Other science areas have also shown the power of extreme digtisation.
@@
 Recently Clearwater et al.
-(private communication) have investigated the advantages of using 1 and 2 bit data when searching for continuous gravitational waves using the Laser Interferometer Gravitational-Wave Observatory (LIGO).
+(private communication) have demonstrated the advantages of using 1 and 2 bit data when searching for continuous gravitational waves using the Laser Interferometer Gravitational-Wave Observatory (LIGO).
 
@@
 Digitisation is a lossy compression technique.
-However, the induced noise depends on the details of the digitisation scheme.
-To minimise the noise induced through this process one must know the nature of the input signal.
+However, the additional noise depends on the details of the digitisation scheme.
+To minimise distortion knowledge of the input signal is essential.
 There is usually little concern around finding an optimal set of digitisation parameters for an input signal as it is common to have access to large numbers of bits to store information, where changes in the digitisation scheme become insignificant to the distortion induced.
@@
 In this case the digitisation scheme must be carefully constructed.
-We review the key of digitisation as edtablished by \cite{max1960} below.
+We review the key aspedcts of designing digitisation schemes as edtablished by \cite{max1960} below.
 
@@
 
-We use the findings of \cite{max1960} to formulate the multi-level functions we use for our 1, 2 and 3 bit digitisation process.
+We use the findings of \cite{max1960} to formulate the multi-level functions we use for 1, 2 and 3 bit digitisation.
 Given an input signal $s(t)$ a digitisation scheme using $N$ bits returns the output $\hat{s}_N(t)$.
@@
 We produce 24 CMB maps in total: three control maps and nine maps with three each obtained from the different digitisation schemes.
-We simulate six observations, producing maps with $800$, $8,000$, $80,000$, $1,024,000$, $10,240,000$ and $102,400,000$ hits in each pixel.
+We simulate six observations with $800$, $8,000$, $80,000$, $1,024,000$, $10,240,000$ and $102,400,000$ hits per map pixel.
 
@@
 
-We use the Spatially Inhomogeneous Correlation Estimator for Temperature and Polarisation \citep[PolSpice;][]{polspice} to compute $\mathrm{TT}, \mathrm{EE}$ and $\mathrm{BB}$ power spectra.
+We use the Spatially Inhomogeneous Correlation Estimator for Temperature and Polarisation \citep[PolSpice;][]{polspice} to compute $\mathrm{TT}$, $\mathrm{EE}$ and $\mathrm{BB}$ power spectra.
 To minimise cut sky effects we apodise the observed path using a cosine mask with $\sigma_{\mathrm{APOD}} = \sqrt{600}/2 \hspace{0.7ex} \mathrm{deg}$.
@@
 To investigate whether the noise added through digitisation has any $\ell$ dependence we subtract the template maps from the simulated observations.
-We use theoretically calculated normalisation constants to minimise calibration error.
-The power spectra of the difference maps do not show any $\ell$ dependence in the added noise as seen in figure \ref{fig:diffpswn}.
+Beyond a calibration error, the power spectra of the difference maps do not show any $\ell$ dependence in the added noise as seen in figure \ref{fig:diffpswn}.
 % This is in agreement with (planck website) but does not match the results of (planck short paper).
@@
 Before analysing $C_\ell^\mathrm{X}/C_\ell^\mathrm{N}$ we rebin the power spectra to $\Delta \ell \approx 100$.
-This guarantees that the points in the noise tail are independent of one another, allowing us to extract an uncertainty for the above quantity.
+This ensures that the points in the noise tail are independent of one another, allowing us to extract an uncertainty for the above quantity.
 
-The deduced additional noise for 1, 2 and 3 bit digitisation schemes are shown in table \ref{tab:extranoisewhite} in the appendix.
+The deduced additional noise for 1, 2 and 3 bit digitisation schemes is shown in table \ref{tab:extranoisewhite} in the appendix.
 As expected 3 Bit digitisation performs the best, followed by 2 Bit and finally 1 Bit digitisation.
-This i because with each additional bit we retain more information.
+This is because with each additional bit we retain more information.
 We observe that $\Delta \sigma / \sigma$ is independent of the hits per pixel in the maps.
@@
 Lastly, the added noise levels are low.
-The optimal 3 Bit digitisaiton scheme devised adds as little as $1.8\%$ to the map noise level.
+The optimal 3 Bit digitisaiton scheme devised adds as little as $1.7\%$ to the map noise level.
 Keeping in mind that CMB detector sensitivity improves in steps of order of magnitude every few years, adding an extra percent-level noise term does not deteriorate the results appreciably.
@@
 We follow up on the compelling results of the white noise case by implementing a more realistic noise profile.
-Noise profiles of ground-based CMB observations have prominent $1/f$ character due to atmospheric effects \citep{lay2000}.
+Ground-based CMB observations have $1/f$ noise spectra due to atmospheric effects \citep{lay2000}.
 We use the Fourier space noise profile
@@
 
-where the first term on the second line produces $1/f$ behaviour and the second term on the same line suppressed the lowest modes.
+where $N\left(0, \sigma \right)$ denotes a normally distributed random variable with mean $0$ and standard deviation $\sigma$.
+the first term on the second line produces $1/f$ behaviour and the second term on the same line suppressed the lowest modes.
 For temperature observations we choose $\lknee = 1000$ and $\lcut = 336$, to ensure that $|N_{1/f}(0.3\lknee)|^2 = |0.95 \sigma|^2$.
@@
 The noise contribution to the map noise level from digitisation for $1/f$ noise is the same as for white noise and does not vary with the number of hits per pixel in the map.
-Temperature and polarisation observations incur the same distortion.
+Temperature and polarisation observations incur the same noise penalty.
 Table \ref{tab:extranoiseoof} in the appendix displays the values of $\Delta \sigma / \sigma$ calculated.
@@
 Through simulated observation we determined the distortion this compression technique causes in temperate and polarisation power spectra.
-We find that an optimal 3 bit digitisation adds as little as $1.8\%$ to the map noise level for temperature and polarisation observations alike.
+We find that an optimal 3 bit digitisation adds as little as $1.7\%$ to the map noise level for temperature and polarisation observations alike.
 No change in the results is observed for maps of different hits per pixel and no $\ell$ dependence was observed in the added noise.
@@
 \hline
-\multirow{3}{*}{$800$}  & TT  & $ 0.252 \pm 0.009 $  & $ 0.064 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
-& EE  & $ 0.252 \pm 0.012 $  & $ 0.064 \pm 0.005 $  & $ 0.017 \pm 0.004 $ \\
-& BB  & $ 0.252 \pm 0.011 $  & $ 0.066 \pm 0.005 $  & $ 0.018 \pm 0.004 $ \\
+\multirow{3}{*}{800}  & TT  & $ 0.252 \pm 0.009 $  & $ 0.064 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
+& EE  & $ 0.253 \pm 0.012 $  & $ 0.064 \pm 0.005 $  & $ 0.017 \pm 0.004 $ \\
+& BB  & $ 0.254 \pm 0.011 $  & $ 0.065 \pm 0.005 $  & $ 0.018 \pm 0.004 $ \\
 \hline
-\multirow{3}{*}{$8,000$}  & TT  & $ 0.252 \pm 0.007 $  & $ 0.064 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
+\multirow{3}{*}{8000}  & TT  & $ 0.252 \pm 0.008 $  & $ 0.064 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
 & EE  & $ 0.252 \pm 0.007 $  & $ 0.064 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
-& BB  & $ 0.25 \pm 0.018 $  & $ 0.064 \pm 0.009 $  & $ 0.018 \pm 0.004 $ \\
+& BB  & $ 0.252 \pm 0.018 $  & $ 0.063 \pm 0.009 $  & $ 0.018 \pm 0.004 $ \\
 \hline
-\multirow{3}{*}{$80,000$}  & TT  & $ 0.249 \pm 0.011 $  & $ 0.064 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
+\multirow{3}{*}{80000}  & TT  & $ 0.249 \pm 0.011 $  & $ 0.064 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
 & EE  & $ 0.25 \pm 0.011 $  & $ 0.064 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
-& BB  & $ 0.247 \pm 0.018 $  & $ 0.064 \pm 0.009 $  & $ 0.018 \pm 0.004 $ \\
+& BB  & $ 0.248 \pm 0.018 $  & $ 0.063 \pm 0.009 $  & $ 0.018 \pm 0.003 $ \\
 \hline
-\multirow{3}{*}{$1,024,000$}  & TT  & $ 0.25 \pm 0.009 $  & $ 0.063 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
+\multirow{3}{*}{1024000}  & TT  & $ 0.25 \pm 0.009 $  & $ 0.063 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
 & EE  & $ 0.252 \pm 0.008 $  & $ 0.065 \pm 0.003 $  & $ 0.018 \pm 0.002 $ \\
-& BB  & $ 0.251 \pm 0.008 $  & $ 0.064 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
+& BB  & $ 0.252 \pm 0.008 $  & $ 0.064 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
 \hline
-\multirow{3}{*}{$10,240,000$}  & TT  & $ 0.186 \pm 0.025 $  & $ 0.047 \pm 0.007 $  & $ 0.013 \pm 0.002 $ \\
-& EE  & $ 0.252 \pm 0.009 $  & $ 0.064 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
-& BB  & $ 0.25 \pm 0.012 $  & $ 0.065 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
+\multirow{3}{*}{10240000}  & TT  & $ 0.186 \pm 0.025 $  & $ 0.046 \pm 0.007 $  & $ 0.013 \pm 0.002 $ \\
+& EE  & $ 0.253 \pm 0.009 $  & $ 0.065 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
+& BB  & $ 0.252 \pm 0.012 $  & $ 0.064 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
 \hline
-\multirow{3}{*}{$102,400,000$}  & TT  & $ 0.246 \pm 0.004 $  & $ 0.063 \pm 0.002 $  & $ 0.017 \pm 0.001 $ \\
-& EE  & $ 0.251 \pm 0.008 $  & $ 0.064 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
-& BB  & $ 0.251 \pm 0.009 $  & $ 0.065 \pm 0.003 $  & $ 0.018 \pm 0.002 $ \\
+\multirow{3}{*}{102400000}  & TT  & $ 0.246 \pm 0.005 $  & $ 0.063 \pm 0.002 $  & $ 0.017 \pm 0.001 $ \\
+& EE  & $ 0.251 \pm 0.008 $  & $ 0.064 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
+& BB  & $ 0.252 \pm 0.009 $  & $ 0.064 \pm 0.003 $  & $ 0.018 \pm 0.002 $ \\
+\hline
+\multicolumn{2}{c}{Average}  & $ 0.248 \pm 0.011 $  & $ 0.063 \pm 0.005 $  & $ 0.017 \pm 0.002 $ \\
 \end{tabular}
 \tablecomments{ 
-Percent addition to the map noise level $\Delta \sigma / \sigma$ by digitisation. Notice the invariance for each digitisation scheme with the number of hits per pixel and channel.
+Percent addition to the map noise level $\Delta \sigma / \sigma$ by digitisation for white detector noise.
+Notice the invariance for each digitisation scheme with the number of hits per pixel and channel.
 } \normalsize
@@
 & BB  & $ 0.251 \pm 0.008 $  & $ 0.064 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
+\hline
+\multicolumn{2}{c}{Average}  & $ 0.248 \pm 0.011 $  & $ 0.063 \pm 0.005 $  & $ 0.017 \pm 0.002 $ \\
 \end{tabular}
 \tablecomments{ 
-Percent addition to the map noise level $\Delta \sigma / \sigma$ by digitisation. Notice the invariance for each digitisation scheme with the number of hits per pixel and channel.
+Percent addition to the map noise level $\Delta \sigma / \sigma$ by digitisation for $1/f$ detector noise.
+The invariance with depth of observation and channel as well as the values match the white noise case.
 } \normalsize

### 2018-09-18 digitisation.tex
Editing

@@
 Unlike lossless compression, this technique introduces additional noise into the data.
-We present a set of 1, 2 and 3 bit digitisation schemes and determine the degradation the compression causes in temperature and polarisation power spectra.
+We present a set of one, two, and three bit digitisation schemes and determine the degradation the compression causes in temperature and polarisation power spectra.
 The added noise has no dependence on the multipole moment $\ell$ for white and $1/f$ detector noise.
-We note that 3 bit digitisation has a percent-level contribution to the map noise level in all cases considered.
+The compression performs equally for temperature and polarisation observations and scales the same way as detector noise with depth of observation.
+Three bit digitisation has a percent-level contribution to the map noise level.
 Such digitisation is a promising strategy for upcoming experiments.
@@
 Upcoming science goals include the discovery of inflationary gravitational waves, construction of precise lensing maps and thorough studies of the Sunyaev-Zeldovich (SZ) effects.
-Additionally, the CMB probes structure formation and reionisation.
-It is a powerful test for dark matter and dark energy models.
+Additionally, the CMB probes structure formation, reionisation and is a powerful test for dark matter and dark energy models.
 Lastly, the relativistic number of species and the neutrino mass sum are measured by CMB experiments \citep{s4sciencebook, core2011, litebird2016, pixie2011}.
 
-The CMB community has developed a variety of compression techniques and computational approaches to handle the increasing volume of data the science goals demand \citep{tristam2007}.
+The CMB community has developed a variety of compression techniques and computational approaches to handle the increasing volume of data \citep{tristam2007}.
 These include the compression of time-ordered data (TOD) into maps \citep{tegmark1997}, bandpower estimation \citep{tegmark1998} and the pseudo $C_l$ method \citep{brown2005}.
@@
 Space-based experiments have employed a combination of lossless and lossy compression techniques, including reduced bits in the TOD \citep{gaztanaga1998, maris2003}.
-Antarctica-based experiments that transmit a portion of their data via a satellite link have downsampled their data in the past to meet their telemetry requirements and have not yet to used few bit digitisation of the TOD.
-As we approach the next generation of ground-based experiments, CMB-S4, and the launch of a new generation of space-based missions (LiteBIRD, PIXIE, COrE+), we must consider potential transmission bottlenecks carefully.
+Antarctica-based experiments that transmit a portion of their data via a satellite link have downsampled their data in the past to meet telemetry allocations and have not yet used few bit digitisation of the TOD.
+As we approach the next generation of ground-based experiments, CMB-S4, and the launch of new space-based missions (LiteBIRD, PIXIE, COrE+), we must consider potential transmission bottlenecks carefully.
 %C: Is there a reference detailing downsampling of SPT data?
 
-In this work we present the method of extreme digitisation, which reduces a many bit (often 32-or 64-bit) signal to a few bits for ground-based experiments.
+In this work we present the method of extreme digitisation, which reduces a many bit (often 32 or 64 bit) signal to a few bits for ground-based experiments.
 We apply extreme digitisation to the TOD and detail the resulting effects on temperature and polarisation power spectra.
-We find that an optimal 3-bit digitisation scheme adds $< 2\%$ to the map noise level.
+We find that an optimal three bit digitisation scheme adds $<1.7\%$ to the map noise level.
 
@@
 
-Excellent observation conditions for CMB observations can be found in space and at various locations on earth, including the South Pole \citep{kovac2007}.
+Excellent conditions for CMB observations can be found in space and at various locations on earth, including the South Pole \citep{kovac2007}.
 Space- and Antarctica-based experiments depend on satellite transmission.
 The South Pole will continue its contribution to CMB science in the next generation of ground-based instruments, which aims to collect 2 million detector years in total.
-During Stage-4 operation the Sout Pole site will face a data influx of $\sim \mathcal{O}(10)\mathrm{Tb/d}$.
-However, the current transmission allocation for SPT3G is at $150\mathrm{Gb/d}$, which will likely see only a moderate increase.
+During CMB-S4 observations at the South Pole site will face a data influx of $\sim \mathcal{O}(10)\mathrm{Tb/d}$.
+However, the current transmission allocation for SPT-3G is $150\mathrm{Gb/d}$, which will likely see only a moderate increase.
 The transmission bottleneck is currently overcome by recovering the full data on hard drives with some latency and transmitting a downsampled version of the data.
-The downsampling process loses high frequency information and continued usage of it will narrow the information window decisively - prohibiting high multipole moment science to be carried out on the transmitted data set.
+The downsampling process loses high frequency information and continued usage of it will narrow the information window decisively, prohibiting high multipole moment science to be carried out on the transmitted data set.
 This also means that any potential faults or errors in the experiment apparent in this range are not visible in the transmitted data.
@@
 It is not clear whether the data compression knowledge developed during the Planck mission will guarantee optimal performance for future satellites.
-Most likely each mission will have to carefully construct a compression algorithm through a combination of lossless and lossy techniques, specific to their requirements.
+Each mission will most likely have to carefully construct a compression algorithm through a combination of lossless and lossy techniques specific to their requirements.
 
@@
 Beyond transmission challenges, mission planning is becoming exceedingly difficult.
-Full simulation of TOD over the entire parameter space of detection scenarios is the desired way to decide on Stage-4 configurations.
+Full simulation of TOD over the entire parameter space of detection scenarios is the desired way to decide on CMB-S4 configurations.
 Given the shear size of TOD expected, this is not possible \citep{s4sciencebook}.
-Space-based missions carry out a similar analysis specific to their instruments to optimise their science output.
 Ground-based observations have to rely on different planning strategies or aim to reduce the size of the TOD in order to maximise the productivity of the design stage.
+Space-based missions carry out a similar analysis specific to their instruments to heighten their science output.
 
@@
 
-Operations on the TOD, such as noise-removal or map-making are a vital part of CMB data analysis.
+Operations on the TOD, such as noise-removal and map-making are a vital part of CMB data analysis.
 While we have not experienced the limitations of the accessible computational assets, the exponential growth of CMB data makes its analysis increasingly expensive.
@@
 Together with already established lossless compression techniques (e.g.
-FLAC, run-length coding, Huffman coding, etc.) this will directly tackle transmission hurdles.
+FLAC, run-length coding, Huffman coding, etc.) this addresses transmission hurdles.
 Beyond that extreme digitisation has the potential of easing issues in planning, analysis and hardware requirements.
 
-The Planck mission has demonstrated the advantages of using few bit CMB data \citep{maris2003}.
-Other science areas have also shown the power of extreme digtisation.
+%what others have done
+
+While the advantages of using few bit TOD in space-based CMB observations have been demonstrated by the Planck mission \citep{maris2003}, other science areas have also shown the power of this tool.
 \cite{jenet1998} explored the application of such compression to radio pulsar timing measurements with success.
 Recently Clearwater et al.
-(private communication) have demonstrated the advantages of using 1 and 2 bit data when searching for continuous gravitational waves using the Laser Interferometer Gravitational-Wave Observatory (LIGO).
+(private communication) have demonstrated the advantages of using one and two bit data when searching for continuous gravitational waves using the Laser Interferometer Gravitational-Wave Observatory (LIGO).
 
@@
 Digitisation is a lossy compression technique.
-However, the additional noise depends on the details of the digitisation scheme.
-To minimise distortion knowledge of the input signal is essential.
-There is usually little concern around finding an optimal set of digitisation parameters for an input signal as it is common to have access to large numbers of bits to store information, where changes in the digitisation scheme become insignificant to the distortion induced.
+However, the additional noise depends on the details of the digitisation scheme and the input signal.
+There is usually little concern around finding an optimal set of digitisation parameters as it is common to have access to large numbers of bits to store information, where changes in the digitisation scheme become insignificant to the distortion induced.
 It is in our interest however to consider extreme, i.e.
-few-bit, digitisation.
+few bit, digitisation.
 In this case the digitisation scheme must be carefully constructed.
@@
 
-Since we wish to minimise the distortion we differentiate the above with respect to $x_i$ and $y_i$ and set the derivatives to zero. We obtain the two equations
+Since we wish to minimise the distortion we differentiate the distortion with respect to $x_i$ and $y_i$ and set the derivatives to zero. We obtain the two equations
 
@@
 To progress further we have to make an assumption about the probability distribution of input signals.
-We suppose CMB observations are low signal to noise and detector noise is Gaussian white.
+We take ground-based CMB observations to operate at low signal to noise and assume detector noise to be gaussian white.
 The input is then purely stochastic and  $p(x) = (1/\sqrt{2\pi\sigma^2}) e^{-x^2/2\sigma^2}$, where $\sigma$ is the standard deviation of the detector noise.
@@
 
-We use the findings of \cite{max1960} to formulate the multi-level functions we use for 1, 2 and 3 bit digitisation.
+We use the findings of \cite{max1960} to formulate the multi-level functions we use for one, two, and three bit digitisation.
 Given an input signal $s(t)$ a digitisation scheme using $N$ bits returns the output $\hat{s}_N(t)$.
-For 1 bit digitisation we apply the sign function
+For one bit digitisation we apply the sign function
 
@@
 
-to the TOD. For 2 bit digitisation we apply the four-level function
+to the TOD. For two bit digitisation we apply the four-level function
 
@@
 
-Finally the optimal 3 bit digitisation is described by the eight-level function
+Finally the optimal three bit digitisation is described by the eight-level function
 
@@
 The fact that the single bit digitisation scheme makes no references to the standard deviation of the signal is irrelevant.
-This affects the normalisation of the signal, which we can account for during later analysis as demonstrated in the appendix.
-What is important is the spacing between digitisation thresholds and relative distance between the output levels chosen.
+This affects the normalisation of the output, which we can account for during later analysis as demonstrated in the appendix.
+What is important is the spacing between digitisation thresholds and between output levels.
 
@@
 To investigate the performance of the derived digitisation schemes we simulate many scans over CMB template maps at the timestream level.
-We obtain control maps that use 64bit TOD and maps that have undergone 1, 2 and 3 bit digitisation at the timestream level.
+We obtain control maps that use 64 bit TOD and maps that have undergone one, two, and three bit digitisation of the TOD.
 We calculate the temperature and polarisation power spectra of each map and determine the noise level.
@@
 We simulate observing a $\sim 600 \mathrm{deg^2}$ patch of the sky with a single detector.
-To do so we perform a number of constant elevation scans (CES), equally spaced in declination (DEC).
+To do so we perform constant elevation scans (CES), equally spaced in declination (DEC).
 We repeat the observation strategy $100$ times with a slight offset in right ascension (RA) and DEC each time, such that each pixel is sampled approximately uniformly.
@@
 We compress the timestreams into maps by averaging all hits falling into the same pixel.
-We produce 24 CMB maps in total: three control maps and nine maps with three each obtained from the different digitisation schemes.
-We simulate six observations with $800$, $8,000$, $80,000$, $1,024,000$, $10,240,000$ and $102,400,000$ hits per map pixel.
+This produces 24 CMB maps in total: three control maps and nine maps with three each obtained from the different digitisation schemes.
+We simulate observations with $800$, $8,000$, $80,000$, $1,024,000$, $10,240,000$ and $102,400,000$ hits per map pixel.
 
@@
 We use the Spatially Inhomogeneous Correlation Estimator for Temperature and Polarisation \citep[PolSpice;][]{polspice} to compute $\mathrm{TT}$, $\mathrm{EE}$ and $\mathrm{BB}$ power spectra.
-To minimise cut sky effects we apodise the observed path using a cosine mask with $\sigma_{\mathrm{APOD}} = \sqrt{600}/2 \hspace{0.7ex} \mathrm{deg}$.
+To minimise cut sky effects we apodise the observed patch using a cosine mask with $\sigma_{\mathrm{APOD}} = \sqrt{600}/2 \hspace{0.7ex} \mathrm{deg}$.
 %C: Should I also mention the multiplication with a cosine mask that I supply in addition to setting this option on polspice?
@@
 The digitisation schemes introduced in section \S\ref{sec:dig} are designed to minimise the distortion of the TOD, not to preserve power.
-We obtain the appropriate normalisation as described in the appendix.
+We obtain the appropriate normalisation constants as described in the appendix.
+All reconstructed maps from few bit TOD produce acceptable power spectra.
+As shown by figure \ref{fig:psrecover} temperature and polarisation alike trace the control case well.
+Variation around the input power spectrum exists since we observe a single CMB realisation and because of cut sky effects.
 
@@
 We again refer to the discussion in the appendix on how to modify digitisation schemes to ensure power preservation.%C: Do you know a good reference for this?
-
-All digitisation schemes are able to recover the input power spectra well. This is shown by figure \ref{fig:psrecover}.
 
@@
   \caption[Current ]{
-  Input plus detector noise power spectra (black) and power spectra using 3-bit TOD (violet) for observations with $\sim 10^6$ hits per map pixel.
-Differences occur because we observe a single CMB realisation.
+  Input power spectra used to generate the CMB template maps plus detector noise level (black) and power spectra using 3 bit TOD (violet).
+Simulations with $1,024,000$ hits per map pixel were used in this plot.
+No appreciable differences are visible for other depths of observation.
 \label{fig:psrecover}
@@
 To investigate whether the noise added through digitisation has any $\ell$ dependence we subtract the template maps from the simulated observations.
-Beyond a calibration error, the power spectra of the difference maps do not show any $\ell$ dependence in the added noise as seen in figure \ref{fig:diffpswn}.
+Beyond a calibration error, the power spectra of difference maps do not show any $\ell$ dependence in the added noise as seen in figure \ref{fig:diffpswn}.
+The plot also reveals that three bit digitisation performs best, followed by two bit and one bit.
+This is expected, because with each additional bit more information is retained.
+
 % This is in agreement with (planck website) but does not match the results of (planck short paper).
@@
   \caption[Current ]{
-  Power spectra of reconstructed minus input maps for 1 Bit (green), 2 Bit (orange), 3 Bit (violet) and the control case (black).
-We do not observe any $\ell$ dependence.
-Simulations with $\sim 10^6$ hits per map pixel are shown.
+  Power spectra of reconstructed minus input maps for one-bit (green), two-bit (orange), three-bit (violet) and the control case (black).
+No $\ell$ dependence is observed for all depths of observation.
+Simulations with $1,024,000$ hits per map pixel are shown.
 \label{fig:diffpswn}
@@
 
-where $\sigma_{\mathrm{map}}^{\mathrm{C}}$ is the map noise level of control observations and $\sigma_{\mathrm{map}}^{\mathrm{D}}$ is the map noise level of observations using few bit TOD.
-Because we have seen no $\ell$ dependence in the additional noise we presume digitisation to add a constant noise term to the power spectrum.
+where $\sigma_{\mathrm{map}}^{\mathrm{C}}$ and $\sigma_{\mathrm{map}}^{\mathrm{D}}$ are the map noise levels of control observations and simulations using few bit TOD respectively.
+Because we have seen no $\ell$ dependence in the power spectra of difference maps we presume digitisation to add a constant noise term.
 In the regime
@@
 Before analysing $C_\ell^\mathrm{X}/C_\ell^\mathrm{N}$ we rebin the power spectra to $\Delta \ell \approx 100$.
-This ensures that the points in the noise tail are independent of one another, allowing us to extract an uncertainty for the above quantity.
+This ensures that points in the noise tail are independent, allowing us to extract an uncertainty for the above quantity.
 
-The deduced additional noise for 1, 2 and 3 bit digitisation schemes is shown in table \ref{tab:extranoisewhite} in the appendix.
-As expected 3 Bit digitisation performs the best, followed by 2 Bit and finally 1 Bit digitisation.
-This is because with each additional bit we retain more information.
+The deduced additional noise for one, two, and three bit digitisation schemes is shown in table \ref{tab:extranoisewhite} in the appendix.
 We observe that $\Delta \sigma / \sigma$ is independent of the hits per pixel in the maps.
 This suggests that the additional noise scales in the same way as detector noise does.
-Furthermore, the compression works equally well for temperature as it does for polarisation observations.
+Furthermore, the compression performs equally well for temperature as it does for polarisation observations.
 Lastly, the added noise levels are low.
-The optimal 3 Bit digitisaiton scheme devised adds as little as $1.7\%$ to the map noise level.
-Keeping in mind that CMB detector sensitivity improves in steps of order of magnitude every few years, adding an extra percent-level noise term does not deteriorate the results appreciably.
-This is impressive, given that the use of an optimal 3 bit digitisation scheme reduces TOD volume by an order of magnitude.
+When averaged over depth of observation and channel one bit digitisation adds $(24.8\pm 1.1)\%$ to the map noise level, two bit $(6.3\pm0.5)\%$ and three bit $(1.7\pm0.2)\%$.
+This is impressive, given that all schemes considered reduce TOD volume by at least an order of magnitude.
 
@@
 where $N\left(0, \sigma \right)$ denotes a normally distributed random variable with mean $0$ and standard deviation $\sigma$.
-the first term on the second line produces $1/f$ behaviour and the second term on the same line suppressed the lowest modes.
+The first term on the second line produces $1/f$ behaviour and the second term on the same line suppresses the lowest modes.
 For temperature observations we choose $\lknee = 1000$ and $\lcut = 336$, to ensure that $|N_{1/f}(0.3\lknee)|^2 = |0.95 \sigma|^2$.
@@
 
-We carry out the same simulations as described in subsection \ref{subsec:method} using the above noise profile and digitisation scheme for maps with $8,000$, $102,400$ and $1,024,000$ hits per pixel.
-HEALPix is used to calculate the spherical harmonics transform coefficients $a\ell m$.
+We carry out the same simulations as described in subsection \ref{subsec:method} using the above noise profile and digitisation schemes for maps with $8,000$, $102,400$ and $1,024,000$ hits per pixel.
+HEALPix is used to calculate the spherical harmonics coefficients $a\ell m$.
 We mask the lowest modes in scan direction and obtain the power spectra.
@@
 We recover similar results to the white noise case.
-As before 3 bit performs the best, followed by 2 bit and 1 bit.
-The excess noise due to digitisation does not show any $\ell$ dependence.
-The noise contribution to the map noise level from digitisation for $1/f$ noise is the same as for white noise and does not vary with the number of hits per pixel in the map.
+As before three-bit performs the best, followed by two-bit and one-bit.
+The excess noise due to digitisation does not show any $\ell$ dependence and does not vary with the number of hits per pixel in the map.
 Temperature and polarisation observations incur the same noise penalty.
 Table \ref{tab:extranoiseoof} in the appendix displays the values of $\Delta \sigma / \sigma$ calculated.
+Averaging over depth of observation and channel we calculate that one bit digitisation adds $(27.7\pm 1.5)\%$ to the map noise level, two bit $(6.2\pm0.5)\%$ and three bit $(1.7\pm0.2)\%$.
+No significant deviations between white and $1/f$ noise results are found.
 
-Extreme digitisation is a viable lossy compression technique when dealing with low signal to noise, large data sets with a signal that is slowly varying with respect to the sampling rate.
+Extreme digitisation is a viable lossy compression technique when dealing with low signal to noise, large data sets, that have a slowly varying signal with respect to the sampling rate.
 These conditions allow us to reconstruct the underlying signal well.
 A small signal will also prevent saturation of the output, i.e.
-the inability of the digitised output to represent numbers larger than $y_N N_{\mathrm{hit}}$ or smaller than $y_1 N_{\mathrm{hit}}$, where $N_{\mathrm{hit}}$ is the number of data points in a given interval.
+the inability of the digitised output to represent numbers larger than $y_N N_{\mathrm{hit}}$ or smaller than $y_1 N_{\mathrm{hit}}$, where $N_{\mathrm{hit}}$ is the number of data points in an interval.
 These conditions are satisfied by CMB data and our findings reflect this.
@@
 
-We derived a set of 1, 2 and 3 bit digitisation schemes assuming white detector noise.
+We derived a set of one, two, and three bit digitisation schemes assuming white detector noise.
 Through simulated observation we determined the distortion this compression technique causes in temperate and polarisation power spectra.
-We find that an optimal 3 bit digitisation adds as little as $1.7\%$ to the map noise level for temperature and polarisation observations alike.
+We find that an optimal three bit digitisation adds as little as $1.7\%$ to the map noise level for temperature and polarisation observations alike.
 No change in the results is observed for maps of different hits per pixel and no $\ell$ dependence was observed in the added noise.
@@
 
-The additional noise is insignificant for ground-based experiments, which make order of mangitude jumps in detector count and hence sensitivity between generations.
+The additional noise is insignificant for ground-based experiments, which increase their detector count by orders of magnitude between generations.
 Space-based experiments which have their own unique issues may consider this noise penalty problematic.
@@
 % result tables
-\subsection{Additional Noise due to Digitisation}
+\subsection{Digitisation Noise Levels}
 \label{subsec:appendixnoisetables}
@@
 \hline
-\multirow{3}{*}{800}  & TT  & $ 0.252 \pm 0.009 $  & $ 0.064 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
+\multirow{3}{*}{$800$}  & TT  & $ 0.252 \pm 0.009 $  & $ 0.064 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
 & EE  & $ 0.253 \pm 0.012 $  & $ 0.064 \pm 0.005 $  & $ 0.017 \pm 0.004 $ \\
@@
 \hline
-\multirow{3}{*}{8000}  & TT  & $ 0.252 \pm 0.008 $  & $ 0.064 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
+\multirow{3}{*}{$8,000$}  & TT  & $ 0.252 \pm 0.008 $  & $ 0.064 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
 & EE  & $ 0.252 \pm 0.007 $  & $ 0.064 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
@@
 \hline
-\multirow{3}{*}{80000}  & TT  & $ 0.249 \pm 0.011 $  & $ 0.064 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
-& EE  & $ 0.25 \pm 0.011 $  & $ 0.064 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
+\multirow{3}{*}{$80,000$}  & TT  & $ 0.249 \pm 0.011 $  & $ 0.064 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
+& EE  & $ 0.250 \pm 0.011 $  & $ 0.064 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
 & BB  & $ 0.248 \pm 0.018 $  & $ 0.063 \pm 0.009 $  & $ 0.018 \pm 0.003 $ \\
 \hline
-\multirow{3}{*}{1024000}  & TT  & $ 0.25 \pm 0.009 $  & $ 0.063 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
+\multirow{3}{*}{$1,024,000$}  & TT  & $ 0.25 \pm 0.009 $  & $ 0.063 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
 & EE  & $ 0.252 \pm 0.008 $  & $ 0.065 \pm 0.003 $  & $ 0.018 \pm 0.002 $ \\
@@
 \hline
-\multirow{3}{*}{10240000}  & TT  & $ 0.186 \pm 0.025 $  & $ 0.046 \pm 0.007 $  & $ 0.013 \pm 0.002 $ \\
+\multirow{3}{*}{$10,240,000$}  & TT  & $ 0.186 \pm 0.025 $  & $ 0.046 \pm 0.007 $  & $ 0.013 \pm 0.002 $ \\
 & EE  & $ 0.253 \pm 0.009 $  & $ 0.065 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
@@
 \hline
-\multirow{3}{*}{102400000}  & TT  & $ 0.246 \pm 0.005 $  & $ 0.063 \pm 0.002 $  & $ 0.017 \pm 0.001 $ \\
+\multirow{3}{*}{$102,400,000$}  & TT  & $ 0.246 \pm 0.005 $  & $ 0.063 \pm 0.002 $  & $ 0.017 \pm 0.001 $ \\
 & EE  & $ 0.251 \pm 0.008 $  & $ 0.064 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
@@
 \tablecomments{ 
-Percent addition to the map noise level $\Delta \sigma / \sigma$ by digitisation for white detector noise.
-Notice the invariance for each digitisation scheme with the number of hits per pixel and channel.
+Percent addition to the map noise level $\Delta \sigma / \sigma$ due to one, two, and three bit digitisation for white detector noise.
+We notice that there is no appreciable variation with depth of observations and channel for a given digitisation scheme.
 } \normalsize
@@
 \hline
-\multirow{3}{*}{$8,000$}  & TT  & $ 0.252 \pm 0.007 $  & $ 0.064 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
-& EE  & $ 0.252 \pm 0.007 $  & $ 0.064 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
-& BB  & $ 0.25 \pm 0.018 $  & $ 0.064 \pm 0.009 $  & $ 0.018 \pm 0.004 $ \\
+\multirow{3}{*}{$8,000$}  & TT  & $ 0.231 \pm 0.017 $  & $ 0.058 \pm 0.005 $  & $ 0.016 \pm 0.002 $ \\
+& EE  & $ 0.300 \pm 0.012 $  & $ 0.064 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
+& BB  & $ 0.294 \pm 0.022 $  & $ 0.062 \pm 0.007 $  & $ 0.017 \pm 0.003 $ \\
 \hline
-\multirow{3}{*}{$102,400$}  & TT  & $ 0.249 \pm 0.011 $  & $ 0.064 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
-& EE  & $ 0.25 \pm 0.011 $  & $ 0.064 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
-& BB  & $ 0.247 \pm 0.018 $  & $ 0.064 \pm 0.009 $  & $ 0.018 \pm 0.004 $ \\
+\multirow{3}{*}{$102,400$}  & TT  & $ 0.235 \pm 0.014 $  & $ 0.06 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
+& EE  & $ 0.299 \pm 0.009 $  & $ 0.064 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
+& BB  & $ 0.298 \pm 0.025 $  & $ 0.064 \pm 0.007 $  & $ 0.018 \pm 0.002 $ \\
 \hline
-\multirow{3}{*}{$1,024,000$}  & TT  & $ 0.25 \pm 0.009 $  & $ 0.063 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
-& EE  & $ 0.252 \pm 0.008 $  & $ 0.065 \pm 0.003 $  & $ 0.018 \pm 0.002 $ \\
-& BB  & $ 0.251 \pm 0.008 $  & $ 0.064 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
+\multirow{3}{*}{$1,024,000$}  & TT  & $ 0.238 \pm 0.013 $  & $ 0.06 \pm 0.005 $  & $ 0.017 \pm 0.002 $ \\
+& EE  & $ 0.299 \pm 0.008 $  & $ 0.063 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
+& BB  & $ 0.299 \pm 0.010 $  & $ 0.064 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
 \hline
-\multicolumn{2}{c}{Average}  & $ 0.248 \pm 0.011 $  & $ 0.063 \pm 0.005 $  & $ 0.017 \pm 0.002 $ \\
+\multicolumn{2}{c}{Average}  & $ 0.277 \pm 0.015 $  & $ 0.062 \pm 0.005 $  & $ 0.017 \pm 0.002 $ \\
 \end{tabular}
 \tablecomments{ 
-Percent addition to the map noise level $\Delta \sigma / \sigma$ by digitisation for $1/f$ detector noise.
-The invariance with depth of observation and channel as well as the values match the white noise case.
+Percent addition to the map noise level $\Delta \sigma / \sigma$ due to one, two, and three bit digitisation for $1/f$ detector noise.
+As in the white noise case we observe no significant trend with the numbers of hits per map pixel or channel for a given digitisation scheme.
+No significant deviation in the values compared to simulations assuming white detector noise exists.
 } \normalsize

### 2018-09-20 digitisation.tex
Implemented comments made and shuffled parts of the appendix around for clarity.

@@
 
-Observations of the Cosmic Microwave Background (CMB) are of significant value to modern cosmology and particle physics.
-The large size of time-ordered-data (TOD) being recorded causes new challenges in hardware design, mission planning and analysis for future CMB experiments.
-These issues are particularly significant for Antarctica- and space-based experiments which depend on satellite links to transmit their data.
-We explore the viability of reducing the TOD to few-bit numbers to address these issues.
-Unlike lossless compression, this technique introduces additional noise into the data.
-We present a set of one, two, and three bit digitisation schemes and determine the degradation the compression causes in temperature and polarisation power spectra.
-The added noise has no dependence on the multipole moment $\ell$ for white and $1/f$ detector noise.
-The compression performs equally for temperature and polarisation observations and scales the same way as detector noise with depth of observation.
-Three bit digitisation has a percent-level contribution to the map noise level.
+Observations of the cosmic microwave background (CMB) are of significant value to modern cosmology and particle physics.
+The large size of the time ordered data (TOD) of future CMB experiments presents challenges for mission planning and data analysis.
+These issues are particularly significant for Antarctica- and space-based experiments, which depend on satellite links to transmit data.
+We explore the viability of reducing the TOD to few bit numbers to address these challenges.
+Unlike lossless compression, few bit digitisation introduces additional noise into the data.
+We present a set of one, two, and three bit digitisation schemes and measure the noise increase due to compression in the CMB temperature and polarisation power spectra.
+The added noise is independent of angular scale and is well-described as a percentage increase on the original detector noise.
+Three bit digitisation increases the map noise level by $< 2\%$, while reducing the data volume by a factor of ten relative to 32 bit floats.
 Such digitisation is a promising strategy for upcoming experiments.
@@
 
-Observations of the Cosmic Microwave Background (CMB) have played a key role in physics since its discovery \citep{penzias1965}.
-Current and future study of the information it contains will put tight constraints on cosmological models.
-Upcoming science goals include the discovery of inflationary gravitational waves, construction of precise lensing maps and thorough studies of the Sunyaev-Zeldovich (SZ) effects.
+Observations of the cosmic microwave background (CMB) have played a key role in cosmology \citep{penzias1965, smoot1992, bennett2013, sptpol2013, planck2018}.
+Current CMB Experiments target science goals such as the discovery of inflationary gravitational waves, construction of precise lensing maps and thorough studies of the Sunyaev-Zeldovich (SZ) effects.
 Additionally, the CMB probes structure formation, reionisation and is a powerful test for dark matter and dark energy models.
-Lastly, the relativistic number of species and the neutrino mass sum are measured by CMB experiments \citep{s4sciencebook, core2011, litebird2016, pixie2011}.
+Lastly, the relativistic number of species and the neutrino mass sum can be measured by CMB experiments \citep{s4sciencebook, core2011, litebird2016, pixie2011}.
 
 The CMB community has developed a variety of compression techniques and computational approaches to handle the increasing volume of data \citep{tristam2007}.
-These include the compression of time-ordered data (TOD) into maps \citep{tegmark1997}, bandpower estimation \citep{tegmark1998} and the pseudo $C_l$ method \citep{brown2005}.
+These include the compression of time-ordered data (TOD) into maps \citep{tegmark1997}, bandpower estimation \citep{tegmark1998}, and the pseudo-$C_l$ method \citep{brown2005}.
 
@@
 We apply extreme digitisation to the TOD and detail the resulting effects on temperature and polarisation power spectra.
-We find that an optimal three bit digitisation scheme adds $<1.7\%$ to the map noise level.
+We find that an optimal three bit digitisation scheme adds $<2\%$ to the map noise level.
 
@@
 The science goals of upcoming CMB experiments depend on achieving substantially faster mapping speeds.
-This has lead to the number of detectors of ground-based experiments following a Moore's law like trend, doubling approximately every 2 years \citep{s4sciencebook, Abazajian2015}.
-The growing detector count translates into an exponential growth in data volume.
+Given CMB detectors are generally photon noise limited, the number of detectors (and data volume) of ground-based experiments has followed a trend like Moore's law, doubling approximately every 2 years \citep{s4sciencebook, Abazajian2015}.
 
-Excellent conditions for CMB observations can be found in space and at various locations on earth, including the South Pole \citep{kovac2007}.
-Space- and Antarctica-based experiments depend on satellite transmission.
-The South Pole will continue its contribution to CMB science in the next generation of ground-based instruments, which aims to collect 2 million detector years in total.
-During CMB-S4 observations at the South Pole site will face a data influx of $\sim \mathcal{O}(10)\mathrm{Tb/d}$.
+The South Pole is one of the best sites for CMB observations on Earth \citep{kovac2007}.
+The next generation CMB experiment, CMB-S4, plans to include several telescopes at the South Pole.
+These telescopes would generate a data influx of $\sim \mathcal{O}(10)\mathrm{Tb/d}$.
 However, the current transmission allocation for SPT-3G is $150\mathrm{Gb/d}$, which will likely see only a moderate increase.
-The transmission bottleneck is currently overcome by recovering the full data on hard drives with some latency and transmitting a downsampled version of the data.
-The downsampling process loses high frequency information and continued usage of it will narrow the information window decisively, prohibiting high multipole moment science to be carried out on the transmitted data set.
-This also means that any potential faults or errors in the experiment apparent in this range are not visible in the transmitted data.
-%C: Is there a paper detailing trasmission limits for SPT and that the data is currently downsampled?
+The transmission bottleneck could be overcome by recovering the full data on hard drives every summer and transmitting a downsampled version of the data.
+Downsampling eliminates high frequency information, which makes it unsuitable for science on small angular scales, such as SZ galaxy clusters.
+The potential delay (i.e.
+only getting the high frequency data once a year) also introduces risks by delaying when potential faults or issues at high frequency are noticed.
+One can avoid some of these issues by running substantial portions of the analysis at the South Pole, but this comes with its own costs and challenges.
 
-The Planck mission has demonstrated the merits of carrying out CMB observations from space \citep{planck2018}.
+% maybe dont include this
+%The Planck mission has demonstrated the merits of carrying out CMB observations from space \citep{planck2018}.
 Upcoming missions aim to exceed the detector count of Planck by at least an order of magnitude \citep{litebird2014, pixie2011, core2018}.
@@
 
-%Planning
+%Planning+Analysis
 
-Beyond transmission challenges, mission planning is becoming exceedingly difficult.
-Full simulation of TOD over the entire parameter space of detection scenarios is the desired way to decide on CMB-S4 configurations.
-Given the shear size of TOD expected, this is not possible \citep{s4sciencebook}.
-Ground-based observations have to rely on different planning strategies or aim to reduce the size of the TOD in order to maximise the productivity of the design stage.
+Beyond transmission challenges, larger data volumes require larger and more expensive disk arrays and make end-to-end simulations of the experiment for the purpose of optimisation on systematics estimation more time consuming.
+Given the shear size of upcoming data sets full end-to-end simulation may prove impossible \citep{s4sciencebook}.
+Operations on the TOD, such as noise removal and map making, are a vital part of CMB data analysis.
+While we have not experienced the limitations of the accessible computational assets for these processes, the exponential growth of CMB data makes its analysis increasingly expensive.
+Few bit digitisation could ameliorate both of these challenges.
+
+%Ground-based observations have to rely on different planning strategies or aim to reduce the size of the TOD in order to maximise the productivity of the design stage.
 Space-based missions carry out a similar analysis specific to their instruments to heighten their science output.
-
-% Analysis
-
-Operations on the TOD, such as noise-removal and map-making are a vital part of CMB data analysis.
-While we have not experienced the limitations of the accessible computational assets, the exponential growth of CMB data makes its analysis increasingly expensive.
-
-%why is it worth considering
-
-Extreme Digitisation would tackle the challenges mentioned above by reducing TOD volume by an order of magnitude.
-Together with already established lossless compression techniques (e.g.
-FLAC, run-length coding, Huffman coding, etc.) this addresses transmission hurdles.
-Beyond that extreme digitisation has the potential of easing issues in planning, analysis and hardware requirements.
 
@@
 
-While the advantages of using few bit TOD in space-based CMB observations have been demonstrated by the Planck mission \citep{maris2003}, other science areas have also shown the power of this tool.
+Few bit digitisation has already been demonstrated in space by the Planck mission \citep{maris2003}, as well as in other areas of physics.
 \cite{jenet1998} explored the application of such compression to radio pulsar timing measurements with success.
@@
 
-Digitisation is a lossy compression technique.
-However, the additional noise depends on the details of the digitisation scheme and the input signal.
-There is usually little concern around finding an optimal set of digitisation parameters as it is common to have access to large numbers of bits to store information, where changes in the digitisation scheme become insignificant to the distortion induced.
-It is in our interest however to consider extreme, i.e.
-few bit, digitisation.
-In this case the digitisation scheme must be carefully constructed.
-We review the key aspedcts of designing digitisation schemes as edtablished by \cite{max1960} below.
+The optimal digitisation scheme to minimise distortion for a fixed number of bits depends on the details of the input signal.
+In most cases this is ignored, since the distortions become vanishingly small as the number of bits increases.
+However, optimal schemes are critical to the success of extreme (few bit) digitisation.
+We review the key aspects of designing digitisation schemes as established by \cite{max1960} below.
 
-Digitisation discretises an input signal by sorting it into $N$ appropriate ranges, such that an input between $x_i$ and $x_{i+1}$ produces an output at $y_i$.
+Digitisation discretises an input signal by sorting it into $N$ appropriate ranges, such that an input between $x_i$ and $x_{i+1}$ produces an output value $y_i$.
 The set of parameters $N, x_i, y_i$ fully specify a digitisation scheme.
-Conventionally one chooses $x_{1} = -\infty$ and $x_{N+1} = \infty$.
+Conventionally one chooses $x_{1} = -\infty$ and $x_{N+1} = \infty$, i.e.
+values beyond some threshold saturate and yield identical output.
 In order to quantify the performance of a given digitisation scheme we define the distortion as
@@
 
-where $s$ is the input and $\hat{s}$ the output signal.
-For an input signal that has a stochastic element we introduce the input amplitude probability density $p(x)$.
-This allows us to rewrite the above as
+where $s$ is the input and $\hat{s}$ the output signal. For a stochastic input signal we can calculate an amplitude probability distribution $p(x)$:
 
@@
 
-Since we wish to minimise the distortion we differentiate the distortion with respect to $x_i$ and $y_i$ and set the derivatives to zero. We obtain the two equations
+Since we wish to minimise the distortion, we set the derivatives with respect to $x_i$ and $y_i$ to zero. We obtain the two equations
 
@@
 \begin{equation} \label{eq:distderiv2}
-\frac{\partial D}{\partial x_j} = -2 \int_{x_i}^{x_{i+1}} \left( x-y_i \right) p(x) dx = 0.
+\frac{\partial D}{\partial y_i} = -2 \int_{x_i}^{x_{i+1}} \left( x-y_i \right) p(x) dx = 0.
 \end{equation}
 
-Rearranging equation \ref{eq:distderiv1} we deduce
+Rearranging equation \ref{eq:distderiv1}, we find
 
@@
 
-which informs us that an output level $y_i$ must lie halfway between its delimiting thresholds $x_i$ and $x_{i+1}$. From equation \ref{eq:distderiv2} we gain the additional condition
+which informs us that the cutoff levels $x_i$ should lie midway between the output levels. Equation \ref{eq:distderiv2} gives the additional condition
 
@@
 
-To progress further we have to make an assumption about the probability distribution of input signals.
-We take ground-based CMB observations to operate at low signal to noise and assume detector noise to be gaussian white.
-The input is then purely stochastic and  $p(x) = (1/\sqrt{2\pi\sigma^2}) e^{-x^2/2\sigma^2}$, where $\sigma$ is the standard deviation of the detector noise.
-Under these assumptions the system of equations can be solved numerically.
-One begins by picking $y_1$ and calculating the remaining $x_i$'s and $y_i$'s using equation \ref{eq:digitequalspacecondition}.
-Afterwards one checks whether this choice of values satisfy the conditions given by equation \ref{eq:digitareacondition}.
-If that is the case, the $x_i$'s and $y_i$'s were chosen appropriately.
+To progress further, we must consider the probability distribution $p(x)$ of the input signal.
+The simplest case, which is also appropriate for ground-based CMB experiments is the low signal to noise regime.
+If the detector noise is Gaussian we have $p(x) = (1/\sqrt{2\pi\sigma^2}) e^{-x^2/2\sigma^2}$.
 
-We use the findings of \cite{max1960} to formulate the multi-level functions we use for one, two, and three bit digitisation.
-Given an input signal $s(t)$ a digitisation scheme using $N$ bits returns the output $\hat{s}_N(t)$.
-For one bit digitisation we apply the sign function
+The solution for this case was given by \cite{max1960}. For one bit digitisation, the result is
 
@@
 
-to the TOD. For two bit digitisation we apply the four-level function
+For two bit digitisation we have
 
@@
 
-The fact that the single bit digitisation scheme makes no references to the standard deviation of the signal is irrelevant.
-This affects the normalisation of the output, which we can account for during later analysis as demonstrated in the appendix.
-What is important is the spacing between digitisation thresholds and between output levels.
+The digitisation schemes listed above do not conserve signal power and need to be re-calibrated to recover the correct power in the map.
+This is not particularly important, since most CMB experiments determine the absolute calibration from their maps.
+However, it is worth noting that one can rescale the schemes laid out by \cite{max1960} to preserve power.
+This is demonstrated in the appendix.
 
@@
 To investigate the performance of the derived digitisation schemes we simulate many scans over CMB template maps at the timestream level.
-We obtain control maps that use 64 bit TOD and maps that have undergone one, two, and three bit digitisation of the TOD.
+We simulate detector noise and apply the digitisation schemes from above to the constructed I, Q, and U TOD.
+Additionally we retain the 64 bit data to construct control maps.
+The different timestreams are compressed into maps for further analysis.
+
+% We obtain control maps that use 64 bit TOD and maps that have undergone one, two, and three bit digitisation of the TOD.
 We calculate the temperature and polarisation power spectra of each map and determine the noise level.
@@
 
-To create the template maps we use the Hierarchical Equal Area isoLatitude Pixelization of a sphere \citep[HEALPix;][]{healpix}\footnote{Available at \url{http://healpix.sourceforge.net/}} framework.
-We generate a realisation of I, Q and U maps with $\mathrm{NSIDE} = 4096$ based on the 2015 results of the Planck collaboration.
-The key cosmological parameters are $\Omega_b \approx 0.049 $, $\Omega_c \approx 0.265$, $\Omega_m \approx 0.316$ and $h \approx 0.67$ \citep{planck2016}.
+To create the template maps we use the HEALPix \citep{healpix}\footnote{Available at \url{http://healpix.sourceforge.net/}.} framework.
+We generate a realisation of I, Q and U maps with resolution $\mathrm{NSIDE} = 4096$, corresponding to a pixel area of approximately $0.74\mathrm{arcmin^2}$.
+This is suitable for the analysis of Stage 3 and CMB-S4 experiments.
+The input power spectrum is based on the 2015 results of the Planck collaboration.
+The key cosmological parameters are $H_0 = 67.8 \mathrm{km \> s^{-1} Mpc^{-1}}$, $\Omega_{\mathrm{m}} = 0.308 $, $n_{\mathrm{s}} = 0.968$, and $\tau = 0.066$ \citep{planck2016}.
 
@@
 
-We simulate observing a $\sim 600 \mathrm{deg^2}$ patch of the sky with a single detector.
-To do so we perform constant elevation scans (CES), equally spaced in declination (DEC).
+We simulate observing a $600 \mathrm{deg^2}$ patch of the sky with a single detector.
+Constant elevation scans (CES) are performed beginning at right ascension (RA) and declination (DEC) $(0^\circ, 0^\circ)$.
+After covering $24^\circ30'$ along RA the detector is reset to a RA of $0^\circ$ and an offset in DEC corresponding to the pixel size.
+Through repetition of CES with constant steps in DEC the survey patch is covered.
+The entire scan strategy is repeated $100$ times with offsets in the starting RA and DEC up to the size of one pixel, ensuring that pixels are sampled uniformly.
+This scan strategy imitates the SPT, which sweeps across the sky along RA with constant steps in DEC \citep{chown2018}.
+
+The detector is read out at $200\mathrm{Hz}$ and the on sky speed is adjusted between simulation to produce the desired number of hits per map pixel.
+The detector noise level for temperature is $500 \mathrm{\mu K \sqrt{s}}$ and for polarisation observations $\sqrt{2} \hspace{0.7ex} 500 \mathrm{\mu K \sqrt{s}}$, i.e.
+the corresponding photon noise limit.
+
+During a CES the pixels being targeted are determined.
+The corresponding values of the input maps are accessed and added to detector noise realisations of the same length as the number of hits falling into each pixel.
+This is done for I, Q, and U maps, producing three timestreams.
+A 64 bit copy of the TOD is processed separately and the above digitisation schemes are applied.
+The four resulting sets of observation data are compressed into maps by averaging all hits falling into the same pixel.
+This produces 24 CMB maps in total: three control maps and nine maps with three each obtained from the different digitisation schemes.
+Additionally we keep hold of the hitmap.
+
+This simulation is implemented in Python.
+We use parallelisation to navigate memory requirements and reach the depth of observation desired.
+Artificial observations with $800$, $8,000$, $80,000$, $1,024,000$, $10,240,000$ and $102,400,000$ hits per map pixel are carried out.
+
+%To do so we perform constant elevation scans (CES), equally spaced in declination (DEC).
 We repeat the observation strategy $100$ times with a slight offset in right ascension (RA) and DEC each time, such that each pixel is sampled approximately uniformly.
@@
 
-While performing each CES the pixels being targeted are determined.
+%While performing each CES the pixels being targeted are determined.
 The corresponding values from the template maps are accessed and added to detector noise realisations.
@@
 
-Having constructed I,Q and U TOD we apply the digitisation schemes derived above.
+%Having constructed I,Q and U TOD we apply the digitisation schemes derived above.
 We compress the timestreams into maps by averaging all hits falling into the same pixel.
@@
 
-We use the Spatially Inhomogeneous Correlation Estimator for Temperature and Polarisation \citep[PolSpice;][]{polspice} to compute $\mathrm{TT}$, $\mathrm{EE}$ and $\mathrm{BB}$ power spectra.
-To minimise cut sky effects we apodise the observed patch using a cosine mask with $\sigma_{\mathrm{APOD}} = \sqrt{600}/2 \hspace{0.7ex} \mathrm{deg}$.
+We use PolSpice \citep{polspice} to compute $\mathrm{TT}$, $\mathrm{EE}$ and $\mathrm{BB}$ power spectra.
+To minimise boundary effects we apodise the observed patch using a cosine mask with $\sigma_{\mathrm{APOD}} = \sqrt{600}/2 \hspace{0.7ex} \mathrm{deg}$.
 %C: Should I also mention the multiplication with a cosine mask that I supply in addition to setting this option on polspice?
@@
 
-%%%%%%%%%%%
-% THIS MIGHT CHANGE WHEN WORKING OUT THEORETICAL NORMALISATION
-%%%%%%%%%%%
-
-Normalisation of the obtained power spectra is necessary.
-The digitisation schemes introduced in section \S\ref{sec:dig} are designed to minimise the distortion of the TOD, not to preserve power.
-We obtain the appropriate normalisation constants as described in the appendix.
-All reconstructed maps from few bit TOD produce acceptable power spectra.
-As shown by figure \ref{fig:psrecover} temperature and polarisation alike trace the control case well.
-Variation around the input power spectrum exists since we observe a single CMB realisation and because of cut sky effects.
-
-%Normalising obtained spectra against previous experiments is a common calibration technique in many CMB experiments (e.g.
-SPT).
-We again refer to the discussion in the appendix on how to modify digitisation schemes to ensure power preservation.%C: Do you know a good reference for this?
+We find the appropriate normalisation constants for power spectra of few bit TOD as described in the appendix.
+All spectra of extremely digitised data recover the input to a satisfying degree.
+Figure \ref{fig:psrecover} shows this for temperature and polarisation observations.
+Differences between few bit and the input power spectra are due to observing a single CMB realisation and residual boundary effects.
 
@@
 Simulations with $1,024,000$ hits per map pixel were used in this plot.
-No appreciable differences are visible for other depths of observation.
+No qualitative differences are visible for other depths of observation.
 \label{fig:psrecover}
@@
 
-To investigate whether the noise added through digitisation has any $\ell$ dependence we subtract the template maps from the simulated observations.
-Beyond a calibration error, the power spectra of difference maps do not show any $\ell$ dependence in the added noise as seen in figure \ref{fig:diffpswn}.
-The plot also reveals that three bit digitisation performs best, followed by two bit and one bit.
+To quantify the distortion caused by extreme digitisation we infer the map noise level from the relevant power spectra and compare them to the map noise level of control observations. We formulate
+
+\begin{equation} \label{eq:extramapnoise}
+\frac{\Delta \sigma}{\sigma} = \frac{\sigma_{\mathrm{map}}^{\mathrm{D}}-\sigma_{\mathrm{map}}^{\mathrm{C}}}{\sigma_{\mathrm{map}}^{\mathrm{C}}},
+\end{equation}
+
+where $\sigma_{\mathrm{map}}^{\mathrm{C}}$ and $\sigma_{\mathrm{map}}^{\mathrm{D}}$ are the map noise levels of control observations and simulations using few bit TOD respectively.
+In the noise dominated regime
+
+\begin{equation} C_{\ell}^\mathrm{D} = C_\ell^\mathrm{N} + C_\ell^\mathrm{X}, \end{equation}
+
+where $C_{\ell}^\mathrm{D}$, $ C_\ell^\mathrm{N}$, and $C_\ell^\mathrm{X}$ are the power spectra of few bit data, the detector noise level, and the additional noise due to digitisation respectively.
+We progress equation \ref{eq:extramapnoise} to
+
+\begin{equation}\frac{\Delta \sigma}{\sigma} = \sqrt{\frac{C_l^\mathrm{D}}{C_l^{\mathrm{N}}}} - 1 = \sqrt{1 + \frac{C_\ell^\mathrm{X}}{C_\ell^{\mathrm{N}}}} - 1. \end{equation}
+
+Before analysing $C_\ell^\mathrm{X}/C_\ell^\mathrm{N}$ we rebin the power spectra to $\Delta \ell \approx 100$.
+This ensures that points in the noise tail are independent, allowing us to extract an uncertainty for the above quantity.
+
+The deduced additional noise for one, two, and three bit digitisation is shown in table \ref{tab:extranoisewhite} in the appendix.
+We see that 3 bit digitisation performs best, followed by two bit and one bit.
 This is expected, because with each additional bit more information is retained.
+We observe that $\Delta \sigma / \sigma$ is independent of the hits per pixel in the maps.
+This suggests that the additional noise scales in the same way as detector noise does.
+Furthermore, the compression performs equally well for temperature as it does for polarisation observations.
+Lastly, the added noise levels are low.
+When averaged over depth of observation and channel one bit digitisation adds $(24.8\pm 1.1)\%$ to the map noise level, two bit $(6.3\pm0.5)\%$ and three bit $(1.7\pm0.2)\%$.
+This is impressive, given that all schemes considered reduce TOD volume by at least an order of magnitude.
+
+To investigate whether the noise added through digitisation has any angular scale dependence we subtract the input maps from simulated observations.
+The power in the difference of maps is plotted in Figure \ref{fig:diffpswn}.
+There is no sign of any angular scale dependence.
 
@@
   Power spectra of reconstructed minus input maps for one-bit (green), two-bit (orange), three-bit (violet) and the control case (black).
-No $\ell$ dependence is observed for all depths of observation.
+No angular scale dependence is observed for all depths of observation.
 Simulations with $1,024,000$ hits per map pixel are shown.
@@
 \end{figure}
-
-To quantify the distortion caused by extreme digitisation we infer the map noise level from the relevant power spectra and compare them to the map noise level of control observations. We formulate
-
-\begin{equation} \label{eq:extramapnoise}
-\frac{\Delta \sigma}{\sigma} = \frac{\sigma_{\mathrm{map}}^{\mathrm{D}}-\sigma_{\mathrm{map}}^{\mathrm{C}}}{\sigma_{\mathrm{map}}^{\mathrm{C}}},
-\end{equation}
-
-where $\sigma_{\mathrm{map}}^{\mathrm{C}}$ and $\sigma_{\mathrm{map}}^{\mathrm{D}}$ are the map noise levels of control observations and simulations using few bit TOD respectively.
-Because we have seen no $\ell$ dependence in the power spectra of difference maps we presume digitisation to add a constant noise term.
-In the regime
-
-\begin{equation} \frac{C_\ell^{\mathrm{N}}}{C_\ell^{\mathrm{S}}} \geq 10, \end{equation}
-
-where $C_l^\mathrm{S}$ is the input CMB signal and $C_l^\mathrm{N}$ the detector noise level, power spectra are assumed to be noise dominated. Under this condition
-
-\begin{equation} C_{\ell}^\mathrm{D} = C_\ell^\mathrm{N} + C_\ell^\mathrm{X}, \end{equation}
-
-where $C_\ell^\mathrm{D}$ is the total power spectrum originating from a digitised timestream and $C_l^\mathrm{X}$ the additional noise induced through digitisation.
-We progress equation \ref{eq:extramapnoise} to
-
-\begin{equation}\frac{\Delta \sigma}{\sigma} = \sqrt{\frac{C_l^\mathrm{D}}{C_l^{\mathrm{N}}}} - 1 = \sqrt{1 + \frac{C_\ell^\mathrm{X}}{C_\ell^{\mathrm{N}}}} - 1. \end{equation}
-
-Before analysing $C_\ell^\mathrm{X}/C_\ell^\mathrm{N}$ we rebin the power spectra to $\Delta \ell \approx 100$.
-This ensures that points in the noise tail are independent, allowing us to extract an uncertainty for the above quantity.
-
-The deduced additional noise for one, two, and three bit digitisation schemes is shown in table \ref{tab:extranoisewhite} in the appendix.
-We observe that $\Delta \sigma / \sigma$ is independent of the hits per pixel in the maps.
-This suggests that the additional noise scales in the same way as detector noise does.
-Furthermore, the compression performs equally well for temperature as it does for polarisation observations.
-Lastly, the added noise levels are low.
-When averaged over depth of observation and channel one bit digitisation adds $(24.8\pm 1.1)\%$ to the map noise level, two bit $(6.3\pm0.5)\%$ and three bit $(1.7\pm0.2)\%$.
-This is impressive, given that all schemes considered reduce TOD volume by at least an order of magnitude.
-
-%communicate any information where in the first and last digitisation range ( $\left(x_N, \infty \right)$ and $\left( -\infty, x_1 \right)$ )the input signal lies.
-
-%Extreme digitisation is a viable lossy compression technique when dealing with large volume, low signal to noise, slowly varying datasets.
-Making these assumptions the signal is small and constant for a considerable number of datapoints.
-For such an interval many samplings of the signal plus noise result in producing appropriate numbers in each output bin.
-The addition of more bits helps by classifying the size of the noise more accurately and hence approaching an improved estimate for the value of the signal faster.
-It is necessary to remain in the low signal to noise regime, as otherwise a dominant signal would saturate the digitisation scheme and for a given scheme only values in the range $\pm |N_{\mathrm{hits}}y_N|$ can be represented, given that we have $N_{\mathrm{hits}}$ datapoints.
-We must demand the data to be slowly varying such that many datapoints can be used to reconstruct a single significant step of the input signal.
-Finally a large data volume is clearly favourable for the reasons illustrated above: it allows for repeated sampling and therefore more accurate reconstruction of the weak signal.
 
@@
 The first term on the second line produces $1/f$ behaviour and the second term on the same line suppresses the lowest modes.
-For temperature observations we choose $\lknee = 1000$ and $\lcut = 336$, to ensure that $|N_{1/f}(0.3\lknee)|^2 = |0.95 \sigma|^2$.
-For polarisation we set $\lknee = 50$ and $\lcut = 17$ for the same reason.
+For temperature observations we choose $\lknee = 1000$ and $\lcut = 336$, to suppress modes below $0.3 \lknee$.
+We set $\lknee = 50$ and $\lcut = 17$ for polarisation for the same reason.
 
@@
 We carry out the same simulations as described in subsection \ref{subsec:method} using the above noise profile and digitisation schemes for maps with $8,000$, $102,400$ and $1,024,000$ hits per pixel.
-HEALPix is used to calculate the spherical harmonics coefficients $a\ell m$.
-We mask the lowest modes in scan direction and obtain the power spectra.
+We apply a cosine mask to the revocered maps to minimise boundary effects before using HEALPix to calculate the spherical harmonic coefficients $a\ell m$.
+We discount modes with $m < \lknee$.
+The scan strategy of the simulated observations is similar to the SPT's in that smaller values of $m$ correspond to larger angular features along the scan direction \citep{chown2018}.
+Giving equal weight to all modes would lead to inaccurate power spectra due to the anisotropic nature of the noise.
+We mask these modes, which is the first order approximation to optimal weighting.
 The analysis procedure of section \ref{subsec:whitenoise} is used to characterise the additional noise.
@@
 We recover similar results to the white noise case.
-As before three-bit performs the best, followed by two-bit and one-bit.
-The excess noise due to digitisation does not show any $\ell$ dependence and does not vary with the number of hits per pixel in the map.
+As before three bit performs the best, followed by two bit, and one bit.
+The excess noise due to digitisation does not show any angular scale dependence and does not vary with the number of hits per pixel in the map.
 Temperature and polarisation observations incur the same noise penalty.
@@
 
-Extreme digitisation is a viable lossy compression technique when dealing with low signal to noise, large data sets, that have a slowly varying signal with respect to the sampling rate.
-These conditions allow us to reconstruct the underlying signal well.
+% find a different home for this
+%Extreme digitisation is a viable lossy compression technique when dealing with low signal to noise, large data sets.
+A signal that varies slowly with respect to the sampling rate can be reconstructed well.
 A small signal will also prevent saturation of the output, i.e.
@@
 
-In this work we have motivated the investigation of extreme digitisation as a technique in combating potential challenges of CMB experiments.
+In this work we have conducted an investigation of extreme digitisation as a technique in combating potential challenges of CMB experiments.
 The reduction of TOD by an order of magnitude addresses potential issues in transmission from remote locations.
@@
 We derived a set of one, two, and three bit digitisation schemes assuming white detector noise.
-Through simulated observation we determined the distortion this compression technique causes in temperate and polarisation power spectra.
-We find that an optimal three bit digitisation adds as little as $1.7\%$ to the map noise level for temperature and polarisation observations alike.
-No change in the results is observed for maps of different hits per pixel and no $\ell$ dependence was observed in the added noise.
+Through simulated observation we determined the distortion this compression technique causes.
+We find that optimal three bit digitisation adds as little as $1.7\%$ to the map noise level for temperature and polarisation observations alike.
+No change in the results is observed for maps of different hits per pixel and no angular scale dependence was observed in the added noise.
 Extending the model to more realistic $1/f$ detector noise did not change these results.
 
-The additional noise is insignificant for ground-based experiments, which increase their detector count by orders of magnitude between generations.
+% restate from first paragraph
+%The additional noise is insignificant for ground-based experiments, which increase their detector count by orders of magnitude between generations.
 Space-based experiments which have their own unique issues may consider this noise penalty problematic.
@@
 Future work investigating this compression technique must aim to understand the nature of the induced noise better.
-It is of great value to find the higher statistical moments, i.e.
+It is of interest to find the higher statistical moments, i.e.
 skewness and kurtosis, of the added noise term.
@@
 
-As mentioned in subsection \ref{subsec:extremedigitisation} the designed digitisation schemes do not conserve power.
+As mentioned in section \ref{subsec:extremedigitisation} the designed digitisation schemes do not conserve power.
 This is because the distortion metric used does not demand this.
-We can ensure power preservation by multiplying the output by a constant $\mathcal{N}$.
+We can ensure power preservation by rescaling the digitised output by a constant, $\mathcal{N}$.
 In the following we will show that this is true and derive an expression for $\mathcal{N}$.
@@
 The probabilities can then be expressed as integrals over the relevant sections of correctly parameterised probability distributions.
-We proceed by assuming that noise is drawn from normal distributions with mean $\mu$ and standard deviation $\sigma$, i.e.
-$N(\mu, \sigma, \xi)$.
+We proceed by assuming that the noise $\xi$ is drawn from normal distributions with mean $\mu$ and standard deviation $\sigma$, denoted by $N(\mu, \sigma, \xi)$.
 
@@
 
-A digitisation scheme designed according to the distortion metric in section \ref{subsec:extremedigitisation} for normally distributed noise will satisfy $y_i = -y_{N-i}$ and $x_i = -x_{N+1-i}$.
-This simply states symmetry of the digitisation thresholds and levels about zero or after a mean has been subtracted.
+Equations \ref{eq:digitequalspacecondition} and \ref{eq:digitareacondition} in \ref{subsec:extremedigitisation} demand that for a Gaussian input distribution $y_i = -y_{N-i}$ and $x_i = -x_{N+1-i}$.
+This simply states symmetry of the digitisation thresholds and levels about zero (or after a global mean has been subtracted).
 With this information we can condense the above to
@@
 \sum_{i,j = 0}^N y_i y_j \psi_i^{k, n} \psi_j^{l, m} = \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} y_i y_j \left( \psi_i^{k, n} \psi_j^{l, m} - \psi_{N-i}^{k, n} \psi_j^{l, m} - \psi_i^{k, n} \psi_{N-j}^{l, m} + \psi_{N-i}^{k, n} \psi_{N-j}^{l, m} \right)
+\end{equation}
+
+The index shift $i \to N-i$ affects $\psi_{i}^{k, n}$ as follows:
+
+\begin{equation}
+\begin{aligned}
+\psi_{N-i}^{k,n} &= x_{N+1-i}^{2n+1-k} - x_{N-i}^{2n+1-k} = (-x_{i})^{2n+1-k} - (-x_{i+1})^{2n+1-k}\\
+&= (-1)^{2n+1-k} (x_{i})^{2n+1-k} - x_{i+1})^{2n+1-k} ) = (-1)^{2n+2-k} (x_{i+1})^{2n+1-k} - x_{i1})^{2n+1-k} )\\
+&= (-1)^{k} \psi_{i}^{k,n}.
+\end{aligned}
+\end{equation}
+
+Therefore equation \ref{eq:ijsum} may be rewritten as
+
+\begin{equation} \label{eq:ijsumpsiflip}
+\sum_{i,j = 0}^N y_i y_j \psi_i^{k, n} \psi_j^{l, m} = \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} y_i y_j \left( \psi_i^{k, n} \psi_j^{l, m} + (-1)^{k+1} \psi_{i}^{k, n} \psi_j^{l, m} + (-1)^{l+1} \psi_i^{k, n} \psi_{j}^{l, m} + (-1)^{k+l} \psi_{N-i}^{k, n} \psi_{N-j}^{l, m} \right).
 \end{equation}
@@
 
-The first two terms in the above will lead to even powers of $\mu$ whereas the last two will lead to odd powers of $\mu$.
-Before we can look at these terms in more detail we need to see how the index shift $i \to N-i$ affects $\psi_{i}^{k, n}$.
-
-\begin{equation}
-\begin{aligned}
-\psi_{N-i}^{k,n} &= x_{N+1-i}^{2n+1-k} - x_{N-i}^{2n+1-k} = (-x_{i})^{2n+1-k} - (-x_{i+1})^{2n+1-k}\\
-&= (-1)^{2n+1-k} (x_{i})^{2n+1-k} - x_{i+1})^{2n+1-k} ) = (-1)^{2n+2-k} (x_{i+1})^{2n+1-k} - x_{i1})^{2n+1-k} )\\
-&= (-1)^{k} \psi_{i}^{k,n}.
-\end{aligned}
-\end{equation}
-
-We are now ready to put all pieces together.
-We start by investigating even powers of $\mu$ in equation \ref{eq:dcompleteklsum} that arise from even $k$ and $l$.
-Under the assumptions made \ref{eq:ijsum} reduces to
+The first two terms in the above will lead to even powers of $\mu$, whereas the last two will lead to odd powers of $\mu$.
+We are now ready to put all pieces together and start by investigating even powers of $\mu$ in equation \ref{eq:dcompleteklsum} that arise from even $k$ and $l$.
+Under the assumptions made \ref{eq:ijsumpsiflip} reduces to
 
@@
 
+% this is handy in practice! do not have to account for different detector sensitivities within detector array and T/pol.
+
 \newpage

### 2018-09-20 digitisation.tex
editing

@@
 We apply a cosine mask to the revocered maps to minimise boundary effects before using HEALPix to calculate the spherical harmonic coefficients $a\ell m$.
-We discount modes with $m < \lknee$.
+We discount modes with $m < \lknee$ before calculating the power spectra.
 The scan strategy of the simulated observations is similar to the SPT's in that smaller values of $m$ correspond to larger angular features along the scan direction \citep{chown2018}.
@@
 The first two terms in the above will lead to even powers of $\mu$, whereas the last two will lead to odd powers of $\mu$.
-We are now ready to put all pieces together and start by investigating even powers of $\mu$ in equation \ref{eq:dcompleteklsum} that arise from even $k$ and $l$.
+We are now ready to put all the pieces together and start by investigating even powers of $\mu$ in equation \ref{eq:dcompleteklsum} that arise from even $k$ and $l$.
 Under the assumptions made \ref{eq:ijsumpsiflip} reduces to

### 2018-09-21 digitisation.tex
removed newpage
Implemented 2nd round of comments
(have not updated plots yet)

--- 2018-09-20 digitisation.tex
@@
 \newcommand{\lknee}{\ensuremath{\ell_{\rm knee}}}
-\newcommand{\lcut}{\ensuremath{\ell_{\rm cut}}}
+\newcommand{\lcut}{\ensuremath{\ell_{\rm min}}}
 \newcommand{\fermilat}{\textit{Fermi}-LAT}
@@
 
-\title{Cosmic Microwave Background Power Spectra from Few Bit Timestreams}
+\title{Extreme Digitisation For Ground-Based Cosmic Microwave Background Experiments}
 \author{L.~Balkenhol\altaffilmark{\Melbourne} and C.~L.~Reichardt\altaffilmark{\Melbourne}}
@@
 Unlike lossless compression, few bit digitisation introduces additional noise into the data.
-We present a set of one, two, and three bit digitisation schemes and measure the noise increase due to compression in the CMB temperature and polarisation power spectra.
-The added noise is independent of angular scale and is well-described as a percentage increase on the original detector noise.
+We present a set of one, two, and three bit digitisation schemes and measure increase in noise in the CMB temperature and polarisation power spectra.
+The digitisation noise is independent of angular scale and is well-described as a constant percentage of the original detector noise.
 Three bit digitisation increases the map noise level by $< 2\%$, while reducing the data volume by a factor of ten relative to 32 bit floats.
@@
 Observations of the cosmic microwave background (CMB) have played a key role in cosmology \citep{penzias1965, smoot1992, bennett2013, sptpol2013, planck2018}.
-Current CMB Experiments target science goals such as the discovery of inflationary gravitational waves, construction of precise lensing maps and thorough studies of the Sunyaev-Zeldovich (SZ) effects.
-Additionally, the CMB probes structure formation, reionisation and is a powerful test for dark matter and dark energy models.
-Lastly, the relativistic number of species and the neutrino mass sum can be measured by CMB experiments \citep{s4sciencebook, core2011, litebird2016, pixie2011}.
+They can measure the relativistic number of species and the neutrino mass sum.
+Current CMB Experiments target science goals such as the discovery of inflationary gravitational waves, measuring the relativistic number of species and the neutrino mass sum, and mapping the large-scale distribution of matter through gravitational lensing and the Sunyaev-Zeldovich (SZ) effects.
+Through lensing and the SZ effects the CMB probes structure formation, reionisation, and is a powerful test for dark matter and dark energy models \citep{s4sciencebook, core2011, litebird2016, pixie2011}.
 
+% CITES FOR PARTICLE PHYSICS WITH CMB, ALSO MAYBE SHUFFLE POSITION OF PARTICLE PHYSICS STUFF, KINDA AWKWARD
+
+Over the last two decades CMB experiments have gone from single detectors to over $10,000$ detectors.
 The CMB community has developed a variety of compression techniques and computational approaches to handle the increasing volume of data \citep{tristam2007}.
@@
 Space-based experiments have employed a combination of lossless and lossy compression techniques, including reduced bits in the TOD \citep{gaztanaga1998, maris2003}.
-Antarctica-based experiments that transmit a portion of their data via a satellite link have downsampled their data in the past to meet telemetry allocations and have not yet used few bit digitisation of the TOD.
+Antarctica-based experiments that transmit a portion of their results via a satellite link downsample their data to meet telemetry allocations, but have not yet used few bit digitisation of the TOD.
 As we approach the next generation of ground-based experiments, CMB-S4, and the launch of new space-based missions (LiteBIRD, PIXIE, COrE+), we must consider potential transmission bottlenecks carefully.
@@
 The science goals of upcoming CMB experiments depend on achieving substantially faster mapping speeds.
-Given CMB detectors are generally photon noise limited, the number of detectors (and data volume) of ground-based experiments has followed a trend like Moore's law, doubling approximately every 2 years \citep{s4sciencebook, Abazajian2015}.
+Given CMB detectors are generally photon noise limited, improving an experiments means adding more detectors.
+As a result the number of detectors (and data volume) of ground-based experiments has followed a trend like Moore's law, doubling approximately every 2 years \citep{s4sciencebook, Abazajian2015}.
 
@@
 The next generation CMB experiment, CMB-S4, plans to include several telescopes at the South Pole.
-These telescopes would generate a data influx of $\sim \mathcal{O}(10)\mathrm{Tb/d}$.
+These telescopes would generate a data influx of $\sim\mathcal{O}(10)\mathrm{Tb/d}$.
 However, the current transmission allocation for SPT-3G is $150\mathrm{Gb/d}$, which will likely see only a moderate increase.
@@
 Beyond transmission challenges, larger data volumes require larger and more expensive disk arrays and make end-to-end simulations of the experiment for the purpose of optimisation on systematics estimation more time consuming.
-Given the shear size of upcoming data sets full end-to-end simulation may prove impossible \citep{s4sciencebook}.
+Given the shear size of upcoming data sets, full end-to-end simulation may prove impossible \citep{s4sciencebook}.
 Operations on the TOD, such as noise removal and map making, are a vital part of CMB data analysis.
-While we have not experienced the limitations of the accessible computational assets for these processes, the exponential growth of CMB data makes its analysis increasingly expensive.
+The exponential growth of CMB data makes these steps increasingly expensive.
 Few bit digitisation could ameliorate both of these challenges.
@@
 The optimal digitisation scheme to minimise distortion for a fixed number of bits depends on the details of the input signal.
-In most cases this is ignored, since the distortions become vanishingly small as the number of bits increases.
+In most cases this optimisation is neglected, since the distortions become vanishingly small as the number of bits increases.
 However, optimal schemes are critical to the success of extreme (few bit) digitisation.
@@
 
-\begin{equation} D = \left\langle  \left( s - \hat{s} \right)^2 \right\rangle, \end{equation}
+\begin{equation}\label{eq:distdef}
+D = \left\langle  \left( s - \hat{s} \right)^2 \right\rangle,
+\end{equation}
 
@@
 
-\begin{equation} D = \sum_{i = 1}^N \int_{x_i}^{x_{i+1}} \left(x-y_i\right)^2 p(x) dx. \end{equation}
+\begin{equation} \label{eq:dist}
+D = \sum_{i = 1}^N \int_{x_i}^{x_{i+1}} \left(x-y_i\right)^2 p(x) dx.
+\end{equation}
 
-Since we wish to minimise the distortion, we set the derivatives with respect to $x_i$ and $y_i$ to zero. We obtain the two equations
+Since we wish to minimise the distortion, we set the derivatives with respect to $x_i$ and $y_i$ to zero. Starting with $x_i$ we obtain
 
 \begin{equation} \label{eq:distderiv1}
-\frac{\partial D}{\partial x_i} = \left(x_i-y_{i-1}\right)^2 p(x_i) - \left(x_i - y_i\right)^2 p(x_i) = 0,
+\frac{\partial D}{\partial x_i} = \left(x_i-y_{i-1}\right)^2 p(x_i) - \left(x_i - y_i\right)^2 p(x_i) = 0.
 \end{equation}
-
-\begin{equation} \label{eq:distderiv2}
-\frac{\partial D}{\partial y_i} = -2 \int_{x_i}^{x_{i+1}} \left( x-y_i \right) p(x) dx = 0.
-\end{equation}
-
-Rearranging equation \ref{eq:distderiv1}, we find
 
@@
 
-which informs us that the cutoff levels $x_i$ should lie midway between the output levels. Equation \ref{eq:distderiv2} gives the additional condition
+which informs us that the cutoff levels $x_i$ should lie midway between the output levels.
+Setting the derivative of equation \ref{eq:dist} with respect to $y_i$  to zero gives the additional condition
+
+\begin{equation} \label{eq:distderiv2}
+\frac{\partial D}{\partial y_i} = -2 \int_{x_i}^{x_{i+1}} \left( x-y_i \right) p(x) dx = 0.
+\end{equation}
 
@@
 To progress further, we must consider the probability distribution $p(x)$ of the input signal.
-The simplest case, which is also appropriate for ground-based CMB experiments is the low signal to noise regime.
-If the detector noise is Gaussian we have $p(x) = (1/\sqrt{2\pi\sigma^2}) e^{-x^2/2\sigma^2}$.
-
-The solution for this case was given by \cite{max1960}. For one bit digitisation, the result is
-
+For Gaussian detector noise, we have $p(x) = (1/\sqrt{2\pi\sigma^2}) e^{-x^2/2\sigma^2}$.
+We also assume that we are in the low signal to noise regime, as is appropriate for ground-based CMB experiments.
+The solution for this case was given by \cite{max1960}.
+For one bit digitisation, the result is
 \begin{equation} \label{eq:1bit}
 \hat{s}_1(t) = \left\{ \begin{array}{lr}
-1, & \text{for } s(t) > 0\\
--1, & \text{for } s(t) \leq 0
+1, & s(t) > 0\\
+-1, & s(t) \leq 0
 \end{array} \right. \end{equation}
@@
 For two bit digitisation we have
-
 \begin{equation}  \label{eq:2bit}
 \hat{s}_2(t) = \left\{ \begin{array}{rl}
-1.51 \sigma, & \text{for } s(t) \geq 0.9816 \sigma\\
-0.4528 \sigma, & \text{for } 0 \leq s(t) < 0.9816 \sigma\\
--0.4528 \sigma, & \text{for } -0.9816 \sigma \leq s(t) < 0\\
--1.51 \sigma, & \text{for } -0.9816 \sigma < s(t)\\
+1.51 \sigma, & s(t) \geq 0.9816 \sigma\\
+0.4528 \sigma, & 0 \leq s(t) < 0.9816 \sigma\\
+-0.4528 \sigma, & -0.9816 \sigma \leq s(t) < 0\\
+-1.51 \sigma, & -0.9816 \sigma < s(t)\\
 \end{array} \right. . \end{equation}
@@
 Finally the optimal three bit digitisation is described by the eight-level function
-
 \begin{equation}  \label{eq:3bit}
 \hat{s}_3(t) = \left\{ \begin{array}{rl}
-2.152 \sigma, & \text{for } s(t) \geq 1.748 \sigma\\
-1.344 \sigma, & \text{for } 1.05 \sigma \leq s(t) < 1.748 \sigma\\
-0.756 \sigma, & \text{for } 0.501 \sigma \leq s(t) < 1.05 \sigma\\
-0.245 \sigma, & \text{for } 0 \leq s(t) < 0.501 \sigma\\
--0.245 \sigma, & \text{for } -0.501 \sigma \leq s(t) < 0\\
--0.756 \sigma, & \text{for } -1.05 \sigma \leq s(t) < -0.501 \sigma\\
--1.344 \sigma, & \text{for } -1.748 \sigma \leq s(t) < -1.05 \sigma\\
--2.152 \sigma, & \text{for } -1.748 \sigma < s(t)\\
+2.152 \sigma, & s(t) \geq 1.748 \sigma\\
+1.344 \sigma, & 1.05 \sigma \leq s(t) < 1.748 \sigma\\
+0.756 \sigma, & 0.501 \sigma \leq s(t) < 1.05 \sigma\\
+0.245 \sigma, & 0 \leq s(t) < 0.501 \sigma\\
+-0.245 \sigma, & -0.501 \sigma \leq s(t) < 0\\
+-0.756 \sigma, & -1.05 \sigma \leq s(t) < -0.501 \sigma\\
+-1.344 \sigma, & -1.748 \sigma \leq s(t) < -1.05 \sigma\\
+-2.152 \sigma, & -1.748 \sigma < s(t)\\
 \end{array} \right. . \end{equation}
@@
 This is not particularly important, since most CMB experiments determine the absolute calibration from their maps.
-However, it is worth noting that one can rescale the schemes laid out by \cite{max1960} to preserve power.
-This is demonstrated in the appendix.
+However, it is worth noting that one can rescale the schemes laid out by \cite{max1960} to preserve power, which is demonstrated in the appendix.
 
@@
 
-To investigate the performance of the derived digitisation schemes we simulate many scans over CMB template maps at the timestream level.
-We simulate detector noise and apply the digitisation schemes from above to the constructed I, Q, and U TOD.
-Additionally we retain the 64 bit data to construct control maps.
-The different timestreams are compressed into maps for further analysis.
+To investigate the performance of the derived digitisation schemes, we simulate many scans over CMB template maps at the timestream level.
+We add realisations of detector noise and apply the digitisation schemes above to the constructed I, Q, and U TOD.
+Additionally, we retain the original 64 bit TOD to construct control maps.
+The different timestreams are binned into maps for a T, E, and B power spectrum analysis.
 
@@
 
-To create the template maps we use the HEALPix \citep{healpix}\footnote{Available at \url{http://healpix.sourceforge.net/}.} framework.
-We generate a realisation of I, Q and U maps with resolution $\mathrm{NSIDE} = 4096$, corresponding to a pixel area of approximately $0.74\mathrm{arcmin^2}$.
+To create the template maps we use the HEALPix \citep{healpix}\footnote{Available at \url{http://healpix.sourceforge.net/}.} framework through the healpy package.
+We generate a realisation of I, Q and U maps with resolution $\mathrm{NSIDE} = 4096$, corresponding to a pixel area of approximately $0.74 \> \mathrm{arcmin^2}$.
 This is suitable for the analysis of Stage 3 and CMB-S4 experiments.
@@
 
-We simulate observing a $600 \mathrm{deg^2}$ patch of the sky with a single detector.
+We simulate observing a $600 \> \mathrm{ deg^2}$ patch of the sky with a single detector.
 Constant elevation scans (CES) are performed beginning at right ascension (RA) and declination (DEC) $(0^\circ, 0^\circ)$.
+The detector noise level for temperature is $500 \mathrm{\mu K \sqrt{s}}$ and for polarisation observations $\sqrt{2} \hspace{0.7ex} 500 \mathrm{\mu K \sqrt{s}}$, i.e.
+the corresponding photon noise limit.
 After covering $24^\circ30'$ along RA the detector is reset to a RA of $0^\circ$ and an offset in DEC corresponding to the pixel size.
@@
 The entire scan strategy is repeated $100$ times with offsets in the starting RA and DEC up to the size of one pixel, ensuring that pixels are sampled uniformly.
-This scan strategy imitates the SPT, which sweeps across the sky along RA with constant steps in DEC \citep{chown2018}.
+This scan strategy imitates the SPT, which sweeps across the sky along RA with constant steps in DEC \citep{schaffer2011}.
 
-The detector is read out at $200\mathrm{Hz}$ and the on sky speed is adjusted between simulation to produce the desired number of hits per map pixel.
+We create naive I,Q, and U maps from the simulated I, Q, and U TOD, i.e.
+the value of a map pixel is the average of all TOD samples lying within that pixel.
+We create four maps of the original 64 bit TOD and one for each of the digitisation schemes.
+We save the ouput maps at $800$, $8,000$, $80,000$, $1,024,000$, $10,240,000$ and $102,400,000$ hits per pixel.
+
+%The detector is read out at $200\mathrm{Hz}$ and the on sky speed is adjusted between simulation to produce the desired number of hits per map pixel.
 The detector noise level for temperature is $500 \mathrm{\mu K \sqrt{s}}$ and for polarisation observations $\sqrt{2} \hspace{0.7ex} 500 \mathrm{\mu K \sqrt{s}}$, i.e.
@@
 
-During a CES the pixels being targeted are determined.
+%During a CES the pixels being targeted are determined.
 The corresponding values of the input maps are accessed and added to detector noise realisations of the same length as the number of hits falling into each pixel.
@@
 
-This simulation is implemented in Python.
+%This simulation is implemented in Python.
 We use parallelisation to navigate memory requirements and reach the depth of observation desired.
@@
 
-We use PolSpice \citep{polspice} to compute $\mathrm{TT}$, $\mathrm{EE}$ and $\mathrm{BB}$ power spectra.
+We use PolSpice \citep{polspice} to compute $\mathrm{TT}$, $\mathrm{EE}$ and $\mathrm{BB}$ power spectra from the I, Q, and U maps above.
 To minimise boundary effects we apodise the observed patch using a cosine mask with $\sigma_{\mathrm{APOD}} = \sqrt{600}/2 \hspace{0.7ex} \mathrm{deg}$.
@@
 We find the appropriate normalisation constants for power spectra of few bit TOD as described in the appendix.
-All spectra of extremely digitised data recover the input to a satisfying degree.
-Figure \ref{fig:psrecover} shows this for temperature and polarisation observations.
+As seen in Figure \ref{fig:psrecover} all spectra of extremely digitised data recover the input to a satisfying degree.
 Differences between few bit and the input power spectra are due to observing a single CMB realisation and residual boundary effects.
@@
   Input power spectra used to generate the CMB template maps plus detector noise level (black) and power spectra using 3 bit TOD (violet).
-Simulations with $1,024,000$ hits per map pixel were used in this plot.
+A simulations with $1,024,000$ hits per map pixel is shown in this plot.
 No qualitative differences are visible for other depths of observation.
@@
 
-To quantify the distortion caused by extreme digitisation we infer the map noise level from the relevant power spectra and compare them to the map noise level of control observations. We formulate
+To quantify the distortion caused by extreme digitisation we infer the fractional increase of the original map noise level, $\Delta \sigma / \sigma$.
+We calculate this from the noise dominated angular scales in the power spectra via
 
-\begin{equation} \label{eq:extramapnoise}
-\frac{\Delta \sigma}{\sigma} = \frac{\sigma_{\mathrm{map}}^{\mathrm{D}}-\sigma_{\mathrm{map}}^{\mathrm{C}}}{\sigma_{\mathrm{map}}^{\mathrm{C}}},
+\begin{equation}
+\frac{\Delta \sigma}{\sigma} = \sqrt{1 + \frac{C_\ell^\mathrm{X}}{C_\ell^{\mathrm{N}}}} - 1,
 \end{equation}
 
-where $\sigma_{\mathrm{map}}^{\mathrm{C}}$ and $\sigma_{\mathrm{map}}^{\mathrm{D}}$ are the map noise levels of control observations and simulations using few bit TOD respectively.
+where $C_\ell^{\mathrm{N}}$ and $C_\ell^\mathrm{X}$ are the detector noise level and the additional digitisation noise respectively.
+Before calculating $C_\ell^\mathrm{X}/C_\ell^\mathrm{N}$ we rebin the power spectra to $\Delta \ell \approx 100$.
+This ensures that points in the noise tail are independent, allowing us to extract an uncertainty for $\Delta \sigma / \sigma$.
+
+%We formulate
+
+%\begin{equation} \label{eq:extramapnoise}
+%\frac{\Delta \sigma}{\sigma} = \frac{\sigma_{\mathrm{map}}^{\mathrm{D}}-\sigma_{\mathrm{map}}^{\mathrm{C}}}{\sigma_{\mathrm{map}}^{\mathrm{C}}},
+%\end{equation}
+
+%where $\sigma_{\mathrm{map}}^{\mathrm{C}}$ and $\sigma_{\mathrm{map}}^{\mathrm{D}}$ are the map noise levels of control observations and simulations using few bit TOD respectively.
 In the noise dominated regime
 
-\begin{equation} C_{\ell}^\mathrm{D} = C_\ell^\mathrm{N} + C_\ell^\mathrm{X}, \end{equation}
+%\begin{equation} C_{\ell}^\mathrm{D} = C_\ell^\mathrm{N} + C_\ell^\mathrm{X}, \end{equation}
 
-where $C_{\ell}^\mathrm{D}$, $ C_\ell^\mathrm{N}$, and $C_\ell^\mathrm{X}$ are the power spectra of few bit data, the detector noise level, and the additional noise due to digitisation respectively.
+%where $C_{\ell}^\mathrm{D}$, $ C_\ell^\mathrm{N}$, and $C_\ell^\mathrm{X}$ are the power spectra of few bit data, the detector noise level, and the additional noise due to digitisation respectively.
 We progress equation \ref{eq:extramapnoise} to
 
-\begin{equation}\frac{\Delta \sigma}{\sigma} = \sqrt{\frac{C_l^\mathrm{D}}{C_l^{\mathrm{N}}}} - 1 = \sqrt{1 + \frac{C_\ell^\mathrm{X}}{C_\ell^{\mathrm{N}}}} - 1. \end{equation}
-
-Before analysing $C_\ell^\mathrm{X}/C_\ell^\mathrm{N}$ we rebin the power spectra to $\Delta \ell \approx 100$.
-This ensures that points in the noise tail are independent, allowing us to extract an uncertainty for the above quantity.
+%\begin{equation}\frac{\Delta \sigma}{\sigma} = \sqrt{\frac{C_l^\mathrm{D}}{C_l^{\mathrm{N}}}} - 1 = \sqrt{1 + \frac{C_\ell^\mathrm{X}}{C_\ell^{\mathrm{N}}}} - 1. \end{equation}
 
@@
 We see that 3 bit digitisation performs best, followed by two bit and one bit.
-This is expected, because with each additional bit more information is retained.
-We observe that $\Delta \sigma / \sigma$ is independent of the hits per pixel in the maps.
-This suggests that the additional noise scales in the same way as detector noise does.
-Furthermore, the compression performs equally well for temperature as it does for polarisation observations.
-Lastly, the added noise levels are low.
-When averaged over depth of observation and channel one bit digitisation adds $(24.8\pm 1.1)\%$ to the map noise level, two bit $(6.3\pm0.5)\%$ and three bit $(1.7\pm0.2)\%$.
+This is expected, since more bits allow a more faithful representation of the input signal.
+We observe that the fractional increase, $\Delta \sigma / \sigma$, is independent of the hits per pixel in the maps and that the compression performs as well for polarisation as is does for temperature spectra.
+Finally, it is striking how little noise is added.
+On average one bit digitisation leads to a $(24.8\pm 1.1)\%$ increase in the map noise level, two bit yields a $(6.3\pm0.5)\%$ increase and three bit adds only $(1.7\pm0.2)\%$.
 This is impressive, given that all schemes considered reduce TOD volume by at least an order of magnitude.
@@
 To investigate whether the noise added through digitisation has any angular scale dependence we subtract the input maps from simulated observations.
-The power in the difference of maps is plotted in Figure \ref{fig:diffpswn}.
+The power in the difference maps is plotted in Figure \ref{fig:diffpswn}.
 There is no sign of any angular scale dependence.
@@
   \caption[Current ]{
-  Power spectra of reconstructed minus input maps for one-bit (green), two-bit (orange), three-bit (violet) and the control case (black).
+  Power spectra of reconstructed minus input maps for one-bit (green), two-bit (orange), three-bit (violet) and the control case (black) rebined to $\Delta \ell = 100$.
 No angular scale dependence is observed for all depths of observation.
-Simulations with $1,024,000$ hits per map pixel are shown.
+A simulation with $1,024,000$ hits per map pixel is shown.
 \label{fig:diffpswn}
@@
 
-We follow up on the compelling results of the white noise case by implementing a more realistic noise profile.
-Ground-based CMB observations have $1/f$ noise spectra due to atmospheric effects \citep{lay2000}.
-We use the Fourier space noise profile
+While digitisation clearly works extremely well for white noise, real ground-based CMB experiments face several sources of low-frequency noise.
+High on the list is atmospheric noise, which may me described by \citep{lay2000}:
 
@@
 \begin{aligned}
-\left| N_{1/f}(\ell) \right|^2 &= \left| N\left(0, \sigma \right) \right|^2\\
+\left| N_{1/f}(\ell) \right|^2 &= \left| \xi \left(0, \sigma \right) \right|^2\\
 &\times \left[ 1 + \left( \frac{\ell}{\lknee} \right)^{-3/2} \right]^2  exp\left[ -\left( \frac{\lcut}{\ell} \right)^{12} \right],
@@
 
-where $N\left(0, \sigma \right)$ denotes a normally distributed random variable with mean $0$ and standard deviation $\sigma$.
-The first term on the second line produces $1/f$ behaviour and the second term on the same line suppresses the lowest modes.
-For temperature observations we choose $\lknee = 1000$ and $\lcut = 336$, to suppress modes below $0.3 \lknee$.
-We set $\lknee = 50$ and $\lcut = 17$ for polarisation for the same reason.
+where $\xi \left(0, \sigma \right)$ denotes a normally distributed random variable with mean $0$ and standard deviation $\sigma$.
+The first term on the second line amplifies modes corresponding to large angular scales and low frequencies \citep{tegmarkdesignersguide1997}.
+The second term on the same line suppresses the lowest modes.
+For temperature we choose $\lknee = 1000$ and $\lcut = 336$ and for polarisation $\lknee = 50$ and $\lcut = 17$.
 
@@
 We carry out the same simulations as described in subsection \ref{subsec:method} using the above noise profile and digitisation schemes for maps with $8,000$, $102,400$ and $1,024,000$ hits per pixel.
-We apply a cosine mask to the revocered maps to minimise boundary effects before using HEALPix to calculate the spherical harmonic coefficients $a\ell m$.
-We discount modes with $m < \lknee$.
-The scan strategy of the simulated observations is similar to the SPT's in that smaller values of $m$ correspond to larger angular features along the scan direction \citep{chown2018}.
-Giving equal weight to all modes would lead to inaccurate power spectra due to the anisotropic nature of the noise.
-We mask these modes, which is the first order approximation to optimal weighting.
-The analysis procedure of section \ref{subsec:whitenoise} is used to characterise the additional noise.
+We use HEALPix to calculate the spherical harmonic coefficients $a\ell m$ and eliminate modes with $m < \lknee$ before calculating the power spectra.
+Because the scan strategy of the simulated observations is similar to the SPT's, smaller values of $m$ correspond to larger angular features along the scan direction \citep{chown2018}.
+Because of the anisotropic nature of the noise, equally weighting all modes leads to inaccurate power spectra.
+Discounting the lowest modes in scan direction is the first order approximation to optimal weighting.
+The analysis procedure of subsection \S\ref{subsec:whitenoise} is used to characterise the digitisation noise.
 
-We recover similar results to the white noise case.
-As before three bit performs the best, followed by two bit, and one bit.
-The excess noise due to digitisation does not show any angular scale dependence and does not vary with the number of hits per pixel in the map.
+We recover similar results to the white noise case, with three bit digitisation performing the best, followed by two bit, and one bit.
 Temperature and polarisation observations incur the same noise penalty.
-Table \ref{tab:extranoiseoof} in the appendix displays the values of $\Delta \sigma / \sigma$ calculated.
-Averaging over depth of observation and channel we calculate that one bit digitisation adds $(27.7\pm 1.5)\%$ to the map noise level, two bit $(6.2\pm0.5)\%$ and three bit $(1.7\pm0.2)\%$.
-No significant deviations between white and $1/f$ noise results are found.
+Table \ref{tab:extranoiseoof} in the appendix displays the fractional increase to the detector noise level due to digitisation, $\Delta \sigma / \sigma$.
+On average one bit digitisation leads to a $(27.7\pm 1.5)\%$ increase in the map noise level, two bit adds $(6.2\pm0.5)\%$, and three bit leads to an increase of only $(1.7\pm0.2)\%$.
+The digitisation noise shows no angular scale dependence.
 
@@
 
-In this work we have conducted an investigation of extreme digitisation as a technique in combating potential challenges of CMB experiments.
-The reduction of TOD by an order of magnitude addresses potential issues in transmission from remote locations.
-Benefits in mission planning, data analysis and hardware requirements are possible.
+In this work we have conducted an investigation of extreme digitisation as a technique in combating the challenges of large data volumes for ground-based CMB experiments.
+In particular the reduction of TOD volume by an order of magnitude decreases the transmission requirements from remote locations.
+The reduction may also streamline TOD simulations.
 
-We derived a set of one, two, and three bit digitisation schemes assuming white detector noise.
-Through simulated observation we determined the distortion this compression technique causes.
-We find that optimal three bit digitisation adds as little as $1.7\%$ to the map noise level for temperature and polarisation observations alike.
+We present a set of one, two, and three bit digitisation schemes.
+In both, the white noise and $1/f$ noise cases, we find that optimal three bit digitisation adds as little as $1.7\%$ to the map noise level for temperature and polarisation observations alike.
 No change in the results is observed for maps of different hits per pixel and no angular scale dependence was observed in the added noise.
-Extending the model to more realistic $1/f$ detector noise did not change these results.
 
@@
 
-Future work investigating this compression technique must aim to understand the nature of the induced noise better.
-It is of interest to find the higher statistical moments, i.e.
-skewness and kurtosis, of the added noise term.
-For this an analysis of the performance of cluster-finding algorithms on the digitised TOD is recommended.
+Future work should focus on finding the higher statistical moments of the digitisation noise.
+Using the digitised TOD for cluster finding can help determine the skewness and kurtosis of the additional noise.
 
@@
 \hline
-\multirow{3}{*}{$1,024,000$}  & TT  & $ 0.25 \pm 0.009 $  & $ 0.063 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
+\multirow{3}{*}{$1,024,000$}  & TT  & $ 0.250 \pm 0.009 $  & $ 0.063 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
 & EE  & $ 0.252 \pm 0.008 $  & $ 0.065 \pm 0.003 $  & $ 0.018 \pm 0.002 $ \\
@@
 \tablecomments{ 
-Percent addition to the map noise level $\Delta \sigma / \sigma$ due to one, two, and three bit digitisation for white detector noise.
+Percent addition to the map noise level, $\Delta \sigma / \sigma$, due to one, two, and three bit digitisation for white detector noise.
 We notice that there is no appreciable variation with depth of observations and channel for a given digitisation scheme.
@@
 \hline
-\multirow{3}{*}{$102,400$}  & TT  & $ 0.235 \pm 0.014 $  & $ 0.06 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
+\multirow{3}{*}{$102,400$}  & TT  & $ 0.235 \pm 0.014 $  & $ 0.060 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
 & EE  & $ 0.299 \pm 0.009 $  & $ 0.064 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
@@
 \tablecomments{ 
-Percent addition to the map noise level $\Delta \sigma / \sigma$ due to one, two, and three bit digitisation for $1/f$ detector noise.
+Percent addition to the map noise level, $\Delta \sigma / \sigma$, due to one, two, and three bit digitisation for $1/f$ detector noise.
 As in the white noise case we observe no significant trend with the numbers of hits per map pixel or channel for a given digitisation scheme.
-No significant deviation in the values compared to simulations assuming white detector noise exists.
+No significant deviation in the values compared to simulations assuming white detector noise is found.
 } \normalsize
@@
 
-As mentioned in section \ref{subsec:extremedigitisation} the designed digitisation schemes do not conserve power.
-This is because the distortion metric used does not demand this.
-We can ensure power preservation by rescaling the digitised output by a constant, $\mathcal{N}$.
-In the following we will show that this is true and derive an expression for $\mathcal{N}$.
-
-We demand equal cross power in input and output, i.e.
+As mentioned in subsection \S\ref{subsec:extremedigitisation}, equation \ref{eq:distdef} did not require signal power to be conserved when constructing a digitisation scheme.
+It is relatively simple to add this requirement however, since it boils down to rescaling the digitised output by a constant, $\gamma$.
+Specifically we will derive an expression for $\gamma$, such that a cross-spectrum of two digitised maps yields an unbiased estimate of the input power spectrum.
+Mathematically, this is equivalent to demanding:
 
 \begin{equation} \label{eq:normcrosspower}
-\langle s_1 s_2 \rangle = \mathcal{N} \langle D(s_1) D(s_2) \rangle,
+\langle (\mu + \xi_1) (\mu + \xi_2) \rangle = \gamma^2 \langle D(\mu + \xi_1) D(\mu + \xi_2) \rangle,
 \end{equation}
 
-where timestreams $s_{1,2}$ sample the same underlying signal, $\mu$, but add different noise realisations $\xi_{1, 2}$.
+where all timestreams sample the same underlying signal, $\mu$, but add different noise realisations $\xi_{1, 2}$.
 The digitisation process is denoted by $D(\dots)$.
-We begin by looking at the crosspower of non-digitised timestreams
+The covariance of two timestreams must be zero.
+This still holds true once the signals have been digitised, as this only discretises the probability distributions at play but does not change the underlying dynamics.
+Therefore
 
 \begin{equation}
-\langle s_1 s_2 \rangle = \langle (\mu + \xi_1) (\mu + \xi_2) \rangle =  \langle \mu^2 \rangle = \mu^2.
+\langle D(\mu + \xi_1) D(\mu + \xi_2) \rangle = \sum_1 \sum_2 D(\mu + \xi_1) D(\mu + \xi_1) P(D(\mu + \xi_1)) P(D(\mu + \xi_2)),
 \end{equation}
 
-This allows us to state
+where $P$ denotes the probability of a certain output level occurring.
+For a small signal $D(\mu + \xi)=D(\xi)$.
+The probabilities $P$ can then be expressed as integrals over probability distributions.
+Assuming Gaussian detector noise we introduce normal distributions $p(\xi)$ with mean $\mu$ and standard deviation $\sigma$.
 
 \begin{equation}
-cov(s_1, s_2) = \langle s_1 s_2 \rangle - \langle s_1 \rangle \langle s_2 \rangle = \mu^2 - \mu^2 = 0.
+\langle D(\mu + \xi_1) D(\mu + \xi_2) \rangle = \sum_{i=0}^N \sum_{j=0}^N  \int_{x_i}^{x_{i+1}} \int_{x_j}^{x_{j+1}} y_i y_j p(\xi_1) p(\xi_2) d\xi_1 d\xi_2.
 \end{equation}
 
-We argue that this still holds true after the digitisation process, since timestreams still have independent stochastic components.
-Digitisation discretises the probability distributions at play, but does not change the underlying dynamics.
+We separate the above sums and recognise the error function.
 
 \begin{equation}
-\langle D(s_1) D(s_2) \rangle = \sum_1 \sum_2 D(\mu + \xi_1) D(\mu + \xi_1) P(D(\mu + \xi_1)) P(D(\mu + \xi_2)),
-\end{equation}
-
-where $P$ denotes the probability of a certain digitisation level occurring.
-We now assume that the underlying signal is small, i.e.
-$| \mu | \ll | \xi |$.
-This is true for ground-based CMB observations.
-In this limit $D(\mu + \xi)$ is digitised to the same value as $D(\xi)$.
-The probabilities can then be expressed as integrals over the relevant sections of correctly parameterised probability distributions.
-We proceed by assuming that the noise $\xi$ is drawn from normal distributions with mean $\mu$ and standard deviation $\sigma$, denoted by $N(\mu, \sigma, \xi)$.
-
-\begin{equation}
-\langle D(s_1) D(s_2) \rangle = \sum_{i=0}^N \sum_{j=0}^N  \int_{x_i}^{x_{i+1}} \int_{x_j}^{x_{j+1}} y_i y_j N(\mu, \sigma, \xi_1) N(\mu, \sigma, \xi_2) d\xi_1 d\xi_2.
-\end{equation}
-
-We separate the above sums temporarily to detail our next steps. We split up the integrals over ranges that allow us to recognise the error function evaluated at different points.
-
-\begin{equation}
-\langle D(s_1) D(s_2) \rangle = \left[ \sum_{i=0}^N  y_i \int_{x_i}^{x_{i+1}} N(\mu, \sigma)  d\xi_1 \right] \left[  \sum_{j=0}^N y_j \int_{x_j}^{x_{j+1}} N(\mu, \sigma)d\xi_2 \right].
+\begin{aligned}
+\langle D(\mu + \xi_1) D(\mu + \xi_2) \rangle &= \left\{ \sum_{i=0}^N  \frac{y_i}{2} \left[ erf \left( \frac{x_{i+1} - \mu}{\sqrt{2}\sigma} \right) - erf \left( \frac{x_{i} - \mu}{\sqrt{2}\sigma} \right) \right] \right\} \\
+&\times \left\{  \sum_{j=0}^N \frac{y_j}{2} \left[ erf \left( \frac{x_{j+1} - \mu}{\sqrt{2}\sigma} \right) - erf \left( \frac{x_{j} - \mu}{\sqrt{2}\sigma} \right) \right] \right\}.
+\end{aligned}
 \end{equation}
@@
 \begin{equation}
-= \left\{ \sum_{i=0}^N  y_i \left[ \int_{-\infty}^{x_{i+1}} - \int_{-\infty}^{x_{i}} \right] N(\mu, \sigma)  d\xi_1 \right\} \left\{  \sum_{j=0}^N y_j \left[ \int_{-\infty}^{x_{j+1}} - \int_{-\infty}^{x_{j}} \right] N(\mu, \sigma)d\xi_2 \right\}.
+\begin{aligned}
+&= \left\{ \sum_{i=0}^N  \frac{y_i}{2} \left[ \frac{2}{\sqrt{\pi}} \sum_{n = 0}^\infty \frac{(-1)^n}{n!
+(2n+1)} \frac{1}{(\sqrt{2}\sigma)^{2n+1}} \left( (x_{i+1}-\mu)^{2n+1} - (x_{i}-\mu)^{2n+1} \right) \right] \right\} \\
+&\times \left\{ \sum_{j=0}^N  \frac{y_j}{2} \left[ \frac{2}{\sqrt{\pi}} \sum_{m = 0}^\infty \frac{(-1)^m}{m!
+(2m+1)} \frac{1}{(\sqrt{2}\sigma)^{2m+1}} \left( (x_{j+1}-\mu)^{2m+1} - (x_{j}-\mu)^{2m+1} \right) \right] \right\}.
+\end{aligned}
 \end{equation}
@@
 \begin{equation}
-= \left\{ \sum_{i=0}^N  \frac{y_i}{2} \left[ erf \left( \frac{x_{i+1} - \mu}{\sqrt{2}\sigma} \right) - erf \left( \frac{x_{i} - \mu}{\sqrt{2}\sigma} \right) \right] \right\} \left\{  \sum_{j=0}^N \frac{y_j}{2} \left[ erf \left( \frac{x_{j+1} - \mu}{\sqrt{2}\sigma} \right) - erf \left( \frac{x_{j} - \mu}{\sqrt{2}\sigma} \right) \right] \right\}.
-\end{equation}
-
-Using the Maclaurin series of the error function
-
-\begin{equation}
 \begin{aligned}
-\langle D(s_1) D(s_2) \rangle &= \left\{ \sum_{i=0}^N  \frac{y_i}{2} \left[ \frac{2}{\sqrt{\pi}} \sum_{n = 0}^\infty \frac{(-1)^n}{n!
-(2n+1)} \frac{1}{(\sqrt{2}\sigma)^{2n+1}} \left( (x_{i+1}-\mu)^{2n+1} - (x_{i}-\mu)^{2n+1} \right) \right] \right\} \\
-&\times \left\{ \sum_{j=0}^N  \frac{y_j}{2} \left[ \frac{2}{\sqrt{\pi}} \sum_{m = 0}^\infty \frac{(-1)^m}{m!
-(2m+1)} \frac{1}{(\sqrt{2}\sigma)^{2m+1}} \left( (x_{j+1}-\mu)^{2m+1} - (x_{j}-\mu)^{2m+1} \right) \right] \right\}
-\end{aligned}.
-\end{equation}
-
-The binomial expansion is now applied to all brackets containing $\mu$ and the sums over $n$ and $m$ are recombined
-
-\begin{equation}
-\begin{aligned}
-\langle D(s_1) D(s_2) \rangle &= \left\{ \sum_{i=0}^N  \frac{y_i}{\sqrt{pi}} \left[ \sum_{n = 0}^\infty \frac{(-1)^n}{n!
+&= \left\{ \sum_{i=0}^N  \frac{y_i}{\sqrt{pi}} \left[ \sum_{n = 0}^\infty \frac{(-1)^n}{n!
 (2n+1)} \frac{1}{(\sqrt{2}\sigma)^{2n+1}} \left( \sum_{k=0}^{2n+1} {2n+1 \choose k} \mu^k ( x_{i+1}^{2n+1-k} - x_{i}^{2n+1-k} ) \right) \right] \right\} \\
@@
 
-\begin{equation}
+Where we have used the Maclaurin series of the error function and the binomial expansion. Combining all possible sums and moving the $i, j$ summation forwards yields
+
+\begin{equation} \label{eq:dcompleteklsum}
 \begin{aligned}
-&= \sum_{i,j=0}^N  \frac{y_i y_j}{\pi} \left[ \sum_{n,m = 0}^\infty \frac{1}{n! (2n+1)} \frac{(-1)^{n+m}}{m! (2m+1)} \frac{1}{(\sqrt{2}\sigma)^{2n+2m+2}} \right. \\
-& \left. \times \left( \sum_{k, l = 0}^{2n+1} {2n+1 \choose k} {2m+1 \choose l} \mu^{k+l} ( x_{i+1}^{2n+1-k} - x_{i}^{2n+1-k} ) ( x_{j+1}^{2m+1-l}- x_{j}^{2m+1-l}) \right) \right].
+\langle D(\mu + \xi_1) D(\mu + \xi_2) \rangle &=  \frac{1}{\pi} \sum_{n,m = 0}^\infty \frac{1}{n! (2n+1)} \frac{(-1)^{n+m}}{m! (2m+1)} \frac{1}{(\sqrt{2}\sigma)^{2n+2m+2}} \\
+&\times \sum_{k, l = 0}^{2n+1} {2n+1 \choose k} {2m+1 \choose l} \mu^{k+l} \sum_{i,j=0}^N y_i y_j \psi^{k, n}_i \psi^{l, m}_j,
 \end{aligned}
@@
 
-\begin{equation} \label{eq:dcompleteklsum}
-=  \frac{1}{\pi} \sum_{n,m = 0}^\infty \frac{1}{n!
-(2n+1)} \frac{(-1)^{n+m}}{m!
-(2m+1)} \frac{1}{(\sqrt{2}\sigma)^{2n+2m+2}} \sum_{k, l = 0}^{2n+1} {2n+1 \choose k} {2m+1 \choose l} \mu^{k+l} \sum_{i,j=0}^N y_i y_j \psi^{k, n}_i \psi^{l, m}_j,
-\end{equation}
-
-where in the last line we have moved $y_i$ and $y_j$ and the sum over $i,j$ as far forward as possible. Additionally, we have introduced the new variable
+with
 
@@
 
-We now split the innermost term such that
+We split the innermost term in equation \ref{eq:dcompleteklsum} and shift indicies such that
 
@@
 \begin{equation}
- = \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} y_i y_j \psi_i^{k, n} \psi_j^{l, m} + \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} y_{N-i} y_j \psi_{N-i}^{k, n} \psi_j^{l, m} + \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} y_i y_{N-j} \psi_i^{k, n} \psi_{N-j}^{l, m} + \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} y_{N-i} y_{N-j} \psi_{N-i}^{k, n} \psi_{N-j}^{l, m},
+= \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} \left( y_i y_j \psi_i^{k, n} \psi_j^{l, m} +  y_{N-i} y_j \psi_{N-i}^{k, n} \psi_j^{l, m} + y_i y_{N-j} \psi_i^{k, n} \psi_{N-j}^{l, m} + y_{N-i} y_{N-j} \psi_{N-i}^{k, n} \psi_{N-j}^{l, m} \right).
 \end{equation}
 
-where we have shifted all indices of sums that do not start at $0$ to $N-i$ and/or $N-j$ as applicable. We recombine the sums such that
-
-\begin{equation}
-\sum_{i,j = 0}^N y_i y_j \psi_i^{k, n} \psi_j^{l, m} = \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} \left( y_i y_j \psi_i^{k, n} \psi_j^{l, m} +  y_{N-i} y_j \psi_{N-i}^{k, n} \psi_j^{l, m} + y_i y_{N-j} \psi_i^{k, n} \psi_{N-j}^{l, m} + y_{N-i} y_{N-j} \psi_{N-i}^{k, n} \psi_{N-j}^{l, m} \right)
-\end{equation}
-
-Equations \ref{eq:digitequalspacecondition} and \ref{eq:digitareacondition} in \ref{subsec:extremedigitisation} demand that for a Gaussian input distribution $y_i = -y_{N-i}$ and $x_i = -x_{N+1-i}$.
-This simply states symmetry of the digitisation thresholds and levels about zero (or after a global mean has been subtracted).
-With this information we can condense the above to
-
-\begin{equation} \label{eq:ijsum}
-\sum_{i,j = 0}^N y_i y_j \psi_i^{k, n} \psi_j^{l, m} = \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} y_i y_j \left( \psi_i^{k, n} \psi_j^{l, m} - \psi_{N-i}^{k, n} \psi_j^{l, m} - \psi_i^{k, n} \psi_{N-j}^{l, m} + \psi_{N-i}^{k, n} \psi_{N-j}^{l, m} \right)
-\end{equation}
-
-The index shift $i \to N-i$ affects $\psi_{i}^{k, n}$ as follows:
-
-\begin{equation}
-\begin{aligned}
-\psi_{N-i}^{k,n} &= x_{N+1-i}^{2n+1-k} - x_{N-i}^{2n+1-k} = (-x_{i})^{2n+1-k} - (-x_{i+1})^{2n+1-k}\\
-&= (-1)^{2n+1-k} (x_{i})^{2n+1-k} - x_{i+1})^{2n+1-k} ) = (-1)^{2n+2-k} (x_{i+1})^{2n+1-k} - x_{i1})^{2n+1-k} )\\
-&= (-1)^{k} \psi_{i}^{k,n}.
-\end{aligned}
-\end{equation}
-
-Therefore equation \ref{eq:ijsum} may be rewritten as
+Equations \ref{eq:digitequalspacecondition} and \ref{eq:digitareacondition} in subsection \S\ref{subsec:extremedigitisation} demand that for a Gaussian input distribution $y_i = -y_{N-i}$ and $x_i = -x_{N+1-i}$.
+This simply states symmetry of the digitisation thresholds and output levels about zero (or after a global mean has been subtracted).
+Therefore
 
@@
 
-We keep hold of this result and move onto the double sum over $k,l$ in equation \ref{eq:dcompleteklsum}.
-Since we assumed that the signal is small we would like to calculate the leading order contributions in $\mu$.
-We split up the double sum such that
+since $\psi_{N-i}^{k,n} = (-1)^{k} \psi_{i}^{k,n}$.
+Since we assume that $\mu$ is small we would like to calculate the leading order terms in the above.
+Since we have power of $\mu^{k+l}$ in equation \ref{eq:dcompleteklsum} we focus on the case of even $k$ and $l$ first.
+This reduces equation \ref{eq:ijsumpsiflip} to
 
-\begin{equation} \label{eq:ksums}
-\sum_{k, l = 0}^{2n+1} = \sum_{\mathrm{even} \> k}^{2n+1} \sum_{\mathrm{even} \> l}^{2n+1}  + \sum_{\mathrm{odd} \> k}^{2n+1} \sum_{\mathrm{odd} \> l}^{2n+1} + \sum_{\mathrm{odd} \> k}^{2n+1} \sum_{\mathrm{even} \> l}^{2n+1}  + \sum_{\mathrm{even} \> k}^{2n+1} \sum_{\mathrm{odd} \> l}^{2n+1}.
+\begin{equation}
+\sum_{i,j = 0}^N y_i y_j \psi_i^{k, n} \psi_j^{l, m} = \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} y_i y_j  \left( \psi_i^{k, n} \psi_j^{l, m} - \psi_{i}^{k, n} \psi_j^{l, m} - \psi_i^{k, n} \psi_{j}^{l, m} + \psi_{i}^{k, n} \psi_{j}^{l, m} \right) = 0,
 \end{equation}
 
-The first two terms in the above will lead to even powers of $\mu$, whereas the last two will lead to odd powers of $\mu$.
-We are now ready to put all pieces together and start by investigating even powers of $\mu$ in equation \ref{eq:dcompleteklsum} that arise from even $k$ and $l$.
-Under the assumptions made \ref{eq:ijsumpsiflip} reduces to
-
-\begin{equation}
-\sum_{i,j = 0}^N y_i y_j \psi_i^{k, n} \psi_j^{l, m} = \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} y_i y_j  \left( \psi_i^{k, n} \psi_j^{l, m} - \psi_{i}^{k, n} \psi_j^{l, m} - \psi_i^{k, n} \psi_{j}^{l, m} + \psi_{i}^{k, n} \psi_{j}^{l, m} \right) = 0.
-\end{equation}
-
-The term vanishes. The other way of producing even powers of $\mu$ in equation \ref{eq:dcompleteklsum} is with odd $k$ and $l$. These assumptions simplify equation \ref{eq:ijsum} to
+The term vanishes. For odd $k$ and $l$ equation \ref{eq:ijsumpsiflip} may be simplified to
 
@@
 
-This term survives and leads to even powers of $\mu$ in equation \ref{eq:dcompleteklsum}. Moving on to odd powers of $\mu$ we consider the case of odd $k$ and even $l$. We observe
+This term survives and leads to even powers of $\mu$ in equation \ref{eq:dcompleteklsum}. Considering the case of odd $k$ and even $l$ we observe
 
@@
 for odd $k$ and even $l$.
-All odd powers of $\mu$ in equation \ref{eq:dcompleteklsum} vanish.
-Moreover we can also state that there is no $\mu$ independent term, since this could only be produced by $k=l=0$, but even $k$ and $l$ terms are shown to vanish.
+Therefore all odd powers of $\mu$ in equation \ref{eq:dcompleteklsum} vanish.
+Moreover there is no $\mu$ independent term, since this could only be produced by $k=l=0$, but even $k$ and $l$ terms are shown to vanish.
 Therefore
@@
 \begin{equation}
-\langle D(s_1) D(s_2) \rangle = c \mu^2 + \mathcal{O}(\mu^4)
+\langle D(\mu + \xi_1) D(\mu + \xi_2) \rangle = c \mu^2 + \mathcal{O}(\mu^4)
 \end{equation}
@@
 where $c$ is some constants that depend on the specific digitisation scheme chosen.
-We now have prove that in the small signal limit and under the assumptions about the digitisation scheme stated normalisation can indeed be achieved by a constant.
+We have proved that in the small signal limit and for Gaussian noise normalisation can be achieved through rescaling by a constant.
 Returning to equation \ref{eq:normcrosspower} we see
@@
 \begin{equation}
-\mathcal{N}  = \frac{\langle s_1 s_2 \rangle}{\langle D(s_1) D(s_2) \rangle} = \frac{\mu^2}{c \mu^2 + \mathcal{O}(\mu^4)} = \frac{1}{c + \mathcal{O}(\mu^2)} = \frac{1}{c},
+\gamma^2  = \frac{\langle (\mu + \xi_1) (\mu + \xi_2) \rangle}{\langle D(\mu + \xi_1) D(\mu + \xi_2) \rangle} = \frac{1}{c + \mathcal{O}(\mu^2)} = \frac{1}{c},
 \end{equation}
 
-where we have ignored higher order contributions. From equation \ref{eq:dcompleteklsum} we obtain an expression for $\mathcal{N}$. We focus on $k=l=1$ terms.
+where we have ignored higher order contributions. From equation \ref{eq:dcompleteklsum} we obtain an expression for $\gamma$, by investigating $k=l=1$ terms.
 
@@
 \begin{equation}
-c = \sum_{i,j=0}^N  \frac{y_i y_j}{\pi} \left[ \sum_{n,m = 0}^\infty \frac{1}{n!
-(2n+1)} \frac{(-1)^{n+m}}{m!
-(2m+1)} \frac{1}{(\sqrt{2}\sigma)^{2n+2m+2}} \left( \frac{(2n+1)!}{(2n)!} \frac{(2m+1)!}{(2m)!} ( x_{i+1}^{2n} - x_{i}^{2n} ) ( x_{j+1}^{2m} - x_{j}^{2m} ) \right) \right]
-\end{equation}
-
-\begin{equation}
 c = \sum_{i,j=0}^N \frac{y_i y_j}{2\pi\sigma^2} \left[ \sum_{n = 0}^\infty \frac{(-1)^n}{n!} \frac{1}{(\sqrt{2}\sigma)^{2n}}  ( x_{i+1}^{2n} - x_{i}^{2n} ) \right] \left[ \sum_{m = 0}^\infty \frac{(-1)^m}{m!} \frac{1}{(\sqrt{2}\sigma)^{2m}} ( x_{j+1}^{2m} - x_{j}^{2m} ) \right].
 \end{equation}
-
-We recognise the exponential function in the above.
 
@@
 \begin{equation}
-\mathcal{N} = \left( \sum_{i,j=0}^N \frac{y_i y_j}{2\pi\sigma^2} \left\{ exp \left[ - \left(\frac{x_{i+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{i}}{\sqrt{2}\sigma} \right)^2 \right] \right\} \left\{ exp \left[ - \left(\frac{x_{j+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{j}}{\sqrt{2}\sigma} \right)^2 \right] \right\} \right)^{-1}
+\gamma = \left( \sum_{i,j=0}^N \frac{y_i y_j}{2\pi\sigma^2} \left\{ exp \left[ - \left(\frac{x_{i+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{i}}{\sqrt{2}\sigma} \right)^2 \right] \right\} \left\{ exp \left[ - \left(\frac{x_{j+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{j}}{\sqrt{2}\sigma} \right)^2 \right] \right\} \right)^{-1/2}
 \end{equation}
 
-This can be calculated for a given digitisation scheme $N, x_i, y_i$ and standard deviation of noise $\sigma$.
+This can be calculated for a given digitisation scheme $N, x_i, y_i$ and standard deviation of the noise $\sigma$.
 Notice that if the digitisation scheme is chosen such that $y_i$ and $x_i$ depend linearly on $\sigma$ the normalisation constant is independent of $\sigma$.

### 2018-09-23 digitisation.tex
Editing and put in preliminary plots

@@
 Observations of the cosmic microwave background (CMB) have played a key role in cosmology \citep{penzias1965, smoot1992, bennett2013, sptpol2013, planck2018}.
-They can measure the relativistic number of species and the neutrino mass sum.
 Current CMB Experiments target science goals such as the discovery of inflationary gravitational waves, measuring the relativistic number of species and the neutrino mass sum, and mapping the large-scale distribution of matter through gravitational lensing and the Sunyaev-Zeldovich (SZ) effects.
@@
 We find the appropriate normalisation constants for power spectra of few bit TOD as described in the appendix.
-As seen in Figure \ref{fig:psrecover} all spectra of extremely digitised data recover the input to a satisfying degree.
+As seen in Figure \ref{fig:psrecover} (appendix) all spectra of extremely digitised data recover the input to a satisfying degree.
 Differences between few bit and the input power spectra are due to observing a single CMB realisation and residual boundary effects.
@@
 %%%%%%%%%%%
-
-\begin{figure}[htb]\centering
-\includegraphics[width=0.5\textwidth,clip]{Plots/psrecovery.pdf}
-  \caption[Current ]{
-  Input power spectra used to generate the CMB template maps plus detector noise level (black) and power spectra using 3 bit TOD (violet).
-A simulations with $1,024,000$ hits per map pixel is shown in this plot.
-No qualitative differences are visible for other depths of observation.
-\label{fig:psrecover}
-}
-\end{figure}
 
@@
   \caption[Current ]{
-  Power spectra of reconstructed minus input maps for one-bit (green), two-bit (orange), three-bit (violet) and the control case (black) rebined to $\Delta \ell = 100$.
+  Power spectra of reconstructed minus input maps for one-bit (green), two-bit (orange), three-bit (violet) and the control case (black) rebined to $\Delta \ell = 400$.
 No angular scale dependence is observed for all depths of observation.
@@
 \appendix
+
+\begin{figure*}[h!t]\centering
+\vspace{-1cm}
+\includegraphics[width=0.8\textwidth,clip]{Plots/psrecovery.pdf}
+  \caption[Current ]{
+  Input power spectra used to generate the CMB template maps plus detector noise level (black) and digitisation noise (gray) and power spectra using three (violet), two (orange), and one (green) bit TOD.
+Spectra are rebined to $\Delta \ell = 100$.
+Main panels are the calculated power spectra binned to $\Delta \ell = 100$, accompanying panels show the difference between the corresponding input plus noise curves and the $C_\ell$ of digitised TOD.
+Variations occur because only observe a single CMB realisation.
+A simulations with $1,024,000$ hits per map pixel is shown in this plot.
+No qualitative differences are visible for other depths of observation.
+\label{fig:psrecover}
+}
+\end{figure*}
 

### 2018-10-08 digitisation.tex
Editing, comments from Nikhel.

@@
 \newcommand{\bicepkeck}{BICEP2/Keck Array}
-\newcommand{\sptnew}{SPT-3G}
+\newcommand{\sptnew}{SPT-3G }
 \newcommand{\pb}{\textsc{Polarbear}}
@@
 \newcommand{\sptpol}{SPTpol}
-\newcommand{\advactpol}{Adv.~ACTpol}
+\newcommand{\advactpol}{Adv.~ACTpol }
 
@@
 
-Observations of the cosmic microwave background (CMB) are of significant value to modern cosmology and particle physics.
-The large size of the time ordered data (TOD) of future CMB experiments presents challenges for mission planning and data analysis.
+The large size of the time ordered data (TOD) of cosmic microwave background (CMB) experiments presents challenges for mission planning and data analysis.
 These issues are particularly significant for Antarctica- and space-based experiments, which depend on satellite links to transmit data.
@@
 Observations of the cosmic microwave background (CMB) have played a key role in cosmology \citep{penzias1965, smoot1992, bennett2013, sptpol2013, planck2018}.
-Current CMB Experiments target science goals such as the discovery of inflationary gravitational waves, measuring the relativistic number of species and the neutrino mass sum, and mapping the large-scale distribution of matter through gravitational lensing and the Sunyaev-Zeldovich (SZ) effects.
+Current ground-based CMB experiments, like \sptnew \citep{benson2014} at the South Pole telescope (SPT) and \advactpol \citep{thornton2016}, at the Atacama cosmology telescope (ACT), target science goals such as the discovery of inflationary gravitational waves, measuring the number of relativistic species, the neutrino mass sum, and mapping the large-scale distribution of matter through gravitational lensing and the Sunyaev-Zeldovich (SZ) effects.
 Through lensing and the SZ effects the CMB probes structure formation, reionisation, and is a powerful test for dark matter and dark energy models \citep{s4sciencebook, core2011, litebird2016, pixie2011}.
@@
 
-Over the last two decades CMB experiments have gone from single detectors to over $10,000$ detectors.
+Over the last two decades, CMB experiments have gone from single detectors to over $10,000$ detectors.
 The CMB community has developed a variety of compression techniques and computational approaches to handle the increasing volume of data \citep{tristam2007}.
@@
 Antarctica-based experiments that transmit a portion of their results via a satellite link downsample their data to meet telemetry allocations, but have not yet used few bit digitisation of the TOD.
-As we approach the next generation of ground-based experiments, CMB-S4, and the launch of new space-based missions (LiteBIRD, PIXIE, COrE+), we must consider potential transmission bottlenecks carefully.
+As we approach the next generation ground-based experiment, CMB-S4, and the launch of new space-based missions (LiteBIRD, PIXIE, COrE+), we must consider potential transmission bottlenecks carefully.
 %C: Is there a reference detailing downsampling of SPT data?
@@
 In this work we present the method of extreme digitisation, which reduces a many bit (often 32 or 64 bit) signal to a few bits for ground-based experiments.
-We apply extreme digitisation to the TOD and detail the resulting effects on temperature and polarisation power spectra.
+We apply extreme digitisation to simulated TOD and detail the resulting effects on temperature and polarisation power spectra.
 We find that an optimal three bit digitisation scheme adds $<2\%$ to the map noise level.
@@
 In \S\ref{sec:dig} we detail the challenges that come with handling large data volumes, introduce the process of extreme digitisation and lay out the framework used to test its performance.
-We describe the power spectrum estimation and present the results in section \S\ref{sec:results}.
+We present results for white and $1/f$ detector noise in section \S\ref{sec:results}.
 We summarise our findings in \S\ref{sec:conclusions}.
@@
 The science goals of upcoming CMB experiments depend on achieving substantially faster mapping speeds.
-Given CMB detectors are generally photon noise limited, improving an experiments means adding more detectors.
-As a result the number of detectors (and data volume) of ground-based experiments has followed a trend like Moore's law, doubling approximately every 2 years \citep{s4sciencebook, Abazajian2015}.
+Given CMB detectors are generally photon noise limited, improving the mapping speed means adding more detectors.
+As a result the number of detectors (and data volume) of ground-based experiments has followed an exponential trend like Moore's law, doubling approximately every 2 years \citep{s4sciencebook, Abazajian2015}.
 
 The South Pole is one of the best sites for CMB observations on Earth \citep{kovac2007}.
-The next generation CMB experiment, CMB-S4, plans to include several telescopes at the South Pole.
-These telescopes would generate a data influx of $\sim\mathcal{O}(10)\mathrm{Tb/d}$.
-However, the current transmission allocation for SPT-3G is $150\mathrm{Gb/d}$, which will likely see only a moderate increase.
+CMB-S4 plans to include several telescopes at the South Pole, which will generate a data influx of $\sim\mathcal{O}(10)\mathrm{Tb/d}$.
+Transferring this data volume via satellite would be expensive.
+For context, the transmission allocation for a current CMB experiment, SPT-3G, is $150\mathrm{Gb/d}$.
 The transmission bottleneck could be overcome by recovering the full data on hard drives every summer and transmitting a downsampled version of the data.
@@
 
-Beyond transmission challenges, larger data volumes require larger and more expensive disk arrays and make end-to-end simulations of the experiment for the purpose of optimisation on systematics estimation more time consuming.
-Given the shear size of upcoming data sets, full end-to-end simulation may prove impossible \citep{s4sciencebook}.
-Operations on the TOD, such as noise removal and map making, are a vital part of CMB data analysis.
-The exponential growth of CMB data makes these steps increasingly expensive.
-Few bit digitisation could ameliorate both of these challenges.
+Beyond transmission challenges, a larger data volume increases the size and cost of disk arrays and makes end-to-end simulations of experiments for the purpose of optimisation on systematics estimation more time consuming.
+Given the sheer size of upcoming data sets, full end-to-end simulations may prove impractical for CMB-S4 \citep{s4sciencebook}.
+The exponential growth of CMB data makes timestream level operations, such as noise removal and map making, increasingly time consuming.
+Few bit digitisation could ameliorate all of these challenges.
 
@@
 
-Few bit digitisation has already been demonstrated in space by the Planck mission \citep{maris2003}, as well as in other areas of physics.
-\cite{jenet1998} explored the application of such compression to radio pulsar timing measurements with success.
+The Planck mission has already demonstrated the success of extreme digitisation for space-based CMB experiment s\citep{maris2003}.
+\cite{jenet1998} explored the application of few bit digitisation to radio pulsar timing measurements.
 Recently Clearwater et al.
@@
 
-Digitisation discretises an input signal by sorting it into $N$ appropriate ranges, such that an input between $x_i$ and $x_{i+1}$ produces an output value $y_i$.
+Digitisation discretises an input signal by sorting it into $N$ ranges, such that an input between $x_i$ and $x_{i+1}$ produces an output value $y_i$.
 The set of parameters $N, x_i, y_i$ fully specify a digitisation scheme.
@@
 
-which informs us that the cutoff levels $x_i$ should lie midway between the output levels.
+which informs us that the threshold levels should lie midway between their adjacent output levels.
 Setting the derivative of equation \ref{eq:dist} with respect to $y_i$  to zero gives the additional condition
@@
 \hat{s}_2(t) = \left\{ \begin{array}{rl}
-1.51 \sigma, & s(t) \geq 0.9816 \sigma\\
-0.4528 \sigma, & 0 \leq s(t) < 0.9816 \sigma\\
--0.4528 \sigma, & -0.9816 \sigma \leq s(t) < 0\\
--1.51 \sigma, & -0.9816 \sigma < s(t)\\
+1.51 \sigma, & s(t) \geq 0.98 \sigma\\
+0.45 \sigma, & 0 \leq s(t) < 0.98 \sigma\\
+-0.45 \sigma, & -0.98 \sigma \leq s(t) < 0\\
+-1.51 \sigma, & -0.98 \sigma < s(t)\\
 \end{array} \right. . \end{equation}
@@
 \hat{s}_3(t) = \left\{ \begin{array}{rl}
-2.152 \sigma, & s(t) \geq 1.748 \sigma\\
-1.344 \sigma, & 1.05 \sigma \leq s(t) < 1.748 \sigma\\
-0.756 \sigma, & 0.501 \sigma \leq s(t) < 1.05 \sigma\\
-0.245 \sigma, & 0 \leq s(t) < 0.501 \sigma\\
--0.245 \sigma, & -0.501 \sigma \leq s(t) < 0\\
--0.756 \sigma, & -1.05 \sigma \leq s(t) < -0.501 \sigma\\
--1.344 \sigma, & -1.748 \sigma \leq s(t) < -1.05 \sigma\\
--2.152 \sigma, & -1.748 \sigma < s(t)\\
+2.15 \sigma, & s(t) \geq 1.75 \sigma\\
+1.34 \sigma, & 1.05 \sigma \leq s(t) < 1.75 \sigma\\
+0.76 \sigma, & 0.50 \sigma \leq s(t) < 1.05 \sigma\\
+0.25 \sigma, & 0 \leq s(t) < 0.50 \sigma\\
+-0.25 \sigma, & -0.50 \sigma \leq s(t) < 0\\
+-0.76 \sigma, & -1.05 \sigma \leq s(t) < -0.50 \sigma\\
+-1.34 \sigma, & -1.75 \sigma \leq s(t) < -1.05 \sigma\\
+-2.15 \sigma, & -1.75 \sigma < s(t)\\
 \end{array} \right. . \end{equation}
@@
 The digitisation schemes listed above do not conserve signal power and need to be re-calibrated to recover the correct power in the map.
-This is not particularly important, since most CMB experiments determine the absolute calibration from their maps.
-However, it is worth noting that one can rescale the schemes laid out by \cite{max1960} to preserve power, which is demonstrated in the appendix.
+We demonstrate that the schemes laid out by \cite{max1960} can be rescaled to conserve power and derive an analytical expression for the normalisation constant in the appendix.
+We use this analytical rescaling for all results in this work.
+
+%This is not particularly important, since most CMB experiments determine the absolute calibration from their maps. However,
+
+%However, it is worth noting that one can rescale the schemes laid out by \cite{max1960} to preserve power, which is demonstrated in the appendix.
 
@@
 
-\subsection{Methods}
+\subsection{From Time Ordered Data To Maps}
 \label{subsec:method}
 
-To investigate the performance of the derived digitisation schemes, we simulate many scans over CMB template maps at the timestream level.
-We add realisations of detector noise and apply the digitisation schemes above to the constructed I, Q, and U TOD.
-Additionally, we retain the original 64 bit TOD to construct control maps.
-The different timestreams are binned into maps for a T, E, and B power spectrum analysis.
+To investigate the performance of the derived digitisation schemes, we simulate timestream level scans over a single CMB realisation.
+We add detector noise and apply the digitisation schemes above to the simulated I, Q, and U TOD.
+We also retain the original 64 bit TOD to construct control maps.
+The different timestreams are binned into HEALPix \citep{healpix}\footnote{Available at \url{http://healpix.sourceforge.net/}.} maps with resolution $\mathrm{NSIDE} = 4096$ for a T, E, and B power spectrum analysis.
 
@@
 
-To create the template maps we use the HEALPix \citep{healpix}\footnote{Available at \url{http://healpix.sourceforge.net/}.} framework through the healpy package.
-We generate a realisation of I, Q and U maps with resolution $\mathrm{NSIDE} = 4096$, corresponding to a pixel area of approximately $0.74 \> \mathrm{arcmin^2}$.
-This is suitable for the analysis of Stage 3 and CMB-S4 experiments.
-The input power spectrum is based on the 2015 results of the Planck collaboration.
-The key cosmological parameters are $H_0 = 67.8 \mathrm{km \> s^{-1} Mpc^{-1}}$, $\Omega_{\mathrm{m}} = 0.308 $, $n_{\mathrm{s}} = 0.968$, and $\tau = 0.066$ \citep{planck2016}.
+We use the healpy Python package to generate a single CMB realisation for the best-fit Planck 2015 cosmology.
+Specifically, the key cosmological parameters are $H_0 = 67.8 \mathrm{km \> s^{-1} Mpc^{-1}}$, $\Omega_{\mathrm{m}} = 0.308 $, $n_{\mathrm{s}} = 0.968$, and $\tau = 0.066$ \citep{planck2016}.
 
@@
 We simulate observing a $600 \> \mathrm{ deg^2}$ patch of the sky with a single detector.
-Constant elevation scans (CES) are performed beginning at right ascension (RA) and declination (DEC) $(0^\circ, 0^\circ)$.
 The detector noise level for temperature is $500 \mathrm{\mu K \sqrt{s}}$ and for polarisation observations $\sqrt{2} \hspace{0.7ex} 500 \mathrm{\mu K \sqrt{s}}$, i.e.
 the corresponding photon noise limit.
+Constant elevation scans (CES) are performed beginning at right ascension (RA) and declination (DEC) $(0^\circ, 0^\circ)$.
 After covering $24^\circ30'$ along RA the detector is reset to a RA of $0^\circ$ and an offset in DEC corresponding to the pixel size.
 Through repetition of CES with constant steps in DEC the survey patch is covered.
+This imitates the scan strategy of the SPT \citep{schaffer2011}.
 The entire scan strategy is repeated $100$ times with offsets in the starting RA and DEC up to the size of one pixel, ensuring that pixels are sampled uniformly.
-This scan strategy imitates the SPT, which sweeps across the sky along RA with constant steps in DEC \citep{schaffer2011}.
 
@@
 
-\section{Results}
-\label{sec:results}
-
 \subsection{Power Spectrum Estimation}
@@
 
-We use PolSpice \citep{polspice} to compute $\mathrm{TT}$, $\mathrm{EE}$ and $\mathrm{BB}$ power spectra from the I, Q, and U maps above.
-To minimise boundary effects we apodise the observed patch using a cosine mask with $\sigma_{\mathrm{APOD}} = \sqrt{600}/2 \hspace{0.7ex} \mathrm{deg}$.
-%C: Should I also mention the multiplication with a cosine mask that I supply in addition to setting this option on polspice?
-This will likely be reworked when using Healpix to get the power spectra though.
+To minimise boundary effects in the subsequent analysis we apodise the observed patch using a cosine mask.
+For the case of white detector noise we use PolSpice \citep{polspice} to compute $\mathrm{TT}$, $\mathrm{EE}$ and $\mathrm{BB}$ power spectra from output maps.
 
-We find the appropriate normalisation constants for power spectra of few bit TOD as described in the appendix.
+For the later considered case of anisotropic noise, we use healpy to calculate the spherical harmonic coefficients $a\ell m$.
+Since the scan strategy of the simulated observations is similar to the SPT's, smaller values of $m$ correspond to larger angular features along the scan direction \citep{chown2018}.
+Because of the anisotropic nature of the noise, equally weighting all modes leads to inaccurate power spectra.
+As a first order approximation to optimal weighting we discount the lowest modes in scan direction.
+
+We rebin the calculated power spectra as mentioned in section \S\ref{sec:results} and assign Fisher matrix error bars \citep{tegmarkdesignersguide1997}.
+
 As seen in Figure \ref{fig:psrecover} (appendix) all spectra of extremely digitised data recover the input to a satisfying degree.
@@
 
-%%%%%%%%%%%
-%SAMPLE POWER SPECTRA WITH THEORY SPECTRA
-% want to show that we recover the PS, including theory PS in background
-%%%%%%%%%%%
+\section{Results}
+\label{sec:results}
 
@@
 
-To quantify the distortion caused by extreme digitisation we infer the fractional increase of the original map noise level, $\Delta \sigma / \sigma$.
-We calculate this from the noise dominated angular scales in the power spectra via
+We analyse the results of the simulated observations to quantify the distortion digitisation causes in the power spectra and investigate whether the additional noise has any angular scale dependence.
+
+%This will show us whether extreme digitisation is viable for the observations targeted by CMBS-4.
+We infer the fractional increase of the original map noise level, $\Delta \sigma / \sigma$, to quantify the distortion caused by extreme digitisation.
+We calculate $\Delta \sigma / \sigma$ from the noise dominated angular scales in the power spectra via
 
@@
 
-The deduced additional noise for one, two, and three bit digitisation is shown in table \ref{tab:extranoisewhite} in the appendix.
+The deduced additional noise for one, two, and three bit digitisation is shown in Table \ref{tab:extranoisewhite} in the appendix.
 We see that 3 bit digitisation performs best, followed by two bit and one bit.
 This is expected, since more bits allow a more faithful representation of the input signal.
-We observe that the fractional increase, $\Delta \sigma / \sigma$, is independent of the hits per pixel in the maps and that the compression performs as well for polarisation as is does for temperature spectra.
+We observe that the fractional increase, $\Delta \sigma / \sigma$, is independent of the hits per pixel in the maps and that the compression performs as well for polarisation as is does for temperature data.
 Finally, it is striking how little noise is added.
-On average one bit digitisation leads to a $(24.8\pm 1.1)\%$ increase in the map noise level, two bit yields a $(6.3\pm0.5)\%$ increase and three bit adds only $(1.7\pm0.2)\%$.
-This is impressive, given that all schemes considered reduce TOD volume by at least an order of magnitude.
+On average one bit digitisation leads to a $(25.1\pm 0.9)\%$ increase in the map noise level, two bit yields a $(6.4\pm0.4)\%$ increase and three bit adds only $(1.8\pm0.2)\%$.
+The uncertainties given are inferred from the standard deviation of rebined points in the $C_\ell^{\mathrm{N}}/C_\ell^{\mathrm{N}}$ noise tail.
+The performance of few bit digitisation is impressive, given that all schemes considered reduce TOD volume by at least an order of magnitude.
 
-To investigate whether the noise added through digitisation has any angular scale dependence we subtract the input maps from simulated observations.
-The power in the difference maps is plotted in Figure \ref{fig:diffpswn}.
+To investigate whether the noise added through digitisation has any angular scale dependence, we subtract the input map from simulated observations.
+The power in the difference map is plotted in Figure \ref{fig:diffpswn}.
 There is no sign of any angular scale dependence.
@@
   \caption[Current ]{
-  Power spectra of reconstructed minus input maps for one-bit (green), two-bit (orange), three-bit (violet) and the control case (black) rebined to $\Delta \ell = 400$.
+  Power spectra of reconstructed maps minus input maps for one-bit (green), two-bit (orange), three-bit (violet), and the control case (black).
+The spectra are bined to $\Delta \ell = 200$.
+Results for a simulation assuming white detector noise with $1,024,000$ hits per map pixel are shown.
+The uncertainties assigned underestimate the variations within the noise in the temperature spectrum slightly.
 No angular scale dependence is observed for all depths of observation.
-A simulation with $1,024,000$ hits per map pixel is shown.
 \label{fig:diffpswn}
@@
 While digitisation clearly works extremely well for white noise, real ground-based CMB experiments face several sources of low-frequency noise.
-High on the list is atmospheric noise, which may me described by \citep{lay2000}:
+High on the list is atmospheric noise, which \cite{lay2000} describe as
 
 \begin{equation} \label{eq:1/fnoiseform}
-\begin{aligned}
-\left| N_{1/f}(\ell) \right|^2 &= \left| \xi \left(0, \sigma \right) \right|^2\\
-&\times \left[ 1 + \left( \frac{\ell}{\lknee} \right)^{-3/2} \right]^2  exp\left[ -\left( \frac{\lcut}{\ell} \right)^{12} \right],
-\end{aligned}
+\left| N_{1/f}(\ell) \right|^2 = \left| N_{\mathrm{white}} \right|^2 \left[ 1 + \left( \frac{\ell}{\lknee} \right)^{-3/2} \right]^2,
 \end{equation}
 
-where $\xi \left(0, \sigma \right)$ denotes a normally distributed random variable with mean $0$ and standard deviation $\sigma$.
-The first term on the second line amplifies modes corresponding to large angular scales and low frequencies \citep{tegmarkdesignersguide1997}.
-The second term on the same line suppresses the lowest modes.
-For temperature we choose $\lknee = 1000$ and $\lcut = 336$ and for polarisation $\lknee = 50$ and $\lcut = 17$.
+%  exp\left[ -\left( \frac{\lcut}{\ell} \right)^{12} \right]
+
+where $N_{\mathrm{white}}$ denotes white detector noise.
+In this expression $\ell$ describes the multipole moment of angular features of corresponding size along the scan direction.
+The scale at which such modes are amplified appreciably by the second term in equation \ref{eq:1/fnoiseform} is set by $\lknee$.
+We choose $\lknee = 1000$ for temperature and $\lknee = 50$ for polarisation.
+In practice CMB experiments suppress power in the lowest modes through a range of techniques, including fitting polynomials, sines, and cosines to the TOD.
+We simulate successful removal of these modes by applying a high pass filter to the noise, such that
+
+\begin{equation} \label{eq:1/fnoiseformfinal}
+\left| N (\ell) \right|^2 = \left| N_{1/f}(\ell) \right|^2 exp\left[ -\left( \frac{\lcut}{\ell} \right)^{12} \right],
+\end{equation}
+
+with $\lcut = 336.3$ for temperature and $\lcut = 16.8$ for polarisation.
 
@@
 
-\begin{equation}
-\sigma^2 = \int_0^\infty \left| N_{1/f}(\ell) \right|^2 d\ell.
+\begin{equation} \label{eq:psvl1/f}
+\sigma^2 = \int_0^\infty \left| N(\ell) \right|^2 d\ell.
 \end{equation}
 
-We carry out the same simulations as described in subsection \ref{subsec:method} using the above noise profile and digitisation schemes for maps with $8,000$, $102,400$ and $1,024,000$ hits per pixel.
-We use HEALPix to calculate the spherical harmonic coefficients $a\ell m$ and eliminate modes with $m < \lknee$ before calculating the power spectra.
-Because the scan strategy of the simulated observations is similar to the SPT's, smaller values of $m$ correspond to larger angular features along the scan direction \citep{chown2018}.
-Because of the anisotropic nature of the noise, equally weighting all modes leads to inaccurate power spectra.
-Discounting the lowest modes in scan direction is the first order approximation to optimal weighting.
+After modifying the digitisation schemes derived in subsection \S\ref{subsec:extremedigitisation} accordingly, we carry out the same simulations as described in subsection \S\ref{subsec:method} with the noise profile given in equation \ref{eq:1/fnoiseformfinal}.
+We save the output maps at $8,000$, $102,400$ and $1,024,000$ hits per pixel.
 The analysis procedure of subsection \S\ref{subsec:whitenoise} is used to characterise the digitisation noise.
@@
 We recover similar results to the white noise case, with three bit digitisation performing the best, followed by two bit, and one bit.
-Temperature and polarisation observations incur the same noise penalty.
 Table \ref{tab:extranoiseoof} in the appendix displays the fractional increase to the detector noise level due to digitisation, $\Delta \sigma / \sigma$.
+While polarisation observations seem to incur a slightly higher noise penalty compared to temperature, deviations do not exceed $3\sigma$.
 On average one bit digitisation leads to a $(27.7\pm 1.5)\%$ increase in the map noise level, two bit adds $(6.2\pm0.5)\%$, and three bit leads to an increase of only $(1.7\pm0.2)\%$.
 The digitisation noise shows no angular scale dependence.
+Thus the effectiveness of few bit digitisation is not limited to the white noise case, but extents to more realistic noise profiles.
 
@@
 In particular the reduction of TOD volume by an order of magnitude decreases the transmission requirements from remote locations.
-The reduction may also streamline TOD simulations.
+The reduction may also streamline the analysis and simulation of TOD.
 
 We present a set of one, two, and three bit digitisation schemes.
-In both, the white noise and $1/f$ noise cases, we find that optimal three bit digitisation adds as little as $1.7\%$ to the map noise level for temperature and polarisation observations alike.
+For white and $1/f$ detector noise alike, we find that optimal three bit digitisation adds as little as $<2\%$ to the map noise level.
+This is true for temperature and polarisation observations.
 No change in the results is observed for maps of different hits per pixel and no angular scale dependence was observed in the added noise.
@@
 
-Future work should focus on finding the higher statistical moments of the digitisation noise.
-Using the digitised TOD for cluster finding can help determine the skewness and kurtosis of the additional noise.
+Given that the digitisation noise remains low at small angular scales, cluster finding algorithms may perform well on few bit TOD.
+Investigating this will also help determine the higher order statistical moments, i.e.
+skewness and kurtosis, of the digitisation noise.
 
@@
 
-We thank the \changed{referee as well as} Srinivasan Raghunathan and Federico Bianchini for valuable feedback on the manuscript. 
+We are grateful for insightful discussions about the prospects of extreme digitisation for CMB experiments with Andrew Melatos and Patrick Clearwater.
+We thank the \changed{referee as well as} Srinivasan Raghunathan, Federico Bianchini, Nikhel Gupta, Andrew Melatos, and Patrick Clearwater for valuable feedback on the manuscript.
 We acknowledge support from an Australian Research Council Future Fellowship (FT150100074), and also from the University of Melbourne. 
@@
 \begin{figure*}[h!t]\centering
-\vspace{-1cm}
+\vspace{-1.5cm}
 \includegraphics[width=0.8\textwidth,clip]{Plots/psrecovery.pdf}
   \caption[Current ]{
-  Input power spectra used to generate the CMB template maps plus detector noise level (black) and digitisation noise (gray) and power spectra using three (violet), two (orange), and one (green) bit TOD.
-Spectra are rebined to $\Delta \ell = 100$.
-Main panels are the calculated power spectra binned to $\Delta \ell = 100$, accompanying panels show the difference between the corresponding input plus noise curves and the $C_\ell$ of digitised TOD.
-Variations occur because only observe a single CMB realisation.
-A simulations with $1,024,000$ hits per map pixel is shown in this plot.
+     Input power spectra used to generate the CMB template map plus detector noise level (black) and digitisation noise (gray); power spectra using three (violet), two (orange), and one (green) bit TOD.
+Main panels show the TT, EE, and BB power spectra; accompanying panels show the residual of the input plus noise curves and the power spectra recovered from few bit TOD.
+We assume white detector noise and show results for maps with $1,024,000$ hits per pixel.
+Spectra are bined to $\Delta \ell = 100$.
+Variations are attributed to observing a single CMB realisation and residual boundary effects.
+Even though small angular scales have small errors in the temperature power spectrum, no significant trend in the residual is found.
 No qualitative differences are visible for other depths of observation.
@@
 \hline
-\multirow{3}{*}{$800$}  & TT  & $ 0.252 \pm 0.009 $  & $ 0.064 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
+\multirow{3}{*}{800}  & TT  & $ 0.252 \pm 0.009 $  & $ 0.064 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
 & EE  & $ 0.253 \pm 0.012 $  & $ 0.064 \pm 0.005 $  & $ 0.017 \pm 0.004 $ \\
@@
 \hline
-\multirow{3}{*}{$8,000$}  & TT  & $ 0.252 \pm 0.008 $  & $ 0.064 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
+\multirow{3}{*}{8,000}  & TT  & $ 0.252 \pm 0.008 $  & $ 0.064 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
 & EE  & $ 0.252 \pm 0.007 $  & $ 0.064 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
@@
 \hline
-\multirow{3}{*}{$80,000$}  & TT  & $ 0.249 \pm 0.011 $  & $ 0.064 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
-& EE  & $ 0.250 \pm 0.011 $  & $ 0.064 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
-& BB  & $ 0.248 \pm 0.018 $  & $ 0.063 \pm 0.009 $  & $ 0.018 \pm 0.003 $ \\
+\multirow{3}{*}{80,000}  & TT  & $ 0.253 \pm 0.008 $  & $ 0.065 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
+& EE  & $ 0.251 \pm 0.009 $  & $ 0.064 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
+& BB  & $ 0.250 \pm 0.015 $  & $ 0.064 \pm 0.007 $  & $ 0.018 \pm 0.003 $ \\
 \hline
-\multirow{3}{*}{$1,024,000$}  & TT  & $ 0.250 \pm 0.009 $  & $ 0.063 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
+\multirow{3}{*}{1,024,000}  & TT  & $ 0.250 \pm 0.009 $  & $ 0.063 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
 & EE  & $ 0.252 \pm 0.008 $  & $ 0.065 \pm 0.003 $  & $ 0.018 \pm 0.002 $ \\
@@
 \hline
-\multirow{3}{*}{$10,240,000$}  & TT  & $ 0.186 \pm 0.025 $  & $ 0.046 \pm 0.007 $  & $ 0.013 \pm 0.002 $ \\
-& EE  & $ 0.253 \pm 0.009 $  & $ 0.065 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
-& BB  & $ 0.252 \pm 0.012 $  & $ 0.064 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
+\multirow{3}{*}{10,240,000}  & TT  & $ 0.248 \pm 0.008 $  & $ 0.063 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
+& EE  & $ 0.253 \pm 0.008 $  & $ 0.065 \pm 0.003 $  & $ 0.018 \pm 0.002 $ \\
+& BB  & $ 0.252 \pm 0.010 $  & $ 0.064 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
 \hline
-\multirow{3}{*}{$102,400,000$}  & TT  & $ 0.246 \pm 0.005 $  & $ 0.063 \pm 0.002 $  & $ 0.017 \pm 0.001 $ \\
+\multirow{3}{*}{102,400,000}  & TT  & $ 0.246 \pm 0.005 $  & $ 0.063 \pm 0.002 $  & $ 0.017 \pm 0.001 $ \\
 & EE  & $ 0.251 \pm 0.008 $  & $ 0.064 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
@@
 \hline
-\multicolumn{2}{c}{Average}  & $ 0.248 \pm 0.011 $  & $ 0.063 \pm 0.005 $  & $ 0.017 \pm 0.002 $ \\
+\multirow{3}{*}{Average}  & TT  & $ 0.25 \pm 0.008 $  & $ 0.064 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
+ & EE  & $ 0.252 \pm 0.009 $  & $ 0.064 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
+ & BB  & $ 0.252 \pm 0.012 $  & $ 0.064 \pm 0.005 $  & $ 0.018 \pm 0.003 $ \\
 \end{tabular}
@@
 Percent addition to the map noise level, $\Delta \sigma / \sigma$, due to one, two, and three bit digitisation for white detector noise.
-We notice that there is no appreciable variation with depth of observations and channel for a given digitisation scheme.
+We use the standard deviation of points in the $C_\ell^\mathrm{X}/C_\ell^\mathrm{N}$ noise tail as the uncertainty.
+There is no appreciable variation with depth of observations and channel for a given digitisation scheme.
 } \normalsize
@@
 \hline
-\multirow{3}{*}{$8,000$}  & TT  & $ 0.231 \pm 0.017 $  & $ 0.058 \pm 0.005 $  & $ 0.016 \pm 0.002 $ \\
+\multirow{3}{*}{8000}  & TT  & $ 0.231 \pm 0.017 $  & $ 0.058 \pm 0.005 $  & $ 0.016 \pm 0.002 $ \\
 & EE  & $ 0.300 \pm 0.012 $  & $ 0.064 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
@@
 \hline
-\multirow{3}{*}{$102,400$}  & TT  & $ 0.235 \pm 0.014 $  & $ 0.060 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
+\multirow{3}{*}{102400}  & TT  & $ 0.235 \pm 0.014 $  & $ 0.060 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
 & EE  & $ 0.299 \pm 0.009 $  & $ 0.064 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
@@
 \hline
-\multirow{3}{*}{$1,024,000$}  & TT  & $ 0.238 \pm 0.013 $  & $ 0.06 \pm 0.005 $  & $ 0.017 \pm 0.002 $ \\
+\multirow{3}{*}{1024000}  & TT  & $ 0.238 \pm 0.013 $  & $ 0.06 \pm 0.005 $  & $ 0.017 \pm 0.002 $ \\
 & EE  & $ 0.299 \pm 0.008 $  & $ 0.063 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
-& BB  & $ 0.299 \pm 0.010 $  & $ 0.064 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
+& BB  & $ 0.299 \pm 0.01 $  & $ 0.064 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
 \hline
-\multicolumn{2}{c}{Average}  & $ 0.277 \pm 0.015 $  & $ 0.062 \pm 0.005 $  & $ 0.017 \pm 0.002 $ \\
+\multirow{3}{*}{Average}  & TT  & $ 0.235 \pm 0.015 $  & $ 0.059 \pm 0.005 $  & $ 0.016 \pm 0.002 $ \\
+ & EE  & $ 0.299 \pm 0.010 $  & $ 0.064 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
+ & BB  & $ 0.297 \pm 0.019 $  & $ 0.063 \pm 0.006 $  & $ 0.017 \pm 0.003 $ \\
 \end{tabular}
@@
 Percent addition to the map noise level, $\Delta \sigma / \sigma$, due to one, two, and three bit digitisation for $1/f$ detector noise.
+Uncertainties are obtained in the same way as for white noise.
 As in the white noise case we observe no significant trend with the numbers of hits per map pixel or channel for a given digitisation scheme.
@@
 
-As mentioned in subsection \S\ref{subsec:extremedigitisation}, equation \ref{eq:distdef} did not require signal power to be conserved when constructing a digitisation scheme.
-It is relatively simple to add this requirement however, since it boils down to rescaling the digitised output by a constant, $\gamma$.
-Specifically we will derive an expression for $\gamma$, such that a cross-spectrum of two digitised maps yields an unbiased estimate of the input power spectrum.
-Mathematically, this is equivalent to demanding:
+As mentioned in subsection \S\ref{subsec:extremedigitisation}, equation \ref{eq:distdef} does not demand not signal power conservation in constructing a digitisation scheme.
+It is relatively simple to impose power conservation, since it boils down to rescaling the digitised output by a constant, $\gamma$.
+We derive an expression for $\gamma$, such that a cross-spectrum of two digitised maps yields an unbiased estimate of the input power spectrum.
+Mathematically, this is equivalent to:
 
@@
 The covariance of two timestreams must be zero.
-This still holds true once the signals have been digitised, as this only discretises the probability distributions at play but does not change the underlying dynamics.
+This remains true for the digitised signals, as digitisation discretises the probability distributions at play but does not change the underlying dynamics.
 Therefore
@@
 
-where $P$ denotes the probability of a certain output level occurring.
-For a small signal $D(\mu + \xi)=D(\xi)$.
-The probabilities $P$ can then be expressed as integrals over probability distributions.
-Assuming Gaussian detector noise we introduce normal distributions $p(\xi)$ with mean $\mu$ and standard deviation $\sigma$.
+where we sum over all output levels of both timestreams $(1, 2)$ and $P$ denotes the probability of an output level occurring.
+The probabilities $P$ can be expressed as integrals over probability distributions.
+Assuming Gaussian detector noise we introduce normal distributions $p(\xi)$ with mean zero and standard deviation $\sigma$.
 
 \begin{equation}
-\langle D(\mu + \xi_1) D(\mu + \xi_2) \rangle = \sum_{i=0}^N \sum_{j=0}^N  \int_{x_i}^{x_{i+1}} \int_{x_j}^{x_{j+1}} y_i y_j p(\xi_1) p(\xi_2) d\xi_1 d\xi_2.
+\langle D(\mu + \xi_1) D(\mu + \xi_2) \rangle = \sum_{i=0}^N \sum_{j=0}^N  \int_{x_i-\mu}^{x_{i+1}-\mu} \int_{x_j-\mu}^{x_{j+1}-\mu} y_i y_j p(\xi_1) p(\xi_2) d\xi_1 d\xi_2.
 \end{equation}
@@
 
-Where we have used the Maclaurin series of the error function and the binomial expansion. Combining all possible sums and moving the $i, j$ summation forwards yields
+Where we have used the Maclaurin series expansion of the error function and the binomial expansion. Moving the $i, j$ summation forwards yields
 
@@
 \begin{equation} \label{eq:ijsumpsiflip}
-\sum_{i,j = 0}^N y_i y_j \psi_i^{k, n} \psi_j^{l, m} = \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} y_i y_j \left( \psi_i^{k, n} \psi_j^{l, m} + (-1)^{k+1} \psi_{i}^{k, n} \psi_j^{l, m} + (-1)^{l+1} \psi_i^{k, n} \psi_{j}^{l, m} + (-1)^{k+l} \psi_{N-i}^{k, n} \psi_{N-j}^{l, m} \right).
+\sum_{i,j = 0}^N y_i y_j \psi_i^{k, n} \psi_j^{l, m} = \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} y_i y_j \left( \psi_i^{k, n} \psi_j^{l, m} + (-1)^{k+1} \psi_{i}^{k, n} \psi_j^{l, m} + (-1)^{l+1} \psi_i^{k, n} \psi_{j}^{l, m} + (-1)^{k+l} \psi_{N-i}^{k, n} \psi_{N-j}^{l, m} \right),
 \end{equation}
@@
 since $\psi_{N-i}^{k,n} = (-1)^{k} \psi_{i}^{k,n}$.
-Since we assume that $\mu$ is small we would like to calculate the leading order terms in the above.
-Since we have power of $\mu^{k+l}$ in equation \ref{eq:dcompleteklsum} we focus on the case of even $k$ and $l$ first.
-This reduces equation \ref{eq:ijsumpsiflip} to
+For ground-based CMB observations $\mu$ is small.
+Therefore we calculate the leading order terms in $\mu$ in equation \ref{eq:dcompleteklsum}.
+Since this equation has powers of $\mu^{k+l}$ we split the sum over $k, l$ as follows
 
@@
 
-The term vanishes. For odd $k$ and $l$ equation \ref{eq:ijsumpsiflip} may be simplified to
+For odd $k$ and $l$ equation \ref{eq:ijsumpsiflip} may be simplified to
 
@@
 
-where $c$ is some constants that depend on the specific digitisation scheme chosen.
-We have proved that in the small signal limit and for Gaussian noise normalisation can be achieved through rescaling by a constant.
-Returning to equation \ref{eq:normcrosspower} we see
+where $c$ is some constants that depend on the specific digitisation scheme chosen. Returning to equation \ref{eq:normcrosspower} we see
 
@@
 
-where we have ignored higher order contributions. From equation \ref{eq:dcompleteklsum} we obtain an expression for $\gamma$, by investigating $k=l=1$ terms.
+where we have ignored terms beyond quadratic order. By investigating $k=l=1$ terms in equation \ref{eq:dcompleteklsum} we obtain an expression for $\gamma$.
 
@@
 \begin{equation}
-c = \sum_{i,j=0}^N \frac{y_i y_j}{2\pi\sigma^2} \left[ \sum_{n = 0}^\infty \frac{(-1)^n}{n!} \frac{1}{(\sqrt{2}\sigma)^{2n}}  ( x_{i+1}^{2n} - x_{i}^{2n} ) \right] \left[ \sum_{m = 0}^\infty \frac{(-1)^m}{m!} \frac{1}{(\sqrt{2}\sigma)^{2m}} ( x_{j+1}^{2m} - x_{j}^{2m} ) \right].
+= \sum_{i,j=0}^N \frac{y_i y_j}{2\pi\sigma^2} \left[ \sum_{n = 0}^\infty \frac{(-1)^n}{n!} \frac{1}{(\sqrt{2}\sigma)^{2n}}  ( x_{i+1}^{2n} - x_{i}^{2n} ) \right] \left[ \sum_{m = 0}^\infty \frac{(-1)^m}{m!} \frac{1}{(\sqrt{2}\sigma)^{2m}} ( x_{j+1}^{2m} - x_{j}^{2m} ) \right].
 \end{equation}
@@
 \begin{equation}
-c = \sum_{i,j=0}^N \frac{y_i y_j}{2\pi\sigma^2} \left\{ exp \left[ - \left(\frac{x_{i+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{i}}{\sqrt{2}\sigma} \right)^2 \right] \right\} \left\{ exp \left[ - \left(\frac{x_{j+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{j}}{\sqrt{2}\sigma} \right)^2 \right] \right\}
+= \sum_{i,j=0}^N \frac{y_i y_j}{2\pi\sigma^2} \left\{ exp \left[ - \left(\frac{x_{i+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{i}}{\sqrt{2}\sigma} \right)^2 \right] \right\} \left\{ exp \left[ - \left(\frac{x_{j+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{j}}{\sqrt{2}\sigma} \right)^2 \right] \right\}
 \end{equation}
@@
 
-\newpage
+%\newpage
 

### 2018-12-20 response letter/response_letter.tex
Updated paper according to referee's report and included response to the reviewer
Included section on few-bit digitisation together with lossless compression (FLAC/LZMA). Edited introdcution/formulation of extreme digitisation. Added insight about central limit theorem.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass{article}
%\documentclass[preprint2]{aastex61}
%\documentclass[12pt,preprint]{aastex}

\usepackage[table,usenames,dvipsnames]{xcolor}
\usepackage{amsmath}
\usepackage{subfigure}
\usepackage[backref,breaklinks,colorlinks,citecolor=blue]{hyperref}
\usepackage{natbib}
%\usepackage{natbib}
\bibliographystyle{fapj}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{soul}

%\newcommand{\jcap}{JCAP}

\newcommand{\sqdeg}{deg$^2$ }
\newcommand{\omb}{\ensuremath{\Omega_b h^2}}
\newcommand{\omc}{\ensuremath{\Omega_c h^2}}
\newcommand{\clpp}{\ensuremath{C_{L}^{\phi\phi}}}
\newcommand{\cpmf}{\ensuremath{C_{\ell}^{\rm PMF}}}

\newcommand{\cpmftens}{\ensuremath{C_{\ell}^{\rm PMF,\,tens}}}
\newcommand{\cpmfvec}{\ensuremath{C_{\ell}^{\rm PMF,\,vec}}}
\newcommand{\apmf}{\ensuremath{A_{\rm PMF}}}
\newcommand{\bpmf}{\ensuremath{B_{\rm 1\,Mpc}}}
\newcommand{\alens}{\ensuremath{A_{\rm lens}}}
\newcommand{\lcdm}{\ensuremath{\Lambda}CDM}
\newcommand{\nrun}{\ensuremath{n_{\rm run}}}
\newcommand{\neff}{\ensuremath{N_{\rm eff}}}
\newcommand{\ho}{H\ensuremath{_0}}
\newcommand{\mnu}{\ensuremath{\sum m_\nu}}
\newcommand{\ukarcmin}{\ensuremath{\mu}{\rm K-arcmin}}
\newcommand{\lknee}{\ensuremath{\ell_{\rm knee}}}
\newcommand{\lcut}{\ensuremath{\ell_{\rm min}}}
\newcommand{\fermilat}{\textit{Fermi}-LAT}

\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\planck}{{\sl Planck}}
\newcommand{\wmap}{{\sl WMAP}}
\newcommand{\bicepkeck}{BICEP2/Keck Array}
\newcommand{\sptnew}{SPT-3G }
\newcommand{\pb}{\textsc{Polarbear}}
\newcommand{\simons}{Simons Array}
\newcommand{\sptpol}{SPTpol}
\newcommand{\advactpol}{Adv.~ACTpol }

\newcommand{\tbd}[1]{\textcolor{Red}{{\bf TBD}: #1}}
\newcommand{\gab}[1]{\textcolor{Orchid}{[{\bf GS}: #1]}}
\newcommand{\changed}[1]{\textcolor{Red}{#1}}
\newcommand{\removed}[1]{\textcolor{Red}{}}
\newcommand{\question}[1]{\textcolor{Blue}{#1}}
\newcommand{\CR}[1]{\textit{\textcolor{Blue}{#1}}}
\newcommand{\CRremove}[1]{\st{#1}}
\include{number_list}

%

\usepackage{fullpage}
\newlength\tindent
\setlength{\tindent}{\parindent}
\setlength{\parindent}{0pt}
\renewcommand{\indent}{\hspace*{\tindent}}

% ref to section \S\ref{sec:label}

%\submitjournal{ApJ}
\def\Melbourne{1}
\def\uci{2}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%\title{MNRAS: MN-18-3671-MJ\\Response to the Referee}

%\maketitle

\begin{center}
\Large MNRAS: MN-18-3671-MJ\\Response to the Referee
\end{center}

Thank you for your comments. We are pleased that you found our manuscript to be a valuable enquiry into data compression techniques for CMB experiments and a detailed investigation of the effects of few-bit digitisation on the CMB power spectra. We found your suggestion of looking at the compressive power of few-bit digitisation together with established lossless compression techniques particularly insightful and exciting to research. Please find a detailed response to your report below. Shown in blue are remarks in the reviewer's report. The authors' reply is given in black and if applicable relevant sections of the manuscript that have been changed in response are appended in red.

\vspace{0.5cm}

\hrule
\vspace{0.1cm}
\hrule

\vspace{0.5cm}

Reviewer's comment:

\indent \question{[...] the article's case is weakened considerably by a lack of comparison to existing lossless approaches to data compression used in the field. [...]}

\vspace{0.25cm}

Authors' reply:

\indent The referee raises a fair point. We contacted the BICEP Array, POLARBEAR, and SPT-3G experiments to enquire about the lossless compression algorithms currently being used. BICEP uses a Lempel-Ziv-Markov chain algorithm (LZMA), while the SPT and POLARBEAR use the free lossless audio codec (FLAC). We applied few-bit digitisation to a simulated timestream and then applied each compression scheme to the few bit timestreams. 
We compared this to the size of files when the two compression schemes were applied to the original 64bit timestreams. 
We found that while using either LZMA or FLAC reduces the compressive power of extreme digitisation, substantial data volume savings are still achieved. We appended the following subsection to section 3 (Results) of the manuscript to include this insight:

\vspace{0.25cm}

Changes to the manuscript:

\indent \textbf{\changed{3.3 Existing Compression Techniques}}

\changed{Current-generation ground-based CMB experiments employ lossless and lossy compression techniques to manage their transmission bottlenecks. To test whether few-bit digitisation is competitive to existing lossy compression techniques, it is necessary to see if it still provides a substantial reduction in data volume if used in conjunction with lossless compression algorithms.}

\changed{We focus on lossless compression via the Lempel-Ziv-Markoc chain algorithm (LZMA), which is used by the Background Imaging of Cosmic Extragalactic Polarization (BICEP) experiment, and the free lossless audio codec (FLAC), which the SPT and POLARBEAR experiments employ (private communication). We examine the compression of a single CES of our simulation, consisting of ca. $10^7$ data points, by applying LZMA and FLAC separately to four timestreams: the original 64-bit TOD, and copies that have undergone 1-, 2-, and 3-bit digitisation.}

\changed{We find that even though the data reduction power of few-bit digitisation is decreased when used in conjunction with lossless compression algorithms, substantial savings are still achieved. Applying few-bit digitisation to the TOD prior to compression with FLAC reduces the final file size by a factor $5.89$ compared to the FLAC compressed 64-bit data. This is the same for all digitisation schemes considered. Working with LZMA, the data volume of the TOD can be reduced with respect to compressed 64-bit data by factors of $17.34$, $26.23$, and $44.48$ for 3-, 2-, and 1-bit digitisation, respectively. Few-bit digitisation still achieves considerable data reduction if used together with established lossless compression algorithms.}

\vspace{0.5cm}

\hrule

\vspace{0.5cm}

Reviewer's comment:

\indent \question{It is unclear from the text what, exactly, the resulting scheme is after equation 13. I found equations 7-9 extremely helpful to understand the approach taken earlier and it would be very helpful to see the analogs (is it just replacing sigma in equations 7-9?)}

\vspace{0.25cm}

Authors' reply:

\indent Yes, it is simply replacing $\sigma$ in equations 7-9. We have changed the wording of the relevant section to make this more clear.

\vspace{0.25cm}

Changes to the manuscript:

\indent \changed{We use the same digitisation schemes as for white noise, but replace the parameter $\sigma$ in equations \ref{eq:1bit}, \ref{eq:2bit} and \ref{eq:3bit} by considering Parseval's theorem on the above noise profile. The appropriate value is obtained by numerically evaluating the integral}

\[ \changed{
\sigma^2 = \int_0^\infty \left| N(\ell) \right|^2 d\ell. }
\]

\vspace{0.5cm}

\hrule

\vspace{0.5cm}

Reviewer's comment:

\indent \question{The last sentence before section 4 has a typo ("extents" rather than "extends"). There are many more such in the appendix, which needs a bit more careful editing.}

\vspace{0.25cm}

Authors' reply:

\indent Thank you for pointing this out. We have fixed this typo and edited the appendix.

\vspace{0.5cm}

\hrule

\vspace{0.5cm}

Reviewer's comment:

\indent \question{In many cases, the authors dive directly into mathematical formalism without explaining the results. For example, it took me several readings before realizing that I could think of the scheme in equations 7-9, to a reasonably high degree of accuracy, as just dropping the high-order 32-n bits of the data. The text would be improved somewhat by a bit of unpacking along these lines, though it is acceptable as-is.}

\vspace{0.25cm}

Authors' reply:

\indent We recognise that the derivation of the digitisation schemes was not as clear as possible. We have edited subsection 2.1 (Extreme Digitisation) to be more readily understood.

\end{document}

### 2018-12-20 digitisation.tex
Updated paper according to referee's report and included response to the reviewer
Included section on few-bit digitisation together with lossless compression (FLAC/LZMA). Edited introdcution/formulation of extreme digitisation. Added insight about central limit theorem.

@@
 \altaffiltext{\Melbourne}{School of Physics, University of Melbourne, Parkville, VIC 3010, Australia}
-\email{christian.reichardt@unimelb.edu.au}
+\email{lbalkenhol@student.unimelb.edu.au}
 
@@
 
-The large size of the time ordered data (TOD) of cosmic microwave background (CMB) experiments presents challenges for mission planning and data analysis.
+The large size of the time ordered data of cosmic microwave background experiments presents challenges for mission planning and data analysis.
 These issues are particularly significant for Antarctica- and space-based experiments, which depend on satellite links to transmit data.
-We explore the viability of reducing the TOD to few bit numbers to address these challenges.
+We explore the viability of reducing the time ordered data to few bit numbers to address these challenges.
 Unlike lossless compression, few bit digitisation introduces additional noise into the data.
-We present a set of one, two, and three bit digitisation schemes and measure increase in noise in the CMB temperature and polarisation power spectra.
+We present a set of one, two, and three bit digitisation schemes and measure \changed{the} increase in noise in the cosmic microwave background temperature and polarisation power spectra.
 The digitisation noise is independent of angular scale and is well-described as a constant percentage of the original detector noise.
-Three bit digitisation increases the map noise level by $< 2\%$, while reducing the data volume by a factor of ten relative to 32 bit floats.
-Such digitisation is a promising strategy for upcoming experiments.
+Three bit digitisation increases the map noise level by $< 2\%$, while reducing the data volume by a factor of \changed{twenty relative to 64}-bit floats.
+\changed{Comparable gains are seen when extreme digitisation is combined with
+established lossless compression algorithms. Extreme} digitisation is a promising strategy for upcoming experiments.
 
@@
 Current ground-based CMB experiments, like \sptnew \citep{benson2014} at the South Pole telescope (SPT) and \advactpol \citep{thornton2016}, at the Atacama cosmology telescope (ACT), target science goals such as the discovery of inflationary gravitational waves, measuring the number of relativistic species, the neutrino mass sum, and mapping the large-scale distribution of matter through gravitational lensing and the Sunyaev-Zeldovich (SZ) effects.
-Through lensing and the SZ effects the CMB probes structure formation, reionisation, and is a powerful test for dark matter and dark energy models \citep{s4sciencebook, core2011, litebird2016, pixie2011}.
+Through lensing and the SZ effects the CMB probes structure formation, reionisation, and is a powerful test for dark matter and dark energy models \citep{s4sciencebook, 2018corecos, litebird2016, pixie2011}.
 
@@
 Antarctica-based experiments that transmit a portion of their results via a satellite link downsample their data to meet telemetry allocations, but have not yet used few bit digitisation of the TOD.
-As we approach the next generation ground-based experiment, CMB-S4, and the launch of new space-based missions (LiteBIRD, PIXIE, COrE+), we must consider potential transmission bottlenecks carefully.
+As we approach the next generation ground-based experiment, CMB-S4 \citep{s4sciencebook}, and the launch of the new space-based missions COrE+ \citep{core2018}, LiteBIRD \citep{litebird2016}, and PIXIE \citep{pixie2011}, we must consider potential transmission bottlenecks carefully.
 %C: Is there a reference detailing downsampling of SPT data?
@@
 This work is structured as follows.
-In \S\ref{sec:dig} we detail the challenges that come with handling large data volumes, introduce the process of extreme digitisation and lay out the framework used to test its performance.
+In section \S\ref{sec:dig} we detail the challenges that come with handling large data volumes, introduce the process of extreme digitisation and lay out the framework used to test its performance.
 We present results for white and $1/f$ detector noise in section \S\ref{sec:results}.
@@
 
-The South Pole is one of the best sites for CMB observations on Earth \citep{kovac2007}.
-CMB-S4 plans to include several telescopes at the South Pole, which will generate a data influx of $\sim\mathcal{O}(10)\mathrm{Tb/d}$.
+The South Pole is one of the best sites for CMB observations on Earth \citep{chamberlin2001, spt2004}.
+CMB-S4 plans to include several telescopes at the South Pole \citep{s4sciencebook, barron2018}, which will generate a data influx of $\sim\mathcal{O}(10)\mathrm{Tb/d}$.
 Transferring this data volume via satellite would be expensive.
@@
 only getting the high frequency data once a year) also introduces risks by delaying when potential faults or issues at high frequency are noticed.
-One can avoid some of these issues by running substantial portions of the analysis at the South Pole, but this comes with its own costs and challenges.
+One can reduce these issues by running substantial portions of the analysis at the South Pole, but this comes with its own costs and challenges.
 
@@
 
-The Planck mission has already demonstrated the success of extreme digitisation for space-based CMB experiment s\citep{maris2003}.
+The Planck mission has already demonstrated the success of extreme digitisation for space-based CMB \changed{experiments} \citep{maris2003}.
 \cite{jenet1998} explored the application of few bit digitisation to radio pulsar timing measurements.
@@
 In order to quantify the performance of a given digitisation scheme we define the distortion as
-
 \begin{equation}\label{eq:distdef}
@@
 \end{equation}
-
-where $s$ is the input and $\hat{s}$ the output signal. For a stochastic input signal we can calculate an amplitude probability distribution $p(x)$:
+where $s$ is the input and $\hat{s}$ the output signal.
+For a stochastic input signal we can calculate an amplitude probability distribution $p(x)$.
+\changed{This allows us to re-express the distortion as a sum over the digitisation levels:}
 
@@
 
-Since we wish to minimise the distortion, we set the derivatives with respect to $x_i$ and $y_i$ to zero. Starting with $x_i$ we obtain
+Since we wish to minimise the distortion, we set \changed{its derivatives} with respect to $x_i$ and $y_i$ to zero. Starting with $x_i$ we obtain
 
@@
 \end{equation}
-
 which informs us that the threshold levels should lie midway between their adjacent output levels.
 Setting the derivative of equation \ref{eq:dist} with respect to $y_i$  to zero gives the additional condition
-
 \begin{equation} \label{eq:distderiv2}
@@
 \end{equation}
-
-This implies that we should choose $y_i$, such that it halves the area underneath $p(x)$ in the interval from $x_i$ to $x_{i+1}$.
+This \changed{condition} implies that we should choose $y_i$, such that it halves the area underneath $p(x)$ in the interval from $x_i$ to $x_{i+1}$.
 
 To progress further, we must consider the probability distribution $p(x)$ of the input signal. 
-For Gaussian detector noise, we have $p(x) = (1/\sqrt{2\pi\sigma^2}) e^{-x^2/2\sigma^2}$.
-We also assume that we are in the low signal to noise regime, as is appropriate for ground-based CMB experiments.
-The solution for this case was given by \cite{max1960}.
-For one bit digitisation, the result is
+\changed{In this work, we are interested in the compression of the TOD from each detector of a ground-based CMB experiment. 
+Thus we can safely assume that the input signal is in the low signal-to-noise regime. 
+Furthermore, we take the simplest noise profile: Gaussian, white detector noise\footnote{\changed{As discussed in the appendix, this treatment also applies to more realistic noise profiles, because the large number of points lying in each map pixel allows the application of the central limit theorem.}}.
+Together these assumptions imply that the input, $x$, is normally distributed, with} $p(x) = (1/\sqrt{2\pi\sigma^2}) e^{-x^2/2\sigma^2}$.
+
+\changed{\citet{max1960} solved Equations \ref{eq:digitequalspacecondition} and \ref{eq:digitareacondition} for 1-, 2- and 3-bit digitisation for this choice of probability distribution, $p(x)$. 
+We quote their results below and refer the reader to the aforementioned work for the derivation.  
+As one would expect given the symmetry in $p(x)$, the 1-bit (N=2 output levels) case is simply:
 \begin{equation} \label{eq:1bit}
@@
 -1, & s(t) \leq 0
-\end{array} \right. \end{equation}
+\end{array} \right. \hspace{0.5cm}.  \end{equation}
+Essentially, the optimal 1-bit digitisation scheme applies the sign function to the TOD. 
+}
 
-For two bit digitisation we have
+\changed{
+For two bit digitisation, we can encode $N=4$ output levels. The optimal set of thresholds and digitisation levels is given by the four-level function:}
 \begin{equation}  \label{eq:2bit}
@@
 -1.51 \sigma, & -0.98 \sigma < s(t)\\
-\end{array} \right. . \end{equation}
+\end{array} \right. \hspace{0.5cm}.  \end{equation}
 %C: should i round to equal significant figures here? Unfortunately these are the exact values that i used...
 
-Finally the optimal three bit digitisation is described by the eight-level function
+\changed{Finally, with 3 bits we may specify $N=8$ output levels. The optimal three bit digitisation is described by the eight-level function:}
 \begin{equation}  \label{eq:3bit}
@@
 -2.15 \sigma, & -1.75 \sigma < s(t)\\
-\end{array} \right. . \end{equation}
+\end{array} \right. \hspace{0.5cm}.  \end{equation}
 
-The digitisation schemes listed above do not conserve signal power and need to be re-calibrated to recover the correct power in the map.
-We demonstrate that the schemes laid out by \cite{max1960} can be rescaled to conserve power and derive an analytical expression for the normalisation constant in the appendix.
+\changed{The digitisation schemes above are designed to minimise the distortion of the TOD as defined in Equation \ref{eq:distdef}.
+This does not ensure that the digitised output has the same signal power as the input.
+However, the output of the schemes given above can be rescaled by a constant to achieve power conservation.
+This is demonstrated in the appendix, where we also derive an analytical expression for the normalisation constant.
 We use this analytical rescaling for all results in this work.
+}
 
@@
 the value of a map pixel is the average of all TOD samples lying within that pixel.
-We create four maps of the original 64 bit TOD and one for each of the digitisation schemes.
+We create four maps: one using the original 64 bit TOD and one for each of the digitisation schemes.
 We save the ouput maps at $800$, $8,000$, $80,000$, $1,024,000$, $10,240,000$ and $102,400,000$ hits per pixel.
@@
 
-We rebin the calculated power spectra as mentioned in section \S\ref{sec:results} and assign Fisher matrix error bars \citep{tegmarkdesignersguide1997}.
+We rebin the calculated power spectra as mentioned \changed{when applicable} and assign Fisher matrix error bars \citep{tegmarkdesignersguide1997}.
 
@@
 We observe that the fractional increase, $\Delta \sigma / \sigma$, is independent of the hits per pixel in the maps and that the compression performs as well for polarisation as is does for temperature data.
-Finally, it is striking how little noise is added.
-On average one bit digitisation leads to a $(25.1\pm 0.9)\%$ increase in the map noise level, two bit yields a $(6.4\pm0.4)\%$ increase and three bit adds only $(1.8\pm0.2)\%$.
-The uncertainties given are inferred from the standard deviation of rebined points in the $C_\ell^{\mathrm{N}}/C_\ell^{\mathrm{N}}$ noise tail.
-The performance of few bit digitisation is impressive, given that all schemes considered reduce TOD volume by at least an order of magnitude.
+\changed{Finally, it is striking how little noise is added.
+On average one-bit digitisation leads to a $(25.13\pm 0.94)\%$ increase in the map noise level, two-bit yields a $(6.40\pm0.41)\%$ increase and three-bit adds only $(1.76\pm0.22)\%$ ($1\sigma$ confidence intervals given).
+This is impressive, given that all schemes considered reduce the TOD volume by at least an order of magnitude.}
 
@@
 \begin{figure}[htb]\centering
-\includegraphics[width=0.5\textwidth,clip]{Plots/diffwn.pdf}
+\includegraphics[width=0.5\textwidth,clip]{diffwn.pdf}
   \caption[Current ]{
-  Power spectra of reconstructed maps minus input maps for one-bit (green), two-bit (orange), three-bit (violet), and the control case (black).
-The spectra are bined to $\Delta \ell = 200$.
-Results for a simulation assuming white detector noise with $1,024,000$ hits per map pixel are shown.
-The uncertainties assigned underestimate the variations within the noise in the temperature spectrum slightly.
-No angular scale dependence is observed for all depths of observation.
+  \changed{(FIGURE HAS BEEN CHANGED) TT, EE, and BB power spectra (upper, middle and lowest panel respectively) of difference maps for one-bit (green marker, high), two-bit (orange marker, medium), and three-bit digitisation (violet marker, low).
+The control case using 64-bit TOD is also shown (black marker).
+All power spectra are rebined to $\Delta \ell = 200$.
+The vertical axes are logarithmic.
+The noise increase due to digitisation appears to be independent of angular scale.
+This plot shows results for a simulation with $1,024,000$ hits per map pixel using Gaussian white detector noise.}
 \label{fig:diffpswn}
@@
 While digitisation clearly works extremely well for white noise, real ground-based CMB experiments face several sources of low-frequency noise.
-High on the list is atmospheric noise, which \cite{lay2000} describe as
+\changed{High on the list is atmospheric noise.
+Modelling the atmosphere as a thin sheet dominated by Kolmogorov turbulence, we can describe fluctuations with the noise profile \citep{lay2000}:}
 
+\changed{
 \begin{equation} \label{eq:1/fnoiseform}
-\left| N_{1/f}(\ell) \right|^2 = \left| N_{\mathrm{white}} \right|^2 \left[ 1 + \left( \frac{\ell}{\lknee} \right)^{-3/2} \right]^2,
+\left| N_{1/f}(\ell_{\mathrm{scan}}) \right|^2 = \left| N_{\mathrm{white}} \right|^2 \left[ 1 + \left( \frac{\ell_{\mathrm{scan}}}{\ell_0} \right)^{-3/2} \right]^2,
 \end{equation}
+}
 
@@
 where $N_{\mathrm{white}}$ denotes white detector noise.
-In this expression $\ell$ describes the multipole moment of angular features of corresponding size along the scan direction.
-The scale at which such modes are amplified appreciably by the second term in equation \ref{eq:1/fnoiseform} is set by $\lknee$.
-We choose $\lknee = 1000$ for temperature and $\lknee = 50$ for polarisation.
-In practice CMB experiments suppress power in the lowest modes through a range of techniques, including fitting polynomials, sines, and cosines to the TOD.
+In this expression \changed{$\ell_{\mathrm{scan}}$} describes the multipole moment of angular features of corresponding size along the scan direction.
+The scale at which such modes are amplified appreciably by the second term in equation \ref{eq:1/fnoiseform} is set by \changed{$\ell_0$}.
+\changed{The exponent of $\ell_{\mathrm{scan}}/\ell_0$ is chosen to model observation at a constant, moderate elevation \citep{lay1997}.
+In line with the observed noise profile of the SPT we choose $\ell_0 = 1000$ for temperature and $\ell_0 = 0$ for polarisation; this means that power is doubled at $\lknee \approx 1800$ and $\lknee \approx 90$ for temperature and polarisation, respectively \citep{henning2018}.} In practice CMB experiments suppress power in the lowest modes through a range of techniques, including fitting polynomials, sines, and cosines to the TOD.
 We simulate successful removal of these modes by applying a high pass filter to the noise, such that
 
+\changed{
 \begin{equation} \label{eq:1/fnoiseformfinal}
-\left| N (\ell) \right|^2 = \left| N_{1/f}(\ell) \right|^2 exp\left[ -\left( \frac{\lcut}{\ell} \right)^{12} \right],
+\left| N (\ell_{\mathrm{scan}}) \right|^2 = \left| N_{1/f}(\ell_{\mathrm{scan}}) \right|^2 \exp\left[ -2\left( \frac{\lcut}{\ell_{\mathrm{scan}}} \right)^{6} \right],
 \end{equation}
+}
 
@@
 
-We use the same digitisation schemes as for white noise, but modify the parameter $\sigma$ in equations \ref{eq:1bit}, \ref{eq:2bit} and \ref{eq:3bit} by considering Parseval's theorem on the above noise profile.
-The appropriate value is obtained by evaluating the integral
+We use the same digitisation schemes as for white noise, but \changed{replace} the parameter $\sigma$ in equations \ref{eq:1bit}, \ref{eq:2bit} and \ref{eq:3bit} by considering Parseval's theorem on the above noise profile.
+The appropriate value is obtained by \changed{numerically} evaluating the integral
 
@@
 
+% ------------ maybe dont include this ------------
+%\changed{Even though the noise profile given in Equation \ref{eq:1/fnoiseformfinal} is anisotropic in Fourier space, the same is not true in real space.
+In fact, projecting the chosen noise profile onto the sky using our scan strategy we expect the same noise level in each map pixel.
+Normalisation of the digitised output by a constant is therefore still appropriate.
+As explained in Section \ref{sec:digitconspow}, the central limit theorem ensures that the noise amplitude probability distribution in each pixel is Gaussian; using the normalisation constant derived in Section \ref{sec:digitconspow} is still appropriate.}
+
 We recover similar results to the white noise case, with three bit digitisation performing the best, followed by two bit, and one bit.
@@
 While polarisation observations seem to incur a slightly higher noise penalty compared to temperature, deviations do not exceed $3\sigma$.
-On average one bit digitisation leads to a $(27.7\pm 1.5)\%$ increase in the map noise level, two bit adds $(6.2\pm0.5)\%$, and three bit leads to an increase of only $(1.7\pm0.2)\%$.
-The digitisation noise shows no angular scale dependence.
-Thus the effectiveness of few bit digitisation is not limited to the white noise case, but extents to more realistic noise profiles.
+\changed{On average one-bit digitisation leads to a $(24.4\pm 1.4)\%$ increase in the map noise level, two-bit adds $(6.21\pm0.50)\%$, and three-bit yields an increase of only $(1.71\pm0.21)\%$ ($1\sigma$ confidence intervals given).
+Thus the effectiveness of few bit digitisation is not limited to the white noise case, but \changed{extends} to more realistic noise profiles.}
+
+%\changed{Performing a $t$-test for independent variables, we find that for a given digitisation scheme it is unlikely that the noise increase variates for Gaussian white and low-frequency detector noise are drawn from different underlying distributions (p-values: 0.19, 0.34, and 0.57 for one, two, and three bit, respectively).
+This result reflects the work of the central limit theorem.}
 
@@
 These conditions are satisfied by CMB data and our findings reflect this.
+
+\subsection{\changed{Existing Compression Techniques}}
+
+\changed{Current-generation ground-based CMB experiments employ lossless and lossy compression techniques to manage their transmission bottlenecks.
+To test whether few-bit digitisation is competitive to existing lossy compression techniques, it is necessary to see if it still provides a substantial reduction in data volume if used in conjunction with lossless compression algorithms.}
+
+\changed{We focus on lossless compression via the Lempel-Ziv-Markoc chain algorithm (LZMA), which is used by the Background Imaging of Cosmic Extragalactic Polarization (BICEP) experiment, and the free lossless audio codec (FLAC), which the SPT and POLARBEAR experiments employ (private communication).
+We examine the compression of a single CES of our simulation, consisting of ca.
+$10^7$ data points, by applying LZMA and FLAC separately to four timestreams: the original 64-bit TOD, and copies that have undergone 1-, 2-, and 3-bit digitisation.}
+
+\changed{We find that even though the data reduction power of few-bit digitisation is decreased when used in conjunction with lossless compression algorithms, substantial savings are still achieved.
+Applying few-bit digitisation to the TOD prior to compression with FLAC reduces the final file size by a factor $5.89$ compared to the FLAC compressed 64-bit data.
+This is the same for all digitisation schemes considered.
+Working with LZMA, the data volume of the TOD can be reduced with respect to compressed 64-bit data by factors of $17.34$, $26.23$, and $44.48$ for 3-, 2-, and 1-bit digitisation, respectively.
+Few-bit digitisation still achieves considerable data reduction if used together with established lossless compression algorithms.}
 
@@
 This is true for temperature and polarisation observations.
-No change in the results is observed for maps of different hits per pixel and no angular scale dependence was observed in the added noise.
-
-% restate from first paragraph
-%The additional noise is insignificant for ground-based experiments, which increase their detector count by orders of magnitude between generations.
-Space-based experiments which have their own unique issues may consider this noise penalty problematic.
+No change in the results is observed for maps \changed{with different numbers of} hits per pixel and no angular scale dependence was observed in the added noise.
+\changed{Combining few-bit digitisation with currently used lossless compression codecs reduces its compression power only slightly.
+Applying three-bit digitisation to TOD prior to lossless compression via FLAC or LZMA reduces the data volume by a further factor of $5.89$ or $17.34$, respectively.}
 
@@
 We are grateful for insightful discussions about the prospects of extreme digitisation for CMB experiments with Andrew Melatos and Patrick Clearwater.
-We thank the \changed{referee as well as} Srinivasan Raghunathan, Federico Bianchini, Nikhel Gupta, Andrew Melatos, and Patrick Clearwater for valuable feedback on the manuscript.
+We thank the \changed{referee as well as Nikhel Gupta, Srinivasan Raghunathan, Federico Bianchini, Andrew Melatos, and Patrick Clearwater for valuable feedback on the manuscript.}
 We acknowledge support from an Australian Research Council Future Fellowship (FT150100074), and also from the University of Melbourne. 
@@
 \vspace{-1.5cm}
-\includegraphics[width=0.8\textwidth,clip]{Plots/psrecovery.pdf}
+\includegraphics[width=0.8\textwidth,clip]{psrecovery.pdf}
   \caption[Current ]{
-     Input power spectra used to generate the CMB template map plus detector noise level (black) and digitisation noise (gray); power spectra using three (violet), two (orange), and one (green) bit TOD.
-Main panels show the TT, EE, and BB power spectra; accompanying panels show the residual of the input plus noise curves and the power spectra recovered from few bit TOD.
-We assume white detector noise and show results for maps with $1,024,000$ hits per pixel.
-Spectra are bined to $\Delta \ell = 100$.
-Variations are attributed to observing a single CMB realisation and residual boundary effects.
-Even though small angular scales have small errors in the temperature power spectrum, no significant trend in the residual is found.
-No qualitative differences are visible for other depths of observation.
+  \changed{(FIGURE HAS BEEN CHANGED) Input power spectra used to generate the CMB template map plus detector noise level (black line) and digitisation noise (grey line); power spectra using three- (violet marker, right), two- (orange marker, middle), and one-bit (green marker, left) TOD.
+Power spectra are rescaled to $D_\ell = \ell (\ell+1) C_\ell /2\pi$ and bined to $\Delta \ell = 100$.
+Main panels show the TT (top), EE (middle), and BB (bottom) power spectra; accompanying panels show the difference between the input plus noise curves and the power spectra recovered from few-bit TOD.
+Spectra constructed from few-bit TOD recover the input well.
+This plot shows results for a simulation with $1,024,000$ hits per map pixel using Gaussian white detector noise.}
 \label{fig:psrecover}
@@
 \begin{center}
-\caption{\label{tab:extranoisewhite} Additional Noise - White Detector Noise}
+\caption{\label{tab:extranoisewhite} \changed{Fractional Noise Increase Due To Digitisation - Gaussian White Detector Noise}}
 \small
@@
 \hline
-\multirow{3}{*}{800}  & TT  & $ 0.252 \pm 0.009 $  & $ 0.064 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
-& EE  & $ 0.253 \pm 0.012 $  & $ 0.064 \pm 0.005 $  & $ 0.017 \pm 0.004 $ \\
-& BB  & $ 0.254 \pm 0.011 $  & $ 0.065 \pm 0.005 $  & $ 0.018 \pm 0.004 $ \\
+\multirow{3}{*}{800}  & TT  & $ 0.252 \pm 0.0091 $  & $ 0.0639 \pm 0.0034 $  & $ 0.0173 \pm 0.0021 $ \\
+& EE  & $ 0.253 \pm 0.012 $  & $ 0.0640 \pm 0.0047 $  & $ 0.0175 \pm 0.0041 $ \\
+& BB  & $ 0.254 \pm 0.011 $  & $ 0.0652 \pm 0.0054 $  & $ 0.0180 \pm 0.0039 $ \\
 \hline
-\multirow{3}{*}{8,000}  & TT  & $ 0.252 \pm 0.008 $  & $ 0.064 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
-& EE  & $ 0.252 \pm 0.007 $  & $ 0.064 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
-& BB  & $ 0.252 \pm 0.018 $  & $ 0.063 \pm 0.009 $  & $ 0.018 \pm 0.004 $ \\
+\multirow{3}{*}{8,000}  & TT  & $ 0.2521 \pm 0.0076 $  & $ 0.0641 \pm 0.0034 $  & $ 0.0174 \pm 0.0017 $ \\
+& EE  & $ 0.2519 \pm 0.0074 $  & $ 0.0636 \pm 0.0040 $  & $ 0.0173 \pm 0.0022 $ \\
+& BB  & $ 0.252 \pm 0.018 $  & $ 0.0634 \pm 0.0089 $  & $ 0.0176 \pm 0.0036 $ \\
 \hline
-\multirow{3}{*}{80,000}  & TT  & $ 0.253 \pm 0.008 $  & $ 0.065 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
-& EE  & $ 0.251 \pm 0.009 $  & $ 0.064 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
-& BB  & $ 0.250 \pm 0.015 $  & $ 0.064 \pm 0.007 $  & $ 0.018 \pm 0.003 $ \\
+\multirow{3}{*}{80,000}  & TT  & $ 0.2526 \pm 0.0080 $  & $ 0.0645 \pm 0.0040 $  & $ 0.0181 \pm 0.0016 $ \\
+& EE  & $ 0.2506 \pm 0.0093 $  & $ 0.0641 \pm 0.0042 $  & $ 0.0177 \pm 0.0019 $ \\
+& BB  & $ 0.250 \pm 0.012 $  & $ 0.0636 \pm 0.0065 $  & $ 0.0179 \pm 0.0032 $ \\
 \hline
-\multirow{3}{*}{1,024,000}  & TT  & $ 0.250 \pm 0.009 $  & $ 0.063 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
-& EE  & $ 0.252 \pm 0.008 $  & $ 0.065 \pm 0.003 $  & $ 0.018 \pm 0.002 $ \\
-& BB  & $ 0.252 \pm 0.008 $  & $ 0.064 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
+\multirow{3}{*}{1,024,000}  & TT  & $ 0.2501 \pm 0.0090 $  & $ 0.0633 \pm 0.0031 $  & $ 0.0173 \pm 0.0016 $ \\
+& EE  & $ 0.2518 \pm 0.0080 $  & $ 0.0646 \pm 0.0033 $  & $ 0.0179 \pm 0.0018 $ \\
+& BB  & $ 0.2520 \pm 0.0082 $  & $ 0.0639 \pm 0.0037 $  & $ 0.0173 \pm 0.0021 $ \\
 \hline
-\multirow{3}{*}{10,240,000}  & TT  & $ 0.248 \pm 0.008 $  & $ 0.063 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
-& EE  & $ 0.253 \pm 0.008 $  & $ 0.065 \pm 0.003 $  & $ 0.018 \pm 0.002 $ \\
-& BB  & $ 0.252 \pm 0.010 $  & $ 0.064 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
+\multirow{3}{*}{10,240,000}  & TT  & $ 0.2480 \pm 0.0078 $  & $ 0.0630 \pm 0.0038 $  & $ 0.0175 \pm 0.0018 $ \\
+& EE  & $ 0.2526 \pm 0.0077 $  & $ 0.0647 \pm 0.0033 $  & $ 0.0183 \pm 0.0018 $ \\
+& BB  & $ 0.252 \pm 0.010 $  & $ 0.0642 \pm 0.0042 $  & $ 0.0177 \pm 0.0020 $ \\
 \hline
-\multirow{3}{*}{102,400,000}  & TT  & $ 0.246 \pm 0.005 $  & $ 0.063 \pm 0.002 $  & $ 0.017 \pm 0.001 $ \\
-& EE  & $ 0.251 \pm 0.008 $  & $ 0.064 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
-& BB  & $ 0.252 \pm 0.009 $  & $ 0.064 \pm 0.003 $  & $ 0.018 \pm 0.002 $ \\
+\multirow{3}{*}{102,400,000}  & TT  & $ 0.2465 \pm 0.0049 $  & $ 0.0629 \pm 0.0019 $  & $ 0.0170 \pm 0.0010 $ \\
+& EE  & $ 0.2511 \pm 0.0078 $  & $ 0.0644 \pm 0.0035 $  & $ 0.0175 \pm 0.0018 $ \\
+& BB  & $ 0.2523 \pm 0.0086 $  & $ 0.0640 \pm 0.0031 $  & $ 0.0179 \pm 0.0017 $ \\
 \hline
-\multirow{3}{*}{Average}  & TT  & $ 0.25 \pm 0.008 $  & $ 0.064 \pm 0.003 $  & $ 0.017 \pm 0.002 $ \\
- & EE  & $ 0.252 \pm 0.009 $  & $ 0.064 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
- & BB  & $ 0.252 \pm 0.012 $  & $ 0.064 \pm 0.005 $  & $ 0.018 \pm 0.003 $ \\
+\multirow{3}{*}{Average}  & TT  & $ 0.2502 \pm 0.0017 $  & $ 0.0636 \pm 0.0017 $  & $ 0.0175 \pm 0.0017 $ \\
+ & EE  & $ 0.2518 \pm 0.0023 $  & $ 0.0643 \pm 0.0023 $  & $ 0.0177 \pm 0.0023 $ \\
+ & BB  & $ 0.2519 \pm 0.0027 $  & $ 0.0640 \pm 0.0027 $  & $ 0.0177 \pm 0.0027 $ \\
 \end{tabular}
 \tablecomments{ 
-Percent addition to the map noise level, $\Delta \sigma / \sigma$, due to one, two, and three bit digitisation for white detector noise.
-We use the standard deviation of points in the $C_\ell^\mathrm{X}/C_\ell^\mathrm{N}$ noise tail as the uncertainty.
-There is no appreciable variation with depth of observations and channel for a given digitisation scheme.
+\changed{(TABLE HAS BEEN CHANGED) Fractional addition to the map noise level, $\Delta \sigma / \sigma$, due to one-, two-, and three-bit digitisation for white detector noise with $1\sigma$ confidence intervals.
+Note that there is no appreciable variation between temperature and polarisation observations, or between different depths of observation.
+It is striking that three-bit digitisation only leads to a percent-level increase in the map noise level.}
 } \normalsize
@@
 \begin{center}
-\caption{\label{tab:extranoiseoof} Additional Noise - $1/f$ Detector Noise}
+\caption{\label{tab:extranoiseoof} \changed{Fractional Noise Increase Due To Digitisation - $1/f$ Detector Noise}}
 \small
@@
 \hline
-\multirow{3}{*}{8000}  & TT  & $ 0.231 \pm 0.017 $  & $ 0.058 \pm 0.005 $  & $ 0.016 \pm 0.002 $ \\
-& EE  & $ 0.300 \pm 0.012 $  & $ 0.064 \pm 0.004 $  & $ 0.018 \pm 0.002 $ \\
-& BB  & $ 0.294 \pm 0.022 $  & $ 0.062 \pm 0.007 $  & $ 0.017 \pm 0.003 $ \\
+\multirow{3}{*}{8,000}  & TT  & $ 0.231 \pm 0.017 $  & $ 0.0584 \pm 0.0052 $  & $ 0.0159 \pm 0.0022 $ \\
+& EE  & $ 0.250 \pm 0.012 $  & $ 0.0641 \pm 0.0043 $  & $ 0.0177 \pm 0.0019 $ \\
+& BB  & $ 0.245 \pm 0.021 $  & $ 0.0623 \pm 0.0067 $  & $ 0.0169 \pm 0.0032 $ \\
 \hline
-\multirow{3}{*}{102400}  & TT  & $ 0.235 \pm 0.014 $  & $ 0.060 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
-& EE  & $ 0.299 \pm 0.009 $  & $ 0.064 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
-& BB  & $ 0.298 \pm 0.025 $  & $ 0.064 \pm 0.007 $  & $ 0.018 \pm 0.002 $ \\
+\multirow{3}{*}{102,400}  & TT  & $ 0.235 \pm 0.014 $  & $ 0.0597 \pm 0.0042 $  & $ 0.0166 \pm 0.0019 $ \\
+& EE  & $ 0.2503 \pm 0.0091 $  & $ 0.0636 \pm 0.0042 $  & $ 0.0174 \pm 0.0016 $ \\
+& BB  & $ 0.249 \pm 0.024 $  & $ 0.0641 \pm 0.0069 $  & $ 0.0178 \pm 0.0025 $ \\
 \hline
-\multirow{3}{*}{1024000}  & TT  & $ 0.238 \pm 0.013 $  & $ 0.06 \pm 0.005 $  & $ 0.017 \pm 0.002 $ \\
-& EE  & $ 0.299 \pm 0.008 $  & $ 0.063 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
-& BB  & $ 0.299 \pm 0.01 $  & $ 0.064 \pm 0.005 $  & $ 0.018 \pm 0.002 $ \\
+\multirow{3}{*}{1,024,000}  & TT  & $ 0.238 \pm 0.013 $  & $ 0.0601 \pm 0.0046 $  & $ 0.0168 \pm 0.0019 $ \\
+& EE  & $ 0.2502 \pm 0.0078 $  & $ 0.0630 \pm 0.0039 $  & $ 0.0173 \pm 0.0019 $ \\
+& BB  & $ 0.2499 \pm 0.0095 $  & $ 0.0636 \pm 0.0049 $  & $ 0.0177 \pm 0.0019 $ \\
 \hline
-\multirow{3}{*}{Average}  & TT  & $ 0.235 \pm 0.015 $  & $ 0.059 \pm 0.005 $  & $ 0.016 \pm 0.002 $ \\
- & EE  & $ 0.299 \pm 0.010 $  & $ 0.064 \pm 0.004 $  & $ 0.017 \pm 0.002 $ \\
- & BB  & $ 0.297 \pm 0.019 $  & $ 0.063 \pm 0.006 $  & $ 0.017 \pm 0.003 $ \\
+\multirow{3}{*}{Average}  & TT  & $ 0.2346 \pm 0.0020 $  & $ 0.0594 \pm 0.0020 $  & $ 0.0164 \pm 0.0020 $ \\
+ & EE  & $ 0.2503 \pm 0.0018 $  & $ 0.0636 \pm 0.0018 $  & $ 0.0175 \pm 0.0018 $ \\
+ & BB  & $ 0.2480 \pm 0.0025 $  & $ 0.0633 \pm 0.0025 $  & $ 0.0175 \pm 0.0025 $ \\
 \end{tabular}
 \tablecomments{ 
-Percent addition to the map noise level, $\Delta \sigma / \sigma$, due to one, two, and three bit digitisation for $1/f$ detector noise.
-Uncertainties are obtained in the same way as for white noise.
-As in the white noise case we observe no significant trend with the numbers of hits per map pixel or channel for a given digitisation scheme.
-No significant deviation in the values compared to simulations assuming white detector noise is found.
+\changed{(TABLE HAS BEEN CHANGED) Fractional addition to the map noise level, $\Delta \sigma / \sigma$, due to one-, two-, and three-bit digitisation for $1/f$ detector noise with $1\sigma$ confidence intervals.
+As in the white noise case, we observe no significant trend between temperature and polarisation or for different depths of observation.
+We note that, even assuming a more realistic noise profile, three-bit digitisation only leads to a percent-level increase in the map noise level.}
 } \normalsize
@@
 
-As mentioned in subsection \S\ref{subsec:extremedigitisation}, equation \ref{eq:distdef} does not demand not signal power conservation in constructing a digitisation scheme.
-It is relatively simple to impose power conservation, since it boils down to rescaling the digitised output by a constant, $\gamma$.
+As mentioned in subsection \S\ref{subsec:extremedigitisation}, equation \ref{eq:distdef} does not demand \changed{that a digitisation scheme conserves power.
+However, it} is relatively simple to impose power conservation, since it boils down to rescaling the digitised output by a constant, $\gamma$.
 We derive an expression for $\gamma$, such that a cross-spectrum of two digitised maps yields an unbiased estimate of the input power spectrum.
@@
 
-where we sum over all output levels of both timestreams $(1, 2)$ and $P$ denotes the probability of an output level occurring.
-The probabilities $P$ can be expressed as integrals over probability distributions.
-Assuming Gaussian detector noise we introduce normal distributions $p(\xi)$ with mean zero and standard deviation $\sigma$.
+\changed{ where we sum over all output levels of both timestreams $(1, 2)$ and $P$ denotes the probability of an output level occurring.
+For current-generation CMB experiments it is common to draw from $>10^5$ samples to obtain an estimate of a map pixel.
+This corresponds to adding many noise realisation on top of the same signal, $\mu$.
+The central limit theorem ensures that the amplitude probability distribution function of the noise in each pixel is Gaussian. 
+The probabilities $P$ in the above expression can therefore be expressed as integrals over normal distributions, $p(\xi)$, with mean zero and standard deviation $\sigma$.}
 
@@
 &\times \left\{ \sum_{j=0}^N  \frac{y_j}{\sqrt{2}} \left[ \sum_{m = 0}^\infty \frac{(-1)^m}{m!
-(2m+1)} \frac{1}{(\sqrt{2}\sigma)^{2m+1}} \left( \sum_{l=0}^{2m+1} {2m+1 \choose l} \mu^l ( x_{j+1}^{2m+1-l} - x_{j}^{2m+1-l} ) \right) \right] \right\}.
+(2m+1)} \frac{1}{(\sqrt{2}\sigma)^{2m+1}} \left( \sum_{l=0}^{2m+1} {2m+1 \choose l} \mu^l ( x_{j+1}^{2m+1-l} - x_{j}^{2m+1-l} ) \right) \right] \right\} \changed{,}
 \end{aligned}
@@
 
-Where we have used the Maclaurin series expansion of the error function and the binomial expansion. Moving the $i, j$ summation forwards yields
+\changed{where} we have used the Maclaurin series expansion of the error function and the binomial expansion. Moving the $i, j$ summation forwards yields
 
@@
 
-since $\psi_{N-i}^{k,n} = (-1)^{k} \psi_{i}^{k,n}$.
-For ground-based CMB observations $\mu$ is small.
-Therefore we calculate the leading order terms in $\mu$ in equation \ref{eq:dcompleteklsum}.
-Since this equation has powers of $\mu^{k+l}$ we split the sum over $k, l$ as follows
+\changed{because} $\psi_{N-i}^{k,n} = (-1)^{k} \psi_{i}^{k,n}$.
+\changed{Since ground-based CMB experiments operate at low signal-to-noise we calculate the leading order terms of the signal, $\mu$, in equation \ref{eq:dcompleteklsum}.
+We split the sum over $k, l$ in said equation as follows:}
 
 \begin{equation}
-\sum_{i,j = 0}^N y_i y_j \psi_i^{k, n} \psi_j^{l, m} = \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} y_i y_j  \left( \psi_i^{k, n} \psi_j^{l, m} - \psi_{i}^{k, n} \psi_j^{l, m} - \psi_i^{k, n} \psi_{j}^{l, m} + \psi_{i}^{k, n} \psi_{j}^{l, m} \right) = 0,
+\sum_{i,j = 0}^N y_i y_j \psi_i^{k, n} \psi_j^{l, m} = \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} y_i y_j  \left( \psi_i^{k, n} \psi_j^{l, m} - \psi_{i}^{k, n} \psi_j^{l, m} - \psi_i^{k, n} \psi_{j}^{l, m} + \psi_{i}^{k, n} \psi_{j}^{l, m} \right) = 0 \changed{.}
 \end{equation}
@@
 
-where $c$ is some constants that depend on the specific digitisation scheme chosen. Returning to equation \ref{eq:normcrosspower} we see
+where $c$ is some constants that \changed{depends} on the specific digitisation scheme chosen. Returning to equation \ref{eq:normcrosspower} we see
 
@@
 (2n+1)} \frac{(-1)^{n+m}}{m!
-(2m+1)} \frac{1}{(\sqrt{2}\sigma)^{2n+2m+2}} \left( {2n+1 \choose 1} {2m+1 \choose 1} ( x_{i+1}^{2n} - x_{i}^{2n} ) ( x_{j+1}^{2m} - x_{j}^{2m} ) \right) \right]
+(2m+1)} \frac{1}{(\sqrt{2}\sigma)^{2n+2m+2}} \left( {2n+1 \choose 1} {2m+1 \choose 1} ( x_{i+1}^{2n} - x_{i}^{2n} ) ( x_{j+1}^{2m} - x_{j}^{2m} ) \right) \right] \changed{.}
 \end{aligned}
@@
 \begin{equation}
-= \sum_{i,j=0}^N \frac{y_i y_j}{2\pi\sigma^2} \left\{ exp \left[ - \left(\frac{x_{i+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{i}}{\sqrt{2}\sigma} \right)^2 \right] \right\} \left\{ exp \left[ - \left(\frac{x_{j+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{j}}{\sqrt{2}\sigma} \right)^2 \right] \right\}
+= \sum_{i,j=0}^N \frac{y_i y_j}{2\pi\sigma^2} \left\{ exp \left[ - \left(\frac{x_{i+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{i}}{\sqrt{2}\sigma} \right)^2 \right] \right\} \left\{ exp \left[ - \left(\frac{x_{j+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{j}}{\sqrt{2}\sigma} \right)^2 \right] \right\} \changed{.}
 \end{equation}
@@
 \begin{equation}
-\gamma = \left( \sum_{i,j=0}^N \frac{y_i y_j}{2\pi\sigma^2} \left\{ exp \left[ - \left(\frac{x_{i+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{i}}{\sqrt{2}\sigma} \right)^2 \right] \right\} \left\{ exp \left[ - \left(\frac{x_{j+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{j}}{\sqrt{2}\sigma} \right)^2 \right] \right\} \right)^{-1/2}
+\gamma = \left( \sum_{i,j=0}^N \frac{y_i y_j}{2\pi\sigma^2} \left\{ exp \left[ - \left(\frac{x_{i+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{i}}{\sqrt{2}\sigma} \right)^2 \right] \right\} \left\{ exp \left[ - \left(\frac{x_{j+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{j}}{\sqrt{2}\sigma} \right)^2 \right] \right\} \right)^{-1/2} \changed{.}
 \end{equation}
 
-This can be calculated for a given digitisation scheme $N, x_i, y_i$ and standard deviation of the noise $\sigma$.
-Notice that if the digitisation scheme is chosen such that $y_i$ and $x_i$ depend linearly on $\sigma$ the normalisation constant is independent of $\sigma$.
-
-% this is handy in practice! do not have to account for different detector sensitivities within detector array and T/pol.
+\changed{This can be calculated for a given digitisation scheme, specified by $N, x_i, y_i$, and standard deviation of the noise, $\sigma$.
+Notice that if the digitisation scheme is chosen such that the digitisation thresholds and output levels depend linearly on the noise level, the normalisation constant is independent of the noise level.}
 

### 2018-12-21 digitisation.tex
random

@@
 We summarise our findings in \S\ref{sec:conclusions}.
+\changed{We report uncertainties based on the symmetric 68\% confidence regions. }
 
@@
 \changed{Finally, it is striking how little noise is added.
-On average one-bit digitisation leads to a $(25.13\pm 0.94)\%$ increase in the map noise level, two-bit yields a $(6.40\pm0.41)\%$ increase and three-bit adds only $(1.76\pm0.22)\%$ ($1\sigma$ confidence intervals given).
+On average one-bit digitisation leads to a $(25.13\pm 0.94)\%$ increase in the map noise level, two-bit yields a $(6.40\pm0.41)\%$ increase and three-bit adds only $(1.76\pm0.22)\%$. 
 This is impressive, given that all schemes considered reduce the TOD volume by at least an order of magnitude.}
@@
   \caption[Current ]{
-  \changed{(FIGURE HAS BEEN CHANGED) TT, EE, and BB power spectra (upper, middle and lowest panel respectively) of difference maps for one-bit (green marker, high), two-bit (orange marker, medium), and three-bit digitisation (violet marker, low).
+  \changed{TT, EE, and BB power spectra (upper, middle and lowest panel respectively) of difference maps for one-bit (green marker, high), two-bit (orange marker, medium), and three-bit digitisation (violet marker, low).
 The control case using 64-bit TOD is also shown (black marker).
@@
 
-While digitisation clearly works extremely well for white noise, real ground-based CMB experiments face several sources of low-frequency noise.
-\changed{High on the list is atmospheric noise.
-Modelling the atmosphere as a thin sheet dominated by Kolmogorov turbulence, we can describe fluctuations with the noise profile \citep{lay2000}:}
+\changed{While digitisation clearly works extremely well for white noise, real experiments face several sources of low-frequency noise. 
+High on the list for ground-based experiments is atmospheric noise. 
+For typical observing conditions in Antarctica, the atmosphere turbulence is well-described by Kolmogorov turbulence in the thin-sheet regime \citep{bussmann2005}. 
+In this regime, we can describe fluctuations with the noise profile \citep{lay2000}:}
 
@@
 While polarisation observations seem to incur a slightly higher noise penalty compared to temperature, deviations do not exceed $3\sigma$.
-\changed{On average one-bit digitisation leads to a $(24.4\pm 1.4)\%$ increase in the map noise level, two-bit adds $(6.21\pm0.50)\%$, and three-bit yields an increase of only $(1.71\pm0.21)\%$ ($1\sigma$ confidence intervals given).
-Thus the effectiveness of few bit digitisation is not limited to the white noise case, but \changed{extends} to more realistic noise profiles.}
+\changed{On average one-bit digitisation leads to a $(24.4\pm 1.4)\%$ increase in the map noise level, two-bit adds $(6.21\pm0.50)\%$, and three-bit yields an increase of only $(1.71\pm0.21)\%$.
+The fractional added noise from few-bit digitisation is consistent with being identical for the white noise and realistic noise profiles.}
 
@@
 
-\subsection{\changed{Existing Compression Techniques}}
+\subsection{\changed{Interaction with Existing Compression Techniques}}
 
-\changed{Current-generation ground-based CMB experiments employ lossless and lossy compression techniques to manage their transmission bottlenecks.
+\changed{Current ground-based CMB experiments already employ 
+ lossless and lossy compression techniques to manage their transmission bottlenecks.
 To test whether few-bit digitisation is competitive to existing lossy compression techniques, it is necessary to see if it still provides a substantial reduction in data volume if used in conjunction with lossless compression algorithms.}

### 2018-12-21 digitisation.tex
app

@@
 \tablecomments{ 
-\changed{(TABLE HAS BEEN CHANGED) Fractional addition to the map noise level, $\Delta \sigma / \sigma$, due to one-, two-, and three-bit digitisation for $1/f$ detector noise with $1\sigma$ confidence intervals.
+\changed{Fractional addition to the map noise level, $\Delta \sigma / \sigma$, due to one-, two-, and three-bit digitisation for $1/f$ detector noise with $1\sigma$ confidence intervals.
 As in the white noise case, we observe no significant trend between temperature and polarisation or for different depths of observation.
@@
 \changed{ where we sum over all output levels of both timestreams $(1, 2)$ and $P$ denotes the probability of an output level occurring. 
-For current-generation CMB experiments it is common to draw from $>10^5$ samples to obtain an estimate of a map pixel.
-This corresponds to adding many noise realisation on top of the same signal, $\mu$.
-The central limit theorem ensures that the amplitude probability distribution function of the noise in each pixel is Gaussian. 
+The final estimate of the value in a map pixel is frequently based on $>10^5$ TOD samples in current CMB experiments, and this number will increase with the number of detectors for future experiments.
+This can is equivalent to have many noise realisations on top of the same signal, $\mu$. 
+Given the number of samples, the central limit theorem ensures that the amplitude probability distribution function of the noise in each pixel is Gaussian. 
 The probabilities $P$ in the above expression can therefore be expressed as integrals over normal distributions, $p(\xi)$, with mean zero and standard deviation $\sigma$.}
@@
 \changed{because} $\psi_{N-i}^{k,n} = (-1)^{k} \psi_{i}^{k,n}$.
-\changed{Since ground-based CMB experiments operate at low signal-to-noise we calculate the leading order terms of the signal, $\mu$, in equation \ref{eq:dcompleteklsum}.
+\changed{Since the TOD of ground-based CMB experiments has low signal-to-noise, we calculate the leading order terms of the signal, $\mu$, in equation \ref{eq:dcompleteklsum}.
 We split the sum over $k, l$ in said equation as follows:}

### 2019-05-17 response letter/response_letter.tex
Add files via upload
Updated for second response.

@@
 
-Thank you for your comments.
-We are pleased that you found our manuscript to be a valuable enquiry into data compression techniques for CMB experiments and a detailed investigation of the effects of few-bit digitisation on the CMB power spectra.
-We found your suggestion of looking at the compressive power of few-bit digitisation together with established lossless compression techniques particularly insightful and exciting to research.
-Please find a detailed response to your report below.
-Shown in blue are remarks in the reviewer's report.
-The authors' reply is given in black and if applicable relevant sections of the manuscript that have been changed in response are appended in red.
+Thank you for taking the time to reassess our manuscript and review our response.
+We appreciate your comments with regards to the compression power of few-bit digitisation in conjunction with existing data reduction methods.
+Please find a detailed response to the points you raised below.
 
@@
 
-\indent \question{[...] the article's case is weakened considerably by a lack of comparison to existing lossless approaches to data compression used in the field. [...]}
+\indent \question{I am a bit shocked that the performance in conjunction with LPC-/PCM-based compression algorithms is as good as claimed here [...] My suspicion here is that the synthetic data under test have much higher entropy [...]}
 
@@
 
-\indent The referee raises a fair point.
-We contacted the BICEP Array, POLARBEAR, and SPT-3G experiments to enquire about the lossless compression algorithms currently being used.
-BICEP uses a Lempel-Ziv-Markov chain algorithm (LZMA), while the SPT and POLARBEAR use the free lossless audio codec (FLAC).
-We applied few-bit digitisation to a simulated timestream and then applied each compression scheme to the few bit timestreams.
-We compared this to the size of files when the two compression schemes were applied to the original 64bit timestreams. 
-We found that while using either LZMA or FLAC reduces the compressive power of extreme digitisation, substantial data volume savings are still achieved.
-We appended the following subsection to section 3 (Results) of the manuscript to include this insight:
+\indent We appreciate the concern raised. To address this issue appropriately we followed the referee's suggestion and repeated the analysis using genuine experimental data.
+We obtain SPT-3G data and use the FLAC compression algorithm used by the SPT.
+We apply few-bit digitisation to the TOD and apply FLAC.
+This allows us to compare the resulting file sizes between the few-bit data and the typical compression rates achieved for SPT-3G data.
+Applying FLAC to few-bit timestreams does not lead to additional data volume reduction. We altered subsection 3.3 to reflect this insight.
 
@@
 
-\changed{Current-generation ground-based CMB experiments employ lossless and lossy compression techniques to manage their transmission bottlenecks.
-To test whether few-bit digitisation is competitive to existing lossy compression techniques, it is necessary to see if it still provides a substantial reduction in data volume if used in conjunction with lossless compression algorithms.}
+\changed{Current-generation ground-based CMB experiments employ a combination of lossly, e.g.
+downsampling, and lossless compression, e.g.
+FLAC, bzip, LZMA, to manage their transmission bottlenecks.
+We test whether the data volume of timestreams that have undergone few-bit digitisation can be reduced further by applying the most popular lossless compression algorithm: FLAC.
+}
 
-\changed{We focus on lossless compression via the Lempel-Ziv-Markoc chain algorithm (LZMA), which is used by the Background Imaging of Cosmic Extragalactic Polarization (BICEP) experiment, and the free lossless audio codec (FLAC), which the SPT and POLARBEAR experiments employ (private communication).
-We examine the compression of a single CES of our simulation, consisting of ca.
-$10^7$ data points, by applying LZMA and FLAC separately to four timestreams: the original 64-bit TOD, and copies that have undergone 1-, 2-, and 3-bit digitisation.}
+\changed{We obtain 2 minutes of SPT-3G TOD and apply the digitisation schemes in equations \ref{eq:1bit}-\ref{eq:3bit} before compressing the data further using FLAC.
+We compare the file size of FLAC compressed few-bit timestreams to the typical data reduction rate achieved for 24-bit SPT-3G data.}
 
-\changed{We find that even though the data reduction power of few-bit digitisation is decreased when used in conjunction with lossless compression algorithms, substantial savings are still achieved.
-Applying few-bit digitisation to the TOD prior to compression with FLAC reduces the final file size by a factor $5.89$ compared to the FLAC compressed 64-bit data.
-This is the same for all digitisation schemes considered.
-Working with LZMA, the data volume of the TOD can be reduced with respect to compressed 64-bit data by factors of $17.34$, $26.23$, and $44.48$ for 3-, 2-, and 1-bit digitisation, respectively.
-Few-bit digitisation still achieves considerable data reduction if used together with established lossless compression algorithms.}
+\changed{We find that using FLAC in conjunction with extreme digitisation does not yield additional data volume reduction.
+For 3-bit digitisation, both the FLAC and raw timestream have $\sim$3 bits per data point.
+In fact, due to the overhead required by FLAC compressed files perform marginally worse than uncompressed few-bit streams.
+This is unsurprising, given that we have already maximally compressed the data using extreme digitisation.
+For comparison, FLAC reduces the full 24-bit  SPT-3G data to $6-8$ bits of entropy per number.
+For this real-world case, three-bit digitisation reduces the data volume by a factor of $2-3$.}
 
@@
 
-\indent \question{It is unclear from the text what, exactly, the resulting scheme is after equation 13.
-I found equations 7-9 extremely helpful to understand the approach taken earlier and it would be very helpful to see the analogs (is it just replacing sigma in equations 7-9?)}
+\indent \question{[...] what is the impact on non-linearity in power spectrum estimation? [...]}
 
@@
 
-\indent Yes, it is simply replacing $\sigma$ in equations 7-9. We have changed the wording of the relevant section to make this more clear.
+\indent Thank you for raising this point.
+Indeed the non-linearity introduced by few-bit digitisation disappears in the power spectrum estimation.
+This is because many TOD points contribute to each map pixel and hence the central limit theorem ensures that the distribution within each pixel approaches a Gaussian.
+We appended a sentence stating this to subsection 2.4 (Power Spectrum Estimation).
 
@@
 
-\indent \changed{We use the same digitisation schemes as for white noise, but replace the parameter $\sigma$ in equations \ref{eq:1bit}, \ref{eq:2bit} and \ref{eq:3bit} by considering Parseval's theorem on the above noise profile.
-The appropriate value is obtained by numerically evaluating the integral}
-
-\[ \changed{
-\sigma^2 = \int_0^\infty \left| N(\ell) \right|^2 d\ell. }
-\]
-
-\vspace{0.5cm}
-
-\hrule
-
-\vspace{0.5cm}
-
-Reviewer's comment:
-
-\indent \question{The last sentence before section 4 has a typo ("extents" rather than "extends"). There are many more such in the appendix, which needs a bit more careful editing.}
-
-\vspace{0.25cm}
-
-Authors' reply:
-
-\indent Thank you for pointing this out. We have fixed this typo and edited the appendix.
-
-\vspace{0.5cm}
-
-\hrule
-
-\vspace{0.5cm}
-
-Reviewer's comment:
-
-\indent \question{In many cases, the authors dive directly into mathematical formalism without explaining the results.
-For example, it took me several readings before realizing that I could think of the scheme in equations 7-9, to a reasonably high degree of accuracy, as just dropping the high-order 32-n bits of the data.
-The text would be improved somewhat by a bit of unpacking along these lines, though it is acceptable as-is.}
-
-\vspace{0.25cm}
-
-Authors' reply:
-
-\indent We recognise that the derivation of the digitisation schemes was not as clear as possible. We have edited subsection 2.1 (Extreme Digitisation) to be more readily understood.
+\indent \changed{Non-linearity introduced by the digitisation process disappears in the final map due to the large number of TOD points that contribute to each map pixel.
+Thus the digitisation effects are not an issue for power spectrum estimation.}
 

### 2019-05-17 digitisation.tex
Add files via upload
Updated for second response.

--- 2018-12-20 digitisation.tex
@@
 Unlike lossless compression, few bit digitisation introduces additional noise into the data.
-We present a set of one, two, and three bit digitisation schemes and measure \changed{the} increase in noise in the cosmic microwave background temperature and polarisation power spectra.
+We present a set of one, two, and three bit digitisation schemes and measure the increase in noise in the cosmic microwave background temperature and polarisation power spectra.
 The digitisation noise is independent of angular scale and is well-described as a constant percentage of the original detector noise.
-Three bit digitisation increases the map noise level by $< 2\%$, while reducing the data volume by a factor of \changed{twenty relative to 64}-bit floats.
-\changed{Comparable gains are seen when extreme digitisation is combined with
-established lossless compression algorithms. Extreme} digitisation is a promising strategy for upcoming experiments.
+Three bit digitisation increases the map noise level by $< 2\%$, while reducing the data volume by a factor of \changed{ten relative to 32-bit floats.
+Lossless compression of the digitised data does not lead to additional data volume reduction.} Extreme digitisation is a promising strategy for upcoming experiments.
 
@@
 
-The Planck mission has already demonstrated the success of extreme digitisation for space-based CMB \changed{experiments} \citep{maris2003}.
+The Planck mission has already demonstrated the success of extreme digitisation for space-based CMB experiments \citep{maris2003}.
 \cite{jenet1998} explored the application of few bit digitisation to radio pulsar timing measurements.
@@
 For a stochastic input signal we can calculate an amplitude probability distribution $p(x)$.
-\changed{This allows us to re-express the distortion as a sum over the digitisation levels:}
+This allows us to re-express the distortion as a sum over the digitisation levels:
 
@@
 
-Since we wish to minimise the distortion, we set \changed{its derivatives} with respect to $x_i$ and $y_i$ to zero. Starting with $x_i$ we obtain
+Since we wish to minimise the distortion, we set its derivatives with respect to $x_i$ and $y_i$ to zero. Starting with $x_i$ we obtain
 
@@
 \end{equation}
-This \changed{condition} implies that we should choose $y_i$, such that it halves the area underneath $p(x)$ in the interval from $x_i$ to $x_{i+1}$.
+This condition implies that we should choose $y_i$, such that it halves the area underneath $p(x)$ in the interval from $x_i$ to $x_{i+1}$.
 
 To progress further, we must consider the probability distribution $p(x)$ of the input signal. 
-\changed{In this work, we are interested in the compression of the TOD from each detector of a ground-based CMB experiment. 
+In this work, we are interested in the compression of the TOD from each detector of a ground-based CMB experiment. 
 Thus we can safely assume that the input signal is in the low signal-to-noise regime. 
-Furthermore, we take the simplest noise profile: Gaussian, white detector noise\footnote{\changed{As discussed in the appendix, this treatment also applies to more realistic noise profiles, because the large number of points lying in each map pixel allows the application of the central limit theorem.}}.
-Together these assumptions imply that the input, $x$, is normally distributed, with} $p(x) = (1/\sqrt{2\pi\sigma^2}) e^{-x^2/2\sigma^2}$.
+Furthermore, we take the simplest noise profile: Gaussian, white detector noise\footnote{As discussed in the appendix, this treatment also applies to more realistic noise profiles, because the large number of points lying in each map pixel allows the application of the central limit theorem.}.
+Together these assumptions imply that the input, $x$, is normally distributed, with $p(x) = (1/\sqrt{2\pi\sigma^2}) e^{-x^2/2\sigma^2}$.
 
-\changed{\citet{max1960} solved Equations \ref{eq:digitequalspacecondition} and \ref{eq:digitareacondition} for 1-, 2- and 3-bit digitisation for this choice of probability distribution, $p(x)$. 
+\citet{max1960} solved Equations \ref{eq:digitequalspacecondition} and \ref{eq:digitareacondition} for 1-, 2- and 3-bit digitisation for this choice of probability distribution, $p(x)$. 
 We quote their results below and refer the reader to the aforementioned work for the derivation.  
@@
 Essentially, the optimal 1-bit digitisation scheme applies the sign function to the TOD.
-}
 
-\changed{
-For two bit digitisation, we can encode $N=4$ output levels. The optimal set of thresholds and digitisation levels is given by the four-level function:}
+For two bit digitisation, we can encode $N=4$ output levels. The optimal set of thresholds and digitisation levels is given by the four-level function:
 \begin{equation}  \label{eq:2bit}
@@
 
-\changed{Finally, with 3 bits we may specify $N=8$ output levels. The optimal three bit digitisation is described by the eight-level function:}
+Finally, with 3 bits we may specify $N=8$ output levels. The optimal three bit digitisation is described by the eight-level function:
 \begin{equation}  \label{eq:3bit}
@@
 
-\changed{The digitisation schemes above are designed to minimise the distortion of the TOD as defined in Equation \ref{eq:distdef}.
+The digitisation schemes above are designed to minimise the distortion of the TOD as defined in Equation \ref{eq:distdef}.
 This does not ensure that the digitised output has the same signal power as the input.
@@
 We use this analytical rescaling for all results in this work.
-}
 
@@
 
-We rebin the calculated power spectra as mentioned \changed{when applicable} and assign Fisher matrix error bars \citep{tegmarkdesignersguide1997}.
+We rebin the calculated power spectra as mentioned when applicable and assign Fisher matrix error bars \citep{tegmarkdesignersguide1997}.
 
@@
 Differences between few bit and the input power spectra are due to observing a single CMB realisation and residual boundary effects.
+\changed{Non-linearity introduced by the digitisation process disappears in the final map due to the large number of TOD points that contribute to each map pixel.
+Thus the digitisation effects are not an issue for power spectrum estimation.}
 
@@
 We observe that the fractional increase, $\Delta \sigma / \sigma$, is independent of the hits per pixel in the maps and that the compression performs as well for polarisation as is does for temperature data.
-\changed{Finally, it is striking how little noise is added.
-On average one-bit digitisation leads to a $(25.13\pm 0.94)\%$ increase in the map noise level, two-bit yields a $(6.40\pm0.41)\%$ increase and three-bit adds only $(1.76\pm0.22)\%$ ($1\sigma$ confidence intervals given).
-This is impressive, given that all schemes considered reduce the TOD volume by at least an order of magnitude.}
+Finally, it is striking how little noise is added.
+On average one-bit digitisation leads to a $(25.13\pm 0.94)\%$ increase in the map noise level, two-bit yields a $(6.40\pm0.41)\%$ increase and three-bit adds only $(1.76\pm0.22)\%$ ($1\sigma$ confidence intervals given.
+This is impressive, given that all schemes considered reduce the TOD volume by at least an order of magnitude.
 
@@
   \caption[Current ]{
-  \changed{(FIGURE HAS BEEN CHANGED) TT, EE, and BB power spectra (upper, middle and lowest panel respectively) of difference maps for one-bit (green marker, high), two-bit (orange marker, medium), and three-bit digitisation (violet marker, low).
+  TT, EE, and BB power spectra (upper, middle and lowest panel respectively) of difference maps for one-bit (green marker, high), two-bit (orange marker, medium), and three-bit digitisation (violet marker, low).
 The control case using 64-bit TOD is also shown (black marker).
@@
 The noise increase due to digitisation appears to be independent of angular scale.
-This plot shows results for a simulation with $1,024,000$ hits per map pixel using Gaussian white detector noise.}
+This plot shows results for a simulation with $1,024,000$ hits per map pixel using Gaussian white detector noise.
 \label{fig:diffpswn}
@@
 While digitisation clearly works extremely well for white noise, real ground-based CMB experiments face several sources of low-frequency noise.
-\changed{High on the list is atmospheric noise.
-Modelling the atmosphere as a thin sheet dominated by Kolmogorov turbulence, we can describe fluctuations with the noise profile \citep{lay2000}:}
+High on the list is atmospheric noise.
+Modelling the atmosphere as a thin sheet dominated by Kolmogorov turbulence, we can describe fluctuations with the noise profile \citep{lay2000}:
 
-\changed{
 \begin{equation} \label{eq:1/fnoiseform}
@@
 \end{equation}
-}
 
@@
 where $N_{\mathrm{white}}$ denotes white detector noise.
-In this expression \changed{$\ell_{\mathrm{scan}}$} describes the multipole moment of angular features of corresponding size along the scan direction.
-The scale at which such modes are amplified appreciably by the second term in equation \ref{eq:1/fnoiseform} is set by \changed{$\ell_0$}.
-\changed{The exponent of $\ell_{\mathrm{scan}}/\ell_0$ is chosen to model observation at a constant, moderate elevation \citep{lay1997}.
-In line with the observed noise profile of the SPT we choose $\ell_0 = 1000$ for temperature and $\ell_0 = 0$ for polarisation; this means that power is doubled at $\lknee \approx 1800$ and $\lknee \approx 90$ for temperature and polarisation, respectively \citep{henning2018}.} In practice CMB experiments suppress power in the lowest modes through a range of techniques, including fitting polynomials, sines, and cosines to the TOD.
+In this expression $\ell_{\mathrm{scan}}$ describes the multipole moment of angular features of corresponding size along the scan direction.
+The scale at which such modes are amplified appreciably by the second term in equation \ref{eq:1/fnoiseform} is set by $\ell_0$.
+The exponent of $\ell_{\mathrm{scan}}/\ell_0$ is chosen to model observation at a constant, moderate elevation \citep{lay1997}.
+In line with the observed noise profile of the SPT we choose $\ell_0 = 1000$ for temperature and $\ell_0 = 0$ for polarisation; this means that power is doubled at $\lknee \approx 1800$ and $\lknee \approx 90$ for temperature and polarisation, respectively \citep{henning2018}.
+In practice CMB experiments suppress power in the lowest modes through a range of techniques, including fitting polynomials, sines, and cosines to the TOD.
 We simulate successful removal of these modes by applying a high pass filter to the noise, such that
 
-\changed{
 \begin{equation} \label{eq:1/fnoiseformfinal}
@@
 
-We use the same digitisation schemes as for white noise, but \changed{replace} the parameter $\sigma$ in equations \ref{eq:1bit}, \ref{eq:2bit} and \ref{eq:3bit} by considering Parseval's theorem on the above noise profile.
-The appropriate value is obtained by \changed{numerically} evaluating the integral
+We use the same digitisation schemes as for white noise, but replace the parameter $\sigma$ in equations \ref{eq:1bit}, \ref{eq:2bit} and \ref{eq:3bit} by considering Parseval's theorem on the above noise profile.
+The appropriate value is obtained by numerically evaluating the integral
 
@@
 While polarisation observations seem to incur a slightly higher noise penalty compared to temperature, deviations do not exceed $3\sigma$.
-\changed{On average one-bit digitisation leads to a $(24.4\pm 1.4)\%$ increase in the map noise level, two-bit adds $(6.21\pm0.50)\%$, and three-bit yields an increase of only $(1.71\pm0.21)\%$ ($1\sigma$ confidence intervals given).
-Thus the effectiveness of few bit digitisation is not limited to the white noise case, but \changed{extends} to more realistic noise profiles.}
+On average one-bit digitisation leads to a $(24.4\pm 1.4)\%$ increase in the map noise level, two-bit adds $(6.21\pm0.50)\%$, and three-bit yields an increase of only $(1.71\pm0.21)\%$ ($1\sigma$ confidence intervals given).
+Thus the effectiveness of few bit digitisation is not limited to the white noise case, but extends to more realistic noise profiles.
 
@@
 
-\subsection{\changed{Existing Compression Techniques}}
+\subsection{Existing Compression Techniques}
 
-\changed{Current-generation ground-based CMB experiments employ lossless and lossy compression techniques to manage their transmission bottlenecks.
-To test whether few-bit digitisation is competitive to existing lossy compression techniques, it is necessary to see if it still provides a substantial reduction in data volume if used in conjunction with lossless compression algorithms.}
+\changed{Current-generation ground-based CMB experiments employ a combination of lossly, e.g.
+downsampling, and lossless compression, e.g.
+FLAC, bzip, LZMA, to manage their transmission bottlenecks.
+We test whether the data volume of timestreams that have undergone few-bit digitisation can be reduced further by applying the most popular lossless compression algorithm: FLAC.
+}
 
-\changed{We focus on lossless compression via the Lempel-Ziv-Markoc chain algorithm (LZMA), which is used by the Background Imaging of Cosmic Extragalactic Polarization (BICEP) experiment, and the free lossless audio codec (FLAC), which the SPT and POLARBEAR experiments employ (private communication).
-We examine the compression of a single CES of our simulation, consisting of ca.
-$10^7$ data points, by applying LZMA and FLAC separately to four timestreams: the original 64-bit TOD, and copies that have undergone 1-, 2-, and 3-bit digitisation.}
+\changed{We obtain 2 minutes of SPT-3G TOD and apply the digitisation schemes in equations \ref{eq:1bit}-\ref{eq:3bit} before compressing the data further using FLAC.
+We compare the file size of FLAC compressed few-bit timestreams to the typical data reduction rate achieved for 24-bit SPT-3G data.}
 
-\changed{We find that even though the data reduction power of few-bit digitisation is decreased when used in conjunction with lossless compression algorithms, substantial savings are still achieved.
-Applying few-bit digitisation to the TOD prior to compression with FLAC reduces the final file size by a factor $5.89$ compared to the FLAC compressed 64-bit data.
-This is the same for all digitisation schemes considered.
-Working with LZMA, the data volume of the TOD can be reduced with respect to compressed 64-bit data by factors of $17.34$, $26.23$, and $44.48$ for 3-, 2-, and 1-bit digitisation, respectively.
-Few-bit digitisation still achieves considerable data reduction if used together with established lossless compression algorithms.}
+\changed{We find that using FLAC in conjunction with extreme digitisation does not yield additional data volume reduction.
+For 3-bit digitisation, both the FLAC and raw timestream have $\sim$3 bits per data point.
+In fact, due to the overhead required by FLAC compressed files perform marginally worse than uncompressed few-bit streams.
+This is unsurprising, given that we have already maximally compressed the data using extreme digitisation.
+For comparison, FLAC reduces the full 24-bit  SPT-3G data to $6-8$ bits of entropy per number.
+For this real-world case, three-bit digitisation reduces the data volume by a factor of $2-3$.}
 
@@
 This is true for temperature and polarisation observations.
-No change in the results is observed for maps \changed{with different numbers of} hits per pixel and no angular scale dependence was observed in the added noise.
-\changed{Combining few-bit digitisation with currently used lossless compression codecs reduces its compression power only slightly.
-Applying three-bit digitisation to TOD prior to lossless compression via FLAC or LZMA reduces the data volume by a further factor of $5.89$ or $17.34$, respectively.}
+No change in the results is observed for maps with different numbers of hits per pixel and no angular scale dependence was observed in the added noise.
+\changed{Combining few-bit digitisation with FLAC compression does not yield additional data volume reduction.}
 
@@
 
+\changed{We are indebted to Nathan Whitehorn for the details about SPT data compression provided by him.}
 We are grateful for insightful discussions about the prospects of extreme digitisation for CMB experiments with Andrew Melatos and Patrick Clearwater.
-We thank the \changed{referee as well as Nikhel Gupta, Srinivasan Raghunathan, Federico Bianchini, Andrew Melatos, and Patrick Clearwater for valuable feedback on the manuscript.}
+We thank the referee as well as Nikhel Gupta, Srinivasan Raghunathan, Federico Bianchini, Andrew Melatos, and Patrick Clearwater for valuable feedback on the manuscript.
 We acknowledge support from an Australian Research Council Future Fellowship (FT150100074), and also from the University of Melbourne. 
@@
   \caption[Current ]{
-  \changed{(FIGURE HAS BEEN CHANGED) Input power spectra used to generate the CMB template map plus detector noise level (black line) and digitisation noise (grey line); power spectra using three- (violet marker, right), two- (orange marker, middle), and one-bit (green marker, left) TOD.
+  Input power spectra used to generate the CMB template map plus detector noise level (black line) and digitisation noise (grey line); power spectra using three- (violet marker, right), two- (orange marker, middle), and one-bit (green marker, left) TOD.
 Power spectra are rescaled to $D_\ell = \ell (\ell+1) C_\ell /2\pi$ and bined to $\Delta \ell = 100$.
@@
 \label{fig:psrecover}
-}
+
 \end{figure*}
@@
 \begin{center}
-\caption{\label{tab:extranoisewhite} \changed{Fractional Noise Increase Due To Digitisation - Gaussian White Detector Noise}}
+\caption{\label{tab:extranoisewhite} Fractional Noise Increase Due To Digitisation - Gaussian White Detector Noise}
 \small
@@
 \tablecomments{ 
-\changed{(TABLE HAS BEEN CHANGED) Fractional addition to the map noise level, $\Delta \sigma / \sigma$, due to one-, two-, and three-bit digitisation for white detector noise with $1\sigma$ confidence intervals.
+Fractional addition to the map noise level, $\Delta \sigma / \sigma$, due to one-, two-, and three-bit digitisation for white detector noise with $1\sigma$ confidence intervals.
 Note that there is no appreciable variation between temperature and polarisation observations, or between different depths of observation.
 It is striking that three-bit digitisation only leads to a percent-level increase in the map noise level.}
-} \normalsize
+ \normalsize
 \end{center}
@@
 \begin{center}
-\caption{\label{tab:extranoiseoof} \changed{Fractional Noise Increase Due To Digitisation - $1/f$ Detector Noise}}
+\caption{\label{tab:extranoiseoof} Fractional Noise Increase Due To Digitisation - $1/f$ Detector Noise}
 \small
@@
 \tablecomments{ 
-\changed{(TABLE HAS BEEN CHANGED) Fractional addition to the map noise level, $\Delta \sigma / \sigma$, due to one-, two-, and three-bit digitisation for $1/f$ detector noise with $1\sigma$ confidence intervals.
+Fractional addition to the map noise level, $\Delta \sigma / \sigma$, due to one-, two-, and three-bit digitisation for $1/f$ detector noise with $1\sigma$ confidence intervals.
 As in the white noise case, we observe no significant trend between temperature and polarisation or for different depths of observation.
 We note that, even assuming a more realistic noise profile, three-bit digitisation only leads to a percent-level increase in the map noise level.}
-} \normalsize
+ \normalsize
 \end{center}
@@
 
-As mentioned in subsection \S\ref{subsec:extremedigitisation}, equation \ref{eq:distdef} does not demand \changed{that a digitisation scheme conserves power.
-However, it} is relatively simple to impose power conservation, since it boils down to rescaling the digitised output by a constant, $\gamma$.
+As mentioned in subsection \S\ref{subsec:extremedigitisation}, equation \ref{eq:distdef} does not demand that a digitisation scheme conserves power.
+However, it is relatively simple to impose power conservation, since it boils down to rescaling the digitised output by a constant, $\gamma$.
 We derive an expression for $\gamma$, such that a cross-spectrum of two digitised maps yields an unbiased estimate of the input power spectrum.
@@
 
-\changed{ where we sum over all output levels of both timestreams $(1, 2)$ and $P$ denotes the probability of an output level occurring.
+where we sum over all output levels of both timestreams $(1, 2)$ and $P$ denotes the probability of an output level occurring.
 For current-generation CMB experiments it is common to draw from $>10^5$ samples to obtain an estimate of a map pixel.
@@
 The central limit theorem ensures that the amplitude probability distribution function of the noise in each pixel is Gaussian. 
-The probabilities $P$ in the above expression can therefore be expressed as integrals over normal distributions, $p(\xi)$, with mean zero and standard deviation $\sigma$.}
+The probabilities $P$ in the above expression can therefore be expressed as integrals over normal distributions, $p(\xi)$, with mean zero and standard deviation $\sigma$.
 
@@
 &\times \left\{ \sum_{j=0}^N  \frac{y_j}{\sqrt{2}} \left[ \sum_{m = 0}^\infty \frac{(-1)^m}{m!
-(2m+1)} \frac{1}{(\sqrt{2}\sigma)^{2m+1}} \left( \sum_{l=0}^{2m+1} {2m+1 \choose l} \mu^l ( x_{j+1}^{2m+1-l} - x_{j}^{2m+1-l} ) \right) \right] \right\} \changed{,}
+(2m+1)} \frac{1}{(\sqrt{2}\sigma)^{2m+1}} \left( \sum_{l=0}^{2m+1} {2m+1 \choose l} \mu^l ( x_{j+1}^{2m+1-l} - x_{j}^{2m+1-l} ) \right) \right] \right\} ,
 \end{aligned}
@@
 
-\changed{where} we have used the Maclaurin series expansion of the error function and the binomial expansion. Moving the $i, j$ summation forwards yields
+where we have used the Maclaurin series expansion of the error function and the binomial expansion. Moving the $i, j$ summation forwards yields
 
@@
 
-\changed{because} $\psi_{N-i}^{k,n} = (-1)^{k} \psi_{i}^{k,n}$.
-\changed{Since ground-based CMB experiments operate at low signal-to-noise we calculate the leading order terms of the signal, $\mu$, in equation \ref{eq:dcompleteklsum}.
-We split the sum over $k, l$ in said equation as follows:}
+because $\psi_{N-i}^{k,n} = (-1)^{k} \psi_{i}^{k,n}$.
+Since ground-based CMB experiments operate at low signal-to-noise we calculate the leading order terms of the signal, $\mu$, in equation \ref{eq:dcompleteklsum}.
+We split the sum over $k, l$ in said equation as follows:
 
 \begin{equation}
-\sum_{i,j = 0}^N y_i y_j \psi_i^{k, n} \psi_j^{l, m} = \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} y_i y_j  \left( \psi_i^{k, n} \psi_j^{l, m} - \psi_{i}^{k, n} \psi_j^{l, m} - \psi_i^{k, n} \psi_{j}^{l, m} + \psi_{i}^{k, n} \psi_{j}^{l, m} \right) = 0 \changed{.}
+\sum_{i,j = 0}^N y_i y_j \psi_i^{k, n} \psi_j^{l, m} = \sum_{i = 0}^{N/2}\sum_{j = 0}^{N/2} y_i y_j  \left( \psi_i^{k, n} \psi_j^{l, m} - \psi_{i}^{k, n} \psi_j^{l, m} - \psi_i^{k, n} \psi_{j}^{l, m} + \psi_{i}^{k, n} \psi_{j}^{l, m} \right) = 0 .
 \end{equation}
@@
 
-where $c$ is some constants that \changed{depends} on the specific digitisation scheme chosen. Returning to equation \ref{eq:normcrosspower} we see
+where $c$ is some constants that depends on the specific digitisation scheme chosen. Returning to equation \ref{eq:normcrosspower} we see
 
@@
 (2n+1)} \frac{(-1)^{n+m}}{m!
-(2m+1)} \frac{1}{(\sqrt{2}\sigma)^{2n+2m+2}} \left( {2n+1 \choose 1} {2m+1 \choose 1} ( x_{i+1}^{2n} - x_{i}^{2n} ) ( x_{j+1}^{2m} - x_{j}^{2m} ) \right) \right] \changed{.}
+(2m+1)} \frac{1}{(\sqrt{2}\sigma)^{2n+2m+2}} \left( {2n+1 \choose 1} {2m+1 \choose 1} ( x_{i+1}^{2n} - x_{i}^{2n} ) ( x_{j+1}^{2m} - x_{j}^{2m} ) \right) \right] .
 \end{aligned}
@@
 \begin{equation}
-= \sum_{i,j=0}^N \frac{y_i y_j}{2\pi\sigma^2} \left\{ exp \left[ - \left(\frac{x_{i+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{i}}{\sqrt{2}\sigma} \right)^2 \right] \right\} \left\{ exp \left[ - \left(\frac{x_{j+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{j}}{\sqrt{2}\sigma} \right)^2 \right] \right\} \changed{.}
+= \sum_{i,j=0}^N \frac{y_i y_j}{2\pi\sigma^2} \left\{ exp \left[ - \left(\frac{x_{i+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{i}}{\sqrt{2}\sigma} \right)^2 \right] \right\} \left\{ exp \left[ - \left(\frac{x_{j+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{j}}{\sqrt{2}\sigma} \right)^2 \right] \right\} .
 \end{equation}
@@
 \begin{equation}
-\gamma = \left( \sum_{i,j=0}^N \frac{y_i y_j}{2\pi\sigma^2} \left\{ exp \left[ - \left(\frac{x_{i+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{i}}{\sqrt{2}\sigma} \right)^2 \right] \right\} \left\{ exp \left[ - \left(\frac{x_{j+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{j}}{\sqrt{2}\sigma} \right)^2 \right] \right\} \right)^{-1/2} \changed{.}
+\gamma = \left( \sum_{i,j=0}^N \frac{y_i y_j}{2\pi\sigma^2} \left\{ exp \left[ - \left(\frac{x_{i+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{i}}{\sqrt{2}\sigma} \right)^2 \right] \right\} \left\{ exp \left[ - \left(\frac{x_{j+1}}{\sqrt{2}\sigma} \right)^2 \right] - exp \left[ - \left(\frac{x_{j}}{\sqrt{2}\sigma} \right)^2 \right] \right\} \right)^{-1/2} .
 \end{equation}
 
-\changed{This can be calculated for a given digitisation scheme, specified by $N, x_i, y_i$, and standard deviation of the noise, $\sigma$.
-Notice that if the digitisation scheme is chosen such that the digitisation thresholds and output levels depend linearly on the noise level, the normalisation constant is independent of the noise level.}
+This can be calculated for a given digitisation scheme, specified by $N, x_i, y_i$, and standard deviation of the noise, $\sigma$.
+Notice that if the digitisation scheme is chosen such that the digitisation thresholds and output levels depend linearly on the noise level, the normalisation constant is independent of the noise level.
 

### 2019-05-29 mnras_template.tex
Add files via upload

--- 2019-05-17 digitisation.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[apj]{emulateapj}
-%\documentclass[preprint2]{aastex61}
-%\documentclass[12pt,preprint]{aastex}
-\graphicspath{{figures/}}
-\DeclareGraphicsExtensions{.jpg,.pdf,.png,.eps,.ps}
+% mnras_template.tex
+%
+% LaTeX template for creating an MNRAS paper
+%
+% v3.0 released 14 May 2015
+% (version numbers match those of mnras.cls)
+%
+% Copyright (C) Royal Astronomical Society 2015
+% Authors:
+% Keith T. Smith (Royal Astronomical Society)
 
-\usepackage[table,usenames,dvipsnames]{xcolor}
-\usepackage{amsmath}
-\usepackage{subfigure}
-\usepackage[backref,breaklinks,colorlinks,citecolor=blue]{hyperref}
-\usepackage{natbib}
-%\usepackage{natbib}
-\bibliographystyle{fapj}
-\usepackage{graphicx}
+% Change log
+%
+% v3.0 May 2015
+%    Renamed to match the new package name
+%    Version number matches mnras.cls
+%    A few minor tweaks to wording
+% v1.0 September 2013
+%    Beta testing only - never publicly released
+%    First version: a simple (ish) template for creating an MNRAS paper
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Basic setup. Most papers should leave these options alone.
+\documentclass[a4paper,fleqn,usenatbib]{mnras}
+
+% MNRAS is set in Times font. If you don't have this installed (most LaTeX
+% installations will be fine) or prefer the old Computer Modern fonts, comment
+% out the following line
+\usepackage{newtxtext,newtxmath}
+% Depending on your LaTeX fonts installation, you might get better results with one of these:
+%\usepackage{mathptmx}
+%\usepackage{txfonts}
+
 \usepackage{multirow}
-\usepackage{soul}
 
-%\newcommand{\jcap}{JCAP}
+% Use vector fonts, so it zooms properly in on-screen viewing software
+% Don't change these lines unless you know what you are doing
+\usepackage[T1]{fontenc}
+\usepackage{ae,aecompl}
 
@@
 \newcommand{\removed}[1]{\textcolor{Red}{}}
-\include{number_list}
 
-%
+%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%
 
-% ref to section \S\ref{sec:label}
+% Only include extra packages if you really need them. Common packages are:
+\usepackage{graphicx}	% Including figure files
+\usepackage{amsmath}	% Advanced maths commands
+\usepackage{amssymb}	% Extra maths symbols
 
-%\submitjournal{ApJ}
-\def\Melbourne{1}
-\def\uci{2}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
+
+% Please keep new commands to a minimum, and use \newcommand not \def to avoid
+% overwriting existing commands. Example:
+%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%
+
+% Title of the paper, and the short title which is used in the headers.
+% Keep the title short and informative.
+\title[Short title, max. 45 characters]{Extreme Digitisation For Ground-Based Cosmic Microwave Background Experiments}
+
+% The list of authors, and the short list which is used in the headers.
+% If you need two or more lines of authors, add an extra line using \newauthor
+\author[L. Balkenhol and C. L. Reichardt.]{
+L. Balkenhol,$^{1}$\thanks{E-mail: lbalkenhol@student.unimelb.edu.au}
+C. L. Reichardt,$^{1}$
+\\
+% List of institutions
+$^{1}$School of Physics, University of Melbourne, Parkville, VIC 3010, Australia\\
+}
+
+% These dates will be filled out by the publisher
+\date{Accepted XXX. Received YYY; in original form ZZZ}
+
+% Enter the current year, for the copyright statements etc.
+\pubyear{2019}
+
+% Don't change these lines
 \begin{document}
+\label{firstpage}
+\pagerange{\pageref{firstpage}--\pageref{lastpage}}
+\maketitle
 
-\title{Extreme Digitisation For Ground-Based Cosmic Microwave Background Experiments}
-\author{L.~Balkenhol\altaffilmark{\Melbourne} and C.~L.~Reichardt\altaffilmark{\Melbourne}}
-\altaffiltext{\Melbourne}{School of Physics, University of Melbourne, Parkville, VIC 3010, Australia}
-\email{lbalkenhol@student.unimelb.edu.au}
-
+% Abstract of the paper
 \begin{abstract}
-
 The large size of the time ordered data of cosmic microwave background experiments presents challenges for mission planning and data analysis.
@@
 The digitisation noise is independent of angular scale and is well-described as a constant percentage of the original detector noise.
-Three bit digitisation increases the map noise level by $< 2\%$, while reducing the data volume by a factor of \changed{ten relative to 32-bit floats.
-Lossless compression of the digitised data does not lead to additional data volume reduction.} Extreme digitisation is a promising strategy for upcoming experiments.
-
+Three bit digitisation increases the map noise level by $< 2\%$, while reducing the data volume by a factor of ten relative to 32-bit floats.
+Extreme digitisation is a promising strategy for upcoming experiments.
 \end{abstract}
 
-%C: maybe sub techniques:misc for methods: data analysis ? unfortunately no data compression tag
-\keywords{cosmic background radiation --- polarization --- techniques: miscellaneous}
+% Select between one and six entries from the list of approved keywords.
+% Don't make up new ones.
+\begin{keywords}
+cosmic background radiation -- polarization -- techniques: miscellaneous
+\end{keywords}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%
+
 \section{Introduction}
@@
 Differences between few bit and the input power spectra are due to observing a single CMB realisation and residual boundary effects.
-\changed{Non-linearity introduced by the digitisation process disappears in the final map due to the large number of TOD points that contribute to each map pixel.
-Thus the digitisation effects are not an issue for power spectrum estimation.}
+Non-linearity introduced by the digitisation process disappears in the final map due to the large number of TOD points that contribute to each map pixel.
+Thus the digitisation effects are not an issue for power spectrum estimation.
 
@@
 
-\changed{Current-generation ground-based CMB experiments employ a combination of lossly, e.g.
+Current-generation ground-based CMB experiments employ a combination of lossy, e.g.
 downsampling, and lossless compression, e.g.
@@
 We test whether the data volume of timestreams that have undergone few-bit digitisation can be reduced further by applying the most popular lossless compression algorithm: FLAC.
-}
 
-\changed{We obtain 2 minutes of SPT-3G TOD and apply the digitisation schemes in equations \ref{eq:1bit}-\ref{eq:3bit} before compressing the data further using FLAC.
-We compare the file size of FLAC compressed few-bit timestreams to the typical data reduction rate achieved for 24-bit SPT-3G data.}
+We obtain 2 minutes of SPT-3G TOD and apply the digitisation schemes in equations \ref{eq:1bit}-\ref{eq:3bit} before compressing the data further using FLAC.
+We compare the file size of FLAC compressed few-bit timestreams to the typical data reduction rate achieved for 24-bit SPT-3G data.
 
-\changed{We find that using FLAC in conjunction with extreme digitisation does not yield additional data volume reduction.
+We find that using FLAC in conjunction with extreme digitisation does not yield additional data volume reduction.
 For 3-bit digitisation, both the FLAC and raw timestream have $\sim$3 bits per data point.
@@
 For comparison, FLAC reduces the full 24-bit  SPT-3G data to $6-8$ bits of entropy per number.
-For this real-world case, three-bit digitisation reduces the data volume by a factor of $2-3$.}
+For this real-world case, three-bit digitisation reduces the data volume by a factor of $2-3$ more than lossless compression.
 
-\section{Conclusion}
+\section{Conclusions}
 \label{sec:conclusions}
@@
 No change in the results is observed for maps with different numbers of hits per pixel and no angular scale dependence was observed in the added noise.
-\changed{Combining few-bit digitisation with FLAC compression does not yield additional data volume reduction.}
+Applying lossless compression techniques like FLAC to few-bit digitised data does not yield further reductions in data volume.
 
@@
 
-\acknowledgments % thank patrick for discussion/starting point? Andrew?
+\section*{Acknowledgements}
 
-% Melbourne CMB group, funding, NERSC.
-
-\changed{We are indebted to Nathan Whitehorn for the details about SPT data compression provided by him.}
+We are indebted to Nathan Whitehorn for the details about SPT data compression provided by him.
 We are grateful for insightful discussions about the prospects of extreme digitisation for CMB experiments with Andrew Melatos and Patrick Clearwater.
@@
 We acknowledge the use of the Legacy Archive for Microwave Background Data Analysis (LAMBDA). Support for LAMBDA is provided by the NASA Office of Space Science.
-
-% all code stuff
-
 This research made use of the NumPy \citep{numpy}, SciPy \citep{scipy}, Matplotlib \citep{matplotlib}, and Astropy \citep{astropy} packages.
 
-\newpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% APPENDIX
+%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%
+
+% The best way to enter references is to use BibTeX:
+
+\bibliographystyle{mnras}
+\bibliography{digitisation} % if your bibtex file is called example.bib
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%
+
+\onecolumn
+
 \appendix
 
-\begin{figure*}[h!t]\centering
+\section*{Appendix}
+
+% result tables
+%\subsection{Digitisation Noise Levels}
+%\label{subsec:appendixnoisetables}
+
+\begin{figure*}\centering
 \vspace{-1.5cm}
@@
 
-% result tables
-\subsection{Digitisation Noise Levels}
-\label{subsec:appendixnoisetables}
-
 \def\arraystretch{1.3}
-\begin{table*}[tbh]
+\begin{table*}
 \begin{center}
-\caption{\label{tab:extranoisewhite} Fractional Noise Increase Due To Digitisation - Gaussian White Detector Noise}
+\caption{ \label{tab:extranoisewhite} Fractional Noise Increase Due To Digitisation - Gaussian White Detector Noise}
 \small
@@
 \end{tabular}
-\tablecomments{ 
+\vspace{1mm}
+\description{ 
 Fractional addition to the map noise level, $\Delta \sigma / \sigma$, due to one-, two-, and three-bit digitisation for white detector noise with $1\sigma$ confidence intervals.
@@
 \def\arraystretch{1.3}
-\begin{table*}[tbh]
+\begin{table*}
 \begin{center}
@@
 \end{tabular}
-\tablecomments{ 
+\vspace{1mm}
+\description{ 
 Fractional addition to the map noise level, $\Delta \sigma / \sigma$, due to one-, two-, and three-bit digitisation for $1/f$ detector noise with $1\sigma$ confidence intervals.
@@
 \begin{equation}
-\begin{aligned}
-\langle D(\mu + \xi_1) D(\mu + \xi_2) \rangle &= \left\{ \sum_{i=0}^N  \frac{y_i}{2} \left[ erf \left( \frac{x_{i+1} - \mu}{\sqrt{2}\sigma} \right) - erf \left( \frac{x_{i} - \mu}{\sqrt{2}\sigma} \right) \right] \right\} \\
-&\times \left\{  \sum_{j=0}^N \frac{y_j}{2} \left[ erf \left( \frac{x_{j+1} - \mu}{\sqrt{2}\sigma} \right) - erf \left( \frac{x_{j} - \mu}{\sqrt{2}\sigma} \right) \right] \right\}.
-\end{aligned}
+\langle D(\mu + \xi_1) D(\mu + \xi_2) \rangle = \left\{ \sum_{i=0}^N  \frac{y_i}{2} \left[ \mathrm{erf} \left( \frac{x_{i+1} - \mu}{\sqrt{2}\sigma} \right) - \mathrm{erf} \left( \frac{x_{i} - \mu}{\sqrt{2}\sigma} \right) \right] \right\} \times \left\{  \sum_{j=0}^N \frac{y_j}{2} \left[ \mathrm{erf} \left( \frac{x_{j+1} - \mu}{\sqrt{2}\sigma} \right) - \mathrm{erf} \left( \frac{x_{j} - \mu}{\sqrt{2}\sigma} \right) \right] \right\}.
 \end{equation}
@@
 
-%\newpage
+% Don't change these lines
+\bsp	% typesetting comment
+\label{lastpage}
+\end{document}
 
-\bibliography{digitisation}
-
-\end{document}
+% End of mnras_template.tex
